### 2016-02-09 sc-principles/software-citation-principles.tex
added mostly blank document for software citation principles

\documentclass[11pt, oneside]{amsart}
\pdfoutput=1

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{longtable}
\usepackage{verbatim}
\usepackage{booktabs,multicol}

% The following three lines are used for displaying footnote in tables.
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage[bookmarks, bookmarksopen, bookmarksnumbered]{hyperref}
\usepackage[all]{hypcap}
\urlstyle{rm}

\title{Software Citation Principles}

\author{}

\date{}

\begin{abstract}
\end{abstract}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction \& motivation}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Use cases}
\label{sec:use_cases}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}
\label{sec:related_work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Software citation principles}
\label{sec:principles}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation examples}
\label{sec:examples}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2016-02-10 sc-principles/software-citation-principles.tex
moving abstract into document to allow compilation to work
adding initial motivation/intro, and initial set of principles can flesh out motivation more once other sections are more in place

@@
 
+\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
+
 \title{Software Citation Principles}
@@
 
+\begin{document}
+\maketitle
+
 \begin{abstract}
 \end{abstract}
-
-\begin{document}
-\maketitle
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+As the process of science has become increasingly digital, scientific outputs and products have
+grown beyond simple papers and books to include software, data, and other electronic
+components.  Scientific knowledge is embedded in these components.  And papers and books
+themselves are also becoming increasingly digital, allowing them to executable
+and reproducible.  As we move towards this future where science is performed in and recorded
+as a variety of linked digital products, the characteristics and properties that developed for
+books and papers need to be applied to all digital products.  Here, we are concerned specifically
+with the citation of software products.
+
+In the next section (\S\ref{sec:use_cases}), we provide some detailed context in which
+software citation is important, by means of use cases.  In \S\ref{sec:related_work}, we
+summarize and analyze a large amount of previous work and thinking in this area.  In
+\S\ref{sec:principles}, we provide set of guiding principles for citation of software within
+scholarly literature, other software dataset, or any other research object.  Finally,
+in \S\ref{sec:examples} we give examples of how these software citation principles
+could be applied, related back to the use cases in \S\ref{sec:use_cases}.
 
@@
 
+\katznote{The initial context of this section is liberally borrowed and loosely adapted from \url{https://www.force11.org/group/joint-declaration-data-citation-principles-final}}
+
+\begin{enumerate}
+\item \textbf{Importance}:
+Software should be considered a legitimate, citable product of research.
+Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
+\item \textbf{Credit and Attribution}
+\begin{enumerate}
+\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.
+\item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.
+\end{enumerate}
+\item \textbf{Unique Identification}
+A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
+\item \textbf{Access}
+Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.
+\item \textbf{Persistence}
+Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
+\item \textbf{Specificity and Verifiability}
+Software citations should facilitate identification of, access to, and verification of the specific software that support a claim.
+Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
+\item \textbf{Interoperability and Flexibility}
+Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
+\end{enumerate}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-02-10 sc-principles/software-citation-principles.tex
adding bib doc and cite to data citation principles, fixing order of abstract and maketitle...

@@
 \begin{document}
-\maketitle
 
@@
 \end{abstract}
+
+\maketitle
 
@@
 
-\katznote{The initial context of this section is liberally borrowed and loosely adapted from \url{https://www.force11.org/group/joint-declaration-data-citation-principles-final}}
+\katznote{The initial context of this section is liberally borrowed and loosely adapted from \cite{data-citation-principles}}
 
@@
 
+\bibliographystyle{abbrv}
+\bibliography{software-citation-principles}
+
 \end{document}

### 2016-02-10 sc-principles/software-citation-principles.tex
initial table with option for presenting use cases

--- 2016-02-09 sc-principles/software-citation-principles.tex
@@
 
+\usepackage{enumitem}
+\setlist{leftmargin=7mm}
+
 \usepackage[bookmarks, bookmarksopen, bookmarksnumbered]{hyperref}
@@
 
-\title{Software Citation Principles}
+% 15 characters / 2.5 cm => 100 characters / line
+% Using 11 pt => 94 characters / line
+\setlength{\paperwidth}{216 mm}
+% 6 lines / 2.5 cm => 55 lines / page
+% Using 11pt => 48 lines / pages
+\setlength{\paperheight}{279 mm}
+\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
+% You can use a baselinestretch of down to 0.9
+\renewcommand{\baselinestretch}{0.96}
+
+\title[]{Software Citation Principles}
 
@@
 
+\begin{document}
 \begin{abstract}
@@
 
-\begin{document}
 \maketitle
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[htbp]
+\caption{Use cases for software citation}
+\centering
+\scriptsize\setlength{\tabcolsep}{2.5pt}
+\begin{tabular}{@{}l l c c c c c c c c@{}}
+\toprule
+ & & \multicolumn{6}{c}{Requirements} \\
+ \cmidrule{3-10}
+\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{DOI} & Indexed & \multirow{2}{*}{Role} \\
+& & name &  &  &  date & repository &  & citations & \\
+\midrule
+Researcher & someone else's software for a paper	& \textbullet	& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & & \\
+Researcher & someone else's software for new software& \textbullet	& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & & \\
+Researcher & contribute to software 				& \textbullet	& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & & \textbullet \\
+Researcher & find citations of software 			& \textbullet &   &   &   &   & \textbullet & \textbullet & \\
+Researcher & credit for software development		& \textbullet & \textbullet &   & \textbullet & \textbullet & \textbullet & & \textbullet \\
+Researcher & ``reproduce'' analysis 				& \textbullet &   & \textbullet & \textbullet & \textbullet & \textbullet & & \\
+Researcher & benchmark software 					& \textbullet &   & \textbullet & \textbullet & \textbullet & \textbullet & & \\
+Researcher & find software to implement task 		& \textbullet & \textbullet &   &   & \textbullet & \textbullet & \textbullet & \\
+Publisher & publish software paper					& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & & \\
+Publisher & publish papers that cite software 		& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \\
+Indexer & build catalog of software 				& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \\
+Domain group & build catalog of software 			& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &  & \\
+Library\slash archive & build catalog of software 	& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &  & \\
+Repository & show scientific impact of holdings 	& \textbullet &   &   &   &   & \textbullet & \textbullet & \\
+Funder & show how funded software has been used		& \textbullet &   &   &   &   & \textbullet & \textbullet & \\
+\bottomrule
+\end{tabular}
+\label{tab:use_cases}
+\end{table}%
+
+Meaning of ``researcher'' includes academic researcher (e.g., postdoc, tenure-track faculty member) and research software engineer.
+``Reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility.
+Examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search.
+Domain groups include ASCL, bioCADDIE.
+
+Alternate metadata
+\begin{itemize}
+\item version number and release date: download date
+\item location\slash respository: contact name\slash email if not publicly available
+\end{itemize}
 

### 2016-02-10 sc-principles/software-citation-principles.tex

--- 2016-02-10 sc-principles/software-citation-principles.tex
@@
 \newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
+\newcommand{\niemeyernote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
+
+% 15 characters / 2.5 cm => 100 characters / line
+% Using 11 pt => 94 characters / line
+\setlength{\paperwidth}{216 mm}
+% 6 lines / 2.5 cm => 55 lines / page
+% Using 11pt => 48 lines / pages
+\setlength{\paperheight}{279 mm}
+\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
+% You can use a baselinestretch of down to 0.9
+\renewcommand{\baselinestretch}{0.96}
 

### 2016-02-11 sc-principles/software-citation-principles.tex
added ref to Google Doc of use cases
minor changes is wording below table 1, also testing sharelatex sync

@@
 \begin{table}[htbp]
-\caption{Use cases for software citation}
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}}
 \centering
@@
 
-Meaning of ``researcher'' includes academic researcher (e.g., postdoc, tenure-track faculty member) and research software engineer.
+``Researcher'' includes academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers.
 ``Reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility.
 Examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search.
-Domain groups include ASCL, bioCADDIE.
+Domain groups include ASCL, bioCADDIE, CIG, etc.
 

### 2016-02-11 sc-principles/software-citation-principles.tex
lots of stuff, mostly related to use case table, also first cut at authors

@@
 
-\author{}
+\author{Force11 Software Citation Working Group (Smith Katz Niemeyer eds)}
 
@@
 
+We have documented and analyzed a set of use cases related to software citation in \cite{SC-Use-Cases}.
+Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case.
+\katznote{not sure if this next long sentence should be here, should be in the table caption, or should be footnotes to the table, or even more than one of these} In the table,
+``researcher'' includes academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers
+``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software; ``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
+examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;  
+domain groups include ASCL, bioCADDIE, CIG, etc.;
+``repository'' is ???; and
+``funder'' is a group that funds software or work using software.
+
 \begin{table}[htbp]
 \caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}}
+\katznote{I changed DOI to UID, meaning unique ID - I think we probably want to be general here}
 \centering
@@
  \cmidrule{3-10}
-\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{DOI} & Indexed & \multirow{2}{*}{Role} \\
+\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
 & & name &  &  &  date & repository &  & citations & \\
@@
 
-``Researcher'' includes academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers.
-``Reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility.
-Examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search.
-Domain groups include ASCL, bioCADDIE, CIG, etc.
-
 Alternate metadata
@@
 
-\katznote{The initial context of this section is liberally borrowed and loosely adapted from \cite{data-citation-principles}}
+\katznote{The current content of this section is liberally borrowed and loosely adapted from \cite{data-citation-principles}.
+We we will want to add some text here explaining that we started with that document, and then modified it to take into account the information from the use cases and the related work, once we do so.}
 

### 2016-02-14 sc-principles/software-citation-principles.tex
cleaned up use cases section a bit

@@
 
-\usepackage{color}
+\usepackage{xcolor}
 \usepackage{dcolumn}
@@
 \newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
-\newcommand{\niemeyernote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
+\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
 
@@
 Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case. 
-\katznote{not sure if this next long sentence should be here, should be in the table caption, or should be footnotes to the table, or even more than one of these} In the table,
-``researcher'' includes academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers
-``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software; ``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
+\katznote{not sure if this next long sentence should be here, should be in the table caption, or should be footnotes to the table, or even more than one of these}
+\niemnote{Not sure either; definitely should include in the text due to length. Maybe an abbreviated version in the caption?}
+In the table,
+``researcher'' includes both academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers;
+``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software;
+``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
 examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search; 
-domain groups include ASCL, bioCADDIE, CIG, etc.;
-``repository'' is ???; and
-``funder'' is a group that funds software or work using software.
+domain groups include bioCADDIE, CIG, etc.;
+``repository'' refers to open-source software repositories such as ASCL, GitHub, Netlib, CRAN, institutional repositories; 
+``funder'' is a group that funds software or work using software; and 
+``UID'' means unique ID, such as DOI or PURL.
 
@@
 \caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}}
-\katznote{I changed DOI to UID, meaning unique ID - I think we probably want to be general here}
 \centering
@@
 
-Alternate metadata
+If certain metadata are not available, alternates may be provided:
 \begin{itemize}

### 2016-02-16 sc-principles/software-citation-principles.tex
tweaks to authors, adding new paragraph in intro

@@
 
-\author{Force11 Software Citation Working Group (Smith Katz Niemeyer eds)}
+\author{Force11 Software Citation Working Group (Editors: Arfon M. Smith, Daniel S. Katz, Kyle E. Niemeyer)}
 
@@
 with the citation of software products.
+
+The Force11 Software Citation Working Group~\cite{f11scwg} has been studying this,
+beginning in ...  with a number of meetings ... and has prepared this document for
+presentation at Force2016~\cite{force2016} and further discussion.  We expect that
+this discussion may lead to a second, final version of this document, and we also
+plan to have a follow-on working group that will work with stakeholders to ensure that
+these principles impact the scientific process.
 

### 2016-02-17 sc-principles/software-citation-principles.tex
Mostly notes describing areas to highlight

@@
 
+With close to 50 working group partipants representing a range of research domains, the working group was tasked to
+document existing practices in their respective communities. A total of 47 documents were submitted by working group
+participants with the life sciences, astrophysics and geosciences being particularly well-represented in the submitted
+resources.
+
+\textbf{Breakdown of resources}
+
+NOTE: Should mention that a number of communities have already focussed their attention on issues around data citation and while
+this isn't exactly the same there are many similarities (what is different exactly?).
+
+\textbf{General community/non domain-specific efforts}
+
+Some of the most actionable work has come from the SSI in the form of blog posts written by their community fellows:
+
+In a blog post from 2012, \cite{ssi-how-to-cite} discussing some of the pitfalls of trying to cite software in publications.
+The author includes useful guidance for when to consider citing software as well as some ways to help `convince' journal editors
+to allow the inclusion of software citations.
+
+In \cite{ssi-citation-files}, Wilson suggests that software authors include a `CITATION' file that documents exactly how the
+authors of the software would like to be cited by others. While this isn't a formal metadata specification (it's not machine
+readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.
+In a later post on the SSI blog, Jackson \cite{ssi-how-shalt-i-cite-thee} gives a good overview of some of the approaches
+package authors have taken to automate the generation of citation entities such as bibtex entries.
+
+NSF/NIH workshops. WSSSPE. JISC reports.
+
+\textbf{Domain-specific community activities}
+
+Astronomy: ASCL, AAS journals announcing they will accept software papers in their journals \cite{aas-sofware-papers}. AAS/GitHub
+workshop \cite{aas-software-index} dedicated to software citation, indexing and discoverability.
+
+....lots more to add here.
+
+\textbf{Existing principles}
+
+Author guidelines in F1000Research, APA authors, SSI `publish or be damned' \cite{ssi-publish-or-be-damned} manifesto.
+
+\textbf{Existing efforts around metadata standards}
+
+AAS workshop (\cite{aas-software-index}), CRediT project \cite{casrai-credit}, others...
+
+\textbf{Studies of author motivations}
+
+Survey of data citation practices \cite{Kratz_2015}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-02-18 sc-principles/software-citation-principles.tex
adding an Arfon-note command and using it

@@
 \newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
+\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}}
 
@@
 
-With close to 50 working group partipants representing a range of research domains, the working group was tasked to
+With close to 50 working group participants representing a range of research domains, the working group was tasked to
 document existing practices in their respective communities. A total of 47 documents were submitted by working group
@@
 
-NOTE: Should mention that a number of communities have already focussed their attention on issues around data citation and while
-this isn't exactly the same there are many similarities (what is different exactly?).
+\asnote{Should mention that a number of communities have already focussed their attention on issues around data citation and while
+this isn't exactly the same there are many similarities (what is different exactly?).}
 

### 2016-02-19 sc-principles/software-citation-principles.tex
Adding some more bibtex entries and notes

@@
 
-\textbf{Breakdown of resources}
+\textbf{Summary of submitted resources}
 
@@
 
-....lots more to add here.
+Life sciences/NIH: Software Discovery Index Meeting Report \cite{software-discovery-index} a cross-disciplinary meeting
+but with a largely NIH-focus. Suggested minimal information about software (MIAS) standard. Discusses how a common set of metadata
+fields are critical for useful indexing - suggested candidates for this list.
+
+Geosciences: Ontosoft \cite{ontosoft} - A Community Software Commons for the Geosciences. Lots of attention given to the metadata
+required to describe, discover and execute research software. NSF geosciences workshop around data lifecycle, management and citation
+\cite{nsf-geo-data}. This report includes many recommendations for data citation.
 
@@
 
-AAS workshop (\cite{aas-software-index}), CRediT project \cite{casrai-credit}, others...
+AAS workshop (\cite{aas-software-index}), CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
 

### 2016-02-23 sc-principles/software-citation-principles.tex
More domain-specific activities

@@
 
-\textbf{General community/non domain-specific efforts}
+\textbf{General community/non domain-specific activities}
 
@@
 
-NSF/NIH workshops. WSSSPE. JISC reports.
+Perhaps in recognition of the broad range of research domains struggling with the challenge of better recognizing the role of sofware, a number of community efforts hosted (and sponsored) by funders and agencies in both the US (NSF, NIH, Alfred P Sloan Foundation) and UK (SFTC, JISC) have run a number of workshops with participants from across a range of disciplines.
+
+\asnote{Should also discuss the WSSSPE \cite{wssspe} series of workshops in this section}
 
@@
 
-Astronomy: ASCL, AAS journals announcing they will accept software papers in their journals \cite{aas-sofware-papers}. AAS/GitHub
+One approach to increasing the `citability' of software is to encourage the submission of papers in standard journals describing a piece of research software.
+While some journals (e.g.
+F1000Research) have traditionally accepted software submissions, the American Astronomical Society (AAS) have recently annouced they will accept software papers in their journals \cite{aas-sofware-papers}.
+
+In astronomy & astrophysics: ASCL \cite{ascl}, AAS journals  AAS/GitHub
 workshop \cite{aas-software-index} dedicated to software citation, indexing and discoverability.
+
+\asnote{Perhaps we should say something here about the potential importance of professional societies and their role in effecting changes in community practice?}
 
@@
 Geosciences: Ontosoft \cite{ontosoft} - A Community Software Commons for the Geosciences. Lots of attention given to the metadata
-required to describe, discover and execute research software. NSF geosciences workshop around data lifecycle, management and citation
-\cite{nsf-geo-data}. This report includes many recommendations for data citation.
+required to describe, discover and execute research software.
+NSF geosciences workshop around data lifecycle, management and citation \cite{nsf-geo-data}.
+This report includes many recommendations for data citation.
 

### 2016-02-23 sc-principles/software-citation-principles.tex
small fix to make this build

@@
 
-In astronomy & astrophysics: ASCL \cite{ascl}, AAS journals  AAS/GitHub
+In astronomy \& astrophysics: ASCL \cite{ascl}, AAS journals  AAS/GitHub
 workshop \cite{aas-software-index} dedicated to software citation, indexing and discoverability.

### 2016-02-23 sc-principles/software-citation-principles.tex
adding a bit more to the intro

@@
 
-The Force11 Software Citation Working Group~\cite{f11scwg} has been studying this,
-beginning in ...  with a number of meetings ... and has prepared this document for
-presentation at Force2016~\cite{force2016} and further discussion.  We expect that
+The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015.
+It gathered members, and then in June 2015 began studying this, with a number of meetings
+and some off-line work by group members to gather materials documenting existing
+practices in member disciplines; gather materials from workshops and other reports;
+review those materials, identifying overlaps and differences; and draft this resulting document.
+This document will be presented and discussed at the Force2016 Conference \cite{force2016}
+in April 2016.  We expect that
 this discussion may lead to a second, final version of this document, and we also

### 2016-02-23 sc-principles/software-citation-principles.tex
Oxfordizing a comma (oxfordising?)

@@
 document existing practices in their respective communities. A total of 47 documents were submitted by working group
-participants with the life sciences, astrophysics and geosciences being particularly well-represented in the submitted
+participants with the life sciences, astrophysics, and geosciences being particularly well-represented in the submitted
 resources.

### 2016-02-24 sc-principles/software-citation-principles.tex
rewording slightly to avoid the multiples of the same word; adding missing words

@@
 practices in member disciplines; gather materials from workshops and other reports;
-review those materials, identifying overlaps and differences; and draft this resulting document.
-This document will be presented and discussed at the Force2016 Conference \cite{force2016}
+review those materials, identifying overlaps and differences; and draft this resulting document,
+which will be presented and discussed at the Force2016 Conference \cite{force2016}
 in April 2016.  We expect that
-this discussion may lead to a second, final version of this document, and we also
-plan to have a follow-on working group that will work with stakeholders to ensure that
+this discussion may lead to a second, final version, and we also
+plan to have a follow-on working group that will interact with stakeholders to ensure that
 these principles impact the scientific process.
@@
 summarize and analyze a large amount of previous work and thinking in this area.  In
-\S\ref{sec:principles}, we provide set of guiding principles for citation of software within
-scholarly literature, other software dataset, or any other research object.  Finally,
+\S\ref{sec:principles}, we provide a set of guiding principles for citation of software within
+scholarly literature, other software, datasets, or any other research object.  Finally,
 in \S\ref{sec:examples} we give examples of how these software citation principles

### 2016-02-24 sc-principles/software-citation-principles.tex
add note

@@
 \newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}}
+\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}}
 
@@
 \katznote{not sure if this next long sentence should be here, should be in the table caption, or should be footnotes to the table, or even more than one of these}
-\niemnote{Not sure either; definitely should include in the text due to length. Maybe an abbreviated version in the caption?}
+\niemnote{Not sure either; definitely should include in the text due to length.
+Maybe an abbreviated version in the caption?}\flnote{If these ``definitions'' are not used anywhere else than in the table (and they don't really seem to be so far), then I tend towards putting this in the caption.
+That's where they are needed (in the table).
+I don't like footnotes for something like this.}
 In the table,

### 2016-02-24 sc-principles/software-citation-principles.tex
add another note

@@
 \begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}}
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{This table is a little hard to read: it is hard to follow which 'dots' belong to which lines.
+What about using a alternating (light) background or such to help?
+Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?}}
 \centering

### 2016-02-25 sc-principles/software-citation-principles.tex
spell out SSI at least once

@@
 this discussion may lead to a second, final version, and we also
-plan to have a follow-on working group that will interact with stakeholders to ensure that
+plan to have a follow-on working group that will work with stakeholders to ensure that
 these principles impact the scientific process.
@@
 
-Some of the most actionable work has come from the SSI in the form of blog posts written by their community fellows:
+Some of the most actionable work has come from the NSF Software Infrastructure for Sustained Innovation (SSI) program in the form of blog posts written by their community fellows:
 

### 2016-03-01 sc-principles/software-citation-principles.tex
Minor corrections

@@
 \item version number and release date: download date
-\item location\slash respository: contact name\slash email if not publicly available
+\item location\slash repository: contact name\slash email if not publicly
+  available
 \end{itemize}
@@
 
-Perhaps in recognition of the broad range of research domains struggling with the challenge of better recognizing the role of sofware, a number of community efforts hosted (and sponsored) by funders and agencies in both the US (NSF, NIH, Alfred P Sloan Foundation) and UK (SFTC, JISC) have run a number of workshops with participants from across a range of disciplines.
+Perhaps in recognition of the broad range of research domains struggling with
+the challenge of better recognizing the role of software, a number of community
+efforts hosted (and sponsored) by funders and agencies in both the US (NSF,
+NIH, Alfred P Sloan Foundation) and UK (SFTC, JISC) have run a number of
+workshops with participants from across a range of disciplines.
 
@@
 
-One approach to increasing the `citability' of software is to encourage the submission of papers in standard journals describing a piece of research software.
-While some journals (e.g.
-F1000Research) have traditionally accepted software submissions, the American Astronomical Society (AAS) have recently annouced they will accept software papers in their journals \cite{aas-sofware-papers}.
+One approach to increasing the `citability' of software is to encourage the
+submission of papers in standard journals describing a piece of research
+software. While some journals (e.g. F1000Research) have traditionally accepted
+software submissions, the American Astronomical Society (AAS) have recently
+announced they will accept software papers in their journals
+\cite{aas-sofware-papers}.
 

### 2016-03-01 sc-principles/software-citation-principles.tex
Add zebra stripes to table 1 using the xcolor package

--- 2016-02-25 sc-principles/software-citation-principles.tex
@@
 
-\usepackage{xcolor}
+\usepackage[table]{xcolor}
 \usepackage{dcolumn}
@@
 
+\rowcolors{5}{gray!25}{white}
 \begin{table}[htbp]
@@
 Researcher            & someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher & someone else's software for new software& \textbullet	& \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & & \\
+Researcher            & someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
 Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\

### 2016-03-01 sc-principles/software-citation-principles.tex
made alternating row colors start with white
removed resolved comment from FL

@@
 
-\rowcolors{5}{gray!25}{white}
+\rowcolors{5}{white}{gray!25}
 \begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{This table is a little hard to read: it is hard to follow which 'dots' belong to which lines.
-What about using a alternating (light) background or such to help?
-Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?}}
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?}}
 \centering

### 2016-03-01 sc-principles/software-citation-principles.tex
minor grammar

@@
 the challenge of better recognizing the role of software, a number of community
-efforts hosted (and sponsored) by funders and agencies in both the US (NSF,
-NIH, Alfred P Sloan Foundation) and UK (SFTC, JISC) have run a number of
+efforts hosted (and sponsored) by funders and agencies in both the US (e.g., NSF,
+NIH, Alfred P. Sloan Foundation) and UK (e.g., SFTC, JISC) have run a number of
 workshops with participants from across a range of disciplines.
@@
 submission of papers in standard journals describing a piece of research
-software. While some journals (e.g. F1000Research) have traditionally accepted
+software. While some journals (e.g., F1000Research) have traditionally accepted
 software submissions, the American Astronomical Society (AAS) have recently

### 2016-03-01 sc-principles/software-citation-principles.tex
Add 2 notes
Add a note on citation styles

@@
 
+\definecolor{darkgreen}{rgb}{0,0.6,0}
 \newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
@@
 \newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}}
+\newcommand{\scnote}[1]{ {\textcolor{darkgreen} { ***SCC: #1 }}} % Sou-Cheng Choi
 
@@
 package authors have taken to automate the generation of citation entities such as bibtex entries.
+\scnote{So does \cite{knepley2013accurately}}
 
@@
 \asnote{Should also discuss the WSSSPE \cite{wssspe} series of workshops in this section}
+\scnote{Section 5.8 and Appendix J in \cite{WSSSPE3} has some preliminary work and discussion on the topic.}
 
@@
 \begin{enumerate}
-\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.
+\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
 \item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.

### 2016-03-02 sc-principles/software-citation-principles.tex
some copyedits and comments

@@
 \newcommand{\scnote}[1]{ {\textcolor{darkgreen} { ***SCC: #1 }}} % Sou-Cheng Choi
+\newcommand{\dmnote}[1]{ {\textcolor{purple} { ***DM: #1 }}} % Daniel Mietchen
 
@@
 components.  Scientific knowledge is embedded in these components.  And papers and books
-themselves are also becoming increasingly digital, allowing them to executable
+themselves are also becoming increasingly digital, allowing them to become executable
 and reproducible.  As we move towards this future where science is performed in and recorded
 as a variety of linked digital products, the characteristics and properties that developed for
-books and papers need to be applied to all digital products.  Here, we are concerned specifically
+books and papers need to be applied to all digital products and possibly adjusted.  Here, we are concerned specifically
 with the citation of software products.
@@
 That's where they are needed (in the table).
-I don't like footnotes for something like this.}
+I don't like footnotes for something like this.}\dmnote{Agree it would be better in the caption.}
 In the table,
@@
 examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
-domain groups include bioCADDIE, CIG, etc.;
+domain groups include bioCADDIE, CIG, etc.;\dmnote{Not sure what CIG stands for, and acronyms should probably be explained or linked}
 ``repository'' refers to open-source software repositories such as ASCL, GitHub, Netlib, CRAN, institutional repositories;
 ``funder'' is a group that funds software or work using software; and
-``UID'' means unique ID, such as DOI or PURL.
+``UID'' means unique ID, such as a DOI or PURL.
 
@@
 \begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?}}
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?} \dmnote{I would leve them separate.
+What about adding a column for license?}}
 \centering
@@
  \cmidrule{3-10}
-\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
+\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to \dmnote{should start with a verb}} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
 & & name &  &  &  date & repository &  & citations & \\
@@
 
-If certain metadata are not available, alternates may be provided:
+If certain metadata are not available, alternatives may be provided:
 \begin{itemize}
@@
 document existing practices in their respective communities. A total of 47 documents were submitted by working group
-participants with the life sciences, astrophysics, and geosciences being particularly well-represented in the submitted
+participants, with the life sciences, astrophysics, and geosciences being particularly well-represented in the submitted
 resources.
@@
 
-In a blog post from 2012, \cite{ssi-how-to-cite} discussing some of the pitfalls of trying to cite software in publications.
+In a blog post from 2012, \cite{ssi-how-to-cite} discusses some of the pitfalls of trying to cite software in publications.
 The author includes useful guidance for when to consider citing software as well as some ways to help `convince' journal editors
@@
 authors of the software would like to be cited by others. While this isn't a formal metadata specification (it's not machine
-readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.
+readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \href{http://doi.org/10.1371/journal.pone.0136631}{10.1371/journal.pone.0136631}.}
 In a later post on the SSI blog, Jackson \cite{ssi-how-shalt-i-cite-thee} gives a good overview of some of the approaches
@@
 Author guidelines in F1000Research, APA authors, SSI `publish or be damned' \cite{ssi-publish-or-be-damned} manifesto.
+\dmnote{There are many other journals publishing software descriptions, some even specialized in software, e.g.
+\href{http://www.openresearchcomputation.com}{Open Research Computation}.}
 

### 2016-03-02 sc-principles/software-citation-principles.tex
adding membership in Appendix A and referring to that in Section 1

@@
 The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015.
-It gathered members, and then in June 2015 began studying this, with a number of meetings
+It gathered members (see Appendix~\ref{app:wg_members}), and then in June 2015 began studying this, with a number of meetings
 and some off-line work by group members to gather materials documenting existing
@@
 
+\appendix
+
+\section{Working Group Membership}
+\label{app:wg_members}
+
+\katznote{should make affiliation styles more consistent}
+
+\katznote{should check to make sure these are all still right, such as for me and Bill Mills}
+
+Alberto Accomazzi, Harvard-Smithsonian CfA
+
+Alice Allen, Astrophysics Source Code Library
+
+Carl Boettiger, UC Berkeley
+
+Sou-Cheng Choi , NORC at the University of Chicago \& Illinois Institute of Technology
+
+Neil Chue Hong, Software Sustainability Institute
+
+Tom Crick, Cardiff Metropolitan University
+
+Martin Fenner, Public Library of Science
+
+Merc\`e Crosas, IQSS, Harvard University
+
+Christopher Erdmann, Harvard-Smithsonian CfA
+
+Ian Gent, University of St Andrews, recomputation.org
+
+Paul Groth, Elsevier Labs
+
+Melissa Haendel, OHSU
+
+Stephanie Hagstrom, Force11
+
+Robert Hanisch, NIST/ODI
+
+Edwin Henneken, Harvard-Smithsonian CfA
+
+Ivan Herman, W3C
+
+James Howison, UTexas
+
+Lorraine Hwang, UC Davis
+
+Thomas Ingraham, F1000Research
+
+Matthew B. Jones, NCEAS, UC Santa Barbara
+
+Catherine Jones, Science and Technology Facilities Council
+
+Daniel S. Katz, U Chicago \& Argonne Natl Lab (co-chair)
+
+Alexander Konovalov, University of St Andrews
+
+John Kratz, California Digital Library
+
+Jennifer Lin, Public Library of Science
+
+Frank L\"offler, Louisiana State University
+
+Brian Matthews, Science and Technology Facilites Council
+
+Abigail Cabunoc Mayes, Mozilla Science Lab
+
+Daniel Mietchen, NIH
+
+Bill Mills, Mozilla Science Lab
+
+Evan Misshula, CUNY Graduate Center
+
+August Muench, American Astronomical Society
+
+Fiona Murphy, Independent Researcher
+
+Lars Holm Nielsen, CERN
+
+Kyle Niemeyer, Oregon State University
+
+Karthik Ram, University of California, Berkeley
+
+Fernando Rios, Johns Hopkins University
+
+Ashley Sands, UCLA Information Studies
+
+Soren Scott , Independent Researcher
+
+Frank J. Seinstra, Netherlands eScience Center
+
+Arfon Smith, GitHub (co-chair)
+
+Kaitlin Thaney, Mozilla Science Lab
+
+Ilian Todorov, STFC
+
+Matt Turk, University of Illinois
+
+Miguel de Val-Borro, Princeton University
+
+Daan Van Hauwermeiren, Ghent University
+
+Stijn Van Hoey, Ghent University
+
+Belinda Weaver, The University of Queensland
+
+Nic Weber, University of Washington iSchool
+
 \bibliographystyle{abbrv}

### 2016-03-02 sc-principles/software-citation-principles.tex
addressing DM comment in table 1 about starting rows with verbs

@@
  \cmidrule{3-10}
-\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to \dmnote{should start with a verb}} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
+\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
 & & name &  &  &  date & repository &  & citations & \\
 \midrule
-Researcher            & someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
+Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
+Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
 Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
 Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
-Researcher            & credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
 Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\

### 2016-03-02 sc-principles/software-citation-principles.tex
adding comments about group and goal based on comment from James

@@
 
-The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015.
-It gathered members (see Appendix~\ref{app:wg_members}), and then in June 2015 began studying this, with a number of meetings
+The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015, with the following mission statement:
+
+\begin{quote}
+The software citation working group is a committee that will leverage the perspectives of a variety of existing initiatives working on software citation to produce a consolidated set of citation principles in order to encourage broad adoption of a consistent policy for software citation across disciplines and venues.
+The working group will review existing efforts and make a set of recommendations.
+These recommendations will be put up for endorsement by the organizations represented by this group and others that play an important role in the community.
+
+The group will produce a set of principles, illustrated with working examples, and a plan for dissemination and distribution.
+This group will not be producing detailed specifications for implementation although it may review and discuss possible technical solutions.\end{quote}
+
+The group gathered members (see Appendix~\ref{app:wg_members}) in April and May 2015, and then in June began work, with a number of meetings
 and some off-line work by group members to gather materials documenting existing
@@
 these principles impact the scientific process.
+
+The principles in this document should guide further development of software citation mechanisms and system, and the reader should be able to look at any particular example of software citation and see if it meets the principles.
 

### 2016-03-02 sc-principles/software-citation-principles.tex
adding note on motivation in intro - need to add cites
adding note about bib in principle 1

@@
 \newcommand{\dmnote}[1]{ {\textcolor{purple} { ***DM: #1 }}} % Daniel Mietchen
+\newcommand{\acmnote}[1]{ {\textcolor{green} { ***ACM: #1 }}} % Abigail Cabunoc Mayes
 
@@
 
+\katznote{add cites to James's study how what people cite software and Matthew Mayernik's similar one as part of motivation}
+
 The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015, with the following mission statement:
@@
 could be applied, related back to the use cases in \S\ref{sec:use_cases}.
+
+ \acmnote{would like to add info on the problem we're trying to solve in academia to motivation}
 
@@
 Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
+\katznote{say specifically that this should be part of the bibliography of a paper?}
 \item \textbf{Credit and Attribution}

### 2016-03-02 sc-principles/software-citation-principles.tex
adding additional sentence in principle 1
Updates from ShareLaTeX

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015.
-It gathered members (see Appendix~\ref{app:wg_members}), and then in June 2015 began studying this, with a number of meetings
+It gathered members, and then in June 2015 began studying this, with a number of meetings
 and some off-line work by group members to gather materials documenting existing
@@
 \begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?} \dmnote{I would leve them separate.
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?} \dmnote{I would leave them separate.
 What about adding a column for license?}}
@@
  \cmidrule{3-10}
-\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
+\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to \dmnote{should start with a verb}} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
 & & name &  &  &  date & repository &  & citations & \\
 \midrule
-Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
+Researcher            & someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
+Researcher            & someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
 Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
 Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
-Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
 Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
@@
 
-\appendix
-
-\section{Working Group Membership}
-\label{app:wg_members}
-
-\katznote{should make affiliation styles more consistent}
-
-\katznote{should check to make sure these are all still right, such as for me and Bill Mills}
-
-Alberto Accomazzi, Harvard-Smithsonian CfA
-
-Alice Allen, Astrophysics Source Code Library
-
-Carl Boettiger, UC Berkeley
-
-Sou-Cheng Choi , NORC at the University of Chicago \& Illinois Institute of Technology
-
-Neil Chue Hong, Software Sustainability Institute
-
-Tom Crick, Cardiff Metropolitan University
-
-Martin Fenner, Public Library of Science
-
-Merc\`e Crosas, IQSS, Harvard University
-
-Christopher Erdmann, Harvard-Smithsonian CfA
-
-Ian Gent, University of St Andrews, recomputation.org
-
-Paul Groth, Elsevier Labs
-
-Melissa Haendel, OHSU
-
-Stephanie Hagstrom, Force11
-
-Robert Hanisch, NIST/ODI
-
-Edwin Henneken, Harvard-Smithsonian CfA
-
-Ivan Herman, W3C
-
-James Howison, UTexas
-
-Lorraine Hwang, UC Davis
-
-Thomas Ingraham, F1000Research
-
-Matthew B. Jones, NCEAS, UC Santa Barbara
-
-Catherine Jones, Science and Technology Facilities Council
-
-Daniel S. Katz, U Chicago \& Argonne Natl Lab (co-chair)
-
-Alexander Konovalov, University of St Andrews
-
-John Kratz, California Digital Library
-
-Jennifer Lin, Public Library of Science
-
-Frank L\"offler, Louisiana State University
-
-Brian Matthews, Science and Technology Facilites Council
-
-Abigail Cabunoc Mayes, Mozilla Science Lab
-
-Daniel Mietchen, NIH
-
-Bill Mills, Mozilla Science Lab
-
-Evan Misshula, CUNY Graduate Center
-
-August Muench, American Astronomical Society
-
-Fiona Murphy, Independent Researcher
-
-Lars Holm Nielsen, CERN
-
-Kyle Niemeyer, Oregon State University
-
-Karthik Ram, University of California, Berkeley
-
-Fernando Rios, Johns Hopkins University
-
-Ashley Sands, UCLA Information Studies
-
-Soren Scott , Independent Researcher
-
-Frank J. Seinstra, Netherlands eScience Center
-
-Arfon Smith, GitHub (co-chair)
-
-Kaitlin Thaney, Mozilla Science Lab
-
-Ilian Todorov, STFC
-
-Matt Turk, University of Illinois
-
-Miguel de Val-Borro, Princeton University
-
-Daan Van Hauwermeiren, Ghent University
-
-Stijn Van Hoey, Ghent University
-
-Belinda Weaver, The University of Queensland
-
-Nic Weber, University of Washington iSchool
-
 \bibliographystyle{abbrv}

### 2016-03-02 sc-principles/software-citation-principles.tex

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
-\katznote{say specifically that this should be part of the bibliography of a paper?}
+Specifically, software citations should be part of the reference list for a research product such as a paper.
 \item \textbf{Credit and Attribution}

### 2016-03-02 sc-principles/software-citation-principles.tex
merging former principle 2b into 1

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
-\katznote{say specifically that this should be part of the bibliography of a paper?}
+Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
 \item \textbf{Credit and Attribution}
-\begin{enumerate}
-\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
-\item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.
-\end{enumerate}
+Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
 \item \textbf{Unique Identification}

### 2016-03-02 sc-principles/software-citation-principles.tex
Updates from ShareLaTeX

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
-\item \textbf{Access}
-Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.
+\item \textbf{Accessibility}
+Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{}
 \item \textbf{Persistence}
@@
 
-Daniel Mietchen, NIH
+Daniel Mietchen, National Institutes of Health
 

### 2016-03-02 sc-principles/software-citation-principles.tex
changing intro to section 4, removing persistence from item 3, and moving item 5 above previous item 4

@@
 
-\katznote{The current content of this section is liberally borrowed and loosely adapted from \cite{data-citation-principles}.
-We we will want to add some text here explaining that we started with that document, and then modified it to take into account the information from the use cases and the related work, once we do so.}
+The following set of software citation principles were originally based on an adaptation of the parallel set of data citation principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
 
@@
 \item \textbf{Unique Identification}
-A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
+A software citation should include a method for identification that is machine actionable, globally unique, and widely used by a community.
+\item \textbf{Persistence}
+Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
 \item \textbf{Accessibility}
 Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{}
-\item \textbf{Persistence}
-Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
 \item \textbf{Specificity and Verifiability}

### 2016-03-02 sc-principles/software-citation-principles.tex
Updates from ShareLaTeX

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
-\item \textbf{Accessibility}
-Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{}
 \item \textbf{Persistence}
 Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
+\item \textbf{Accessibility}
+Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
+Should we say anything about environments?}
 \item \textbf{Specificity and Verifiability}
@@
 Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
+\item \textbf{Anything missing?}
+Interactive software? Screenshots? \ldots ?
 \end{enumerate}
@@
 
-Sou-Cheng Choi , NORC at the University of Chicago \& Illinois Institute of Technology
+Sou-Cheng Choi, NORC at the University of Chicago \& Illinois Institute of Technology
 

### 2016-03-02 sc-principles/software-citation-principles.tex

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 \item \textbf{Accessibility}
-Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{}
+Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
+Should we say anything about environments?}
 \item \textbf{Specificity and Verifiability}

### 2016-03-02 sc-principles/software-citation-principles.tex
Updates from ShareLaTeX
change to principle 6 based on group telecon

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{}
-\item \textbf{Specificity and Verifiability}
-Software citations should facilitate identification of, access to, and verification of the specific software that support a claim.
-Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
+\item \textbf{Specificity}
+Software citations should facilitate identification of and access to the specific version of software that was used.
+Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
 \item \textbf{Interoperability and Flexibility}

### 2016-03-02 sc-principles/software-citation-principles.tex

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 Should we say anything about environments?}
-\item \textbf{Specificity and Verifiability}
-Software citations should facilitate identification of, access to, and verification of the specific software that support a claim.
-Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
+\item \textbf{Specificity}
+Software citations should facilitate identification of and access to the specific version of software that was used.
+Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
 \item \textbf{Interoperability and Flexibility}

### 2016-03-02 sc-principles/software-citation-principles.tex
Updates from ShareLaTeX

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 
-The following set of software citation principles were originally based on an adaptation of the parallel set of data citation principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
+\katznote{The current content of this section is liberally borrowed and loosely adapted from \cite{data-citation-principles}.
+We we will want to add some text here explaining that we started with that document, and then modified it to take into account the information from the use cases and the related work, once we do so.}
 
@@
 \item \textbf{Unique Identification}
-A software citation should include a method for identification that is machine actionable, globally unique, and widely used by a community.
+A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
 \item \textbf{Persistence}
@@
 \item \textbf{Accessibility}
-Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
+Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
 Should we say anything about environments?}
@@
 \end{enumerate}
+\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}
 

### 2016-03-02 sc-principles/software-citation-principles.tex

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 Should we say anything about environments?}
-\item \textbf{Specificity}
-Software citations should facilitate identification of and access to the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
+\item \textbf{Specificity and Verifiability}
+Software citations should facilitate identification of, access to, and verification of the specific software that support a claim.
+Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
 \item \textbf{Interoperability and Flexibility}

### 2016-03-02 sc-principles/software-citation-principles.tex
adding colons

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
-\item \textbf{Credit and Attribution}
+\item \textbf{Credit and Attribution}:
 Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
-\item \textbf{Unique Identification}
+\item \textbf{Unique Identification}:
 A software citation should include a method for identification that is machine actionable, globally unique, and widely used by a community.
-\item \textbf{Persistence}
+\item \textbf{Persistence}:
 Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
-\item \textbf{Accessibility}
+\item \textbf{Accessibility}:
 Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
 Should we say anything about environments?}
-\item \textbf{Specificity}
+\item \textbf{Specificity}:
 Software citations should facilitate identification of and access to the specific version of software that was used.
 Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
-\item \textbf{Interoperability and Flexibility}
+\item \textbf{Interoperability and Flexibility}:
 Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
 \end{enumerate}
+\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}
 

### 2016-03-02 sc-principles/software-citation-principles.tex
fixes to affils

@@
 
-James Howison, UTexas
+James Howison, University of Texas
 
-Lorraine Hwang, UC Davis
+Lorraine Hwang, University of California,  Davis
 
@@
 
-Daniel S. Katz, U Chicago \& Argonne Natl Lab (co-chair)
+Daniel S. Katz, University of Chicago \& Argonne National Laboratory (co-chair)
 
@@
 
-Brian Matthews, Science and Technology Facilites Council
+Brian Matthews, Science and Technology Facilities Council
 
@@
 
-Soren Scott , Independent Researcher
+Soren Scott, Independent Researcher
 
@@
 
-Ilian Todorov, STFC
+Ilian Todorov, Science and Technology Facilities Council
 

### 2016-03-02 sc-principles/software-citation-principles.tex
removing principle 7 and adding interoperable to 3

@@
 \item \textbf{Unique Identification}:
-A software citation should include a method for identification that is machine actionable, globally unique, and widely used by a community.
+A software citation should include a method for identification that is machine actionable, globally unique, interoperable, and widely used by a community.
 \item \textbf{Persistence}:
@@
 Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
-\item \textbf{Interoperability and Flexibility}:
-Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
+%\item \textbf{Interoperability and Flexibility}:
+%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
 \end{enumerate}

### 2016-03-02 sc-principles/software-citation-principles.tex

@@
 with the citation of software products.
+
+Currently software appears in publications in very inconsistent ways.
+For example, Howison and Bullard \cite{howison2015jasist} examined a random sample of 90 articles in the biology literature and found seven different ways that software was mentioned, from simple names in the full-text, URLs in footnote, Instrument-like in-text mentions, to three different kinds of mentions in references lists: cites to project names or websites, cites to user manuals, and cites to publications that describe or introduce the software.
+Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered.
+Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format.
+Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses.
+
+\rowcolors{2}{white}{gray!25}
+\begin{table}[htbp]
+\caption{Varieties of software mentions in publicatons, from  \cite{howison2015jasist}} 
+\centering
+\scriptsize\setlength{\tabcolsep}{2.5pt}
+\begin{tabular}{@{}l l l p{0.6\textwidth}@{}}
+\toprule
+Mention Type & \multicolumn{2}{c}{Count (n=286)} & Example \\
+\midrule
+Cite to publication & 105 & 37\% & \ldots was calculated using biosys (Swofford \& Selander 1981).
+\newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics.
+\emph{Journal of Heredity}, 72, 281–283.\\
+Cite to users manual & 6 & 2\% & \ldots as analyzed by the BIAevaluation software (Biacore, 1997).
+\newline Ref: Biacore, I.
+(1997).
+BIAevaluation Software Handbook,version 3.0 (Uppsala, Sweden: Biacore, Inc) \\
+Cite to name or website & 15 & 5\% & \ldots using the program Autodecay version 4.0.29 PPC (Eriksson 1998). \newline Ref: ERIKSSON, T. 1998. Autodecay, vers. 4.0.29 Stockholm:Department of Botany. \\
+Instrument-like & 53 & 19\% & \ldots calculated by t-test using the Prism 3.0 software (GraphPad Software, San Diego, CA, USA). \\
+URL in text & 13 & 5\% & \ldots freely available from http://www.cibiv.at/software/pda/. \\
+In-text name only & 90 & 31\% & \ldots were analyzed using MapQTL (4.0) software. \\
+Not even name & 4 & 1\% & \ldots was carried out using software implemented in the Java programming language. \\
+\bottomrule
+\end{tabular}
+\label{tab:mentions}
+\end{table}%
+
+In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
+In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
 

### 2016-03-02 sc-principles/software-citation-principles.tex
moving note in intro and fixing to recognize what James has already added

@@
 
+\katznote{as part of motivation, James added cite to his study how what people cite software.  Dan to add cite of Matthew Mayernik's similar one, and perhaps trim text/table for space.}
+
 Currently software appears in publications in very inconsistent ways.
@@
 In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
-
-\katznote{add cites to James's study how what people cite software and Matthew Mayernik's similar one as part of motivation}
 

### 2016-03-03 sc-principles/software-citation-principles.tex
some trimming of new text in 1 done, still need to add Mayernik cite and discuss

--- 2016-03-02 sc-principles/software-citation-principles.tex
@@
 
-\katznote{add cites to James's study how what people cite software and Matthew Mayernik's similar one as part of motivation}
+\katznote{as part of motivation, James added cite to his study how what people cite software.  Dan to add cite of Matthew Mayernik's similar one, and perhaps trim text/table for space.}
+
+Software and other digital resources currently appears in publications in very inconsistent ways.
+For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}.
+Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered.
+%Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format.
+Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses.
+In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
+% In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
+
+\rowcolors{2}{white}{gray!25}
+\begin{table}[htbp]
+\caption{Varieties of software mentions in publications, from  \cite{howison2015jasist}} 
+\centering
+\scriptsize\setlength{\tabcolsep}{2.5pt}
+\begin{tabular}{@{}l l l } % p{0.6\textwidth}@{}}
+\toprule
+Mention Type & \multicolumn{2}{c}{Count (n=286)} \\ % & Example \\
+\midrule
+Cite to publication & 105 & 37\% \\ %& \ldots was calculated using biosys (Swofford \& Selander 1981).
+\newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics.
+\emph{Journal of Heredity}, 72, 281–283.\\
+Cite to users manual & 6 & 2\% \\ %& \ldots as analyzed by the BIAevaluation software (Biacore, 1997).
+\newline Ref: Biacore, I.
+(1997).
+BIAevaluation Software Handbook, version 3.0 (Uppsala, Sweden: Biacore, Inc) \\
+Cite to name or website & 15 & 5\% \\ %& \ldots using the program Autodecay version 4.0.29 PPC (Eriksson 1998).
+\newline Ref: ERIKSSON, T.
+1998.
+Autodecay, vers.
+4.0.29 Stockholm:Department of Botany.
+\\
+Instrument-like & 53 & 19\% \\ %& \ldots calculated by t-test using the Prism 3.0 software (GraphPad Software, San Diego, CA, USA). \\
+URL in text & 13 & 5\% \\ %& \ldots freely available from http://www.cibiv.at/software/pda/. \\
+In-text name only & 90 & 31\% \\ %& \ldots were analyzed using MapQTL (4.0) software. \\
+Not even name & 4 & 1\% \\ %& \ldots was carried out using software implemented in the Java programming language. \\
+\bottomrule
+\end{tabular}
+\label{tab:mentions}
+\end{table}%
 

### 2016-03-03 sc-principles/software-citation-principles.tex
spelling fix and note shortening in 1

@@
 
-\katznote{as part of motivation, James added cite to his study how what people cite software.  Dan to add cite of Matthew Mayernik's similar one, and perhaps trim text/table for space.}
-
-Software and other digital resources currently appears in publications in very inconsistent ways.
+\katznote{to add cite of Matthew Mayernik's similar work} 
+Software and other digital resources currently appear in publications in very inconsistent ways.
 For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}.

### 2016-03-03 sc-principles/software-citation-principles.tex
adding new refs for motivation, shortening old motivation text and adding next text for these refs, fixing affils in appendix

@@
 
-\katznote{to add cite of Matthew Mayernik's similar work} 
 Software and other digital resources currently appear in publications in very inconsistent ways.
@@
 Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses.
-In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
+%In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
+Studies on data and facility citation have shown similar results~\cite{parsons_duerr_minster, mayernik_poster}.
 % In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
@@
 \item \textbf{Accessibility}:
-Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
+Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
 Should we say anything about environments?}
@@
 
-\katznote{should make affiliation styles more consistent}
-
-\katznote{should check to make sure these are all still right, such as for me and Bill Mills}
+\katznote{should check to make sure these are all still right}
 
@@
 
-Carl Boettiger, UC Berkeley
+Carl Boettiger, University of California,  Berkeley
 
@@
 
-Melissa Haendel, OHSU
+Melissa Haendel, Oregon Health and Science University
 
-Stephanie Hagstrom, Force11
+Stephanie Hagstrom, FORCE11
 
-Robert Hanisch, NIST/ODI
+Robert Hanisch, National Institute of Standards and Technology, One Degree Imager
 
@@
 
-Ivan Herman, W3C
+Ivan Herman, World Wide Web Consortium (W3C)
 
@@
 
-Matthew B. Jones, NCEAS, UC Santa Barbara
+Matthew B. Jones, NCEAS, University of California,  Santa Barbara
 
@@
 
-Daniel S. Katz, University of Chicago \& Argonne National Laboratory (co-chair)
+Daniel S. Katz, University of Illinois (co-chair)
 
@@
 
-Ashley Sands, UCLA Information Studies
+Ashley Sands, University of California, Los Angeles
 

### 2016-03-03 sc-principles/software-citation-principles.tex
Add single space before \cite command and minor edits

--- 2016-03-03 sc-principles/software-citation-principles.tex
@@
 
-\katznote{to add cite of Matthew Mayernik's similar work} 
-Software and other digital resources currently appear in publications in very inconsistent ways.
-For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}.
-Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered.
-%Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format.
-Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses.
-In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
-% In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
-
-\rowcolors{2}{white}{gray!25}
-\begin{table}[htbp]
-\caption{Varieties of software mentions in publications, from  \cite{howison2015jasist}} 
-\centering
-\scriptsize\setlength{\tabcolsep}{2.5pt}
-\begin{tabular}{@{}l l l } % p{0.6\textwidth}@{}}
-\toprule
-Mention Type & \multicolumn{2}{c}{Count (n=286)} \\ % & Example \\
-\midrule
-Cite to publication & 105 & 37\% \\ %& \ldots was calculated using biosys (Swofford \& Selander 1981).
-\newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics.
-\emph{Journal of Heredity}, 72, 281–283.\\
-Cite to users manual & 6 & 2\% \\ %& \ldots as analyzed by the BIAevaluation software (Biacore, 1997).
-\newline Ref: Biacore, I.
-(1997).
-BIAevaluation Software Handbook, version 3.0 (Uppsala, Sweden: Biacore, Inc) \\
-Cite to name or website & 15 & 5\% \\ %& \ldots using the program Autodecay version 4.0.29 PPC (Eriksson 1998).
-\newline Ref: ERIKSSON, T.
-1998.
-Autodecay, vers.
-4.0.29 Stockholm:Department of Botany.
-\\
-Instrument-like & 53 & 19\% \\ %& \ldots calculated by t-test using the Prism 3.0 software (GraphPad Software, San Diego, CA, USA). \\
-URL in text & 13 & 5\% \\ %& \ldots freely available from http://www.cibiv.at/software/pda/. \\
-In-text name only & 90 & 31\% \\ %& \ldots were analyzed using MapQTL (4.0) software. \\
-Not even name & 4 & 1\% \\ %& \ldots was carried out using software implemented in the Java programming language. \\
-\bottomrule
-\end{tabular}
-\label{tab:mentions}
-\end{table}%
+\katznote{add cites to James's study how what people cite software and Matthew Mayernik's similar one as part of motivation}
 
@@
 domain groups include bioCADDIE, CIG, etc.;\dmnote{Not sure what CIG stands for, and acronyms should probably be explained or linked}
-``repository'' refers to open-source software repositories such as ASCL, GitHub, Netlib, CRAN, institutional repositories;
+``repository'' refers to public software repositories such as ASCL,
+GitHub, Netlib, CRAN, institutional repositories;
 ``funder'' is a group that funds software or work using software; and
@@
 
-AAS workshop (\cite{aas-software-index}), CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
+AAS workshop \cite{aas-software-index}, CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
 
@@
 Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
-\item \textbf{Credit and Attribution}:
-Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
-\item \textbf{Unique Identification}:
-A software citation should include a method for identification that is machine actionable, globally unique, interoperable, and widely used by a community.
-\item \textbf{Persistence}:
+\item \textbf{Credit and Attribution}
+Software citations should facilitate giving scholarly credit and normative
+and legal attribution to all contributors to the software, recognizing
+that a single style or mechanism of attribution may not be applicable to
+all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for
+software citation are illustrated with examples in~\cite{lipson2011cite}.}
+\item \textbf{Unique Identification}
+A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
+\item \textbf{Accessibility}
+Software citations should facilitate access to the software themselves and
+to such associated metadata, documentation, data, and other materials, as
+are necessary for both humans and machines to make informed use of the
+referenced software.
+\item \textbf{Persistence}
 Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
-\item \textbf{Accessibility}:
-Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
-Should we say anything about environments?}
-\item \textbf{Specificity}:
-Software citations should facilitate identification of and access to the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
-%\item \textbf{Interoperability and Flexibility}:
-%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
+\item \textbf{Specificity and Verifiability}
+Software citations should facilitate identification of, access to, and verification of the specific software that support a claim.
+Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
+\item \textbf{Interoperability and Flexibility}
+Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
 \end{enumerate}
-\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}
 
@@
 
-Sou-Cheng Choi, NORC at the University of Chicago \& Illinois Institute of Technology
+Sou-Cheng Choi , NORC at the University of Chicago \& Illinois Institute of Technology
 
@@
 
-James Howison, University of Texas
+James Howison, UTexas
 
-Lorraine Hwang, University of California,  Davis
+Lorraine Hwang, UC Davis
 
@@
 
-Daniel S. Katz, University of Chicago \& Argonne National Laboratory (co-chair)
+Daniel S. Katz, U Chicago \& Argonne Natl Lab (co-chair)
 
@@
 
-Brian Matthews, Science and Technology Facilities Council
+Brian Matthews, Science and Technology Facilites Council
 
@@
 
-Soren Scott, Independent Researcher
+Soren Scott , Independent Researcher
 
@@
 
-Ilian Todorov, Science and Technology Facilities Council
+Ilian Todorov, STFC
 

### 2016-03-03 sc-principles/software-citation-principles.tex
fixing an affiliation
repository definition in table 1
even if the code repository is available it may not be released under
an open-source license

@@
 
-AAS workshop \cite{aas-software-index}, CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
+AAS workshop (\cite{aas-software-index}), CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
 
@@
 \item \textbf{Credit and Attribution}
-Software citations should facilitate giving scholarly credit and normative
-and legal attribution to all contributors to the software, recognizing
-that a single style or mechanism of attribution may not be applicable to
-all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for
-software citation are illustrated with examples in~\cite{lipson2011cite}.}
+Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
 \item \textbf{Unique Identification}
@@
 \item \textbf{Accessibility}
-Software citations should facilitate access to the software themselves and
-to such associated metadata, documentation, data, and other materials, as
-are necessary for both humans and machines to make informed use of the
-referenced software.
+Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{}
 \item \textbf{Persistence}

### 2016-03-03 sc-principles/software-citation-principles.tex

--- 2016-03-03 sc-principles/software-citation-principles.tex
@@
 \item \textbf{Credit and Attribution}:
-Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
+Software citations should facilitate giving scholarly credit and normative
+and legal attribution to all contributors to the software, recognizing
+that a single style or mechanism of attribution may not be applicable to
+all software. \scnote{At least four styles (AMS, APA, Chicago, MLA) for
+software citation are illustrated with examples in ~\cite{lipson2011cite}.}
 \item \textbf{Unique Identification}:
-A software citation should include a method for identification that is machine actionable, globally unique, interoperable, and widely used by a community.
+A software citation should include a method for identification that is
+machine actionable, globally unique, interoperable, and widely used by
+a community.
 \item \textbf{Persistence}:
@@
 
-Bill Mills, Mozilla Science Lab
+Bill Mills, TRIUMF
 

### 2016-03-04 sc-principles/software-citation-principles.tex
use non-breaking space

@@
 Software and other digital resources currently appear in publications in very inconsistent ways.
-For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}.
-Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered.
+For example, a random sample of 90 articles in the biology literature
+found seven different ways that software was mentioned, including simple
+names in the full-text, URLs in footnote, and different kinds of mentions
+in references lists: project names or websites, user manuals, publications
+that describe or introduce the software~\cite{howison2015jasist}.
+Table~\ref{tab:mentions} shows examples of these varied forms of software
+mentions and the frequency with which they were encountered. 
 %Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format.
@@
 all software. \scnote{At least four styles (AMS, APA, Chicago, MLA) for
-software citation are illustrated with examples in ~\cite{lipson2011cite}.}
+software citation are illustrated with examples in~\cite{lipson2011cite}.}
 \item \textbf{Unique Identification}:

### 2016-03-04 sc-principles/software-citation-principles.tex
Added heading for proportion column in table 1

@@
 \toprule
-Mention Type & \multicolumn{2}{c}{Count (n=286)} \\ % & Example \\
+Mention Type & Count (n=286) & Percentage\\ % & Example \\
 \midrule
-Cite to publication & 105 & 37\% \\ %& \ldots was calculated using biosys (Swofford \& Selander 1981).
+Cite to publication     & 105 & 37\% \\ % & \ldots was calculated using biosys (Swofford \& Selander 1981).
 \newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics.
-\emph{Journal of Heredity}, 72, 281–283.\\
-Cite to users manual & 6 & 2\% \\ %& \ldots as analyzed by the BIAevaluation software (Biacore, 1997).
+\emph{Journal of Heredity}, 72, 281–283.
+\\
+Cite to users manual    & 6   & 2\%  \\ % & \ldots as analyzed by the BIAevaluation software (Biacore, 1997).
 \newline Ref: Biacore, I.
@@
 BIAevaluation Software Handbook, version 3.0 (Uppsala, Sweden: Biacore, Inc) \\
-Cite to name or website & 15 & 5\% \\ %& \ldots using the program Autodecay version 4.0.29 PPC (Eriksson 1998).
+Cite to name or website & 15  & 5\%  \\ % & \ldots using the program Autodecay version 4.0.29 PPC (Eriksson 1998).
 \newline Ref: ERIKSSON, T.
@@
 \\
-Instrument-like & 53 & 19\% \\ %& \ldots calculated by t-test using the Prism 3.0 software (GraphPad Software, San Diego, CA, USA). \\
-URL in text & 13 & 5\% \\ %& \ldots freely available from http://www.cibiv.at/software/pda/. \\
-In-text name only & 90 & 31\% \\ %& \ldots were analyzed using MapQTL (4.0) software. \\
-Not even name & 4 & 1\% \\ %& \ldots was carried out using software implemented in the Java programming language. \\
+Instrument-like         & 53  & 19\% \\ % & \ldots calculated by t-test using the Prism 3.0 software (GraphPad Software, San Diego, CA, USA). \\
+URL in text             & 13  & 5\%  \\ % & \ldots freely available from http://www.cibiv.at/software/pda/. \\
+In-text name only       & 90  & 31\% \\ % & \ldots were analyzed using MapQTL (4.0) software. \\
+Not even name           & 4   & 1\%  \\ % & \ldots was carried out using software implemented in the Java programming language. \\
 \bottomrule

### 2016-03-07 sc-principles/software-citation-principles.tex
reformatted use case table with vertical column headers; added license, evaluator use case

--- 2016-03-03 sc-principles/software-citation-principles.tex
@@
 \usepackage{booktabs,multicol}
+\usepackage{rotating}
 
@@
 ``funder'' is a group that funds software or work using software; and
-``UID'' means unique ID, such as a DOI or PURL.
+``unique ID'' refers to identifiers such as a DOI or PURL.
 
-\rowcolors{5}{white}{gray!25}
+%
+\newcommand*\rot[1]{\begin{turn}{90} #1 \end{turn}}%
+\rowcolors{4}{gray!25}{white}
 \begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?} \dmnote{I would leave them separate.
-What about adding a column for license?}}
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}}
 \centering
 \scriptsize\setlength{\tabcolsep}{2.5pt}
-\begin{tabular}{@{}l l c c c c c c c c@{}}
+\begin{tabular}{@{}l l c c c c c c c c c@{}}
 \toprule
- & & \multicolumn{6}{c}{Requirements} \\
- \cmidrule{3-10}
-\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
-& & name &  &  &  date & repository &  & citations & \\
+ & & \multicolumn{7}{c}{Requirements} \\
+ \cmidrule{3-11}
+Stakeholder & Use\slash wants to 	&  \rot{Software name} & \rot{Author(s)} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Unique identifier} & \rot{Indexed citations} & \rot{Contributor role} & \rot{Software license}  \\
 \midrule
-Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
-Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & find software to implement task          & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             \\
-Publisher             & publish software paper                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Publisher             & publish papers that cite software        & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             \\
-Indexer               & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             \\
-Domain group          & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Library\slash archive & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
-Repository            & show scientific impact of holdings       & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
-Funder                & show how funded software has been used   & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
+Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             & \textbullet \\
+Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             & \textbullet \\
+Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet\\
+Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &            & \\
+Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet  \\
+Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
+Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet \\
+Researcher            & find software to implement task          & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Publisher                  & publish software paper                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             & \\
+Publisher                  & publish papers that cite software        & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \\
+Indexer                     & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet \\
+Domain group           & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet \\
+Library\slash archive & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet \\
+Repository                 & show scientific impact of holdings       & \textbullet &             &             &             &             & \textbullet & \textbullet &             &             \\
+Funder                       & show how funded software has been used   & \textbullet &             &             &             &             & \textbullet & \textbullet &             &             \\
+Evaluator                   & evaluate contributions of researcher   &                  & \textbullet &             & \textbullet &             & \textbullet & \textbullet & \textbullet &             \\
 \bottomrule

### 2016-03-07 sc-principles/software-citation-principles.tex
moved use case defs into caption, and added refs

@@
 Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case.
-\katznote{not sure if this next long sentence should be here, should be in the table caption, or should be footnotes to the table, or even more than one of these}
-\niemnote{Not sure either; definitely should include in the text due to length.
-Maybe an abbreviated version in the caption?}\flnote{If these ``definitions'' are not used anywhere else than in the table (and they don't really seem to be so far), then I tend towards putting this in the caption.
-That's where they are needed (in the table).
-I don't like footnotes for something like this.}\dmnote{Agree it would be better in the caption.}
-In the table,
-``researcher'' includes both academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers;
-``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software;
-``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
-examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
-domain groups include bioCADDIE, CIG, etc.;\dmnote{Not sure what CIG stands for, and acronyms should probably be explained or linked}
-``repository'' refers to public software repositories such as ASCL,
-GitHub, Netlib, CRAN, institutional repositories;
-``funder'' is a group that funds software or work using software; and
-``unique ID'' refers to identifiers such as a DOI or PURL.
 
@@
 \begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}}
+\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}.
+``Researcher'' includes both academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers;
+``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software;
+``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
+examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
+domain groups include bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics~\cite{CIG}, etc.;
+``repository'' refers to public software repositories such as ASCL~\cite{ascl},
+GitHub, Netlib, CRAN, institutional repositories;
+``funder'' is a group that funds software or work using software; and
+``unique ID'' refers to identifiers such as a DOI, ARK, or PURL.}
 \centering

### 2016-03-07 sc-principles/software-citation-principles.tex
added author names to ref in Table 1 caption
capitalized FORCE11 on first page

@@
 \begin{table}[htbp]
-\caption{Varieties of software mentions in publications, from  \cite{howison2015jasist}} 
+\caption{Varieties of software mentions in publications, from Howison and Bullard~\cite{howison2015jasist}.} 
 \centering
@@
 
-The Force11 Software Citation Working Group~\cite{f11scwg} was created in April 2015, with the following mission statement:
+The FORCE11 Software Citation Working Group~\cite{f11scwg} was created in April 2015, with the following mission statement:
 

### 2016-03-07 sc-principles/software-citation-principles.tex
responding to comments in intro and principles

@@
 
+There are many reasons this lack of both software citations in general and standard practices
+for software citation are of concern:
+\begin{itemize}
+\item Software is a product of research, and by not citing it, we leave holes in the record of of scientific progress.
+\item Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
+credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
+\end{itemize}
+
+\acmnote{would like to add info on the problem we're trying to solve in academia to motivation} 
+\katznote{added list above - others can add more items if appropriate.}
+
 The FORCE11 Software Citation Working Group~\cite{f11scwg} was created in April 2015, with the following mission statement:
@@
 could be applied, related back to the use cases in \S\ref{sec:use_cases}.
-
- \acmnote{would like to add info on the problem we're trying to solve in academia to motivation}
 
@@
 Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
-Should we say anything about environments?}
+Should we say anything about environments?} \katznote{If we provide access, I think that includes the location.
+I don't think we need to say anything about environments here, since it is in the next principles (platforms).
+Daniel, if this is ok with you, please comment out this set of comments}
 \item \textbf{Specificity}:
@@
 \dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}
+
+\katznote{perhaps we should have some discussion of these here, so that we could include Sou-Cheng's note and others?
+I would like to have the principles be as brief as possible, but in some cases, more may be needed.
+Or perhaps that comment goes into the next section?}
 

### 2016-03-07 sc-principles/software-citation-principles.tex
using ref from related work in intro

@@
 %In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
-Studies on data and facility citation have shown similar results~\cite{parsons_duerr_minster, mayernik_poster}.
+Studies on data and facility citation have shown similar results~\cite{10.1371/journal.pone.0136631, mayernik_poster, parsons_duerr_minster}.
 % In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
@@
 authors of the software would like to be cited by others. While this isn't a formal metadata specification (it's not machine
-readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \href{http://doi.org/10.1371/journal.pone.0136631}{10.1371/journal.pone.0136631}.}
+readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \cite{10.1371/journal.pone.0136631}.} \katznote{adding this ref to the introduction - not sure if it is still needed here too}
 In a later post on the SSI blog, Jackson \cite{ssi-how-shalt-i-cite-thee} gives a good overview of some of the approaches

### 2016-03-07 sc-principles/software-citation-principles.tex
adding bullet headings

@@
 \begin{itemize}
-\item Software is a product of research, and by not citing it, we leave holes in the record of of scientific progress.
-\item Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
+\item Understanding Science: Software is a product of research, and by not citing it, we leave holes in the record of of scientific progress.
+\item Academic Credit: Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
 credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.

### 2016-03-07 sc-principles/software-citation-principles.tex
adding new bullet
various changes made by group during 7 March call

@@
 
-\definecolor{darkgreen}{rgb}{0,0.6,0}
 \newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
@@
 \newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}}
+\definecolor{darkgreen}{rgb}{0,0.6,0}
 \newcommand{\scnote}[1]{ {\textcolor{darkgreen} { ***SCC: #1 }}} % Sou-Cheng Choi
@@
 \newcommand{\acmnote}[1]{ {\textcolor{green} { ***ACM: #1 }}} % Abigail Cabunoc Mayes 
+\definecolor{fuschsia}{rgb}{1.0, 0.58, 1.0}
+\newcommand{\LJHnote}[1]{ {\textcolor{fuschsia} { ***LJH: #1 }}} % Lorraine Hwang
+\definecolor{bondiblue}{rgb}{0.0, 0.58, 0.71}
+\newcommand{\ssnote}[1]{ {\textcolor{bondiblue} { ***SS: #1}}} % Soren Scott
+\newcommand{\GMnote}[1]{ {\textcolor{violet} { ***GM: #1}}} % August (Gus) Muench
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction \& motivation}
+\section{Software citation principles}
+\label{sec:principles}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The principles in this section are written fairly concisely, and further discussed later in this document (in \S\ref{sec:examples}).
+For example, in this section, we are not defining what software should be cited, but how it should be cited.
+In the next section, we discuss what how we think such decisions might be made.
+
+\katznote{maybe should change the language in the principles to be more requirements-like -- e.g.
+use ``shall'' a lot} \katznote{if the principles are machine readable, that might also imply something about the language used}
+
+\begin{enumerate}
+\item \textbf{Importance}:
+Software should be considered a legitimate, citable product of research.
+Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
+Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product.
+\item \textbf{Credit and Attribution}:
+Software citations should facilitate giving scholarly credit and normative
+and legal attribution to all contributors to the software, recognizing
+that a single style or mechanism of attribution may not be applicable to
+all software. 
+\item \textbf{Unique Identification}:
+A software citation should include a method for identification that is
+machine actionable, globally unique, interoperable, and recognized by
+a community.
+\item \textbf{Persistence}:
+Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
+\item \textbf{Accessibility}:
+Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.
+\item \textbf{Specificity}:
+Software citations should facilitate identification of, and access to, the specific version of software that was used.
+Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
+%\item \textbf{Interoperability and Flexibility}:
+%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
+\end{enumerate}
+
+These software citation principles were originally based on an adaptation of the parallel set of data citation principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
+The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data \ldots \katznote{Dan to write a little more here}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Motivation}
 \label{sec:intro}
@@
 with the citation of software products.
+This is more general than just textual citations of software in a paper, but more general identification of software within the research process.
+\katznote{other comments on this last sentence are welcome - is this a good enough description of what we mean by software citation?}
 
@@
 found seven different ways that software was mentioned, including simple
-names in the full-text, URLs in footnote, and different kinds of mentions
+names in the full-text, URLs in footnotes, and different kinds of mentions
 in references lists: project names or websites, user manuals, publications
@@
 Studies on data and facility citation have shown similar results~\cite{10.1371/journal.pone.0136631, mayernik_poster, parsons_duerr_minster}.
+% Add Hwang et al. from WSSSPE3 LJH
 % In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.
@@
 \begin{itemize}
-\item Understanding Science: Software is a product of research, and by not citing it, we leave holes in the record of of scientific progress.
+\item Understanding Science: Software is a product of research, and by not citing it, we leave holes in the record of scientific progress.
 \item Academic Credit: Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
 credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
+\item Finding Software - citations enable software to be discovered \LJHnote{and reused}.
+\katznote{dan to flesh out more} \ssnote{is it worth framing as software(cite \& find)->researcher(credit)->science (reuse) areas?}
+\item Reproducibility - citations enable specific software that was
+used to be identified, though they are not sufficient to make reproducibility happen \katznote{better language welcome}
 \end{itemize}
@@
 This report includes many recommendations for data citation.
+\LJHnote{ The Seismological Research Letters, Electronic Seismologists publishes papers on software.} %text needs to be cleaned up LJH%
 
@@
 AAS workshop \cite{aas-software-index}, CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
+\acmnote{ CodeMeta? }
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Software citation principles}
-\label{sec:principles}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and implementation examples}
+\label{sec:examples}
 
-The following set of software citation principles were originally based on an adaptation of the parallel set of data citation principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
+\subsection{Discussion of principles}
 
-\begin{enumerate}
-\item \textbf{Importance}:
-Software should be considered a legitimate, citable product of research.
-Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
-Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
-\item \textbf{Credit and Attribution}:
-Software citations should facilitate giving scholarly credit and normative
-and legal attribution to all contributors to the software, recognizing
-that a single style or mechanism of attribution may not be applicable to
-all software. \scnote{At least four styles (AMS, APA, Chicago, MLA) for
+\paragraph{importance}
+
+\katznote{what about derived software?
+What metadata should be preserved from the original?
+What is done for data today?
+should this be in the principles?
+tried to add to principle 1 and can discuss in next section.
+maybe also discuss somewhere else, such as in the list of stuff we recognize but don't address (currently in a GitHub issue)}
+\LJHnote{this is analogous to built upon research}
+
+\paragraph{what software to cite}
+
+\katznote{put discussion of what software to cite here -- something about what type of software we think should be cited and shouldn't not, e.g.
+excel} \LJHnote{sometimes excel could be considered materiel \ldots}
+\niemnote{Example of using Excel to simply store and plot data, vs.\ using it for statistical analysis}
+\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
+\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
+\paragraph{examples of citation styles}  \scnote{At least four styles (AMS, APA, Chicago, MLA) for
 software citation are illustrated with examples in~\cite{lipson2011cite}.}
-\item \textbf{Unique Identification}:
-A software citation should include a method for identification that is
-machine actionable, globally unique, interoperable, and widely used by
-a community.
-\item \textbf{Persistence}:
-Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
-\item \textbf{Accessibility}:
-Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available.
-Should we say anything about environments?} \katznote{If we provide access, I think that includes the location.
-I don't think we need to say anything about environments here, since it is in the next principles (platforms).
-Daniel, if this is ok with you, please comment out this set of comments}
-\item \textbf{Specificity}:
-Software citations should facilitate identification of and access to the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
-%\item \textbf{Interoperability and Flexibility}:
-%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
-\end{enumerate}
-\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}
 
-\katznote{perhaps we should have some discussion of these here, so that we could include Sou-Cheng's note and others?
-I would like to have the principles be as brief as possible, but in some cases, more may be needed.
-Or perhaps that comment goes into the next section?}
+\paragraph{unique identifiers}  \katznote{recognize that a UID points to a specific software, but multiple UIDs might point to the same software} \LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation examples}
-\label{sec:examples}
+\paragraph{talk about access to software - maybe discuss location}
+\LJHnote{recognize that some software is proprietary, not all is open source}
+
+\subsection{Reproducibility}  \katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}
+
+\subsection{Updates}
+\dmnote{Any provisions for updating the principles?}
+\LJHnote{may require feedback from "Implementation Group" and some time to put into practice to find gaps.
+Most of the time revisions are frequent in the beginning of the lifecycle to cover gaps and become less frequent.}
+
+\subsection{Implementation examples}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-03-08 sc-principles/software-citation-principles.tex
adding a bit on differences between software and data in section 1

@@
 
-These software citation principles were originally based on an adaptation of the parallel set of data citation principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
-The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data \ldots \katznote{Dan to write a little more here}
+These software citation principles were originally based on an adaptation of the FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
+The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts and it is executable.
+\katznote{are there any other difference between software and data that should be stated here?}  Also, while software can be considered a type of data, the converse is not generally true.
 

### 2016-03-08 sc-principles/software-citation-principles.tex
edit in motivation of why software citation is not just a style issue
editing in list of motivations in section 2

@@
 with the citation of software products.
-This is more general than just textual citations of software in a paper, but more general identification of software within the research process.
+The challenge is not just the textual citation of software in a paper, but the more general identification of software within the research process.
 \katznote{other comments on this last sentence are welcome - is this a good enough description of what we mean by software citation?}
@@
 There are many reasons this lack of both software citations in general and standard practices
-for software citation are of concern:
+for software citation are of concern: \katznote{Any more items to be added to list?}
+
 \begin{itemize}
@@
 credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
-\item Finding Software - citations enable software to be discovered \LJHnote{and reused}.
-\katznote{dan to flesh out more} \ssnote{is it worth framing as software(cite \& find)->researcher(credit)->science (reuse) areas?}
-\item Reproducibility - citations enable specific software that was
-used to be identified, though they are not sufficient to make reproducibility happen \katznote{better language welcome}
+\item Finding Software: Citations enable the specific software used in a research product to be discovered. Additional researchers can then use the same software for different purposes, leading
+to credit for those responsible for the software.
+\item Reproducibility: Citation of specific software used is necessary for reproducibility, but is not sufficient.  Additional information such as configurations and platform issues are also needed.
 \end{itemize}
-
-\acmnote{would like to add info on the problem we're trying to solve in academia to motivation} 
-\katznote{added list above - others can add more items if appropriate.}
 

### 2016-03-08 sc-principles/software-citation-principles.tex
small style stuff and abbreviation expansion in 3 and 4

@@
 \renewcommand{\baselinestretch}{0.96}
+
+\hyphenation{bioCADDIE}
 
@@
 examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
-domain groups include bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics~\cite{CIG}, etc.;
-``repository'' refers to public software repositories such as ASCL~\cite{ascl},
-GitHub, Netlib, CRAN, institutional repositories;
+domain groups include bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc.;
+``repository'' refers to public software repositories such as Astronomy Source Code Library (ASCL)~\cite{ascl},
+GitHub, Netlib, Comprehensive R Archive Network (CRAN), institutional repositories;
 ``funder'' is a group that funds software or work using software; and
@@
 
-In a blog post from 2012, \cite{ssi-how-to-cite} discusses some of the pitfalls of trying to cite software in publications.
-The author includes useful guidance for when to consider citing software as well as some ways to help `convince' journal editors
+In a blog post from 2012, Jackson discusses some of the pitfalls of trying to cite software in publications~\cite{ssi-how-to-cite}.
+He includes useful guidance for when to consider citing software as well as some ways to help `convince' journal editors
 to allow the inclusion of software citations.
 
-In \cite{ssi-citation-files}, Wilson suggests that software authors include a `CITATION' file that documents exactly how the
-authors of the software would like to be cited by others. While this isn't a formal metadata specification (it's not machine
-readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \cite{10.1371/journal.pone.0136631}.} \katznote{adding this ref to the introduction - not sure if it is still needed here too}
-In a later post on the SSI blog, Jackson \cite{ssi-how-shalt-i-cite-thee} gives a good overview of some of the approaches
-package authors have taken to automate the generation of citation entities such as bibtex entries.
-\scnote{So does \cite{knepley2013accurately}}
+Wilson suggests that software authors include a `CITATION' file that documents exactly how the
+authors of the software would like to be cited by others~\cite{ssi-citation-files}. While this isn't a formal metadata specification (it's not machine
+readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \cite{10.1371/journal.pone.0136631}.} \katznote{added this ref to the introduction - not sure if it is still needed here too}
+In a later post on the SSI blog, Jackson gives a good overview of some of the approaches
+package authors have taken to automate the generation of citation entities such as bibtex entries~\cite{ssi-how-shalt-i-cite-thee},
+and Knepley at al. do similarly~\cite{knepley2013accurately}.
 
@@
 
-\asnote{Should also discuss the WSSSPE \cite{wssspe} series of workshops in this section}
+\asnote{Should also discuss the WSSSPE~\cite{wssspe} series of workshops in this section}
 \scnote{Section 5.8 and Appendix J in \cite{WSSSPE3} has some preliminary work and discussion on the topic.}
@@
 
-AAS workshop \cite{aas-software-index}, CRediT project \cite{casrai-credit}, Ontosoft \cite{ontosoft}
+AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}
 \acmnote{ CodeMeta? }

### 2016-03-08 sc-principles/software-citation-principles.tex
broke 5 into 5 (discussion) and 6 (implementation examples) started writing in 5, on derived software

@@
 \begin{enumerate}
-\item \textbf{Importance}:
+\item \textbf{Importance}: \label{principle:importance}
 Software should be considered a legitimate, citable product of research.
@@
 software citation is important, by means of use cases.  In \S\ref{sec:related_work}, we
-summarize and analyze a large amount of previous work and thinking in this area.  In
-\S\ref{sec:principles}, we provide a set of guiding principles for citation of software within
-scholarly literature, other software, datasets, or any other research object.  Finally,
+summarize and analyze a large amount of previous work and thinking in this area.  In \S\ref{sec:discussion}, we discuss issues related to the principles stated in \S\ref{sec:principles}, and finally,
 in \S\ref{sec:examples} we give examples of how these software citation principles
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion and implementation examples}
-\label{sec:examples}
+\section{Discussion}
+\label{sec:discussion}
+
+In this section, we discuss some the principles stated in Section~\ref{sec:principles} and some related concerns.
 
@@
 
-\paragraph{importance}
+\paragraph{Derived Software}
 
-\katznote{what about derived software?
-What metadata should be preserved from the original?
-What is done for data today?
-should this be in the principles?
-tried to add to principle 1 and can discuss in next section.
-maybe also discuss somewhere else, such as in the list of stuff we recognize but don't address (currently in a GitHub issue)}
-\LJHnote{this is analogous to built upon research}
+The goals of software include the linked ideas of crediting those responsible for software and understanding the dependencies of scientific products on specific software.
+In the Importance principle (\ref{principle:importance}), we state that
+``software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product.''
+In the case of one code that is derived from another code, citing the derived software may appear to not credit those responsible for the original software, nor recognize its role in the work that used the derived software.
+However, this is really analogous to how any research builds on other research, where each research product just cites those products that it directly builds on, not those that it indirectly builds on.
+Understanding these chains of knowledge and credit have been part of history of science field for some time, though more recent work is suggesting more nuanced evaluation of the credit chains~\cite{casrai-credit, transitive_credit_json-ld}.
 
@@
 
-\subsection{Implementation examples}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation examples}
+\label{sec:examples}
 

### 2016-03-08 sc-principles/software-citation-principles.tex
more writing in 5

@@
 
-The principles in this section are written fairly concisely, and further discussed later in this document (in \S\ref{sec:examples}).
-For example, in this section, we are not defining what software should be cited, but how it should be cited.
-In the next section, we discuss what how we think such decisions might be made.
+The principles in this section are written fairly concisely, and discussed further later in this document (\S\ref{sec:discussion}).
+For example, in this section we do not define what software should be cited, but how it should be cited.
+we talk about how we think such decisions might be made in the discussion section.
 
@@
 
-\subsection{Discussion of principles}
+\textbf{What software to cite}
 
-\paragraph{Derived Software}
+In the principles, we did not define what software should be cited, but how it should be cited.
+What software should be cited is the decision of the author(s) of the research work, in the context of community norms and practices, and in most scientific communities, these are currently in flux.
+In general, we believe that any software on which a research product directly depends should be cited, if it significantly contributes to the research product.
+Again, the specific decision of what is significant must be made by the author(s) of the product.
+However, an illustrative example is the use of Microsoft Excel in research.
+We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
+\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
+\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
+
+\textbf{Derived software}
 
@@
 
-\paragraph{what software to cite}
+\textbf{Citations in text}
 
-\katznote{put discussion of what software to cite here -- something about what type of software we think should be cited and shouldn't not, e.g.
-excel} \LJHnote{sometimes excel could be considered materiel \ldots}
-\niemnote{Example of using Excel to simply store and plot data, vs.\ using it for statistical analysis}
-\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
-\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
-\paragraph{examples of citation styles}  \scnote{At least four styles (AMS, APA, Chicago, MLA) for
-software citation are illustrated with examples in~\cite{lipson2011cite}.}
+In terms of how citations should be expressed as text, such
+as in journals or other human readable material, there are currently at least four styles (AMS, APA, Chicago, MLA). These have been illustrated with examples by Lipson~\cite{lipson2011cite}.
+This is another area that is in flux at this time.
 
-\paragraph{unique identifiers}  \katznote{recognize that a UID points to a specific software, but multiple UIDs might point to the same software} \LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}
+\textbf{Unique identifiers}
 
-\paragraph{talk about access to software - maybe discuss location}
+A UID points to a specific software, but multiple UIDs might point to the same software.
+\LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}
+
+\textbf{talk about access to software - maybe discuss location}
+
 \LJHnote{recognize that some software is proprietary, not all is open source}
 
-\subsection{Reproducibility}  \katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}
+\textbf{Reproducibility}
 
-\subsection{Updates}
+\katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}
+
+\textbf{Updates}
+
 \dmnote{Any provisions for updating the principles?}

### 2016-03-09 sc-principles/software-citation-principles.tex

@@
 
-\author{Force11 Software Citation Working Group (Editors: Arfon M. Smith, Daniel S. Katz, Kyle E. Niemeyer)}
+\author{Force11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}
 
@@
 Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case.
+\scnote{Perhaps the long caption of Table 2 can be mostly moved here and be replaced with a shorter version there.}
 
@@
 package authors have taken to automate the generation of citation entities such as bibtex entries~\cite{ssi-how-shalt-i-cite-thee},
-and Knepley at al. do similarly~\cite{knepley2013accurately}.
+and Knepley at \scnote{$\rightarrow$ et} al. do similarly~\cite{knepley2013accurately}.
 
@@
 \dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
+\scnote{\emph{Do not cite standard office software (e.g.
+Word, Excel) or programming languages.
+Provide references only for specialized software.
+}---Source: \url{https://owl.english.purdue.edu/owl/resource/560/10/}.
+Likewise,  on what software not to cite,  generally speaking, we should not cite general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g.
+Skype).}
 
@@
 
-Sou-Cheng Choi, NORC at the University of Chicago \& Illinois Institute of Technology
+Sou-Cheng T.~Choi, NORC at the University of Chicago \& Illinois Institute of Technology
 
@@
 
-Matthew B. Jones, NCEAS, University of California,  Santa Barbara
+Matthew B.~Jones, NCEAS, University of California,  Santa Barbara
 
@@
 
-Daniel S. Katz, University of Illinois (co-chair)
+Daniel S.~Katz, University of Illinois (co-chair)
 
@@
 
-Frank J. Seinstra, Netherlands eScience Center
+Frank J.~Seinstra, Netherlands eScience Center
 

### 2016-03-09 sc-principles/software-citation-principles.tex
made change suggested by scc

@@
 package authors have taken to automate the generation of citation entities such as bibtex entries~\cite{ssi-how-shalt-i-cite-thee},
-and Knepley at \scnote{$\rightarrow$ et} al. do similarly~\cite{knepley2013accurately}.
+and Knepley et al.\ do similarly~\cite{knepley2013accurately}.
 

### 2016-03-10 sc-principles/software-citation-principles.tex

@@
 
-\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
-\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
-\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}}
-\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}}
+\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}} % Dan 
+\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}} % Kyle
+\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}} %Arfon
+\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}} %Frank Lšffler
 \definecolor{darkgreen}{rgb}{0,0.6,0}
@@
 
-The principles in this section are written fairly concisely, and discussed further later in this document (\S\ref{sec:discussion}).
-For example, in this section we do not define what software should be cited, but how it should be cited.
-we talk about how we think such decisions might be made in the discussion section.
+The principles in this section are written fairly concisely, and discussed
+further later in this document (\S\ref{sec:discussion}). For example, in this
+section we do not define what software should be cited, but how it should be
+cited.  We talk about how we think such decisions might be made in the
+discussion section.
 
@@
 \newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics.
-\emph{Journal of Heredity}, 72, 281–283.
+\emph{Journal of Heredity}, 72, 281-283.
 \\
@@
 
-There are many reasons this lack of both software citations in general and standard practices
-for software citation are of concern: \katznote{Any more items to be added to list?}
+There are many reasons why this lack of both software citations in general and
+standard practices for software citation are of concern: \katznote{Any more
+items to be added to list?}
 
@@
 \toprule
- & & \multicolumn{7}{c}{Requirements} \\
+ & & \multicolumn{9}{c}{Requirements} \\
  \cmidrule{3-11}
@@
 
-In astronomy \& astrophysics: ASCL \cite{ascl}, AAS journals  AAS/GitHub
-workshop \cite{aas-software-index} dedicated to software citation, indexing and discoverability.
+In astronomy \& astrophysics: ASCL \cite{ascl}, AAS journals  and the joint
+AAS/GitHub workshop \cite{aas-software-index} dedicated to software citation,
+indexing and discoverability.
 
@@
 
-Author guidelines in F1000Research, APA authors, SSI `publish or be damned' \cite{ssi-publish-or-be-damned} manifesto.
-\dmnote{There are many other journals publishing software descriptions, some even specialized in software, e.g.
-\href{http://www.openresearchcomputation.com}{Open Research Computation}.}
+Author guidelines in F1000Research, APA authors, SSI `publish or be damned'
+\cite{ssi-publish-or-be-damned} manifesto, workflow for publishing code using
+GitHub and Zenodo \cite{github-citable-code-guide}.
+\dmnote{There are many other journals publishing software descriptions, some even specialized in software, e.g. \href{http://www.openresearchcomputation.com}{Open Research Computation}.}
 

### 2016-03-16 sc-principles/software-citation-principles.tex
another change in a commented out line :)
updating comments in 1 and 2, completing draft of text in 5

@@
 
-\author{Force11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}
+\author{FORCE11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}
 
@@
 
-\katznote{maybe should change the language in the principles to be more requirements-like -- e.g.
-use ``shall'' a lot} \katznote{if the principles are machine readable, that might also imply something about the language used}
+\katznote{consider changing the language in the principles to be more requirements-like -- e.g.
+use ``shall'' a lot; related: consider making the principles machine-readable (has implications on language}
 
@@
 a community.
-\item \textbf{Persistence}:
+\item \textbf{Persistence}: \label{principle:persistence}
 Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
-\item \textbf{Accessibility}:
+\item \textbf{Accessibility}: \label{principle:accessibility}
 Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.
@@
 books and papers need to be applied to all digital products and possibly adjusted.  Here, we are concerned specifically
-with the citation of software products.
-The challenge is not just the textual citation of software in a paper, but the more general identification of software within the research process.
-\katznote{other comments on this last sentence are welcome - is this a good enough description of what we mean by software citation?}
+with the citation of software products.  The challenge is not just the textual citation of software in a paper, but the more general identification of software used within the research process.
 
@@
 \newline Ref: Swofford DL, Selander RB (1981) biosys-1: a Fortran program for the comprehensive analyses of electrophoretic data in population genetics and systematics.
-\emph{Journal of Heredity}, 72, 281-283.
+\emph{Journal of Heredity}, 72, 281--283.
 \\
@@
 There are many reasons why this lack of both software citations in general and
-standard practices for software citation are of concern: \katznote{Any more
-items to be added to list?}
+standard practices for software citation are of concern: \katznote{Are there any more
+items that should be added to this list?}
 
@@
 
-In this section, we discuss some the principles stated in Section~\ref{sec:principles} and some related concerns.
+In this section we discuss some the issues and concerns related to the principles stated in Section~\ref{sec:principles}.
 
@@
 We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
-\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
-\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
-\scnote{\emph{Do not cite standard office software (e.g.
+Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.''
+This recommendation matches that of the Purdue Online Writing Lab: ``Do not cite standard office software (e.g.
 Word, Excel) or programming languages.
-Provide references only for specialized software.
-}---Source: \url{https://owl.english.purdue.edu/owl/resource/560/10/}.
-Likewise,  on what software not to cite,  generally speaking, we should not cite general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g.
-Skype).}
+Provide references only for specialized software.''~\cite{powl-citing-software}
+
+Note that some software which is or could be captured as part of data provenance may not be cited.
+Citation is a a record of software that is important to the research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
+This implies that for a data research product, provenance data will include all cited software, but not necessarily vice versa.
+Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.
+And the data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.  
+These statements may also be true for software products.
+In general, we intend the software citation principles to cover what is minimum of what is necessary for software citation for the purpose of software identification.  
+Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
+%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
+%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
 
@@
 
-\textbf{Unique identifiers}
+\textbf{Unique identifiers} \katznote{Is this discussion topic needed?  I would like to delete it.}
 
@@
 
-\textbf{talk about access to software - maybe discuss location}
+\textbf{Access to software}
 
-\LJHnote{recognize that some software is proprietary, not all is open source}
+The Accessibility principle (\ref{principle:accessibility}) states that ``software citations should permit and facilitate access to the software itself.''
+This does not mean that the software must be freely available.
+Rather, the metadata should provide enough information that the software can be accessed.
+If the software is free, the metadata will likely provide an identifier that can be resolved to a URL that contains the specific version of the software that is being cited.
+For commercial software, the metadata should still provide information on how to access the specific software, but this may be a link to a web site that allows the software be purchased or a company's product number.
+As stated in the Persistence principle (\ref{principle:persistence}), we recognize that the commercial software version may no longer be available, but it still should be cited along with information about how it was accessed.
 
-\textbf{Reproducibility}
+%\LJHnote{recognize that some software is proprietary, not all is open source}
 
-\katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}
+%\textbf{Reproducibility}
+%
+%\katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}
 
-\textbf{Updates}
+\textbf{Updates to this document}
 
-\dmnote{Any provisions for updating the principles?}
-\LJHnote{may require feedback from "Implementation Group" and some time to put into practice to find gaps.
+As this set of software citation principles has been created by the FORCE11 Software Citation Working Group, which will cease work and dissolve after these principles have been published,
+any updates will require a different FORCE11 working group to make them.
+We expect a follow-on working group to be established to promote the implementation of
+these principles, and it is possible that this group might find items that need correction or addition
+in these principles.
+We recommend that this Software Citation Implementation Working Group be charged, in part,
+with updating these principles during its lifetime, and that FORCE11 should listen to community requests for later updates and respond by creating a new working group.
+
+%\dmnote{Any provisions for updating the principles?}
+%\LJHnote{may require feedback from "Implementation Group" and some time to put into practice to find gaps.
 Most of the time revisions are frequent in the beginning of the lifecycle to cover gaps and become less frequent.}
@@
 
-\katznote{should check to make sure these are all still right}
+\katznote{please check to make sure these are all still right}
 

### 2016-03-16 sc-principles/software-citation-principles.tex
adding a paragraph in 5 on software papers

@@
 
+\textbf{Software papers}
+
+Currently, and for the foreseeable future, software papers are being published and cited,
+in addition to software itself being cited, as many community norms and practices are
+oriented towards citation of papers.
+Given a software paper and software, the question of which should be cited and when is
+complicated, and there is no simple answer.  
+As discussed in the Importance principle (\ref{principle:importance}), \textit{the software
+itself should also be cited if it is directly used and was important to the work}.
+If a software paper exists and it contains results (performance, validation, etc.) that are
+important to the work, then the software paper should also be cited.
+In addition, if the software authors ask that a paper should be cited, that should be respected.
+Note that there are situations where it's appropriate to cite the software, the software paper, or both.
+
 \textbf{Derived software}

### 2016-03-16 sc-principles/software-citation-principles.tex
adding a paragraph on software peer review in 5

@@
 
+\textbf{Software peer review}
+
+This document does not discuss peer review of software, which is an important issue but is
+also mostly out of scope in the context of software citation principles.
+Since the goal of software citation is to identify the software
+that has been used in a scholarly product, whether or not that software has been peer-reviewed
+is irrelevant.  One possible exception would be if the peer-review status of the software should
+be part of the metadata, but the working group does not believe this to be part of the
+minimal metadata needed to identify the software.
+
 \textbf{Citations in text}

### 2016-03-17 sc-principles/software-citation-principles.tex
updating the discussion on UIDs

@@
 
-\textbf{Unique identifiers} \katznote{Is this discussion topic needed?  I would like to delete it.}
+\textbf{Unique identifiers}
 
-A UID points to a specific software, but multiple UIDs might point to the same software.
-\LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}
+The ``unique'' in a UID means that it points to a unique, specific software.
+However, multiple UIDs might point to the same software.
+This is not recommended, but is possible.
+We strongly recommend that if there is already a UID for a version of software, no additional UID should be
+created.  Multiple UIDs can lead to split credit, which does not help anyone.  In particular, we
+urge the use of standard DOIs as the UIDs for software.
+
+%\LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}
 

### 2016-03-17 sc-principles/software-citation-principles.tex
smoothing some writing in discussion section
adding note about identifier discussion in data citation implementation paper

@@
 
-In the principles, we did not define what software should be cited, but how it should be cited.
+The software citation principles do not define what software should be cited, but rather, how software should be cited.
 What software should be cited is the decision of the author(s) of the research work, in the context of community norms and practices, and in most scientific communities, these are currently in flux.
@@
 created.  Multiple UIDs can lead to split credit, which does not help anyone.  In particular, we
-urge the use of standard DOIs as the UIDs for software.
+urge the use of standard DOIs as the UIDs for software. \katznote{This topic is discussed in detail in the `Unique Identification' section of \cite{10.7717/peerj-cs.1}}
 

### 2016-03-17 sc-principles/software-citation-principles.tex
fixed some unintentional bullets in table 2

--- 2016-03-16 sc-principles/software-citation-principles.tex
@@
 Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
-Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet \\
+Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
 Researcher            & find software to implement task          & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
@@
 Publisher                  & publish papers that cite software        & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \\
-Indexer                     & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet \\
-Domain group           & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet \\
-Library\slash archive & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet \\
+Indexer                     & build catalog of software               & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &            & \textbullet \\
+Domain group           & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
+Library\slash archive  & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
 Repository                 & show scientific impact of holdings       & \textbullet &             &             &             &             & \textbullet & \textbullet &             &             \\

### 2016-03-18 sc-principles/software-citation-principles.tex
updates in unique identification discussion is section 5

@@
 Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product.
-\item \textbf{Credit and Attribution}:
+\item \textbf{Credit and Attribution}: \label{principle:credit}
 Software citations should facilitate giving scholarly credit and normative
@@
 all software. 
-\item \textbf{Unique Identification}:
+\item \textbf{Unique Identification}: \label{principle:uid}
 A software citation should include a method for identification that is
@@
 
-\textbf{Unique identifiers}
+\textbf{Unique identification}
 
-The ``unique'' in a UID means that it points to a unique, specific software.
-However, multiple UIDs might point to the same software.
+The Unique identification principle (\ref{principle:uid}) calls for ``a method for identification that is machine actionable, globally unique, interoperable, and recognized by a community.''
+What this means for data is discussed in detail in the `Unique Identification' section of a report by the FORCE11 Data Citation Implementation Group (DCIG)~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
+This report also lists examples of identifiers that match these criteria including DOIs, PURLs, Handles, ARKS, and NBNs.
+For software, we recommend the use of DOIs as the unique identifier due to their common usage and acceptance, particularly as they are the standard for other digital products such as publications.
+
+Note that the ``unique'' in a UID means that it points to a unique, specific software. However, multiple UIDs might point to the same software.
 This is not recommended, but is possible.
-We strongly recommend that if there is already a UID for a version of software, no additional UID should be
-created.  Multiple UIDs can lead to split credit, which does not help anyone.  In particular, we
-urge the use of standard DOIs as the UIDs for software. \katznote{This topic is discussed in detail in the `Unique Identification' section of \cite{10.7717/peerj-cs.1}}
+We strongly recommend that if there is already a UID for a version of software, no additional UID should be created.
+Multiple UIDs can lead to split credit, which goes against the Credit and Attribution principle (\ref{principle:credit}).
 

### 2016-03-19 sc-principles/software-citation-principles.tex
fixing duplicate a

@@
 Note that some software which is or could be captured as part of data provenance may not be cited.
-Citation is a a record of software that is important to the research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
+Citation is a record of software that is important to the research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
 This implies that for a data research product, provenance data will include all cited software, but not necessarily vice versa.

### 2016-03-22 sc-principles/software-citation-principles.tex
utf-8 encoding

@@
 \newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}} %Arfon
-\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}} %Frank Lšffler
+\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}} %Frank Löffler
 \definecolor{darkgreen}{rgb}{0,0.6,0}
@@
 These software citation principles were originally based on an adaptation of the FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
-The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts and it is executable.
+The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts, it is updated more frequently, and it is executable.
 \katznote{are there any other difference between software and data that should be stated here?}  Also, while software can be considered a type of data, the converse is not generally true.
@@
 There are many reasons why this lack of both software citations in general and
-standard practices for software citation are of concern: \katznote{Are there any more
-items that should be added to this list?}
+standard practices for software citation are of concern:
 
@@
 
-The group gathered members (see Appendix~\ref{app:wg_members}) in April and May 2015, and then in June began work, with a number of meetings
+The group gathered members (see Appendix~\ref{app:wg_members}) in April and May 2015,
+and then in June began work, with a number of meetings
 and some off-line work by group members to gather materials documenting existing
@@
 
-The principles in this document should guide further development of software citation mechanisms and system, and the reader should be able to look at any particular example of software citation and see if it meets the principles.
+The principles in this document should guide further development of software citation
+mechanisms and system, and the reader should be able to look at any particular example
+of software citation and see if it meets the principles.
 
@@
 
-With close to 50 working group participants representing a range of research domains, the working group was tasked to
+With close to 50 working group participants (see Appendix~\ref{app:wg_members}) representing a range of research domains, the working group was tasked to
 document existing practices in their respective communities. A total of 47 documents were submitted by working group
@@
 
-\textbf{Summary of submitted resources}
+\subsection{General community/non domain-specific activities}
 
-\asnote{Should mention that a number of communities have already focussed their attention on issues around data citation and while
-this isn't exactly the same there are many similarities (what is different exactly?).}
-
-\textbf{General community/non domain-specific activities}
-
-Some of the most actionable work has come from the NSF Software Infrastructure for Sustained Innovation (SSI) program in the form of blog posts written by their community fellows:
+Some of the most actionable work has come from the UK Software Sustainability Institute (SSI) in the form of blog posts written by their community fellows:
 
@@
 authors of the software would like to be cited by others~\cite{ssi-citation-files}. While this isn't a formal metadata specification (it's not machine
-readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \cite{10.1371/journal.pone.0136631}.} \katznote{added this ref to the introduction - not sure if it is still needed here too}
+readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors and as noted in the introduction (\S\ref{sec:intro}), there is evidence that authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
+
 In a later post on the SSI blog, Jackson gives a good overview of some of the approaches
@@
 
-\asnote{Should also discuss the WSSSPE~\cite{wssspe} series of workshops in this section}
-\scnote{Section 5.8 and Appendix J in \cite{WSSSPE3} has some preliminary work and discussion on the topic.}
+Most noteable of the community efforts are those of WSSSPE~\cite{wssspe} and
+SSI~\cite{ssi-workshops} who between them have run a series of workshops aimed
+at gathering together community members with an interest in improving the
+status of software in academia. \asnote{Not sure this is a sufficient one-line
+summary but couldn't think of anything better.} In each of the three years that
+WSSSPE workshops have run thus far, the participants have produced a report
+documenting the topics covered. Section 5.8 and Appendix J in \cite{WSSSPE3}
+has some preliminary work and discussion particularly relevant to this working
+group.
 
-\textbf{Domain-specific community activities}
+\subsection{Domain-specific community activities}
 
@@
 submission of papers in standard journals describing a piece of research
-software. While some journals (e.g., F1000Research) have traditionally accepted
+software, often known as software papers (see \S\ref{sec:software_papers}).
+While some journals (e.g., Transactions on Mathematical Software [TOMS], Bioinformatics, F1000Research) have traditionally accepted
 software submissions, the American Astronomical Society (AAS) have recently
@@
 \cite{aas-sofware-papers}. 
+Professional societies are in a good position to change their respective communities, as the publishers of journals and conveners of domain-specific conferences; as publishers they can change editorial policies (as AAS has done) and conferences are an opportunity to communicate and discuss these changes with their communities.
 
@@
 
-\asnote{Perhaps we should say something here about the potential importance of professional societies and their role in effecting changes in community practice?}
-
 Life sciences/NIH: Software Discovery Index Meeting Report \cite{software-discovery-index} a cross-disciplinary meeting
-but with a largely NIH-focus. Suggested minimal information about software (MIAS) standard. Discusses how a common set of metadata
+but with a largely biomedical-focus. Suggested minimal information about software (MIAS) standard. Discusses how a common set of metadata
 fields are critical for useful indexing - suggested candidates for this list.
@@
 
-\textbf{Existing principles}
+\subsection{Existing principles}
 
@@
 
-\textbf{What software to cite}
+\subsection{What software to cite}
 
@@
 
-\textbf{Software papers}
+\subsection{Software papers} \label{sec:software_papers}
 
@@
 
-\textbf{Derived software}
+\subsection{Derived software}
 
@@
 
-\textbf{Software peer review}
+\subsection{Software peer review}
 
@@
 
-\textbf{Citations in text}
+\subsection{Citations in text}
 
@@
 
-\textbf{Unique identification}
+\subsection{Unique identification}
 
@@
 
-\textbf{Access to software}
+\subsection{Access to software}
 
@@
 
-\textbf{Updates to this document}
+\subsection{Updates to this document}
 
@@
 Merc\`e Crosas, IQSS, Harvard University
+
+Scott Edmunds, GigaScience, BGI Hong Kong
 

### 2016-03-22 sc-principles/software-citation-principles.tex

@@
 AAS/GitHub workshop \cite{aas-software-index} dedicated to software citation,
-indexing and discoverability.
+indexing and discoverability, and the report from a Birds of a Feather session
+organized by the ASCL at the Astronomical Data Analysis Software and Systems
+(ADASS) XXV conference \cite{2015arXiv151207919A}.
 

### 2016-03-22 sc-principles/software-citation-principles.tex
adding tildas

--- 2016-03-22 sc-principles/software-citation-principles.tex
@@
 software submissions, the American Astronomical Society (AAS) have recently
-announced they will accept software papers in their journals
-\cite{aas-sofware-papers}. 
+announced they will accept software papers in their journals~\cite{aas-sofware-papers}. 
 Professional societies are in a good position to change their respective communities, as the publishers of journals and conveners of domain-specific conferences; as publishers they can change editorial policies (as AAS has done) and conferences are an opportunity to communicate and discuss these changes with their communities.
 
-In astronomy \& astrophysics: ASCL \cite{ascl}, AAS journals  and the joint
-AAS/GitHub workshop \cite{aas-software-index} dedicated to software citation,
-indexing and discoverability.
+In astronomy \& astrophysics: ASCL~\cite{ascl}, AAS journals  and the joint
+AAS/GitHub workshop~\cite{aas-software-index} dedicated to software citation,
+indexing and discoverability, and the report from a Birds of a Feather session
+organized by the ASCL at the Astronomical Data Analysis Software and Systems
+(ADASS) XXV conference~\cite{2015arXiv151207919A}.
 
-Life sciences/NIH: Software Discovery Index Meeting Report \cite{software-discovery-index} a cross-disciplinary meeting
+Life sciences/NIH: Software Discovery Index Meeting Report~\cite{software-discovery-index} a cross-disciplinary meeting
 but with a largely biomedical-focus. Suggested minimal information about software (MIAS) standard. Discusses how a common set of metadata
@@
 
-Geosciences: Ontosoft \cite{ontosoft} - A Community Software Commons for the Geosciences. Lots of attention given to the metadata
+Geosciences: Ontosoft~\cite{ontosoft} - A Community Software Commons for the Geosciences. Lots of attention given to the metadata
 required to describe, discover and execute research software.
-NSF geosciences workshop around data lifecycle, management and citation \cite{nsf-geo-data}.
+NSF geosciences workshop around data lifecycle, management and citation~\cite{nsf-geo-data}.
 This report includes many recommendations for data citation.

### 2016-03-22 sc-principles/software-citation-principles.tex
removing "commercial" when describing software that may no longer be available

@@
 For commercial software, the metadata should still provide information on how to access the specific software, but this may be a link to a web site that allows the software be purchased or a company's product number.
-As stated in the Persistence principle (\ref{principle:persistence}), we recognize that the commercial software version may no longer be available, but it still should be cited along with information about how it was accessed.
+As stated in the Persistence principle (\ref{principle:persistence}), we recognize that the software version may no longer be available, but it still should be cited along with information about how it was accessed.
 

### 2016-03-22 sc-principles/software-citation-principles.tex
adding new member who got missed
fixing two typos as explained in issue 79

@@
 
-Most noteable of the community efforts are those of WSSSPE~\cite{wssspe} and
+Most notable of the community efforts are those of WSSSPE~\cite{wssspe} and
 SSI~\cite{ssi-workshops} who between them have run a series of workshops aimed
@@
 However, this is really analogous to how any research builds on other research, where each research product just cites those products that it directly builds on, not those that it indirectly builds on.
-Understanding these chains of knowledge and credit have been part of history of science field for some time, though more recent work is suggesting more nuanced evaluation of the credit chains~\cite{casrai-credit, transitive_credit_json-ld}.
+Understanding these chains of knowledge and credit have been part of the history of science field for some time, though more recent work is suggesting more nuanced evaluation of the credit chains~\cite{casrai-credit, transitive_credit_json-ld}.
 
@@
 Alice Allen, Astrophysics Source Code Library
+
+Jay Jay Billings, Oak Ridge National Laboratory
 

### 2016-03-22 sc-principles/software-citation-principles.tex
added mention of Van de Sampel work; address issue #63

@@
 
+Van de Sompel et al.~\cite{VandeSompel2004} argued for registration of all units of scholarly communication, including software.\niemnote{suggesting this reference go here}
 Author guidelines in F1000Research, APA authors, SSI `publish or be damned'

### 2016-03-22 sc-principles/software-citation-principles.tex
removed orphan quote marks

@@
 We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
-Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.''
+Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.
 This recommendation matches that of the Purdue Online Writing Lab: ``Do not cite standard office software (e.g.

### 2016-03-22 sc-principles/software-citation-principles.tex
wording fix in subsection 5.1

@@
 These statements may also be true for software products.
-In general, we intend the software citation principles to cover what is minimum of what is necessary for software citation for the purpose of software identification.
+In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification.
 Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).

### 2016-03-22 sc-principles/software-citation-principles.tex
changed it's to it is in subsection 5.2
added Comput Phys Comm to subsection 4.2

@@
 software, often known as software papers (see \S\ref{sec:software_papers}).
-While some journals (e.g., Transactions on Mathematical Software [TOMS], Bioinformatics, F1000Research) have traditionally accepted
+While some journals (e.g., Transactions on Mathematical Software [TOMS], Bioinformatics, Computer Physics Communications, F1000Research) have traditionally accepted
 software submissions, the American Astronomical Society (AAS) have recently
@@
 In addition, if the software authors ask that a paper should be cited, that should be respected.
-Note that there are situations where it's appropriate to cite the software, the software paper, or both.
+Note that there are situations where it is appropriate to cite the software, the software paper, or both.
 

### 2016-03-22 sc-principles/software-citation-principles.tex
some formatting cleanup in various locations

@@
 In a blog post from 2012, Jackson discusses some of the pitfalls of trying to cite software in publications~\cite{ssi-how-to-cite}.
-He includes useful guidance for when to consider citing software as well as some ways to help `convince' journal editors
+He includes useful guidance for when to consider citing software as well as some ways to help ``convince'' journal editors
 to allow the inclusion of software citations.
 
-Wilson suggests that software authors include a `CITATION' file that documents exactly how the
-authors of the software would like to be cited by others~\cite{ssi-citation-files}. While this isn't a formal metadata specification (it's not machine
-readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors and as noted in the introduction (\S\ref{sec:intro}), there is evidence that authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
+Wilson suggests that software authors include a \texttt{CITATION} file that documents exactly how the authors of the software would like to be cited by others~\cite{ssi-citation-files}.
+While this is not a formal metadata specification (e.g., it is not machine readable) this does offer a solution for authors wishing to give explicit instructions to potential citing authors and as noted in the introduction (\S\ref{sec:intro}), there is evidence that authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
 
 In a later post on the SSI blog, Jackson gives a good overview of some of the approaches
-package authors have taken to automate the generation of citation entities such as bibtex entries~\cite{ssi-how-shalt-i-cite-thee},
+package authors have taken to automate the generation of citation entities such as \textsc{Bib}\TeX\ entries~\cite{ssi-how-shalt-i-cite-thee},
 and Knepley et al.\ do similarly~\cite{knepley2013accurately}.
@@
 efforts hosted (and sponsored) by funders and agencies in both the US (e.g., NSF,
-NIH, Alfred P. Sloan Foundation) and UK (e.g., SFTC, JISC) have run a number of
+NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC) have run a number of
 workshops with participants from across a range of disciplines.
@@
 
-One approach to increasing the `citability' of software is to encourage the
+One approach to increasing the ``citability'' of software is to encourage the
 submission of papers in standard journals describing a piece of research
@@
 The Unique identification principle (\ref{principle:uid}) calls for ``a method for identification that is machine actionable, globally unique, interoperable, and recognized by a community.''
-What this means for data is discussed in detail in the `Unique Identification' section of a report by the FORCE11 Data Citation Implementation Group (DCIG)~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
+What this means for data is discussed in detail in the ``Unique Identification'' section of a report by the FORCE11 Data Citation Implementation Group (DCIG)~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
 This report also lists examples of identifiers that match these criteria including DOIs, PURLs, Handles, ARKS, and NBNs.

### 2016-03-24 sc-principles/software-citation-principles.tex
added my middle initial
Fleshing out the NIH section

@@
 
-Life sciences/NIH: Software Discovery Index Meeting Report~\cite{software-discovery-index} a cross-disciplinary meeting
-but with a largely biomedical-focus. Suggested minimal information about software (MIAS) standard. Discusses how a common set of metadata
-fields are critical for useful indexing - suggested candidates for this list.
+The Life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers.
+The primary outcome of this workshop was the Software Discovery Index Meeting Report~\cite{software-discovery-index}.
+The authors of the report cite discuss what framework would be required for supporting a Software Discovery Index including the need for unique identifiers, how citations to these would be handled by publishers and the critical need for metadata to describe software packages.
 
@@
 
-\textbf{Existing efforts around metadata standards}
+\subsection{Existing efforts around metadata standards}
 
@@
 
-\textbf{Studies of author motivations}
+\subsection{Studies of author motivations}
 
@@
 
-Kyle Niemeyer, Oregon State University
+Kyle E.~Niemeyer, Oregon State University
 

### 2016-03-24 sc-principles/software-citation-principles.tex
Re-working astro section slightly

@@
 
-In astronomy \& astrophysics: ASCL~\cite{ascl}, AAS journals  and the joint
-AAS/GitHub workshop~\cite{aas-software-index} dedicated to software citation,
-indexing and discoverability, and the report from a Birds of a Feather session
-organized by the ASCL at the Astronomical Data Analysis Software and Systems
-(ADASS) XXV conference~\cite{2015arXiv151207919A}.
+In astronomy \& astrophysics: The Astronomy Source Code Library (ASCL) ~\cite{ascl}, is a website dedicated to the curation and indexing of software used in the astronomy-based literature.
+In 2015, the AAS and GitHub co-hosted a workshop~\cite{aas-software-index} dedicated to software citation,
+indexing and discoverability and more recently, a Birds of a Feather session was held at the Astronomical Data Analysis Software and Systems (ADASS) XXV conference~\cite{2015arXiv151207919A} which included discussion of software citation.
 
@@
 
-Geosciences: Ontosoft~\cite{ontosoft} - A Community Software Commons for the Geosciences. Lots of attention given to the metadata
+Geosciences: The Ontosoft~\cite{ontosoft} project describes itself as `A Community Software Commons for the Geosciences'. Lots of attention given to the metadata
 required to describe, discover and execute research software.

### 2016-03-24 sc-principles/software-citation-principles.tex
Unbreakable space before \cite{ascl}

@@
 
-In astronomy \& astrophysics: The Astronomy Source Code Library (ASCL) ~\cite{ascl}, is a website dedicated to the curation and indexing of software used in the astronomy-based literature.
-In 2015, the AAS and GitHub co-hosted a workshop~\cite{aas-software-index} dedicated to software citation,
-indexing and discoverability and more recently, a Birds of a Feather session was held at the Astronomical Data Analysis Software and Systems (ADASS) XXV conference~\cite{2015arXiv151207919A} which included discussion of software citation.
+In astronomy \& astrophysics: The Astronomy Source Code Library
+(ASCL)~\cite{ascl}, is a website dedicated to the curation and indexing of
+software used in the astronomy-based literature. In 2015, the AAS and GitHub
+co-hosted a workshop~\cite{aas-software-index} dedicated to software citation,
+indexing and discoverability and more recently, a Birds of a Feather session
+was held at the Astronomical Data Analysis Software and Systems (ADASS) XXV
+conference~\cite{2015arXiv151207919A} which included discussion of software
+citation.
 

### 2016-03-25 sc-principles/software-citation-principles.tex
Metadata section

@@
 
-AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}
-\acmnote{ CodeMeta? }
+Producing detailed specifications and recommendations for possible metadata standards to support software citation was not within the scope of this working group.
+However some discussion on the topic did occur and there is significant interest in the wider community to produce standards for describing research software.
+
+Notable examples of community efforts around metadata include the AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta}
 

### 2016-03-28 sc-principles/software-citation-principles.tex
adding new member, fixing alphabetization
Reworking subsections

@@
 
+While not usually expressed as software citation priciples, a number of groups have developed community guidelines around software and data citation.
+Van de Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of scholarly communication, including software.
+In a `Publish or be damned?
+An alternative impact manifesto for research software'~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as part of `The Research Software Impact Manifesto'.
+
 Perhaps in recognition of the broad range of research domains struggling with
@@
 group.
+
+In addition, a number of academic publishers such as APA~\cite{APA-guidelines} have recommendations for submitting authors on how to cite software and journals such as F1000Research~\cite{F1000Research}, SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow for submissions entirely focussed on research software.
 
@@
 
-\subsection{Existing principles}
-
-Van de Sompel et al.~\cite{VandeSompel2004} argued for registration of all units of scholarly communication, including software.\niemnote{suggesting this reference go here}
-Author guidelines in F1000Research, APA authors, SSI `publish or be damned'
-\cite{ssi-publish-or-be-damned} manifesto, workflow for publishing code using
-GitHub and Zenodo \cite{github-citable-code-guide}.
-\dmnote{There are many other journals publishing software descriptions, some even specialized in software, e.g. \href{http://www.openresearchcomputation.com}{Open Research Computation}.}
-
 \subsection{Existing efforts around metadata standards}
@@
 
-Notable examples of community efforts around metadata include the AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta}
-
-\subsection{Studies of author motivations}
-
-Survey of data citation practices \cite{Kratz_2015}
+Notable examples of community efforts around metadata include the AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta}.
 
@@
 
-Martin Fenner, Public Library of Science
-
 Merc\`e Crosas, IQSS, Harvard University
@@
 Christopher Erdmann, Harvard-Smithsonian CfA
+
+Martin Fenner, Public Library of Science
+
+Darel Finkbeiner, OSTI
 

### 2016-03-28 sc-principles/software-citation-principles.tex
Updating reference

@@
 group.
-
-In addition, a number of academic publishers such as APA~\cite{APA-guidelines} have recommendations for submitting authors on how to cite software and journals such as F1000Research~\cite{F1000Research}, SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow for submissions entirely focussed on research software.
+Up
+In addition, a number of academic publishers such as APA~\cite{APA-guidelines} have recommendations for submitting authors on how to cite software and journals such as F1000Research~\cite{F1000}, SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow for submissions entirely focussed on research software.
 

### 2016-03-29 sc-principles/software-citation-principles.tex
edits in a few scattered places - removing some old comments

@@
 The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts, it is updated more frequently, and it is executable.
-\katznote{are there any other difference between software and data that should be stated here?}  Also, while software can be considered a type of data, the converse is not generally true.
+% \katznote{are there any other difference between software and data that should be stated here?}  
+Also, while software can be considered a type of data, the converse is not generally true.
 
@@
 efforts hosted (and sponsored) by funders and agencies in both the US (e.g., NSF,
-NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC) have run a number of
+NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC, Wellcome Trust) have run a number of
 workshops with participants from across a range of disciplines.
@@
 Most notable of the community efforts are those of WSSSPE~\cite{wssspe} and
-SSI~\cite{ssi-workshops} who between them have run a series of workshops aimed
-at gathering together community members with an interest in improving the
-status of software in academia. \asnote{Not sure this is a sufficient one-line
-summary but couldn't think of anything better.} In each of the three years that
+SSI~\cite{ssi-workshops}, who between them have run a series of workshops aimed
+at gathering together community members with an interest in 1) defining the set of problems
+related to the role of software and associated people in research settings, particularly academia,
+2) discussing potential solutions to those problems, and 3) beginning to work on implementing
+some of those solutions.
+%\asnote{Not sure this is a sufficient one-line
+%summary but couldn't think of anything better.} \katznote{I've edited it a bit}
+In each of the three years that
 WSSSPE workshops have run thus far, the participants have produced a report
@@
 
-The Life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers.
+The life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers.
 The primary outcome of this workshop was the Software Discovery Index Meeting Report~\cite{software-discovery-index}.
@@
 
-Kyle E.~Niemeyer, Oregon State University
+Kyle E.~Niemeyer, Oregon State University (co-chair)
 

### 2016-03-30 sc-principles/software-citation-principles.tex
fixing Martin Fenner's affiliation
updating important principle text

@@
 Software should be considered a legitimate, citable product of research. 
-Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
-Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product.
+Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be part of the standard reference list for that research product and should not be separated.
+Software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
 \item \textbf{Credit and Attribution}: \label{principle:credit}
@@
 
-Martin Fenner, Public Library of Science
+Martin Fenner, DataCite
 

### 2016-03-31 sc-principles/software-citation-principles.tex
fix typo
edits to discussion of software papers based on #84

@@
 
-While not usually expressed as software citation priciples, a number of groups have developed community guidelines around software and data citation.
+While not usually expressed as software citation principles, a number of groups have developed community guidelines around software and data citation.
 Van de Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of scholarly communication, including software.
@@
 oriented towards citation of papers.
-Given a software paper and software, the question of which should be cited and when is
+Where software should be cited, the question of whether to cite the software itself or a software paper is
 complicated, and there is no simple answer.
@@
 important to the work, then the software paper should also be cited.
-In addition, if the software authors ask that a paper should be cited, that should be respected.
+In addition, if the software authors ask that a paper should be cited, that should typically be respected and the paper cited in addition to the software. 
+However, the software itself should still be cited.
 Note that there are situations where it is appropriate to cite the software, the software paper, or both.

### 2016-03-31 sc-principles/software-citation-principles.tex
Added references to DOAP and schema.org

@@
 
-Notable examples of community efforts around metadata include the AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta}.
+Notable examples around metadata include DOAP~\cite{DOAP}, an early metadata term set used by the Open Source Community, as well as more recent community efforts like the AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta}, or the terms and classes defined at \url{schema.org} related to the \url{https://schema.org/SoftwareApplication} class.
 

### 2016-03-31 sc-principles/software-citation-principles.tex
small change to reflect section reordering a few weeks ago

@@
 Wilson suggests that software authors include a \texttt{CITATION} file that documents exactly how the authors of the software would like to be cited by others~\cite{ssi-citation-files}.
-While this is not a formal metadata specification (e.g., it is not machine readable) this does offer a solution for authors wishing to give explicit instructions to potential citing authors and as noted in the introduction (\S\ref{sec:intro}), there is evidence that authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
+While this is not a formal metadata specification (e.g., it is not machine readable) this does offer a solution for authors wishing to give explicit instructions to potential citing authors and as noted in the motivation section (\S\ref{sec:intro}), there is evidence that authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
 

### 2016-03-31 sc-principles/software-citation-principles.tex
adding other WSSSPE report cites

@@
 Van de Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of scholarly communication, including software.
-In a `Publish or be damned?
-An alternative impact manifesto for research software'~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as part of `The Research Software Impact Manifesto'.
+In ``Publish or be damned?
+An alternative impact manifesto for research software''~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as part of `The Research Software Impact Manifesto'.
 
@@
 In each of the three years that
-WSSSPE workshops have run thus far, the participants have produced a report
-documenting the topics covered. Section 5.8 and Appendix J in \cite{WSSSPE3}
+WSSSPE workshops have run thus far, the participants have produced a report~\cite{WSSSPE1,WSSSPE2,WSSSPE3}
+documenting the topics covered. Section 5.8 and Appendix J in the WSSSPE3 report~\cite{WSSSPE3}
 has some preliminary work and discussion particularly relevant to this working

### 2016-03-31 sc-principles/software-citation-principles.tex
small changes to text in 4

@@
 
-In astronomy \& astrophysics: The Astronomy Source Code Library
+In astronomy and astrophysics: The Astronomy Source Code Library
 (ASCL)~\cite{ascl}, is a website dedicated to the curation and indexing of
@@
 
-The life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers.
+In the life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers.
 The primary outcome of this workshop was the Software Discovery Index Meeting Report~\cite{software-discovery-index}.
@@
 
-Geosciences: The Ontosoft~\cite{ontosoft} project describes itself as `A Community Software Commons for the Geosciences'. Lots of attention given to the metadata
+In the geosciences: The Ontosoft~\cite{ontosoft} project describes itself as `A Community Software Commons for the Geosciences'. Lots of attention given to the metadata
 required to describe, discover and execute research software.

### 2016-03-31 sc-principles/software-citation-principles.tex
changed in 5 to match changes in definition of importance in 1

@@
 The software citation principles do not define what software should be cited, but rather, how software should be cited.
-What software should be cited is the decision of the author(s) of the research work, in the context of community norms and practices, and in most scientific communities, these are currently in flux.
-In general, we believe that any software on which a research product directly depends should be cited, if it significantly contributes to the research product.
-Again, the specific decision of what is significant must be made by the author(s) of the product.
+What software should be cited is the decision of the author(s) of the research work in the context of community norms and practices, and in most scientific communities, these are currently in flux.
+In general, we believe that software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
+Again, the specific decision of what is appropriate must be made by the author(s) of the product.
 However, an illustrative example is the use of Microsoft Excel in research.
@@
 Currently, and for the foreseeable future, software papers are being published and cited,
-in addition to software itself being cited, as many community norms and practices are
+in addition to software itself being published and cited, as many community norms and practices are
 oriented towards citation of papers.
-Where software should be cited, the question of whether to cite the software itself or a software paper is
-complicated, and there is no simple answer.
-As discussed in the Importance principle (\ref{principle:importance}), \textit{the software
-itself should also be cited if it is directly used and was important to the work}.
+As discussed in the Importance principle (\ref{principle:importance}) and the discussion above, \textit{the software
+itself should be cited on the same basis as any other research product; authors should cite the appropriate set of software products}.
 If a software paper exists and it contains results (performance, validation, etc.) that are
 important to the work, then the software paper should also be cited.
-In addition, if the software authors ask that a paper should be cited, that should typically be respected and the paper cited in addition to the software.
-However, the software itself should still be cited.
-Note that there are situations where it is appropriate to cite the software, the software paper, or both.
+In addition, if the software authors ask that a paper should be cited, that should typically be respected and the paper cited \textit{in addition to} the software being cited.
+%There are situations where it is appropriate to cite the software, the software paper, or both.
 
@@
 In the Importance principle (\ref{principle:importance}), we state that
-``software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product.''
+``software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.''
 In the case of one code that is derived from another code, citing the derived software may appear to not credit those responsible for the original software, nor recognize its role in the work that used the derived software.

### 2016-04-05 sc-principles/software-citation-principles.tex
Section 1 suggested changes.

@@
 cited.  We talk about how we think such decisions might be made in the
-discussion section.
+discussion section~(\S\ref{sec:discussion}).
 
-\katznote{consider changing the language in the principles to be more requirements-like -- e.g.
-use ``shall'' a lot; related: consider making the principles machine-readable (has implications on language}
+\katznote{consider changing the language in the principles to be more
+requirements-like -- e.g. use ``shall'' a lot; related: consider making the
+principles machine-readable (has implications on language}
 
 \begin{enumerate}
-\item \textbf{Importance}: \label{principle:importance}
-Software should be considered a legitimate, citable product of research.
-Software citations should be accorded the same importance in the scholarly record as citations of other research products, such as publications and data; they should be part of the standard reference list for that research product and should not be separated.
-Software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
+\item \textbf{Importance}: \label{principle:importance} Software should be
+considered a legitimate and citable product of research. Software citations should
+be accorded the same importance in the scholarly record as citations of other
+research products, such as publications and data; they should be part of the
+standard reference list for a research product and should not be omitted or separated.
+Software should be cited on the same basis as any other research products such as
+papers or books, that is, authors should cite the appropriate set of software
+products just as they cite the appropriate set of papers.
+
 \item \textbf{Credit and Attribution}: \label{principle:credit}
-Software citations should facilitate giving scholarly credit and normative
+Software citations should facilitate giving scholarly credits and normative
 and legal attribution to all contributors to the software, recognizing
@@
 all software.
+
 \item \textbf{Unique Identification}: \label{principle:uid}
@@
 machine actionable, globally unique, interoperable, and recognized by
-a community.
+at least a community of the corresponding domain experts, and preferably by 
+general public researchers.
+
 \item \textbf{Persistence}: \label{principle:persistence}
-Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
-\item \textbf{Accessibility}: \label{principle:accessibility}
-Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.
-\item \textbf{Specificity}:
-Software citations should facilitate identification of, and access to, the specific version of software that was used.
-Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
+Unique identifiers and metadata describing the software and its disposition
+should persist~-- even beyond the lifespan of the software they describe.
+
+\item \textbf{Accessibility}: \label{principle:accessibility} Software citations
+should permit and facilitate access to the software itself and to its
+associated metadata, documentation, data, and other materials necessary
+for both humans and machines to make informed use of the referenced software.
+
+\item \textbf{Specificity}: Software citations should facilitate identification
+of, and access to, the specific version of software that was used. Software
+identification should be as specific as necessary, such as using version
+numbers, revision numbers, or variants such as platforms.
 %\item \textbf{Interoperability and Flexibility}:
@@
 
-These software citation principles were originally based on an adaptation of the FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
-The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts, it is updated more frequently, and it is executable.
+These software citation principles were originally based on an adaptation of the
+FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were
+modified based on discussions of the FORCE11 Software Citation Working 
+Group~(see \S\ref{app:wg_members} for members), the information from the use
+cases in~\S\ref{sec:use_cases}, and the related work in~\S\ref{sec:related_work}. 
+The adaptations and modifications have been made
+because software, while similar to data in terms of not traditionally having
+been cited in publications, is also different than data in that it can be used
+to express or explain concepts, it is updated more frequently, and it is
+executable.
 % \katznote{are there any other difference between software and data that should be stated here?}
-Also, while software can be considered a type of data, the converse is not generally true.
+Also, while software can be considered a type of data, the converse is not
+generally true.
 

### 2016-04-05 sc-principles/software-citation-principles.tex
undoing one SCTC change in section 1

@@
 \item \textbf{Credit and Attribution}: \label{principle:credit}
-Software citations should facilitate giving scholarly credits and normative
+Software citations should facilitate giving scholarly credit and normative
 and legal attribution to all contributors to the software, recognizing

### 2016-04-05 sc-principles/software-citation-principles.tex
Section 2 suggested changes.

--- 2016-04-05 sc-principles/software-citation-principles.tex
@@
 \usepackage{rotating}
+\usepackage[inline]{enumitem}
 
@@
 \newcommand{\GMnote}[1]{ {\textcolor{violet} { ***GM: #1}}} % August (Gus) Muench
+
+\newenvironment{italicquotes}
+{\begin{quote}\itshape}
+{\end{quote}}
 
@@
 
-As the process of science has become increasingly digital, scientific outputs and products have
-grown beyond simple papers and books to include software, data, and other electronic
-components.  Scientific knowledge is embedded in these components.  And papers and books
-themselves are also becoming increasingly digital, allowing them to become executable
-and reproducible.  As we move towards this future where science is performed in and recorded
-as a variety of linked digital products, the characteristics and properties that developed for
-books and papers need to be applied to all digital products and possibly adjusted.  Here, we are concerned specifically
-with the citation of software products.  The challenge is not just the textual citation of software in a paper, but the more general identification of software used within the research process.
+As the process of science has become increasingly digital, scientific outputs
+and products have grown beyond simply papers and books to include software,
+data, and other electronic components such as presentation slides, posters, 
+(interactive)  graphs, maps, websites (e.g., blogs and forums),  multimedia 
+(e.g., audio and  video lectures).  Scientific knowledge is embedded in these
+components. And papers and books themselves are also becoming increasingly
+digital, allowing them to become executable and reproducible. As we move towards
+this future where science is performed in and recorded as a variety of linked
+digital products, the characteristics and properties that developed for books
+and papers need to be applied to all digital products and possibly adjusted.
+Here, we are concerned specifically with the citation of software products. The
+challenge is not just the textual citation of software in a paper, but the more
+general identification of software used within the research process.
 
-Software and other digital resources currently appear in publications in very inconsistent ways.
-For example, a random sample of 90 articles in the biology literature
-found seven different ways that software was mentioned, including simple
-names in the full-text, URLs in footnotes, and different kinds of mentions
-in references lists: project names or websites, user manuals, publications
-that describe or introduce the software~\cite{howison2015jasist}.
+Software and other digital resources currently appear in publications in very
+inconsistent ways. For example, a random sample of 90 articles in the biology
+literature found seven different ways that software was mentioned, including
+simple names in the full-text, URLs in footnotes, and different kinds of
+mentions in references lists: project names or websites, user manuals,
+publications that describe or introduce the software~\cite{howison2015jasist}.
 Table~\ref{tab:mentions} shows examples of these varied forms of software
@@
 %Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format.
-Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses.
+Many of these kinds of mentions fail to perform the functions needed of
+citations, and their very diversity and frequent informality undermines the
+integration of software work into bibliometrics and other analyses.
 %In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
-Studies on data and facility citation have shown similar results~\cite{10.1371/journal.pone.0136631, mayernik_poster, parsons_duerr_minster}.
+Studies on data and facility citation have shown similar
+results~\cite{10.1371/journal.pone.0136631, mayernik_poster,
+parsons_duerr_minster}.
 % Add Hwang et al. from WSSSPE3 LJH
@@
 \rowcolors{2}{white}{gray!25}
-\begin{table}[htbp]
+\begin{table}[tb]
 \caption{Varieties of software mentions in publications, from Howison and Bullard~\cite{howison2015jasist}.}
@@
 \begin{itemize}
-\item Understanding Science: Software is a product of research, and by not citing it, we leave holes in the record of scientific progress.
-\item Academic Credit: Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
-credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
-\item Finding Software: Citations enable the specific software used in a research product to be discovered. Additional researchers can then use the same software for different purposes, leading
-to credit for those responsible for the software.
-\item Reproducibility: Citation of specific software used is necessary for reproducibility, but is not sufficient.  Additional information such as configurations and platform issues are also needed.
+\item Understanding Science: Software is a product of research, and by not
+citing it, we leave holes in the record of scientific progress.
+
+\item Academic Credit: Academic researchers at all levels, including students,
+postdocs, faculty, and staff, should be credited for the software products they
+develop and contribute to, particularly when those products enable or further
+research done by others.
+
+\item Discovering Software: Citations enable the specific software used in a
+research product to be found. Additional researchers can then use the same
+software for different purposes, leading to credit for those responsible for the
+software.
+
+\item Reproducibility: Citation of specific software used is necessary for
+reproducibility, but is not sufficient. Additional information such as
+configurations and platform issues are also needed.
+
 \end{itemize}
 
-The FORCE11 Software Citation Working Group~\cite{f11scwg} was created in April 2015, with the following mission statement:
+The FORCE11 Software Citation Working Group~\cite{f11scwg} was created in April
+2015 with the following mission statement:
+% following quote updated from webpage https://www.force11.org/group/software-citation-working-group
+\advance\leftmargini -1em
+\begin{italicquotes} 
+The software citation working group is a cross-team committee leveraging the
+perspectives from a variety of existing initiatives working on software citation
+to produce a consolidated set of citation principles in order to encourage broad
+adoption of a consistent policy for software citation across disciplines and
+venues. The working group will review existing efforts and make a set of
+recommendations. These recommendations will be put of for endorsement by the
+organizations represented by this group and others that play an important role
+in the community.
 
-\begin{quote}
-The software citation working group is a committee that will leverage the perspectives of a variety of existing initiatives working on software citation to produce a consolidated set of citation principles in order to encourage broad adoption of a consistent policy for software citation across disciplines and venues.
-The working group will review existing efforts and make a set of recommendations.
-These recommendations will be put up for endorsement by the organizations represented by this group and others that play an important role in the community.
-
-The group will produce a set of principles, illustrated with working examples, and a plan for dissemination and distribution.
-This group will not be producing detailed specifications for implementation although it may review and discuss possible technical solutions.\end{quote}
+The group will produce a set of principles, illustrated with working examples,
+and a plan for dissemination and distribution. This group will not be producing
+detailed specifications for implementation although it may review and discuss
+possible technical solutions.
+\end{italicquotes}
 
 The group gathered members (see Appendix~\ref{app:wg_members}) in April and May 2015,
-and then in June began work, with a number of meetings
+and then began work in June, with a number of meetings
 and some off-line work by group members to gather materials documenting existing
 practices in member disciplines; gather materials from workshops and other reports;
-review those materials, identifying overlaps and differences; and draft this resulting document,
-which will be presented and discussed at the Force2016 Conference \cite{force2016}
-in April 2016.  We expect that
-this discussion may lead to a second, final version, and we also
+review those materials, identifying overlaps and differences; and subsequently draft this resulting document,
+which will be presented and discussed at the Force2016 Conference~\cite{force2016}
+in April 2016.  We expect that this discussion may lead to a second, final version, and we also
 plan to have a follow-on working group that will work with stakeholders to ensure that
@@
 The principles in this document should guide further development of software citation
-mechanisms and system, and the reader should be able to look at any particular example
+mechanisms and systems, and the reader should be able to look at any particular example
 of software citation and see if it meets the principles.
 
-In the next section (\S\ref{sec:use_cases}), we provide some detailed context in which
-software citation is important, by means of use cases.  In \S\ref{sec:related_work}, we
-summarize and analyze a large amount of previous work and thinking in this area.  In \S\ref{sec:discussion}, we discuss issues related to the principles stated in \S\ref{sec:principles}, and finally,
-in \S\ref{sec:examples} we give examples of how these software citation principles
+In the next section~(\S\ref{sec:use_cases}), we provide some detailed context in
+which software citation is important, by means of use cases. In
+\S\ref{sec:related_work}, we summarize and analyze a large amount of previous
+work and thinking in this area. In \S\ref{sec:discussion}, we discuss issues
+related to the principles stated in \S\ref{sec:principles}, and finally, in
+\S\ref{sec:examples} we give examples of how these software citation principles
 could be applied, related back to the use cases in \S\ref{sec:use_cases}.

### 2016-04-05 sc-principles/software-citation-principles.tex
Section 3 suggested changes.

@@
 
-We have documented and analyzed a set of use cases related to software citation in \cite{SC-Use-Cases}.
-Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case.
-\scnote{Perhaps the long caption of Table 2 can be mostly moved here and be replaced with a shorter version there.}
+We have documented and analyzed a set of use cases related to software citation
+in \cite{SC-Use-Cases}. Table~\ref{tab:use_cases} summarizes these use cases and
+makes clear what the requirements are for software citation in each case. In the following list,
+we clarify a few notions used in the table.
+\begin{itemize}
 
-%
+\item ``Researcher'' includes both academic researchers (e.g., postdoc,
+tenure-track faculty member) and research software engineers;
+
+\item ``Publisher'' includes both traditional publishers that publish text
+and\slash or software papers as well as archives such as Zenodo that directly
+publish software;
+
+\item ``Repository'' refers to public software repositories such as Astronomy
+Source Code Library (ASCL)~\cite{ascl}, GitHub, Netlib, Comprehensive R Archive
+Network (CRAN), institutional repositories;
+
+\item ``Funder'' is a group that funds software or work using software;
+
+\item Examples of ``Indexer'' include Scopus, Web of Science, Google Scholar,
+and Microsoft Academic Search;
+
+\item ``Domain group'' includes bioCADDIE~\cite{bioCADDIE}, Computational
+Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc.;
+
+\item ``Unique identifier'' refers to identifiers such as a DOI, ARK, or PURL;
+
+\item``reproduce'' can mean reproduction, replication, verification, validation,
+repeatability, and\slash or utility.
+\end{itemize}
+
 \newcommand*\rot[1]{\begin{turn}{90} #1 \end{turn}}%
 \rowcolors{4}{gray!25}{white}
-\begin{table}[htbp]
-\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases}.
-``Researcher'' includes both academic researchers (e.g., postdoc, tenure-track faculty member) and research software engineers;
-``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software;
-``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
-examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
-domain groups include bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc.;
-``repository'' refers to public software repositories such as Astronomy Source Code Library (ASCL)~\cite{ascl},
-GitHub, Netlib, Comprehensive R Archive Network (CRAN), institutional repositories;
-``funder'' is a group that funds software or work using software; and
-``unique ID'' refers to identifiers such as a DOI, ARK, or PURL.}
+\begin{table}[tbhp]
+\caption{Use cases for software citation, adapted from~\cite{SC-Use-Cases}.}
+%``Researcher'' includes both academic researchers (e.g., postdoc, tenure-track
+%faculty member) and research software engineers;
+%``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software;
+%``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
+%examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
+%domain groups include bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc.;
+%``repository'' refers to public software repositories such as Astronomy Source Code Library (ASCL)~\cite{ascl},
+%GitHub, Netlib, Comprehensive R Archive Network (CRAN), institutional repositories;
+%``funder'' is a group that funds software or work using software; and
+%``unique ID'' refers to identifiers such as a DOI, ARK, or PURL.
+%}
 \centering
@@
  \cmidrule{3-11}
-Stakeholder & Use\slash wants to 	&  \rot{Software name} & \rot{Author(s)} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Unique identifier} & \rot{Indexed citations} & \rot{Contributor role} & \rot{Software license}  \\
+Stakeholder & wants to 	&  \rot{Software name} & \rot{Author(s)} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Unique identifier} & \rot{Indexed citations} & \rot{Contributor role} & \rot{Software license}  \\
 \midrule

### 2016-04-05 sc-principles/software-citation-principles.tex
Section 4 suggested changes.

@@
 
-With close to 50 working group participants (see Appendix~\ref{app:wg_members}) representing a range of research domains, the working group was tasked to
-document existing practices in their respective communities. A total of 47 documents were submitted by working group
-participants, with the life sciences, astrophysics, and geosciences being particularly well-represented in the submitted
-resources.
+With approximately 50 working group participants (see Appendix~\ref{app:wg_members})
+representing a range of research domains, the working group was tasked to
+document existing practices in their respective communities. A total of 47
+documents were submitted by working group participants, with the life sciences,
+astrophysics, and geosciences being particularly well-represented in the
+submitted resources.
 
@@
 
-Some of the most actionable work has come from the UK Software Sustainability Institute (SSI) in the form of blog posts written by their community fellows:
+Some of the most actionable work has come from the UK Software Sustainability
+Institute (SSI) in the form of blog posts written by their community fellows:
 
-In a blog post from 2012, Jackson discusses some of the pitfalls of trying to cite software in publications~\cite{ssi-how-to-cite}.
-He includes useful guidance for when to consider citing software as well as some ways to help ``convince'' journal editors
-to allow the inclusion of software citations.
+In a blog post from 2012, Jackson discusses some of the pitfalls of trying to
+cite software in publications~\cite{ssi-how-to-cite}. He includes useful
+guidance for when to consider citing software as well as some ways to help
+``convince'' journal editors to allow the inclusion of software citations.
 
-Wilson suggests that software authors include a \texttt{CITATION} file that documents exactly how the authors of the software would like to be cited by others~\cite{ssi-citation-files}.
-While this is not a formal metadata specification (e.g., it is not machine readable) this does offer a solution for authors wishing to give explicit instructions to potential citing authors and as noted in the motivation section (\S\ref{sec:intro}), there is evidence that authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
+Wilson suggests that software authors include a \texttt{CITATION} file that
+documents exactly how the authors of the software would like to be cited by
+others~\cite{ssi-citation-files}. While this is not a formal metadata
+specification (e.g., it is not machine readable) this does offer a solution for
+authors wishing to give explicit instructions to potential citing authors and as
+noted in the motivation section (\S\ref{sec:intro}), there is evidence that
+authors follow instructions if they exist~\cite{10.1371/journal.pone.0136631}.
 
-In a later post on the SSI blog, Jackson gives a good overview of some of the approaches
-package authors have taken to automate the generation of citation entities such as \textsc{Bib}\TeX\ entries~\cite{ssi-how-shalt-i-cite-thee},
-and Knepley et al.\ do similarly~\cite{knepley2013accurately}.
+In a later post on the SSI blog, Jackson gives a good overview of some of the
+approaches package authors have taken to automate the generation of citation
+entities such as \textsc{Bib}\TeX\ entries~\cite{ssi-how-shalt-i-cite-thee}, and
+Knepley et al.\ do similarly~\cite{knepley2013accurately}.
 
-While not usually expressed as software citation principles, a number of groups have developed community guidelines around software and data citation.
-Van de Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of scholarly communication, including software.
-In ``Publish or be damned?
-An alternative impact manifesto for research software''~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as part of `The Research Software Impact Manifesto'.
+While not usually expressed as software citation principles, a number of groups
+have developed community guidelines around software and data citation. Van de
+Sompel et al.~\cite{VandeSompel2004} argue for registration of all units of
+scholarly communication, including software. In ``Publish or be damned? An
+alternative impact manifesto for research
+software''~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as
+part of ``The Research Software Impact Manifesto.''
 
@@
 the challenge of better recognizing the role of software, a number of community
-efforts hosted (and sponsored) by funders and agencies in both the US (e.g., NSF,
-NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC, Wellcome Trust) have run a number of
-workshops with participants from across a range of disciplines.
+efforts hosted (and sponsored) by funders and agencies in both the US (e.g.,
+NSF, NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC, Wellcome Trust)
+have run a number of workshops with participants from across a range of
+disciplines.
 
@@
 SSI~\cite{ssi-workshops}, who between them have run a series of workshops aimed
-at gathering together community members with an interest in 1) defining the set of problems
-related to the role of software and associated people in research settings, particularly academia,
-2) discussing potential solutions to those problems, and 3) beginning to work on implementing
-some of those solutions.
+at gathering together community members with an interest in
+\begin{enumerate*}[series=InlineList, before=\hspace{-0.6ex}]
+    \item defining the set
+of problems related to the role of software and associated people in research
+settings, particularly academia,
+    \item discussing potential solutions to those
+problems, 
+\item beginning to work on implementing some of those solutions.
+  \end{enumerate*}
 %\asnote{Not sure this is a sufficient one-line
 %summary but couldn't think of anything better.} \katznote{I've edited it a bit}
-In each of the three years that
-WSSSPE workshops have run thus far, the participants have produced a report~\cite{WSSSPE1,WSSSPE2,WSSSPE3}
-documenting the topics covered. Section 5.8 and Appendix J in the WSSSPE3 report~\cite{WSSSPE3}
-has some preliminary work and discussion particularly relevant to this working
-group.
-Up
-In addition, a number of academic publishers such as APA~\cite{APA-guidelines} have recommendations for submitting authors on how to cite software and journals such as F1000Research~\cite{F1000}, SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow for submissions entirely focussed on research software.
+In each of the three years that WSSSPE workshops have run thus far, the
+participants have produced a report~\cite{WSSSPE1,WSSSPE2,WSSSPE3} documenting
+the topics covered. Section~5.8 and Appendix~J in the WSSSPE3
+report~\cite{WSSSPE3} has some preliminary work and discussion particularly
+relevant to this working group.  In addition, a number of academic publishers
+such as APA~\cite{APA-guidelines} have recommendations for submitting authors on
+how to cite software and journals such as F1000Research~\cite{F1000},
+SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow
+for submissions entirely focussed on research software.
 
@@
 
-One approach to increasing the ``citability'' of software is to encourage the
+One approach to increasing software ``citability''   is to encourage the
 submission of papers in standard journals describing a piece of research
 software, often known as software papers (see \S\ref{sec:software_papers}).
-While some journals (e.g., Transactions on Mathematical Software [TOMS], Bioinformatics, Computer Physics Communications, F1000Research) have traditionally accepted
-software submissions, the American Astronomical Society (AAS) have recently
-announced they will accept software papers in their journals~\cite{aas-sofware-papers}.
-Professional societies are in a good position to change their respective communities, as the publishers of journals and conveners of domain-specific conferences; as publishers they can change editorial policies (as AAS has done) and conferences are an opportunity to communicate and discuss these changes with their communities.
+While some journals (e.g., Transactions on Mathematical Software (TOMS),
+Bioinformatics, Computer Physics Communications, F1000Research) have
+traditionally accepted software submissions, the American Astronomical Society
+(AAS) has recently announced they will accept software papers in their
+journals~\cite{aas-sofware-papers}. Professional societies are in a good
+position to change their respective communities, as the publishers of journals
+and conveners of domain-specific conferences; as publishers they can change
+editorial policies (as AAS has done) and conferences are an opportunity to
+communicate and discuss these changes with their communities.
 
@@
 co-hosted a workshop~\cite{aas-software-index} dedicated to software citation,
-indexing and discoverability and more recently, a Birds of a Feather session
-was held at the Astronomical Data Analysis Software and Systems (ADASS) XXV
+indexing and discoverability in astrophysics. More recently, a Birds of a Feather session was
+held at the Astronomical Data Analysis Software and Systems (ADASS) XXV
 conference~\cite{2015arXiv151207919A} which included discussion of software
@@
 
-In the life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers.
-The primary outcome of this workshop was the Software Discovery Index Meeting Report~\cite{software-discovery-index}.
-The authors of the report cite discuss what framework would be required for supporting a Software Discovery Index including the need for unique identifiers, how citations to these would be handled by publishers and the critical need for metadata to describe software packages.
+In the life sciences: In May 2014, the NIH held a workshop aimed at helping the
+biomedical community discover, cite, and reuse software written by their peers.
+The primary outcome of this workshop was the Software Discovery Index Meeting
+Report~\cite{software-discovery-index}. The authors of the report discuss
+what framework would be required for supporting a Software Discovery Index
+including the need for unique identifiers, how citations to these would be
+handled by publishers, and the critical need for metadata to describe software
+packages.
 
-In the geosciences: The Ontosoft~\cite{ontosoft} project describes itself as `A Community Software Commons for the Geosciences'. Lots of attention given to the metadata
-required to describe, discover and execute research software.
-NSF geosciences workshop around data lifecycle, management and citation~\cite{nsf-geo-data}.
-This report includes many recommendations for data citation.
-\LJHnote{ The Seismological Research Letters, Electronic Seismologists publishes papers on software.} %text needs to be cleaned up LJH%
+In the geosciences: The Ontosoft~\cite{ontosoft} project describes itself as ``A
+Community Software Commons for the Geosciences.'' Much attention was given to the
+metadata required to describe, discover, and execute research software. NSF sponsored
+Geo-Data Workshop 2011~\cite{nsf-geo-data} revolved around data lifecycle, management, and
+citation. The workshop report includes many recommendations for data
+citation. \LJHnote{ The Seismological Research Letters, Electronic Seismologists
+publishes papers on software.} %text needs to be cleaned up LJH% dy
 
@@
 
-Producing detailed specifications and recommendations for possible metadata standards to support software citation was not within the scope of this working group.
-However some discussion on the topic did occur and there is significant interest in the wider community to produce standards for describing research software.
+Producing detailed specifications and recommendations for possible metadata
+standards to support software citation was not within the scope of this working
+group. However some discussion on the topic did occur and there was significant
+interest in the wider community to produce standards for describing research
+software metadata.
 
-Notable examples around metadata include DOAP~\cite{DOAP}, an early metadata term set used by the Open Source Community, as well as more recent community efforts like the AAS workshop~\cite{aas-software-index}, Project CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta}, or the terms and classes defined at \url{schema.org} related to the \url{https://schema.org/SoftwareApplication} class.
+Notable examples around metadata include DOAP~\cite{DOAP}, an early metadata
+term set used by the Open Source Community, as well as more recent community
+efforts like the AAS workshop~\cite{aas-software-index}, Project
+CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta},
+or the terms and classes defined at \href{https://schema.org}{www.schema.org} related to the
+\href{https://schema.org/SoftwareApplication}{\texttt{SoftwareApplication}} class.
 

### 2016-04-05 sc-principles/software-citation-principles.tex
further formatting changes in Section 3.

--- 2016-04-05 sc-principles/software-citation-principles.tex
@@
 in \cite{SC-Use-Cases}. Table~\ref{tab:use_cases} summarizes these use cases and
-makes clear what the requirements are for software citation in each case. In the following list,
-we clarify a few notions used in the table.
+makes clear what the requirements are for software citation in each case. In that table,
+we use the following definitions:
 \begin{itemize}
@@
 \item ``Researcher'' includes both academic researchers (e.g., postdoc,
-tenure-track faculty member) and research software engineers;
+tenure-track faculty member) and research software engineers.
 
@@
 and\slash or software papers as well as archives such as Zenodo that directly
-publish software;
+publish software.
 
@@
 Source Code Library (ASCL)~\cite{ascl}, GitHub, Netlib, Comprehensive R Archive
-Network (CRAN), institutional repositories;
+Network (CRAN), institutional repositories.
 
-\item ``Funder'' is a group that funds software or work using software;
+\item ``Funder'' is a group that funds software or work using software.
 
-\item Examples of ``Indexer'' include Scopus, Web of Science, Google Scholar,
-and Microsoft Academic Search;
+\item ``Indexer'' examples include Scopus, Web of Science, Google Scholar,
+and Microsoft Academic Search.
 
 \item ``Domain group'' includes bioCADDIE~\cite{bioCADDIE}, Computational
-Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc.;
+Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc..
 
-\item ``Unique identifier'' refers to identifiers such as a DOI, ARK, or PURL;
+\item ``Unique identifier'' refers to identifiers such as a DOI, ARK, or PURL.
 
-\item``reproduce'' can mean reproduction, replication, verification, validation,
+\item``Reproduce'' can mean reproduction, replication, verification, validation,
 repeatability, and\slash or utility.
@@
 
-If certain metadata are not available, alternatives may be provided:
-\begin{itemize}
-\item version number and release date: download date
-\item location\slash repository: contact name\slash email if not publicly
-  available
-\end{itemize}
+In some cases, if particular metadata are not available, alternatives may be provided.
+For example, if the version number and release date are not available, the download date can be used.
+And the contact name\slash email is an alternative to the location\slash repository.
 

### 2016-04-05 sc-principles/software-citation-principles.tex
changes to changes in 4

@@
 co-hosted a workshop~\cite{aas-software-index} dedicated to software citation,
-indexing and discoverability in astrophysics. More recently, a Birds of a Feather session was
+indexing, and discoverability in astrophysics. More recently, a Birds of a Feather session was
 held at the Astronomical Data Analysis Software and Systems (ADASS) XXV
-conference~\cite{2015arXiv151207919A} which included discussion of software
+conference~\cite{2015arXiv151207919A} that included discussion of software
 citation.

### 2016-04-05 sc-principles/software-citation-principles.tex
update to importance principle as per Martin Fenner in issue 99

@@
 be accorded the same importance in the scholarly record as citations of other
-research products, such as publications and data; they should be part of the
-standard reference list for a research product and should not be omitted or separated.
+research products, such as publications and data; they should be included in the
+metadata of the citing work, for example in the reference list of a journal article,
+and should not be omitted or separated.
 Software should be cited on the same basis as any other research products such as
@@
 products just as they cite the appropriate set of papers.
+
+software that was cited should be included in the metadata of the citing work, for example in the reference list of a journal article
 

### 2016-04-05 sc-principles/software-citation-principles.tex
changes from during call

@@
 
-\katznote{consider changing the language in the principles to be more
-requirements-like -- e.g. use ``shall'' a lot; related: consider making the
-principles machine-readable (has implications on language}
+%\katznote{consider changing the language in the principles to be more
+%requirements-like -- e.g. use ``shall'' a lot; related: consider making the
+%principles machine-readable (has implications on language}
 
@@
 products just as they cite the appropriate set of papers.
-
-software that was cited should be included in the metadata of the citing work, for example in the reference list of a journal article
 
@@
 data, and other electronic components such as presentation slides, posters, 
-(interactive)  graphs, maps, websites (e.g., blogs and forums),  multimedia 
+(interactive)  graphs, maps, websites (e.g., blogs and forums), and multimedia 
 (e.g., audio and  video lectures).  Scientific knowledge is embedded in these
@@
 
-\katznote{please check to make sure these are all still right}
-
 Alberto Accomazzi, Harvard-Smithsonian CfA
@@
 Carl Boettiger, University of California,  Berkeley
+
+Jed Brown, University of Colorado Boulder
 

### 2016-04-05 sc-principles/software-citation-principles.tex
changes to generally use research in place of science, to make it more clear we are trying to impact other areas, such as engineering, humanities, etc.

@@
 
-As the process of science has become increasingly digital, scientific outputs
+As the process of research\footnote{We use the term `research' in this document to include work intended to increase human knowledge and benefit society, in science, engineering, humanities, and other areas.} has become increasingly digital, research outputs
 and products have grown beyond simply papers and books to include software,
@@
 (interactive)  graphs, maps, websites (e.g., blogs and forums), and multimedia 
-(e.g., audio and  video lectures).  Scientific knowledge is embedded in these
+(e.g., audio and  video lectures).  Research knowledge is embedded in these
 components. And papers and books themselves are also becoming increasingly
 digital, allowing them to become executable and reproducible. As we move towards
-this future where science is performed in and recorded as a variety of linked
+this future where research is performed in and recorded as a variety of linked
 digital products, the characteristics and properties that developed for books
@@
 \begin{itemize}
-\item Understanding Science: Software is a product of research, and by not
-citing it, we leave holes in the record of scientific progress.
+\item Understanding Research Fields: Software is a product of research, and by not
+citing it, we leave holes in the record of research of progress in those fields.
 
@@
 plan to have a follow-on working group that will work with stakeholders to ensure that
-these principles impact the scientific process.
+these principles impact the research process.
 
@@
 The software citation principles do not define what software should be cited, but rather, how software should be cited.
-What software should be cited is the decision of the author(s) of the research work in the context of community norms and practices, and in most scientific communities, these are currently in flux.
+What software should be cited is the decision of the author(s) of the research work in the context of community norms and practices, and in most research communities, these are currently in flux.
 In general, we believe that software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
@@
 We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
-Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.
+Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), research presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.
 This recommendation matches that of the Purdue Online Writing Lab: ``Do not cite standard office software (e.g.
@@
 
-The goals of software include the linked ideas of crediting those responsible for software and understanding the dependencies of scientific products on specific software.
+The goals of software include the linked ideas of crediting those responsible for software and understanding the dependencies of research products on specific software.
 In the Importance principle (\ref{principle:importance}), we state that

### 2016-04-05 sc-principles/software-citation-principles.tex
edits to use cases table to match Google Doc, addressing #105

@@
  \cmidrule{3-11}
-Stakeholder & wants to 	&  \rot{Software name} & \rot{Author(s)} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Unique identifier} & \rot{Indexed citations} & \rot{Contributor role} & \rot{Software license}  \\
+Stakeholder & wants to 	& \rot{Unique identifier} &  \rot{Software name} & \rot{Author(s)} & \rot{Contributor role} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Indexed citations} & \rot{Software license}  \\
 \midrule
-Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             & \textbullet \\
-Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             & \textbullet \\
-Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet\\
-Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &            & \\
-Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet  \\
-Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
-Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
-Researcher            & find software to implement task          & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Publisher                  & publish software paper                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             & \\
-Publisher                  & publish papers that cite software        & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \\
-Indexer                     & build catalog of software               & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &            & \textbullet \\
-Domain group           & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
-Library\slash archive  & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &            & \textbullet \\
-Repository                 & show scientific impact of holdings       & \textbullet &             &             &             &             & \textbullet & \textbullet &             &             \\
-Funder                       & show how funded software has been used   & \textbullet &             &             &             &             & \textbullet & \textbullet &             &             \\
-Evaluator                   & evaluate contributions of researcher   &                  & \textbullet &             & \textbullet &             & \textbullet & \textbullet & \textbullet &             \\
+Researcher            & 1.\ use someone else's software for a paper      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             \\
+Researcher            & 2.\ use someone else's software for new software & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & 3.\ contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & 4.\ find citations of software               & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             \\
+Researcher            & 5.\ get credit for software development      & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & 6.\ ``reproduce'' analysis                   & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & 7.\ benchmark software                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
+Researcher            & 8.\ find software to implement task          & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textbullet \\
+Publisher             & 9.\ publish software paper                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             \\
+Publisher             & 10.\ publish papers that cite software        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             \\
+Indexer               & 11.\ build catalog of software                & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet \\
+Domain group\slash library\slash archive          & 12.\ build software catalog\slash registry                & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             \\
+Repository            & 13.\ show scientific impact of holdings       & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             \\
+Funder                & 14.\ show how funded software has been used   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             \\
+Evaluator             & 15.\ evaluate contributions of researcher     & \textbullet &             & \textbullet & \textbullet &             & \textbullet &             & \textbullet &             \\
+Citation manager      & 16.\ store software entry        			 & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             \\
 \bottomrule

### 2016-04-05 sc-principles/software-citation-principles.tex

@@
 
-\item ``Repository'' refers to public software repositories such as Astronomy
-Source Code Library (ASCL)~\cite{ascl}, GitHub, Netlib, Comprehensive R Archive
-Network (CRAN), institutional repositories.
-
 \item ``Funder'' is a group that funds software or work using software.
@@
 
-\item ``Domain group'' includes bioCADDIE~\cite{bioCADDIE}, Computational
-Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc..
+\item ``Domain group\slash library\slash archive'' includes the Astronomy Source Code Library (ASCL)~\cite{ascl}, 
+bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics (CIG)~\cite{CIG}, libraries, institutional archives, etc.
 
-\item ``Unique identifier'' refers to identifiers such as a DOI, ARK, or PURL.
+\item ``Repository'' refers to public software repositories such as GitHub, Netlib, Comprehensive R Archive
+Network (CRAN), and institutional repositories.
 
-\item``Reproduce'' can mean reproduction, replication, verification, validation,
-repeatability, and\slash or utility.
+\item ``Unique identifier'' refers to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL.
+
+\item ``Reproduce'' can mean actions focused on reproduction, replication, verification, validation, repeatability, and\slash or utility.
+
+\item ``Citation manager`` refers to people and organizations that create scholarly reference 
+management software and websites including Zotero, Mendeley, EndNote, RefWorks, BibDesk, etc., 
+that manage citation information and semi-automatically insert those citations into research products.
+
 \end{itemize}

### 2016-04-05 sc-principles/software-citation-principles.tex
changes in 5 to match some suggestions from SCTC.

@@
 oriented towards citation of papers.
-As discussed in the Importance principle (\ref{principle:importance}) and the discussion above, \textit{the software
+As discussed in the Importance principle~(\ref{principle:importance}) and the discussion above, \textit{the software
 itself should be cited on the same basis as any other research product; authors should cite the appropriate set of software products}.
@@
 The goals of software include the linked ideas of crediting those responsible for software and understanding the dependencies of research products on specific software.
-In the Importance principle (\ref{principle:importance}), we state that
+In the Importance principle~(\ref{principle:importance}), we state that
 ``software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.''
@@
 
-The Unique identification principle (\ref{principle:uid}) calls for ``a method for identification that is machine actionable, globally unique, interoperable, and recognized by a community.''
+The Unique Identification principle~(\ref{principle:uid}) calls for ``a method for identification that is machine actionable, globally unique, interoperable, and recognized by a community.''
 What this means for data is discussed in detail in the ``Unique Identification'' section of a report by the FORCE11 Data Citation Implementation Group (DCIG)~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
@@
 Rather, the metadata should provide enough information that the software can be accessed.
-If the software is free, the metadata will likely provide an identifier that can be resolved to a URL that contains the specific version of the software that is being cited.
-For commercial software, the metadata should still provide information on how to access the specific software, but this may be a link to a web site that allows the software be purchased or a company's product number.
-As stated in the Persistence principle (\ref{principle:persistence}), we recognize that the software version may no longer be available, but it still should be cited along with information about how it was accessed.
+If the software is free, the metadata will likely provide an identifier that can be resolved to a URL pointing to the specific version of the software being cited.
+For commercial software, the metadata should still provide information on how to access the specific software, but this may be a company's product number or a link to a web site that allows the software be purchased.
+As stated in the Persistence principle~(\ref{principle:persistence}), we recognize that the software version may no longer be available, but it still should be cited along with information about how it was accessed.
 

### 2016-04-07 sc-principles/software-citation-principles.tex
added clarification statement about use cases
adding a line to discussion on what software to cite

@@
 in \cite{SC-Use-Cases}. Table~\ref{tab:use_cases} summarizes these use cases and
-makes clear what the requirements are for software citation in each case. In that table,
+makes clear what the requirements are for software citation in each case. 
+Each example represents a particular stakeholder performing an activity related to citing software, with the given metadata as information needed to do that.
+In that table,
 we use the following definitions:
@@
 Word, Excel) or programming languages.
-Provide references only for specialized software.''~\cite{powl-citing-software}
+Provide references only for specialized software.''~\cite{powl-citing-software}  In other words, if using different software could produce different data or results, then the software used should be cited.
 

### 2016-04-15 sc-principles/software-citation-principles.tex
Adding Carole Goble to participants
First pass at an abstract

@@
 \begin{abstract}
+Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation.
+
+Inspired by the activities of the Force11 working group focussed on data citation (ref), this document summarises the recommendations of the Force11 Software Citation Working Group and their activities beween June 2015 and April 2016.
+Based on a review of existing community practices, the goal of the working group was to produce a consolidated set of citation principles that may encourage broad adoption of a consistent policy for software citation across disciplines and venues.
+
+Our work is presented here as a set of software citation principles, a discussion of the motivations for developing the princples, reviews of existing community practise and a discussion of the requirements these principles would place upon different stakeholders.
+
+Working examples and possible technical solutions for how these princples can be implemented will be disussed in a separate paper.
+
 \end{abstract}
@@
 
+Carole Goble, The University of Manchester, Software Sustainability Institute
+
 Paul Groth, Elsevier Labs

### 2016-04-15 sc-principles/software-citation-principles.tex
minor fixes to abstract

--- 2016-04-07 sc-principles/software-citation-principles.tex
@@
 \begin{abstract}
+Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation.
+
+Inspired by the activities of the FORCE11 working group focussed on data citation, this document summarizes the recommendations of the FORCE11 Software Citation Working Group and their activities between June 2015 and April 2016.
+Based on a review of existing community practices, the goal of the working group was to produce a consolidated set of citation principles that may encourage broad adoption of a consistent policy for software citation across disciplines and venues.
+
+Our work is presented here as a set of software citation principles, a discussion of the motivations for developing the principles, reviews of existing community practice and a discussion of the requirements these principles would place upon different stakeholders.
+
+Working examples and possible technical solutions for how these principles can be implemented will be discussed in a separate paper.
+
 \end{abstract}

### 2016-04-15 sc-principles/software-citation-principles.tex
abstract into single paragraph

@@
 \begin{abstract}
+
 Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation.
-
 Inspired by the activities of the FORCE11 working group focussed on data citation, this document summarizes the recommendations of the FORCE11 Software Citation Working Group and their activities between June 2015 and April 2016.
 Based on a review of existing community practices, the goal of the working group was to produce a consolidated set of citation principles that may encourage broad adoption of a consistent policy for software citation across disciplines and venues.
-
 Our work is presented here as a set of software citation principles, a discussion of the motivations for developing the principles, reviews of existing community practice and a discussion of the requirements these principles would place upon different stakeholders.
-
 Working examples and possible technical solutions for how these principles can be implemented will be discussed in a separate paper.

### 2016-04-15 sc-principles/software-citation-principles.tex
very minor changes in abstract

@@
 Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation.
-Inspired by the activities of the FORCE11 working group focussed on data citation, this document summarizes the recommendations of the FORCE11 Software Citation Working Group and their activities between June 2015 and April 2016.
+Inspired by the activities of the FORCE11 working group focussed on data citation, this document summarizes the recommendations of the FORCE11 Software Citation Working Group and its activities between June 2015 and April 2016.
 Based on a review of existing community practices, the goal of the working group was to produce a consolidated set of citation principles that may encourage broad adoption of a consistent policy for software citation across disciplines and venues.
-Our work is presented here as a set of software citation principles, a discussion of the motivations for developing the principles, reviews of existing community practice and a discussion of the requirements these principles would place upon different stakeholders.
+Our work is presented here as a set of software citation principles, a discussion of the motivations for developing the principles, reviews of existing community practice, and a discussion of the requirements these principles would place upon different stakeholders.
 Working examples and possible technical solutions for how these principles can be implemented will be discussed in a separate paper.

### 2016-04-17 sc-principles/software-citation-principles.tex
added a reference to trec as support

@@
 develop and contribute to, particularly when those products enable or further
-research done by others.
+research done by others.\footnote{Providing recognition of software can have tremendous economic impact as demonstrated by the role of TREC in information retrieval \cite{trec-economic-impact}}
 

### 2016-04-17 sc-principles/software-citation-principles.tex
adding MIcah Altman to participant list in Appendix

@@
 
-Notable examples around metadata include DOAP~\cite{DOAP}, an early metadata
+Content specifications for software metadata vary across communities, and include 
+include DOAP~\cite{DOAP}, an early metadata
 term set used by the Open Source Community, as well as more recent community
 efforts like the AAS workshop~\cite{aas-software-index}, Project
-CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, CodeMeta~\cite{codemeta},
+CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, 
 or the terms and classes defined at \href{https://schema.org}{www.schema.org} related to the
-\href{https://schema.org/SoftwareApplication}{\texttt{SoftwareApplication}} class.
+\href{https://schema.org/SoftwareApplication}{\texttt{SoftwareApplication}} class.   In addition, 
+language-specific software metadata schemes are in widespread use, including the Debian 
+package format~\cite{Debian_policy}, Python package descriptions~\cite{pypi}, and R package descriptions~\cite{wickham_r_2015}, but these are typically
+conceived for software build, packaging,and distribution rather than citation.  CodeMeta~\cite{codemeta} 
+has created a crosswalk among these software metadata schemes and an exchange format that 
+allows software repositories to effectively interoperate.
 
@@
 
+Micah Altman, MIT
+
 Jay Jay Billings, Oak Ridge National Laboratory

### 2016-04-17 sc-principles/software-citation-principles.tex
Clarified metadata for use case data are basic/minimal.
Closes issue #123. Also removed some old comments.

@@
 \begin{table}[tbhp]
-\caption{Use cases for software citation, adapted from~\cite{SC-Use-Cases}.}
-%``Researcher'' includes both academic researchers (e.g., postdoc, tenure-track
-%faculty member) and research software engineers;
-%``publisher'' includes both traditional publishers that publish text and\slash or software papers as well as archives such as Zenodo that directly publish software;
-%``reproduce'' can mean reproduction, replication, verification, validation, repeatability, and\slash or utility;
-%examples of indexers include Scopus, Web of Science, Google Scholar, and Microsoft Academic Search;
-%domain groups include bioCADDIE~\cite{bioCADDIE}, Computational Infrastructure for Geodynamics (CIG)~\cite{CIG}, etc.;
-%``repository'' refers to public software repositories such as Astronomy Source Code Library (ASCL)~\cite{ascl},
-%GitHub, Netlib, Comprehensive R Archive Network (CRAN), institutional repositories;
-%``funder'' is a group that funds software or work using software; and
-%``unique ID'' refers to identifiers such as a DOI, ARK, or PURL.
-%}
+\caption{Use cases and basic metadata requirements for software citation, adapted from~\cite{SC-Use-Cases}.}
 \centering

### 2016-04-17 sc-principles/software-citation-principles.tex
Reordered use case table columns. Addresses #126.
Shifted stakeholder column to far right. Also clarified that stakeholders are examples.

@@
 \scriptsize\setlength{\tabcolsep}{2.5pt}
-\begin{tabular}{@{}l l c c c c c c c c c@{}}
+\begin{tabular}{@{}l l c c c c c c c c l@{}}
 \toprule
- & & \multicolumn{9}{c}{Requirements} \\
- \cmidrule{3-11}
-Stakeholder & wants to 	& \rot{Unique identifier} &  \rot{Software name} & \rot{Author(s)} & \rot{Contributor role} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Indexed citations} & \rot{Software license}  \\
+ & \multicolumn{9}{c}{Requirements} & \\
+ \cmidrule{2-10}
+Use case 	& \rot{Unique identifier} &  \rot{Software name} & \rot{Author(s)} & \rot{Contributor role} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Indexed citations} & \rot{Software license} & Example stakeholder(s) \\
 \midrule
-Researcher            & 1.\ use someone else's software for a paper      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             \\
-Researcher            & 2.\ use someone else's software for new software & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & 3.\ contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & 4.\ find citations of software               & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             \\
-Researcher            & 5.\ get credit for software development      & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & 6.\ ``reproduce'' analysis                   & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & 7.\ benchmark software                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
-Researcher            & 8.\ find software to implement task          & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textbullet \\
-Publisher             & 9.\ publish software paper                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             \\
-Publisher             & 10.\ publish papers that cite software        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             \\
-Indexer               & 11.\ build catalog of software                & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet \\
-Domain group\slash library\slash archive          & 12.\ build software catalog\slash registry                & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             \\
-Repository            & 13.\ show scientific impact of holdings       & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             \\
-Funder                & 14.\ show how funded software has been used   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             \\
-Evaluator             & 15.\ evaluate contributions of researcher     & \textbullet &             & \textbullet & \textbullet &             & \textbullet &             & \textbullet &             \\
-Citation manager      & 16.\ store software entry        			 & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             \\
+1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & Researcher \\
+2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
+3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
+4.\ Find citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             & Researcher \\
+5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet &             & \textbullet & Researcher \\
+6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
+7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
+8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textbullet & Researcher \\
+9.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & Publisher \\
+10.\ Publish papers that cite software           & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             & Publisher \\
+11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & Indexer \\
+12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & Domain group\slash library\slash archive \\
+13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             & Repository \\
+14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             & Funder \\
+15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textbullet &             & \textbullet &             & \textbullet &             & Evaluator \\
+16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             & Citation manager \\
 \bottomrule

### 2016-04-17 sc-principles/software-citation-principles.tex
minor fixes to footnotes on second page

@@
 
-As the process of research\footnote{We use the term `research' in this document to include work intended to increase human knowledge and benefit society, in science, engineering, humanities, and other areas.} has become increasingly digital, research outputs
+As the process of research\footnote{We use the term ``research'' in this document to include work intended to increase human knowledge and benefit society, in science, engineering, humanities, and other areas.} has become increasingly digital, research outputs
 and products have grown beyond simply papers and books to include software,
@@
 develop and contribute to, particularly when those products enable or further
-research done by others.\footnote{Providing recognition of software can have tremendous economic impact as demonstrated by the role of TREC in information retrieval \cite{trec-economic-impact}}
+research done by others.\footnote{Providing recognition of software can have tremendous economic impact as demonstrated by the role of Text REtrieval Conference (TREC) in information retrieval~\cite{trec-economic-impact}.}
 

### 2016-04-17 sc-principles/software-citation-principles.tex
change in 5.1 to address https://github.com/force11/force11-scwg/issues/129

@@
 What software should be cited is the decision of the author(s) of the research work in the context of community norms and practices, and in most research communities, these are currently in flux.
-In general, we believe that software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
-Again, the specific decision of what is appropriate must be made by the author(s) of the product.
+In general, we believe that software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers, perhaps following the
+FORCE11 Data Citation Working Group principles, which state, ``In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited.''~\cite{data-citation-principles}
+
+Again, the specific decision of what is appropriate to cite must be made by the author(s) of the product.
 However, an illustrative example is the use of Microsoft Excel in research.

### 2016-04-17 sc-principles/software-citation-principles.tex
adding new subsection in 5 on types of software to address https://github.com/force11/force11-scwg/issues/116

@@
 
+\subsection{Types of software}
+
+The principles and discussion in this document have generally been written to focus on software
+as source code.
+However, we recognize that some software is only available as an executable or as a container, while other software may be available as a service.
+We believe the principles apply to all of these forms of software, though the implementation of them will certainly differ based on software type.
+When software exists as both source code and another type, we recommend that the source code be cited.
+
 \subsection{Access to software}

### 2016-04-17 sc-principles/software-citation-principles.tex
removing duplicate word

@@
 
-Content specifications for software metadata vary across communities, and include 
+Content specifications for software metadata vary across communities, and 
 include DOAP~\cite{DOAP}, an early metadata

### 2016-04-17 sc-principles/software-citation-principles.tex
changing "the" to "a"

@@
 Note that some software which is or could be captured as part of data provenance may not be cited.
-Citation is a record of software that is important to the research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
+Citation is a record of software that is important to a research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
 This implies that for a data research product, provenance data will include all cited software, but not necessarily vice versa.

### 2016-04-17 sc-principles/software-citation-principles.tex
adding an unnumbered subsubsection in 5 to discuss the relationship of identifiers to software versions, as discussed in https://github.com/force11/force11-scwg/issues/98 and https://github.com/force11/force11-scwg/issues/115

@@
 
+\subsubsection*{Software versions and identifiers}
+
+There are at least three different potential relationships between identifiers and versions of software.
+\begin{enumerate}
+\item An identifier can point to a specific version of a piece of software.
+\item An identifier can point to the piece of software, effectively all versions of the software.
+\item An identifier can point to the latest version of a piece of software.
+\end{enumerate}
+It is possible that a given piece of software may have identifiers of all three types.  And in addition,
+there may be one or more software papers, each with an identifier.
+
+While we often need to cite a specific version of software, we may also need a way to cite the
+software in general, or the latest version, and to link multiple releases together, perhaps for
+the purpose of understand citations to the software.  The principles in \S\ref{sec:principles} are
+intended to be applicable at all levels, and to all types of identifiers, such as DOIs, RRIDs, etc.,
+though we again recommend when possible the use of DOIs that identify specific versions of
+source code.
+
 \subsection{Types of software}

### 2016-04-17 sc-principles/software-citation-principles.tex
adding a note on RRIDs as discussed in https://github.com/force11/force11-scwg/issues/114

@@
 though we again recommend when possible the use of DOIs that identify specific versions of
-source code.
+source code.  We note that RRIDs were developed by the FORCE11 Resource Identification Initiative
+and have been discussed for use to identify software packages (not specific versions),
+thought the FORCE11 Resource Identification Technical Specifications Working
+Group~\cite{f11rrridtswg} says ``Information resources like software are better suited to
+the Software Citation WG.''
+There is currently a lack of consensus on the use of RRIDs for software.
 

### 2016-04-18 sc-principles/software-citation-principles.tex
adding citation limits subsection in 5 to address https://github.com/force11/force11-scwg/issues/122

@@
 
+\subsection{Citations limits}
+
+This set of software citation principles, if followed, will cause the number of software citations
+in scholarly products to increase, thus causing the number of overall citations to increase.
+Some scholarly products that are or were once printed on paper, such as journal articles,
+may have strict
+limits on the number of citations they permit, partially imposed to reduce production costs.
+Such limits are counter to our recommendation.
+\textit{Citation limits, if absolutely required, should accommodate rather than stifle software citation} (and resource citation, and data citation.)
+
 \subsection{Unique identification}

### 2016-04-18 sc-principles/software-citation-principles.tex
updating text citation discussion in 5 re https://github.com/force11/force11-scwg/issues/100

@@
 
-In terms of how citations should be expressed as text, such
-as in journals or other human readable material, there are currently at least four styles (AMS, APA, Chicago, MLA). These have been illustrated with examples by Lipson~\cite{lipson2011cite}.
-This is another area that is in flux at this time.
+Citations in the scholarly literature are formatted according to the citation style (e.g., AMS, APA, Chicago, MLA) used by that publication.
+(Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
+a) a label indicating that this is software, e.g.\ [Computer program], and b) support for version information, e.g.\ Version 1.8.7.}
 

### 2016-04-18 sc-principles/software-citation-principles.tex
update to citation limit paragraph to address https://github.com/force11/force11-scwg/issues/122

@@
 
-This set of software citation principles, if followed, will cause the number of software citations
-in scholarly products to increase, thus causing the number of overall citations to increase.
-Some scholarly products that are or were once printed on paper, such as journal articles,
-may have strict
-limits on the number of citations they permit, partially imposed to reduce production costs.
-Such limits are counter to our recommendation.
-\textit{Citation limits, if absolutely required, should accommodate rather than stifle software citation} (and resource citation, and data citation.)
+This set of software citation principles, if followed, will cause the number of software citations in scholarly products to increase, thus causing the number of overall citations to increase.
+Some scholarly products, such as journal articles, may have strict limits on the number of citations they permit.
+Such limits are counter to our recommendation, and \textit{we recommend that publishers using strict limits for the number of citations add specific instructions regarding software citations to their author guidelines to not disincentivize software citation.}
 

### 2016-04-18 sc-principles/software-citation-principles.tex
further clarified that metadata requirements are basic
Introduced recommended metadata to address issue #108
Introduced open bullet (\textopenbullet) for recommended metadata.
Switched license and contributor role to recommended, and added (recommended) description and keywords columns.

@@
 \usepackage[inline]{enumitem}
+\usepackage{textcomp} % added for textopenbullet
 
@@
 
+\item ``Description'' refers to some description of the software such as an abstract, README, or other text description.
+
+\item ``Keywords'' refers to keywords or tags used to categorize the software.
+
 \item ``Reproduce'' can mean actions focused on reproduction, replication, verification, validation, repeatability, and\slash or utility.
@@
 \scriptsize\setlength{\tabcolsep}{2.5pt}
-\begin{tabular}{@{}l l c c c c c c c c l@{}}
+\begin{tabular}{@{}l l c c c c c c c c c c l@{}}
 \toprule
- & \multicolumn{9}{c}{Requirements} & \\
- \cmidrule{2-10}
-Use case 	& \rot{Unique identifier} &  \rot{Software name} & \rot{Author(s)} & \rot{Contributor role} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Indexed citations} & \rot{Software license} & Example stakeholder(s) \\
+ & \multicolumn{11}{c}{Basic requirements} & \\
+ \cmidrule{2-12}
+Use case 	& \rot{Unique identifier} &  \rot{Software name} & \rot{Author(s)} & \rot{Contributor role} & \rot{Version number} & \rot{Release date} & \rot{Location\slash repository} & \rot{Indexed citations} & \rot{Software license} & \rot{Description} & \rot{Keywords} & Example stakeholder(s) \\
 \midrule
-1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & Researcher \\
-2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
-3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
-4.\ Find citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             & Researcher \\
-5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet &             & \textbullet & Researcher \\
-6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
-7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textbullet & Researcher \\
-8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textbullet & Researcher \\
-9.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & Publisher \\
-10.\ Publish papers that cite software           & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             & Publisher \\
-11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & Indexer \\
-12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & Domain group\slash library\slash archive \\
-13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             & Repository \\
-14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             & Funder \\
-15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textbullet &             & \textbullet &             & \textbullet &             & Evaluator \\
-16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             & Citation manager \\
+1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher \\
+2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher \\
+3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textopenbullet & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
+4.\ Find citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher \\
+5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textopenbullet &             & \textbullet & \textbullet &             &             &             &             & Researcher \\
+6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
+7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
+8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Researcher \\
+9.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             &             &             & Publisher \\
+10.\ Publish papers that cite software           & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Publisher \\
+11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Indexer \\
+12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & \textopenbullet & \textopenbullet & Domain group\slash library\slash archive \\
+13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Repository \\
+14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder \\
+15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textopenbullet &             & \textbullet &             & \textbullet &             &             &             & Evaluator \\
+16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Citation manager \\
 \bottomrule

### 2016-04-18 sc-principles/software-citation-principles.tex
adding a sentence on references in page limits re https://github.com/force11/force11-scwg/issues/122#issuecomment-211573142

@@
 This set of software citation principles, if followed, will cause the number of software citations in scholarly products to increase, thus causing the number of overall citations to increase.
-Some scholarly products, such as journal articles, may have strict limits on the number of citations they permit.
-Such limits are counter to our recommendation, and \textit{we recommend that publishers using strict limits for the number of citations add specific instructions regarding software citations to their author guidelines to not disincentivize software citation.}
+Some scholarly products, such as journal articles, may have strict limits on the number of citations they permit, or page limits that include reference sections.
+Such limits are counter to our recommendation, and \textit{we recommend that publishers using strict limits for the number of citations add specific instructions regarding software citations to their author guidelines to not disincentivize software citation.} Similarly, publishers should not include references in the content counted against page
+limits.
 

### 2016-04-18 sc-principles/software-citation-principles.tex
Adding sciencecodemanifesto

@@
 software''~\cite{ssi-publish-or-be-damned}, Chue Hong lists nine principles as
-part of ``The Research Software Impact Manifesto.''
+part of ``The Research Software Impact Manifesto.'' In the
+``Science Code Manifesto'`~\cite{sciencecodemanifesto}, the founding signatories
+cite five core principles (Code, Copyright, Citation, Credit, Curation) for
+scientific software.
 
@@
   \end{enumerate*}
-%\asnote{Not sure this is a sufficient one-line
-%summary but couldn't think of anything better.} \katznote{I've edited it a bit}
 In each of the three years that WSSSPE workshops have run thus far, the

### 2016-04-18 sc-principles/software-citation-principles.tex
NIH feedback

@@
 The primary outcome of this workshop was the Software Discovery Index Meeting
-Report~\cite{software-discovery-index}. The authors of the report discuss
+Report~\cite{software-discovery-index} which was shared with the community for
+public comment and feedback. The authors of the report discuss
 what framework would be required for supporting a Software Discovery Index

### 2016-04-18 sc-principles/software-citation-principles.tex
fixing bib entry

@@
 part of ``The Research Software Impact Manifesto.'' In the
-``Science Code Manifesto'`~\cite{sciencecodemanifesto}, the founding signatories
+``Science Code Manifesto''~\cite{sciencecodemanifesto}, the founding signatories
 cite five core principles (Code, Copyright, Citation, Credit, Curation) for

### 2016-04-18 sc-principles/software-citation-principles.tex
Adding additional metadata efforts

--- 2016-04-18 sc-principles/software-citation-principles.tex
@@
 term set used by the Open Source Community, as well as more recent community
-efforts like the AAS workshop~\cite{aas-software-index}, Project
+efforts like Research Objects~\cite{Bechhofer2013599}, The Software Ontology~\cite{Malone2014}, EDAM Ontology~\cite{Ison15052013}, Project
 CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft},

### 2016-04-19 sc-principles/software-citation-principles.tex
added explanation of bullets to Table 2 caption

@@
 \begin{table}[tbhp]
-\caption{Use cases and basic metadata requirements for software citation, adapted from~\cite{SC-Use-Cases}.}
+\caption{
+Use cases and basic metadata requirements for software citation, adapted from~\cite{SC-Use-Cases}.
+Solid circles (\textbullet) indicate required metadata, while open circles (\textopenbullet) indicate recommended metadata.
+}
 \centering

### 2016-04-19 sc-principles/software-citation-principles.tex
Future work

--- 2016-04-18 sc-principles/software-citation-principles.tex
@@
 related to the principles stated in \S\ref{sec:principles}, and finally, in
-\S\ref{sec:examples} we give examples of how these software citation principles
-could be applied, related back to the use cases in \S\ref{sec:use_cases}.
+\S\ref{sec:examples} we outline our future plans for discussing how these
+software citation principles may be applied.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation examples}
-\label{sec:examples}
+\section{Future work}
+\label{sec:futurework}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Software citation principles without clear worked-through examples are of
+limited value to potential implementers and so in addition to the principles
+document the final deliverable of this working group will be an implementation
+paper outlining working examples for each of the use cases listed in ~\S\ref{sec:use_cases}.
+
+Following on from these efforts, a new Force11 working group is expected to
+begin with the goal of supporting potential implementers of the software citation
+principles as well as developing potential metadata standards.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-04-19 sc-principles/software-citation-principles.tex
added explanation of use cases prereq, addressing issue #130

--- 2016-04-19 sc-principles/software-citation-principles.tex
@@
 
+All use cases assume the existence of a citable software object, typically created by the authors\slash developers of the software.
+Developers can achieve this by, e.g., uploading a software release to Figshare or Zenodo~\cite{github-citable-code-guide} to obtain a DOI.
+Necessary metadata should then be included in a \texttt{CITATION} file~\cite{ssi-citation-files} or machine-readable \texttt{CITATION.jsonld} file~\cite{transitive_credit_json-ld}.
+When software is not freely available (e.g., commercial software), alternative means may be used to create citable objects via identifiers such as RRIDs, although these topics are discussed more in \S\ref{sec:identification} and \S\ref{sec:access}.
+
 \newcommand*\rot[1]{\begin{turn}{90} #1 \end{turn}}%
@@
 \subsection{Unique identification}
+\label{sec:identification}
 
@@
 \subsection{Access to software}
+\label{sec:access}
 

### 2016-04-19 sc-principles/software-citation-principles.tex
improved use case table caption

@@
 Use cases and basic metadata requirements for software citation, adapted from~\cite{SC-Use-Cases}.
-Solid circles (\textbullet) indicate required metadata, while open circles (\textopenbullet) indicate recommended metadata.
+Solid circles (\textbullet) indicate that the use case depends on that metadata, while open circles (\textopenbullet) indicate that the use case would benefit from that metadata if available.
 }

### 2016-04-19 sc-principles/software-citation-principles.tex
fixed section 6 label
resolved comment by @ljhwang

@@
 related to the principles stated in \S\ref{sec:principles}, and finally, in
-\S\ref{sec:examples} we outline our future plans for discussing how these
+\S\ref{sec:futurework} we outline our future plans for discussing how these
 software citation principles may be applied.
@@
 While some journals (e.g., Transactions on Mathematical Software (TOMS),
-Bioinformatics, Computer Physics Communications, F1000Research) have
+Bioinformatics, Computer Physics Communications, F1000Research, 
+Seismological Research Letters, Electronic Seismologist) have
 traditionally accepted software submissions, the American Astronomical Society
@@
 citation. The workshop report includes many recommendations for data
-citation. \LJHnote{ The Seismological Research Letters, Electronic Seismologists
-publishes papers on software.} %text needs to be cleaned up LJH% dy
+citation.
 

### 2016-04-19 sc-principles/software-citation-principles.tex
edits to final section

@@
 Software citation principles without clear worked-through examples are of
-limited value to potential implementers and so in addition to the principles
+limited value to potential implementers, and so in addition to this principles
 document the final deliverable of this working group will be an implementation
 paper outlining working examples for each of the use cases listed in ~\S\ref{sec:use_cases}.
-
-Following on from these efforts, a new Force11 working group is expected to
+Following these efforts, a new FORCE11 working group is expected to
 begin with the goal of supporting potential implementers of the software citation

### 2016-04-19 sc-principles/software-citation-principles.tex
change in prereq use case text to generalize

@@
 Necessary metadata should then be included in a \texttt{CITATION} file~\cite{ssi-citation-files} or machine-readable \texttt{CITATION.jsonld} file~\cite{transitive_credit_json-ld}.
-When software is not freely available (e.g., commercial software), alternative means may be used to create citable objects via identifiers such as RRIDs, although these topics are discussed more in \S\ref{sec:identification} and \S\ref{sec:access}.
+When software is not freely available (e.g., commercial software) or when there is no clear identifier
+to use, alternative means may be used to create citable objects as discussed in \S\ref{sec:access}.
 

### 2016-04-19 sc-principles/software-citation-principles.tex
changes in future work section and updates section to improve flow

@@
 related to the principles stated in \S\ref{sec:principles}, and finally, in
-\S\ref{sec:futurework} we outline our future plans for discussing how these
-software citation principles may be applied.
+\S\ref{sec:futurework} we discuss the work needed to lead to these
+software citation principles being applied.
 
@@
 any updates will require a different FORCE11 working group to make them.
-We expect a follow-on working group to be established to promote the implementation of
+As mentioned in \S\ref{sec:futurework}, we expect a follow-on working group to be established to promote the implementation of
 these principles, and it is possible that this group might find items that need correction or addition
@@
 limited value to potential implementers, and so in addition to this principles
-document the final deliverable of this working group will be an implementation
+document, the final deliverable of this working group will be an implementation
 paper outlining working examples for each of the use cases listed in ~\S\ref{sec:use_cases}.
-Following these efforts, a new FORCE11 working group is expected to
-begin with the goal of supporting potential implementers of the software citation
-principles as well as developing potential metadata standards.
+
+Following these efforts, we expect that FORCE11 will start a new working group
+with the goal of supporting potential implementers of the software citation
+principles as well as developing potential metadata standards, following the model
+of the FORCE11 Data Citation Working Group.
 

### 2016-04-19 sc-principles/software-citation-principles.tex
added funder to evaluate contributions use case
switched slashes to commas in use cases table

@@
 11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Indexer \\
-12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & \textopenbullet & \textopenbullet & Domain group\slash library\slash archive \\
+12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & \textopenbullet & \textopenbullet & Domain group, library, archive \\
 13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Repository \\
 14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder \\
-15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textopenbullet &             & \textbullet &             & \textbullet &             &             &             & Evaluator \\
+15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textopenbullet &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
 16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Citation manager \\

### 2016-04-19 sc-principles/software-citation-principles.tex
added policy maker to funder use case, addressing issue #112
moving table 1 down slightly
added statement on credit system to address issue #131

@@
 of software citation and see if it meets the principles.
+Please note that while we strive to offer practical guidelines that acknowledge the current incentive system of academic citation, a more modern system of assigning credit is sorely needed.
+It is not that academic software needs a separate system from academic papers, but that the need for credit for application software underscores the need to overhaul the system of credit for all research products.
 
@@
 13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Repository \\
-14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder \\
+14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder, policy maker \\
 15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textopenbullet &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
@@
 of the FORCE11 Data Citation Working Group.
+Beyond the efforts of this new working group, additional efforts should be focused on updating the academic credit system that focuses solely on citation.
 

### 2016-04-20 sc-principles/software-citation-principles.tex
slight wording change in final sentence
expanded citation use case slightly, closing #124

@@
 \rowcolors{2}{white}{gray!25}
-\begin{table}[tb]
+\begin{table}[htb]
 \caption{Varieties of software mentions in publications, from Howison and Bullard~\cite{howison2015jasist}.}
@@
 3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textopenbullet & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
-4.\ Find citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher \\
+4.\ Determine use\slash citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher \\
 5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textopenbullet &             & \textbullet & \textbullet &             &             &             &             & Researcher \\
@@
 of the FORCE11 Data Citation Working Group.
-Beyond the efforts of this new working group, additional efforts should be focused on updating the academic credit system that focuses solely on citation.
+Beyond the efforts of this new working group, additional effort should be focused on updating the overall academic credit\slash citation system.
 

### 2016-04-25 sc-principles/software-citation-principles.tex
added JISC guidelines ref

@@
 efforts like Research Objects~\cite{Bechhofer2013599}, The Software Ontology~\cite{Malone2014}, EDAM Ontology~\cite{Ison15052013}, Project
-CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft},
-or the terms and classes defined at \href{https://schema.org}{www.schema.org} related to the
+CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, RRR\slash JISC guidelines~\cite{JISC2015}, 
+or the terms and classes defined at \href{https://schema.org}{Schema.org} related to the
 \href{https://schema.org/SoftwareApplication}{\texttt{SoftwareApplication}} class.   In addition,

### 2016-04-25 sc-principles/software-citation-principles.tex
minor changes to make page 1 fit on 1 page - not sure it's actually important

@@
 
+\vspace{-0.2cm}
 Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation.
@@
 Working examples and possible technical solutions for how these principles can be implemented will be discussed in a separate paper.
+\vspace{-1.0cm}
 
@@
 The principles in this section are written fairly concisely, and discussed
-further later in this document (\S\ref{sec:discussion}). For example, in this
-section we do not define what software should be cited, but how it should be
-cited.  We talk about how we think such decisions might be made in the
+further later in this document (\S\ref{sec:discussion}). Here, for example,
+we do not define what software should be cited, but how it should be
+cited, and we talk about how such decisions might be made in the
 discussion section~(\S\ref{sec:discussion}).
@@
 modified based on discussions of the FORCE11 Software Citation Working
-Group~(see \S\ref{app:wg_members} for members), the information from the use
-cases in~\S\ref{sec:use_cases}, and the related work in~\S\ref{sec:related_work}.
-The adaptations and modifications have been made
+Group (see Appendix~\ref{app:wg_members} for members), information from the use
+cases in \S\ref{sec:use_cases}, and the related work in \S\ref{sec:related_work}.
+The adaptations have been made
 because software, while similar to data in terms of not traditionally having

### 2016-04-28 sc-principles/software-citation-principles.tex
Added use case for mixed data and software packages.
slight edit to recently added use case
Fixing citation.

@@
 16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Citation manager \\
+17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet & \textopenbullet & Repository \\
 \bottomrule
@@
 how to cite software and journals such as F1000Research~\cite{F1000},
-SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow
+SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{ORC} allow
 for submissions entirely focussed on research software.

### 2016-05-09 sc-principles/software-citation-principles.tex
add library to stakeholders of mixed packages
It's a use case that is very relevant to all data archives and one that
I'm trying to figure out how to best handle

@@
 16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Citation manager \\
-17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet & \textopenbullet & Repository \\
+17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet & \textopenbullet & Repository, library, archive \\
 \bottomrule

### 2016-05-09 sc-principles/software-citation-principles.tex
fix typo

@@
 software in general, or the latest version, and to link multiple releases together, perhaps for
-the purpose of understand citations to the software.  The principles in \S\ref{sec:principles} are
+the purpose of understanding citations to the software.  The principles in \S\ref{sec:principles} are
 intended to be applicable at all levels, and to all types of identifiers, such as DOIs, RRIDs, etc.,

### 2016-05-09 sc-principles/software-citation-principles.tex
Reword 5.4 software peer review
The wording of this section made it sound like citation was completely
irrelevant to the peer review process. I don't think it is entirely so,
therefore I replaced the first sentence and reworded the rest
accordingly

@@
 
-This document does not discuss peer review of software, which is an important issue but is
-also mostly out of scope in the context of software citation principles.
-Since the goal of software citation is to identify the software
-that has been used in a scholarly product, whether or not that software has been peer-reviewed
-is irrelevant.  One possible exception would be if the peer-review status of the software should
+Adherence to the software citation principles enables better peer reviews through improved reproducibility.
+However, since the primary goal of software citation is to identify the software that has been used in a scholarly product, the peer review of software is mostly out of scope in the context of software citation principles.
+For instance, when identifying a particular software artifact that has been used in a scholarly product, whether or not that software has been peer-reviewed is irrelevant.
+One possible exception would be if the peer-review status of the software should
 be part of the metadata, but the working group does not believe this to be part of the

### 2016-05-09 sc-principles/software-citation-principles.tex
adding a word in 5.4 and adding line breaks

@@
 
-Adherence to the software citation principles enables better peer reviews through improved reproducibility.
-However, since the primary goal of software citation is to identify the software that has been used in a scholarly product, the peer review of software is mostly out of scope in the context of software citation principles.
-For instance, when identifying a particular software artifact that has been used in a scholarly product, whether or not that software has been peer-reviewed is irrelevant.
-One possible exception would be if the peer-review status of the software should
+Adherence to the software citation principles enables better peer reviews through improved
+reproducibility. However, since the primary goal of software citation is to identify the
+software that has been used in a scholarly product, the peer review of software itself is
+mostly out of scope in the context of software citation principles. For instance, when
+identifying a particular software artifact that has been used in a scholarly product,
+whether or not that software has been peer-reviewed is irrelevant. One possible exception
+would be if the peer-review status of the software should
 be part of the metadata, but the working group does not believe this to be part of the

### 2016-05-09 sc-principles/software-citation-principles.tex
fixing grammar in the importance principle as suggested by Maryann Martone through Hypothesis

@@
 and should not be omitted or separated.
-Software should be cited on the same basis as any other research products such as
-papers or books, that is, authors should cite the appropriate set of software
+Software should be cited on the same basis as any other research product such as
+a paper or a book, that is, authors should cite the appropriate set of software
 products just as they cite the appropriate set of papers.
@@
 In the Importance principle~(\ref{principle:importance}), we state that
-``software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.''
+``software should be cited on the same basis as any other research product such as a paper or a book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.''
 In the case of one code that is derived from another code, citing the derived software may appear to not credit those responsible for the original software, nor recognize its role in the work that used the derived software.

### 2016-05-09 sc-principles/software-citation-principles.tex
adding RII cite and bibtex entry

@@
 though we again recommend when possible the use of DOIs that identify specific versions of
-source code.  We note that RRIDs were developed by the FORCE11 Resource Identification Initiative
+source code.  We note that RRIDs were developed by the FORCE11 Resource Identification Initiative~\cite{f11rii}
 and have been discussed for use to identify software packages (not specific versions),

### 2016-05-20 sc-principles/software-citation-principles.tex
remove double space

--- 2016-05-09 sc-principles/software-citation-principles.tex
@@
 \usepackage{booktabs,multicol}
-\usepackage{rotating}
 \usepackage[inline]{enumitem}
@@
 
-\item ``Citation manager`` refers to people and organizations that create scholarly reference
+\item ``Citation manager'' refers to people and organizations that create scholarly reference
 management software and websites including Zotero, Mendeley, EndNote, RefWorks, BibDesk, etc.,
@@
 though we again recommend when possible the use of DOIs that identify specific versions of
-source code.  We note that RRIDs were developed by the FORCE11 Resource Identification Initiative
+source code.  We note that RRIDs were developed by the FORCE11 Resource Identification Initiative~\cite{f11rii}
 and have been discussed for use to identify software packages (not specific versions),
-thought the FORCE11 Resource Identification Technical Specifications Working
+though the FORCE11 Resource Identification Technical Specifications Working
 Group~\cite{f11rrridtswg} says ``Information resources like software are better suited to
@@
 document, the final deliverable of this working group will be an implementation
-paper outlining working examples for each of the use cases listed in ~\S\ref{sec:use_cases}.
+paper outlining working examples for each of the use cases listed in~\S\ref{sec:use_cases}.
 

### 2016-06-13 sc-principles/software-citation-principles.tex
adding comment about credit for non-academic researchers

--- 2016-05-20 sc-principles/software-citation-principles.tex
@@
 
-\item Academic Credit: Academic researchers at all levels, including students,
+\item Credit: Academic researchers at all levels, including students,
 postdocs, faculty, and staff, should be credited for the software products they
@@
 research done by others.\footnote{Providing recognition of software can have tremendous economic impact as demonstrated by the role of Text REtrieval Conference (TREC) in information retrieval~\cite{trec-economic-impact}.}
+Non-academic researchers should also be credited for their software
+work, though the specific forms of credit are different than for academic researchers.
 

### 2016-06-22 sc-principles/software-citation-principles.tex

@@
 efforts like Research Objects~\cite{Bechhofer2013599}, The Software Ontology~\cite{Malone2014}, EDAM Ontology~\cite{Ison15052013}, Project
-CRediT~\cite{casrai-credit}, Ontosoft~\cite{ontosoft}, RRR\slash JISC guidelines~\cite{JISC2015}, 
+CRediT~\cite{casrai-credit}, the OpenRIF Contribution Role Ontology~\cite{Gutzman2016}, Ontosoft~\cite{ontosoft}, RRR\slash JISC guidelines~\cite{JISC2015}, 
 or the terms and classes defined at \href{https://schema.org}{Schema.org} related to the

### 2016-06-23 sc-principles/software-citation-principles.tex
change future action to show that two things will happen concurrently in response to a comment during the RFC period

--- 2016-06-13 sc-principles/software-citation-principles.tex
@@
 Following these efforts, we expect that FORCE11 will start a new working group
-with the goal of supporting potential implementers of the software citation
-principles as well as developing potential metadata standards, following the model
+with the goals of supporting potential implementers of the software citation
+principles and concurrently developing potential metadata standards, loosely following the model
 of the FORCE11 Data Citation Working Group.

### 2016-06-23 sc-principles/software-citation-principles.tex
adding VM words in 5.8 in response to a comment during the RFC period

@@
 as source code.
-However, we recognize that some software is only available as an executable or as a container, while other software may be available as a service.
+However, we recognize that some software is only available as an executable, a container, or a virtual machine image, while other software may be available as a service.
 We believe the principles apply to all of these forms of software, though the implementation of them will certainly differ based on software type.

### 2016-06-23 sc-principles/software-citation-principles.tex
remove "permit and" to close issue #138

@@
 \item \textbf{Accessibility}: \label{principle:accessibility} Software citations
-should permit and facilitate access to the software itself and to its
+should facilitate access to the software itself and to its
 associated metadata, documentation, data, and other materials necessary

### 2016-06-23 sc-principles/software-citation-principles.tex
changing header of 5.5
added software engineer to use cases, closes #152

--- 2016-06-23 sc-principles/software-citation-principles.tex
@@
 1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher \\
-2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher \\
-3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textopenbullet & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
-4.\ Determine use\slash citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher \\
-5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textopenbullet &             & \textbullet & \textbullet &             &             &             &             & Researcher \\
+2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher, software engineer \\
+3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textopenbullet & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher, software engineer \\
+4.\ Determine use\slash citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher, software engineer \\
+5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textopenbullet &             & \textbullet & \textbullet &             &             &             &             & Researcher, software engineer \\
 6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
-7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
-8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Researcher \\
+7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher, software engineer \\
+8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Researcher, software engineer \\
 9.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             &             &             & Publisher \\
@@
 Following these efforts, we expect that FORCE11 will start a new working group
-with the goals of supporting potential implementers of the software citation
-principles and concurrently developing potential metadata standards, loosely following the model
+with the goal of supporting potential implementers of the software citation
+principles as well as developing potential metadata standards, following the model
 of the FORCE11 Data Citation Working Group.

### 2016-06-23 sc-principles/software-citation-principles.tex

@@
 
-\subsection{Citations in text}
+\subsection{Citation format in reference list}
 
-Citations in the scholarly literature are formatted according to the citation style (e.g., AMS, APA, Chicago, MLA) used by that publication.
+Citations in references in the scholarly literature are formatted according to the citation style (e.g., AMS, APA, Chicago, MLA) used by that publication.
 (Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:

### 2016-06-23 sc-principles/software-citation-principles.tex
remove second paragraph of 5.1 - closes #151

@@
 
-Again, the specific decision of what is appropriate to cite must be made by the author(s) of the product.
-However, an illustrative example is the use of Microsoft Excel in research.
-We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
+%Again, the specific decision of what is appropriate to cite must be made by the author(s) of the product.
+%However, an illustrative example is the use of Microsoft Excel in research.
+%We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
 Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), research presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.
-This recommendation matches that of the Purdue Online Writing Lab: ``Do not cite standard office software (e.g.
+%This recommendation matches that of the Purdue Online Writing Lab: ``Do not cite standard office software (e.g.
 Word, Excel) or programming languages.

### 2016-06-23 sc-principles/software-citation-principles.tex
changing [Computer Program] to [Software] etc in 5.5, closes #150

@@
 (Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
-a) a label indicating that this is software, e.g.\ [Computer program], and b) support for version information, e.g.\ Version 1.8.7.}
+a) a label indicating that this is software, e.g.\ [Software], potentially with more information such as [Software: Source Code], [Software: Executable], or [Software: Container], and b) support for version information, e.g.\ Version 1.8.7.}
 

### 2016-06-23 sc-principles/software-citation-principles.tex
italicized unique identification recommendation

--- 2016-06-23 sc-principles/software-citation-principles.tex
@@
 This report also lists examples of identifiers that match these criteria including DOIs, PURLs, Handles, ARKS, and NBNs.
-For software, we recommend the use of DOIs as the unique identifier due to their common usage and acceptance, particularly as they are the standard for other digital products such as publications.
+For software, \textit{we recommend the use of DOIs as the unique identifier due to their common usage and acceptance, particularly as they are the standard for other digital products such as publications.}
 

### 2016-06-23 sc-principles/software-citation-principles.tex
adding info about citation formats, closes #153

@@
 Citations in references in the scholarly literature are formatted according to the citation style (e.g., AMS, APA, Chicago, MLA) used by that publication.
-(Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
+(Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}; the follow-on Software Citation Implementation Group will provide suggested examples.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
 a) a label indicating that this is software, e.g.\ [Software], potentially with more information such as [Software: Source Code], [Software: Executable], or [Software: Container], and b) support for version information, e.g.\ Version 1.8.7.}

### 2016-06-23 sc-principles/software-citation-principles.tex
added commas after e.g.

--- 2016-06-23 sc-principles/software-citation-principles.tex
@@
 (Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
-a) a label indicating that this is software, e.g.\ [Software], potentially with more information such as [Software: Source Code], [Software: Executable], or [Software: Container], and b) support for version information, e.g.\ Version 1.8.7.}
+a) a label indicating that this is software, e.g., [Software], potentially with more information such as [Software: Source Code], [Software: Executable], or [Software: Container], and b) support for version information, e.g., Version 1.8.7.}
 

### 2016-06-23 sc-principles/software-citation-principles.tex

@@
 Citations in references in the scholarly literature are formatted according to the citation style (e.g., AMS, APA, Chicago, MLA) used by that publication.
-(Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
+(Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}; the follow-on Software Citation Implementation Group will provide suggested examples.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
 a) a label indicating that this is software, e.g., [Software], potentially with more information such as [Software: Source Code], [Software: Executable], or [Software: Container], and b) support for version information, e.g., Version 1.8.7.}

### 2016-06-23 sc-principles/software-citation-principles.tex
taking a stab at subsection on what identifiers should resolve to

@@
 All use cases assume the existence of a citable software object, typically created by the authors\slash developers of the software.
-Developers can achieve this by, e.g., uploading a software release to Figshare or Zenodo~\cite{github-citable-code-guide} to obtain a DOI.
+Developers can achieve this by, e.g., uploading a software release to Figshare~\cite{figshare} or Zenodo~\cite{github-citable-code-guide} to obtain a DOI.
 Necessary metadata should then be included in a \texttt{CITATION} file~\cite{ssi-citation-files} or machine-readable \texttt{CITATION.jsonld} file~\cite{transitive_credit_json-ld}.
@@
 
+\subsection{What identifier should resolve to}
+
+While citing an identifier that points to, e.g., a GitHub repository can satisfy the principles of Unique Identification (3), Accessibility (5), and Specificity (6), such a repository cannot guarantee Persistence (4).
+\textit{Therefore, we recommend that the software identifier should resolve to a persistent landing page that shows metadata and a link to the software itself, rather than directly to the source code files, repository, or executable.}
+This ensures longevity of the software metadata---even perhaps beyond the lifespan of the software they describe.
+This is currently offered by services such as Figshare~\cite{figshare} and Zenodo~\cite{github-citable-code-guide}, which both generate persistent DataCite DOIs for submitted software.
+In addition, such landing pages can contain both human-readable metadata (e.g., the types shown by Table~\ref{tab:use_cases}) as well as content-negotiable formats such as RDF or DOAP~\cite{DOAP}.
+
+% issue 144
+
 %\LJHnote{recognize that some software is proprietary, not all is open source}

### 2016-06-24 sc-principles/software-citation-principles.tex
changed Figshare to figshare

@@
 All use cases assume the existence of a citable software object, typically created by the authors\slash developers of the software.
-Developers can achieve this by, e.g., uploading a software release to Figshare~\cite{figshare} or Zenodo~\cite{github-citable-code-guide} to obtain a DOI.
+Developers can achieve this by, e.g., uploading a software release to figshare~\cite{figshare} or Zenodo~\cite{github-citable-code-guide} to obtain a DOI.
 Necessary metadata should then be included in a \texttt{CITATION} file~\cite{ssi-citation-files} or machine-readable \texttt{CITATION.jsonld} file~\cite{transitive_credit_json-ld}.
@@
 This ensures longevity of the software metadata---even perhaps beyond the lifespan of the software they describe.
-This is currently offered by services such as Figshare~\cite{figshare} and Zenodo~\cite{github-citable-code-guide}, which both generate persistent DataCite DOIs for submitted software.
+This is currently offered by services such as figshare~\cite{figshare} and Zenodo~\cite{github-citable-code-guide}, which both generate persistent DataCite DOIs for submitted software.
 In addition, such landing pages can contain both human-readable metadata (e.g., the types shown by Table~\ref{tab:use_cases}) as well as content-negotiable formats such as RDF or DOAP~\cite{DOAP}.

### 2016-06-24 sc-principles/software-citation-principles.tex
updates in new para - 5.10 - mostly tex fixes

@@
 
-\item \textbf{Specificity}: Software citations should facilitate identification
+\item \textbf{Specificity}: \label{principle:specificity} Software citations should facilitate identification
 of, and access to, the specific version of software that was used. Software
@@
 
-\subsection{What identifier should resolve to}
+\subsection{What an identifier should resolve to}
 
-While citing an identifier that points to, e.g., a GitHub repository can satisfy the principles of Unique Identification (3), Accessibility (5), and Specificity (6), such a repository cannot guarantee Persistence (4).
-\textit{Therefore, we recommend that the software identifier should resolve to a persistent landing page that shows metadata and a link to the software itself, rather than directly to the source code files, repository, or executable.}
+While citing an identifier that points to, e.g., a GitHub repository can satisfy the principles of Unique Identification (\ref{principle:uid}), Accessibility (\ref{principle:accessibility}), and Specificity (\ref{principle:specificity}), such a repository cannot guarantee Persistence (\ref{principle:persistence}).
+\textit{Therefore, we recommend that the software identifier should resolve to a persistent landing page that contains metadata and a link to the software itself, rather than directly to the source code files, repository, or executable.}
 This ensures longevity of the software metadata---even perhaps beyond the lifespan of the software they describe.

### 2016-06-24 sc-principles/software-citation-principles.tex
italicizing more recommendation statements

@@
 What software should be cited is the decision of the author(s) of the research work in the context of community norms and practices, and in most research communities, these are currently in flux.
-In general, we believe that software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers, perhaps following the
+In general, \textit{we believe that software should be cited on the same basis as any other research product such as a paper or book}; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers, perhaps following the
 FORCE11 Data Citation Working Group principles, which state, ``In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited.''~\cite{data-citation-principles}
@@
 Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.
-And the data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.
+The data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.
 These statements may also be true for software products.
@@
 We believe the principles apply to all of these forms of software, though the implementation of them will certainly differ based on software type.
-When software exists as both source code and another type, we recommend that the source code be cited.
+\textit{When software exists as both source code and another type, we recommend that the source code be cited.}
 
@@
 in these principles.
-We recommend that this Software Citation Implementation Working Group be charged, in part,
-with updating these principles during its lifetime, and that FORCE11 should listen to community requests for later updates and respond by creating a new working group.
+\textit{We recommend that this Software Citation Implementation Working Group be charged, in part,
+with updating these principles during its lifetime, and that FORCE11 should listen to community requests for later updates and respond by creating a new working group.}
 

### 2016-06-24 sc-principles/software-citation-principles.tex
final italicized statement

@@
 
-The goals of software include the linked ideas of crediting those responsible for software and understanding the dependencies of research products on specific software.
+The goals of software citation include the linked ideas of crediting those responsible for software and understanding the dependencies of research products on specific software.
 In the Importance principle~(\ref{principle:importance}), we state that
@@
 This is not recommended, but is possible.
-We strongly recommend that if there is already a UID for a version of software, no additional UID should be created.
+\textit{We strongly recommend that if there is already a UID for a version of software, no additional UID should be created.}
 Multiple UIDs can lead to split credit, which goes against the Credit and Attribution principle (\ref{principle:credit}).

### 2016-06-24 sc-principles/software-citation-principles.tex
breaking paragraph up onto multiple lines
breaking another paragraph up in source
one more recommendation in itals

@@
 
-Citations in references in the scholarly literature are formatted according to the citation style (e.g., AMS, APA, Chicago, MLA) used by that publication.
-(Examples illustrating these styles have been published by Lipson~\cite{lipson2011cite}; the follow-on Software Citation Implementation Group will provide suggested examples.)  As these citations are typically sent to publishers as text formatted in that citation style, not as structured metadata, and because the citation style dictates how the human reader sees the software citation, \textit{we recommend that all text citation styles support the following:
-a) a label indicating that this is software, e.g., [Software], potentially with more information such as [Software: Source Code], [Software: Executable], or [Software: Container], and b) support for version information, e.g., Version 1.8.7.}
+Citations in references in the scholarly literature are formatted according to the citation 
+style (e.g., AMS, APA, Chicago, MLA) used by that publication. (Examples illustrating these 
+styles have been published by Lipson~\cite{lipson2011cite}; the follow-on Software Citation 
+Implementation Group will provide suggested examples.)  As these citations are typically sent 
+to publishers as text formatted in that citation style, not as structured metadata, and because 
+the citation style dictates how the human reader sees the software citation, \textit{we recommend 
+that all text citation styles support the following:
+a) a label indicating that this is software, e.g., [Software], potentially with more information 
+such as [Software: Source Code], [Software: Executable], or [Software: Container], 
+and b) support for version information, e.g., Version 1.8.7.}
 
@@
 Some scholarly products, such as journal articles, may have strict limits on the number of citations they permit, or page limits that include reference sections.
-Such limits are counter to our recommendation, and \textit{we recommend that publishers using strict limits for the number of citations add specific instructions regarding software citations to their author guidelines to not disincentivize software citation.} Similarly, publishers should not include references in the content counted against page
-limits.
+Such limits are counter to our recommendation, and \textit{we recommend that publishers using strict limits for the number of citations add specific instructions regarding software citations to their author guidelines to not disincentivize software citation.}
+Similarly,  \textit{publishers should not include references in the content counted against page limits.}
 

### 2016-06-24 sc-principles/software-citation-principles.tex
removed comment I introduced earlier
added line numbers
switched to Times font
removed unnecessary fontenc option
reformatted authors and added affiliations

@@
-\documentclass[11pt, oneside]{amsart}
+\documentclass[12pt, oneside]{amsart}
 \pdfoutput=1
@@
 \usepackage{graphicx}
+\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
 \usepackage{lmodern}
@@
 \usepackage[inline]{enumitem}
+
+% add line numbers for submission
+\usepackage{lineno}
+\linenumbers
+
+% change to Times font
+\usepackage{newtxtext,newtxmath}
+
 \usepackage{textcomp} % added for textopenbullet
@@
 % Using 11 pt => 94 characters / line
-\setlength{\paperwidth}{216 mm}
+%\setlength{\paperwidth}{216 mm}
 % 6 lines / 2.5 cm => 55 lines / page
 % Using 11pt => 48 lines / pages
-\setlength{\paperheight}{279 mm}
+%\setlength{\paperheight}{279 mm}
 \usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
@@
 
-\author{FORCE11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}
+\author{Arfon M.~Smith$^1$, Daniel S.~Katz$^2$, Kyle E.~Niemeyer$^3$, and the FORCE11 Software Citation Working Group}
+
+\thanks{Corresponding author: Daniel S.~Katz$^2$, \href{mailto:d.katz@ieee.org}{d.katz@ieee.org}}
+
+\thanks{{}$^{1}$ GitHub, Inc., San Francisco, CA, USA}
+\thanks{{}$^{2}$ National Center for Supercomputing Applications \& Graduate School of Library and Information Science, University of Illinois at Urbana--Champaign, Urbana, IL, USA}
+\thanks{{}$^{3}$ School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
 
@@
 
-% issue 144
-
 %\LJHnote{recognize that some software is proprietary, not all is open source}

### 2016-06-24 sc-principles/software-citation-principles.tex
fixed Dan affiliation

@@
 \thanks{{}$^{1}$ GitHub, Inc., San Francisco, CA, USA}
-\thanks{{}$^{2}$ National Center for Supercomputing Applications \& Graduate School of Library and Information Science, University of Illinois at Urbana--Champaign, Urbana, IL, USA}
+\thanks{{}$^{2}$ National Center for Supercomputing Applications \& Electrical and Computer Engineering Department \& School of Information Sciences, University of Illinois at Urbana--Champaign, Urbana, IL, USA}
 \thanks{{}$^{3}$ School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}

### 2016-06-24 sc-principles/software-citation-principles.tex
added missing space

@@
 package format~\cite{Debian_policy}, Python package descriptions~\cite{pypi}, and R package descriptions~\cite{wickham_r_2015}, but these are typically
-conceived for software build, packaging,and distribution rather than citation.  CodeMeta~\cite{codemeta}
+conceived for software build, packaging, and distribution rather than citation.  CodeMeta~\cite{codemeta}
 has created a crosswalk among these software metadata schemes and an exchange format that

### 2016-06-30 sc-principles/software-citation-principles.tex
grammatical fix
added workshop references and discussion

--- 2016-06-24 sc-principles/software-citation-principles.tex
@@
-\documentclass[12pt, oneside]{amsart}
+\documentclass[11pt, oneside]{amsart}
 \pdfoutput=1
@@
 \usepackage{graphicx}
+\usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
 \usepackage{lmodern}
@@
 \usepackage[inline]{enumitem}
-
-% add line numbers for submission
-\usepackage{lineno}
-\linenumbers
-
-% change to Times font
-\usepackage{newtxtext,newtxmath}
-
 \usepackage{textcomp} % added for textopenbullet
@@
 % Using 11 pt => 94 characters / line
-%\setlength{\paperwidth}{216 mm}
+\setlength{\paperwidth}{216 mm}
 % 6 lines / 2.5 cm => 55 lines / page
 % Using 11pt => 48 lines / pages
-%\setlength{\paperheight}{279 mm}
+\setlength{\paperheight}{279 mm}
 \usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
@@
 
-\author{Arfon M.~Smith$^1$, Daniel S.~Katz$^2$, Kyle E.~Niemeyer$^3$, and the FORCE11 Software Citation Working Group}
-
-\thanks{Corresponding author: Daniel S.~Katz$^2$, \href{mailto:d.katz@ieee.org}{d.katz@ieee.org}}
-
-\thanks{{}$^{1}$ GitHub, Inc., San Francisco, CA, USA}
-\thanks{{}$^{2}$ National Center for Supercomputing Applications \& Graduate School of Library and Information Science, University of Illinois at Urbana--Champaign, Urbana, IL, USA}
-\thanks{{}$^{3}$ School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
+\author{FORCE11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}
 
@@
 Perhaps in recognition of the broad range of research domains struggling with
-the challenge of better recognizing the role of software, a number of community
-efforts hosted (and sponsored) by funders and agencies in both the US (e.g.,
-NSF, NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC, Wellcome Trust)
-have run a number of workshops with participants from across a range of
-disciplines.
+the challenge of better recognizing the role of software, funders and agencies 
+in both the US (e.g., NSF, NIH, Alfred P.\ Sloan Foundation) and UK (e.g., 
+SFTC, JISC, Wellcome Trust) have sponsored or hosted a number of workshops with 
+participants from across a range of
+disciplines, specifically aimed at discussing issues around software 
+citation~\cite{Sufi:2014en,Ahalt:2015ve,ssisoftwarecredit,aas-software-index,saga,Allen:2015ub}.
+In many cases these workshops produced strong recommendations for their 
+respective communities on how best to proceed. In addition, a number of common 
+themes arose in these workshops, including (1) the critical need for making 
+software more ``citable'' (and therefore actions authors and publishers 
+should take to improve the status quo), (2) how to better measure the impact 
+of software (and therefore attract appropriate funding), and (3) how to 
+properly archive software (where, how, and how often) and how this affects 
+what to cite and when.
 
@@
 held at the Astronomical Data Analysis Software and Systems (ADASS) XXV
-conference~\cite{2015arXiv151207919A} that included discussion of software
+conference~\cite{Allen:2015ub} that included discussion of software
 citation.

### 2016-07-02 sc-principles/software-citation-principles.tex
added more journals to those that accept software-only

@@
 such as APA~\cite{APA-guidelines} have recommendations for submitting authors on
-how to cite software and journals such as F1000Research~\cite{F1000},
-SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{ORC} allow
-for submissions entirely focussed on research software.
+how to cite software, and journals such as \emph{F1000Research}~\cite{F1000},
+\emph{SoftwareX}~\cite{softwareX}, \emph{Open Research Computation}~\cite{ORC}, and
+the \emph{Journal of Open Research Software} allow for submissions entirely 
+focussed on research software.
 

### 2016-07-03 sc-principles/software-citation-principles.tex
americanizing a spelling

@@
 the \emph{Journal of Open Research Software} allow for submissions entirely 
-focussed on research software.
+focused on research software.
 

### 2016-07-03 sc-principles/software-citation-principles.tex

@@
 
-\author{FORCE11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}
+\author{Arfon M.~Smith$^1$, Daniel S.~Katz$^2$, Kyle E.~Niemeyer$^3$, and the FORCE11 Software Citation Working Group}
+
+\thanks{Corresponding author: Daniel S.~Katz$^2$, \href{mailto:d.katz@ieee.org}{d.katz@ieee.org}}
+
+\thanks{{}$^{1}$ GitHub, Inc., San Francisco, CA, USA}
+\thanks{{}$^{2}$ National Center for Supercomputing Applications \& Electrical and Computer Engineering Department \& School of Information Sciences, University of Illinois at Urbana--Champaign, Urbana, IL, USA}
+\thanks{{}$^{3}$ School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
 
@@
 Community Software Commons for the Geosciences.'' Much attention was given to the
-metadata required to describe, discover, and execute research software. NSF sponsored
+metadata required to describe, discover, and execute research software. The NSF-sponsored
 Geo-Data Workshop 2011~\cite{nsf-geo-data} revolved around data lifecycle, management, and

### 2016-07-13 sc-principles/rebuttal/rebuttal.tex
added initial rebuttal to reviewer comments

\pdfminorversion=4
\documentclass{article}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx, caption}
\usepackage{latexsym,url,amscd, amssymb, amsmath}
\usepackage{epstopdf}

\usepackage{multicol}
\usepackage{booktabs}
\usepackage{fancyvrb}

\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}

\usepackage[english]{babel}
\usepackage{textcomp}

\usepackage{siunitx}
\sisetup{group-separator={,},
		 detect-all,
		 binary-units,
		 list-units = single,
		 range-units = single,
		 tophrase = --, 
		 per-mode = symbol-or-fraction,
		 separate-uncertainty = true
%		 scientific-notation = fixed
}
\DeclareSIUnit\atm{atm}
\DeclareSIUnit\CA{CA}
\DeclareSIUnit\rpm{rpm}

%to highlight text
\usepackage{soul}
\usepackage[usenames, dvipsnames]{xcolor}
\newcommand{\hly}[1]{{\sethlcolor{yellow}\hl{#1}}}
\newcommand{\hlb}[1]{{\sethlcolor{SkyBlue}\hl{#1}}}
\newcommand{\hlg}[1]{{\sethlcolor{green}\hl{#1}}}

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{}{;}
\usepackage[linkcolor=blue, citecolor=blue, colorlinks=true, breaklinks=true]{hyperref}

\usepackage{url,breakurl}

\usepackage{todonotes}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

%\date{}                                           % Activate to display a given date or no date

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.4cm}

%%% BEGIN DOCUMENT
\begin{document}

We thank the editor and reviewers for their helpful comments, and have addressed their requests and questions here.
Specific changes or additions in the marked revision in response to the Editor, Reviewer \#1, and Reviewer \#2 are respectively highlighted in the marked manuscript in \hly{yellow}, \hlb{blue}, and \hlg{green}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Editor:}

\emph{First of all, I would like to thank the reviewers, who have done a fantastic job and provided very insightful comments for improving the paper. I hope that the authors will take them into serious consideration in the revision.}

\emph{From my personal perspective, this is a very good contribution, which addresses one of the most important topics in the ``citation'' domain, i.e. software citation. I totally agree with the reviewers about the readability of the paper, and I think this paper will deserve publication in PeerJ CS.}

\emph{However, there are some aspects that should be clarified in the revision. In particular:}

\textbf{Comment:}
\emph{I would like to see a new section that describes the presentation of these principles at FORCE 2016, and includes also a summary of the most important outcomes derived from the discussion the authors have done during the conference. Did such outcomes change some of the principles and/or other claims in this paper?}

\textbf{Our response:}

\textbf{Comment:}
\emph{The link between the principles and the requirements should be introduced explicitly, since it is not entirely clear how ones are connected to the others.}

\textbf{Our response:}

\textbf{Comment:}
\emph{A clear research question should be announced explicitly from the very beginning, and accompanied by a text that explains how the authors would like to address it.}

\textbf{Our response:}

\textbf{Comment:}
\emph{I suggest to include one or more appendixes with all the details about the use cases presented. This would make the paper more self-contained.}

\textbf{Our response:}

\textbf{Comment:}
\emph{The reviewers did not agree on some aspects described in the paper, e.g. "if the software authors ask that a paper should be cited, that should typically be respected". I would like to see an appropriate discussion on these points, so as to clearly address these disagreements.}

\textbf{Our response:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Reviewer \#1 (Tobias Kuhn):}
\emph{This paper argues for the adoption of software citations, basically for the same reasons that data citations have been advocated: Because software packages have become so important in science, they should be appropriately credited and referenced, which can be achieved if we extend the concept of citations in scientific articles.}

\emph{The paper tackles an important problem and provides a very valuable foundation for further discussions and future implementations. In particular, the six software citation principles presented in the beginning are very valuable.}

\emph{While I think this is a very useful paper, I see the following problems with the current version:}

\textbf{Reviewer comment:}
\emph{The logical connections between the different parts of the paper are not clear. For example, how do the ``basic requirements'' of Table 2 map to the ``principles'' in Section 1? And are the principles presented in the beginning the *outcome* of this research (this is what lines 29+ seem to suggest) or are they *assumptions* of the approach to be discussed and evaluated (this is how they seem to be treated in Section 5).}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{It is not clear whether the authors consider software citations to be just the first step towards a long-term vision, or whether they consider software citations the overall final solution. It seems that the first is the case, but the long-term vision is not discussed.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{This paper doesn't have an experimental design in the strict sense, as it is more of a position paper than a research article. The PeerJ guidelines state that "the submission should clearly define the research question", which is currently not the case for this paper. However, I see a research question between the lines, which could be phrased as "how can citations be applied to software products to account for their increasing importance in scientific research?". By making such a research question explicit, the PeerJ requirements can be met.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{The use cases described in Section 3 form the main method to answer the implicit research question. Given the importance of these use cases for this paper, I think reference [18] with the details about these use cases should be published in a more official way, either on a site like ArXiv or FigShare, or maybe even as supplemental material or appendix for this paper. I think having these use cases in the appendix of this paper would be the best solution, which would also increase the impact of these use cases and would strengthen the paper in general.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{As noted above, the logical structure of the paper is unclear, and therefore it is not clear what exactly the main findings are. I think the main findings should be the principles stated in the beginning, and the use cases would then form the main method to arrive at these findings. Alternatively, one could label the principles as "assumptions" and apply the use case study to validate these assumptions. In any case, I think the findings are valid but they should be labeled as preliminary, as it is not possible to arrive at a final conclusion for such kinds of questions. Only the future can tell whether software citations really work out.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{line 88: ``which will be presented ...'' > ``which was presented ...''}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{line 89: ``We expect that this discussion may lead to a second, final version'': So this paper is not the final version? Please clarify.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{line 97: ``the need for credit for application software underscores the need to overhaul the system of credit for all research products.'': Very true and very important. But then maybe the concept of a ``citation'' is obsolete altogether, and we shouldn't be advocating ``software citations'', but something new altogether? Some more discussion on this would be valuable, I think. This relates to my comment above about a long-term vision.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{line 238: ``Citation is a record of software that is important to a research outcome'': This is another important issue, but not much more is said about this. What makes a software ``important'' to a research outcome? Or, more generally, what is the semantics of a software citation?}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{line 254: ``In addition, if the software authors ask that a paper should be cited, that should typically be respected'': I think I disagree with that. Whether a paper should be cited or not should *not* depend on whether the given software package tells you to do so. Such pointers can be useful as suggestions, but not more than that in my opinion.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Lines 308+: You don't imply that this list is complete, but wouldn't an identifier of a software *branch* be another important type of identifier? For example, the Windows ``version'' of a software package. Also, I see in what situation one would use (1) or (2) with respect to software citations, but what is the purpose of (3) in the context of citations?}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 325: ``other software may be available as a service'': This makes a huge difference, doesn't it? I think some more discussion on this would be helpful. For example, with software services it can be hard to find out what the official name of the software is, let alone the developers or version numbers. Also, does a software that is only available as a service (e.g. REST API) violate the Accessibility principle (5)?}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{I find it a bit weird to quote passages from the same paper like ``The Accessibility principle (5) states that ?....?'' (line 329 and others).}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{In the discussion about Unique identification (lines 292+), I miss the discussion of hash values as version identifiers, as applied by Git for instance. As the first author is affiliated with GitHub, I assume this aspect was left out intentionally, but why? As basically all software is nowadays maintained with version control systems such as Git, these identifiers come for free and are by design unique. So, I would think they are an important aspect to discuss for software citations.}

\textbf{Our response:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2 (Graham Klyne):}

\textbf{Reviewer comment:}
\emph{The two tables are relevant and helpful, though I did have a slight problem with the legibility of my printed copy of Table 2 (page 5). The distinction between solid and open circles was not immediately clear to me, and I might suggest using a slightly larger version of these symbols.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{The paper does not present a specific research finding, and as such does not feature a research question and experimental procedure that are expected of such papers. However, I think it makes in important contribution to the acceptance of software as part of the process and output of research activity, and reflects a consensus formed among a significant community of researchers. As such, I believe it justifies a place in the formally published literature.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Lines 32--35: as a justification for treating software differently from data, I found this passage raised more questions than answers, and that the comments here were uncompelling and unsupported. For example, some published data is updated on a frequency comparable with some software (e.g. the bi-monthly release schedule of FlyBase is comparable with many software projects - \url{http://flybase.org/static_pages/docs/release_schedule.html}). Also, I think the distinction between data and software is less clear-cut than the authors suggest (e.g. is a schema or stylesheet data or software?). I would see the difference between software and data as more akin to the difference between research methods and research findings, but that is just a personal opinion. My suggestion would be that this justification is not really necessary to the overall purpose of the paper, and might be dropped without significant loss of value.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 57, et seq: [Nit] I would find this easier to read if the sentence introducing the list of reasons were moved to the same page as the list itself.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 88: ``will be presented'' refers to an event now in the past: change to ``was presented''?}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 113: [Nit] ``work using software'' appears to be very broadly scoped, and could apply to just about any activity using a computer. Suggest ``research using software''.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Table 2, page 5: I had a problem with the legibility of my printed copy of this table. The distinction between solid and open circles was not immediately clear to me. I might suggest using a slightly larger version of these symbols, or maybe more distinct symbol shapes.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Table 2, page 5: [Comment] This table contains a lot of information, and much food for debate. I was surprised by some omissions, which may mean that I wasn't fully understanding the concepts being presented:\\
\\
(a) I find it hard to imagine any of the use cases not benefiting from availability of a ``Description'', especially 1, 2, 9, and 16.\\
(b) I would have expected ``Indexed citations'' to be useful for 5: getting credit.\\
(c) For publishing a software paper, I would expect all of the basic requirements other than ``Indexed citations'' to be at least beneficial.\\
(d) For 12: build a software catalog/registry, would there not be a role for indexed citations?\\
(e) For 16: Store software try, I'd expect pretty much all of the Basic requirements other than ``Indexed citations'' to be relevant, and particularly ``Description'' and ``Keywords''.\\
(f) I think there is one use-case missing, which I think is commonly faced by researchers: viz. ``Evaluate suitability of software for a given task'' - I think this is distinct from discovery, and typically involves making a choice between different available software packages, or to use an existing package, modify an existing package or develop something new.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 175: [Nit] ``in(1)'' Missing space?}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Lines 241--242: ``Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.'' I found this non sequitur (or am not seeing the logical connection). I can imagine reasons to cite software that does not form part of the provenance chain of results presented (e.g. alternatives considered).}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Lines 245--247: ``In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification. Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).'' This statement feels to me to be at odds with table 2, which presents ``Use cases and basic metadata requirements for software citation'', which includes ``reproducibility'' as a use-case. This leaves me wondering if the use cases in table 2 are going beyond what is strictly needed for the purposes of software citation.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 327--328: ``When software exists as both source code and another type, we recommend that the source code be cited''. The term ``exists'' here doesn't seem quite right to me. E.g., proprietary software may ``exist'' as source code, but be unavailable. Suggest ``exists and is accessible'' or just ``is accessible''?}

\textbf{Our response:}

\textbf{Reviewer comment:}
\emph{Line 348: ``Updates to this document''. This paper is being presented for consideration as a part of the formal academic record - as such, I don't think ``this document'' is something which may be updated. Suggest ``Updates to these principles''.}

\textbf{Our response:}

%\bibliography{}
%\bibliographystyle{}

\end{document}

### 2016-07-14 sc-principles/software-citation-principles.tex
edit to Section 5.11 title re: reviewer 2 comment 14

@@
 
-\subsection{Updates to this document}
+\subsection{Updates to these principles}
 
-As this set of software citation principles has been created by the FORCE11 Software Citation Working Group, which will cease work and dissolve after these principles have been published,
+As this set of software citation principles has been created by the FORCE11 Software Citation Working Group, which will cease work and dissolve after publication of these principles,
 any updates will require a different FORCE11 working group to make them.
@@
 with updating these principles during its lifetime, and that FORCE11 should listen to community requests for later updates and respond by creating a new working group.}
-
-%\dmnote{Any provisions for updating the principles?}
-%\LJHnote{may require feedback from "Implementation Group" and some time to put into practice to find gaps.
-Most of the time revisions are frequent in the beginning of the lifecycle to cover gaps and become less frequent.}
 

### 2016-07-14 sc-principles/software-citation-principles.tex
edit to Section 5.8 re: reviewer 2 comment 13
fixed spacing in 4.1 re: reviewer 2 comment 10

@@
 SSI~\cite{ssi-workshops}, who between them have run a series of workshops aimed
-at gathering together community members with an interest in
-\begin{enumerate*}[series=InlineList, before=\hspace{-0.6ex}]
+at gathering together community members with an interest in~
+\begin{enumerate*}[series=InlineList]
     \item defining the set
@@
 We believe the principles apply to all of these forms of software, though the implementation of them will certainly differ based on software type.
-\textit{When software exists as both source code and another type, we recommend that the source code be cited.}
+\textit{When software is accessible as both source code and another type, we recommend that the source code be cited.}
 

### 2016-07-14 sc-principles/software-citation-principles.tex
replaced open bullet in table with + sign, re: reviewer 2 comment 8

@@
 Use cases and basic metadata requirements for software citation, adapted from~\cite{SC-Use-Cases}.
-Solid circles (\textbullet) indicate that the use case depends on that metadata, while open circles (\textopenbullet) indicate that the use case would benefit from that metadata if available.
+Solid circles (\textbullet) indicate that the use case depends on that metadata, while plus signs (+) indicate that the use case would benefit from that metadata if available.
 }
@@
 \midrule
-1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher \\
-2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet &             &             & Researcher, software engineer \\
-3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & \textopenbullet & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher, software engineer \\
+1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + &             &             & Researcher \\
+2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + &             &             & Researcher, software engineer \\
+3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & + & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher, software engineer \\
 4.\ Determine use\slash citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher, software engineer \\
-5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & \textopenbullet &             & \textbullet & \textbullet &             &             &             &             & Researcher, software engineer \\
-6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher \\
-7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet &             & Researcher, software engineer \\
-8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Researcher, software engineer \\
+5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & + &             & \textbullet & \textbullet &             &             &             &             & Researcher, software engineer \\
+6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher \\
+7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher, software engineer \\
+8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & + & + & + & Researcher, software engineer \\
 9.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             &             &             & Publisher \\
 10.\ Publish papers that cite software           & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Publisher \\
-11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & \textopenbullet & \textopenbullet & \textopenbullet & Indexer \\
-12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & \textopenbullet & \textopenbullet & Domain group, library, archive \\
+11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & + & + & + & Indexer \\
+12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & + & + & Domain group, library, archive \\
 13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Repository \\
 14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder, policy maker \\
-15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & \textopenbullet &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
+15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & + &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
 16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Citation manager \\
-17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textopenbullet & \textopenbullet & \textopenbullet & Repository, library, archive \\
+17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + & + & + & Repository, library, archive \\
 \bottomrule

### 2016-07-14 sc-principles/software-citation-principles.tex
work using software -> research using software, re: reviewer 2 comment 7
past tense for FORCE2016 activities, re: R1 #6-7 and R2 #6

@@
 review those materials, identifying overlaps and differences; and subsequently draft this resulting document,
-which will be presented and discussed at the Force2016 Conference~\cite{force2016}
-in April 2016.  We expect that this discussion may lead to a second, final version, and we also
+which was presented and discussed at the Force2016 Conference~\cite{force2016}
+in April 2016.  This discussion led to a second, final version, and we also
 plan to have a follow-on working group that will work with stakeholders to ensure that
@@
 
-\item ``Funder'' is a group that funds software or work using software.
+\item ``Funder'' is a group that funds software or research using software.
 

### 2016-07-14 sc-principles/rebuttal/rebuttal.tex
responses to addressed reviewer comments

@@
 
+\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}} % Dan
+\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}} % Kyle
+\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}} %Arfon
+
 \usepackage{todonotes}
@@
 We thank the editor and reviewers for their helpful comments, and have addressed their requests and questions here.
-Specific changes or additions in the marked revision in response to the Editor, Reviewer \#1, and Reviewer \#2 are respectively highlighted in the marked manuscript in \hly{yellow}, \hlb{blue}, and \hlg{green}.
+Specific changes or additions can be found in the marked revision.
+% in response to the Editor, Reviewer \#1, and Reviewer \#2 are respectively highlighted in the marked manuscript in \hly{yellow}, \hlb{blue}, and \hlg{green}.
 
@@
 
-\textbf{Comment:}
+\begin{enumerate}
+\item  \textbf{Comment:}
+%\textbf{1. Comment:}
 \emph{I would like to see a new section that describes the presentation of these principles at FORCE 2016, and includes also a summary of the most important outcomes derived from the discussion the authors have done during the conference.
@@
 
-\textbf{Comment:}
+\item \textbf{Comment:}
 \emph{The link between the principles and the requirements should be introduced explicitly, since it is not entirely clear how ones are connected to the others.}
@@
 
-\textbf{Comment:}
+\item \textbf{Comment:}
 \emph{A clear research question should be announced explicitly from the very beginning, and accompanied by a text that explains how the authors would like to address it.}
@@
 
-\textbf{Comment:}
+\item \textbf{Comment:}
 \emph{I suggest to include one or more appendixes with all the details about the use cases presented. This would make the paper more self-contained.}
@@
 \textbf{Our response:}
+\niemnote{I will add the full use cases text as an appendix}
 
-\textbf{Comment:}
+\item \textbf{Comment:}
 \emph{The reviewers did not agree on some aspects described in the paper, e.g.
@@
 \textbf{Our response:}
+
+\end{enumerate}
 
@@
 
-\textbf{Reviewer comment:}
+\begin{enumerate}
+\item \textbf{Reviewer comment:}
 \emph{The logical connections between the different parts of the paper are not clear.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{It is not clear whether the authors consider software citations to be just the first step towards a long-term vision, or whether they consider software citations the overall final solution.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{This paper doesn't have an experimental design in the strict sense, as it is more of a position paper than a research article.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{The use cases described in Section 3 form the main method to answer the implicit research question.
@@
 \textbf{Our response:}
+\niemnote{I will add the full use cases text as an appendix}
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{As noted above, the logical structure of the paper is unclear, and therefore it is not clear what exactly the main findings are.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{line 88: ``which will be presented ...'' > ``which was presented ...''}
@@
 \textbf{Our response:}
+We agree, and have made this change.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{line 89: ``We expect that this discussion may lead to a second, final version'': So this paper is not the final version? Please clarify.}
@@
 \textbf{Our response:}
+Thank you for pointing this error out---we have corrected this statement to indicate that the past discussion led to a second version (what was submitted).
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{line 97: ``the need for credit for application software underscores the need to overhaul the system of credit for all research products.'': Very true and very important.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{line 238: ``Citation is a record of software that is important to a research outcome'': This is another important issue, but not much more is said about this.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{line 254: ``In addition, if the software authors ask that a paper should be cited, that should typically be respected'': I think I disagree with that.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Lines 308+: You don't imply that this list is complete, but wouldn't an identifier of a software *branch* be another important type of identifier?
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 325: ``other software may be available as a service'': This makes a huge difference, doesn't it?
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{I find it a bit weird to quote passages from the same paper like ``The Accessibility principle (5) states that ?....?'' (line 329 and others).}
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{In the discussion about Unique identification (lines 292+), I miss the discussion of hash values as version identifiers, as applied by Git for instance.
@@
 
+\end{enumerate}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\textbf{Reviewer comment:}
+\begin{enumerate}
+
+\item \textbf{Reviewer comment:}
 \emph{The two tables are relevant and helpful, though I did have a slight problem with the legibility of my printed copy of Table 2 (page 5).
@@
 \textbf{Our response:}
+We agree, and have changed the open symbol to a plus sign to improve legibility.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{The paper does not present a specific research finding, and as such does not feature a research question and experimental procedure that are expected of such papers.
@@
 \textbf{Our response:}
+We agree with the reviewer's description of our paper, and appreciate the view that it should be formally published.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Lines 32--35: as a justification for treating software differently from data, I found this passage raised more questions than answers, and that the comments here were uncompelling and unsupported.
@@
 
-\textbf{Our response:}
+\item \textbf{Our response:}
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 57, et seq: [Nit] I would find this easier to read if the sentence introducing the list of reasons were moved to the same page as the list itself.}
@@
 \textbf{Our response:}
+\niemnote{The formatting will change in the final published paper, so this might not need to be addressed. Or, we can bring in some of the commented statements from above the table}
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 88: ``will be presented'' refers to an event now in the past: change to ``was presented''?}
@@
 \textbf{Our response:}
+We agree, and have made this change.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 113: [Nit] ``work using software'' appears to be very broadly scoped, and could apply to just about any activity using a computer. Suggest ``research using software''.}
@@
 \textbf{Our response:}
+We agree, and have made this change.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Table 2, page 5: I had a problem with the legibility of my printed copy of this table.
@@
 \textbf{Our response:}
+We agree, and have replaced the open bullet with a plus sign for better legibility.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Table 2, page 5: [Comment] This table contains a lot of information, and much food for debate.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 175: [Nit] ``in(1)'' Missing space?}
@@
 \textbf{Our response:}
+This has been fixed.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Lines 241--242: ``Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.'' I found this non sequitur (or am not seeing the logical connection).
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Lines 245--247: ``In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification.
@@
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 327--328: ``When software exists as both source code and another type, we recommend that the source code be cited''.
@@
 \textbf{Our response:}
+Agreed; this change has been been.
 
-\textbf{Reviewer comment:}
+\item \textbf{Reviewer comment:}
 \emph{Line 348: ``Updates to this document''.
@@
 \textbf{Our response:}
+Agreed; this change has been been.
+
+\end{enumerate}
 

### 2016-07-15 sc-principles/software-citation-principles.tex
added appendix with use cases from Google Doc; re editor #4 & reviewer 1 #4

@@
 
+\section{Software citation use cases}
+\label{app:usecases}
+
+This appendix records an edited, extended description of the use cases discussed in~\S\ref{sec:use_cases}, originally found in~\cite{SC-Use-Cases}.
+
+\subsection{Researcher who uses someone else's software for a paper}
+
+One of the most common use cases may be researchers who use someone else's software and want to cite it in a technical paper. 
+This will be similar to existing practices for citing research artifacts in papers.
+
+``Requirements'' for researcher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Software version number and release date, or download date
+\item Location/repository, or contact name/email (if not publicly available)
+\item Citable DOI of software
+\item Format for citing software in text and in bibliography
+\end{itemize}
+
+Possible steps:
+\begin{enumerate} \setcounter{enumi}{0}
+\item Software developers create CITATION file and associate with source code release/repository.
+\item Researcher finds and uses software for research paper.
+\item Researcher identifies citation metadata file (e.g., ``CITATION'' file) associated with downloaded/installed software source code or in online repository/published location.
+CITATION file includes necessary citation metadata.
+CITATION file may include BibTeX entry, suggested citation format
+\item Researcher cites software appropriately, e.g. in methodology section; reference included in bibliography.
+\end{enumerate}
+
+\subsection{Researcher who uses someone else's software for new software}
+
+In this case, a researcher develops new software that incorporates or depends on existing software. 
+In order to credit the developer(s), the researcher will include citations in his/her source code, documentation, or other metadata in a similar manner to papers
+
+Requirements for researcher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Software version number and release date
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in source code, documentation, or citation metadata file
+\end{itemize}
+
+Possible steps:
+\begin{enumerate} \setcounter{enumi}{0}
+\item Assume that software developers have created a CITATION file and associated with the source code release/repository.
+\item Researcher finds and uses software in the development of new software.
+\item Researcher identifies citation metadata file (e.g., ''CITATION'' file) associated with downloaded/installed software source code or in online repository/published location.
+CITATION file includes necessary citation metadata.
+CITATION file may include BibTeX entry, suggested citation format.
+\item Researcher cites software in source code, documentation, or other metadata-containing file.
+\end{enumerate}
+
+\subsection{Researcher who contributes to someone else's software (open source project)}
+
+maybe a variant of above\niemnote{add more to this?}
+
+\subsection{Researcher who wants to know who uses the researcher's software}
+
+This case is similar to a researcher who wants to find other papers/publications that cite a particular paper.
+A researcher wants to gauge the usage of her software within or across communities and measure its impact on research for both credit and funding.
+
+Requirements:
+\begin{itemize}
+\item Uniquely identify software
+\item Indexed citations of software 
+\item Indexed papers that use software
+\end{itemize}
+
+Steps:
+\begin{enumerate}
+\item Researcher finds software official name or unique DOI in metadata associated with downloaded/installed source code or in online repository/published location.
+\item Researcher searches for software, may use online indexer (e.g., Scopus, Web of Science, Google Scholar) using software name or DOI.
+\item Online indexer presents entry for software with list of citations, if any.
+Ideally, entry will also include metadata contained in software CITATION file and citation example.
+\end{enumerate}
+
+\subsection{Researcher who wants to publish about a piece of software}
+
+The research wants to publish about a version of software they have produced.
+A key part of this use case is to be able to connect the given narrative to a specific version of the software in questions and connect that in large story.
+
+Requirements:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of Software
+\item Links to older versions of software
+\end{itemize}
+
+\subsection{Publisher who wants to publish papers that cite software}
+
+\niemnote{add description?}
+
+Requirements for publisher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in, e.g., JATS, as well as references in the text itself
+\end{itemize}
+
+\subsection{Indexer (e.g., Scopus, WoS, Scholar, MS Academic Search) who wants to build a catalog of software}
+
+Provide an index over the software that is used within the research domain. Track how that software is being used by different groups of researchers and to what ends.
+
+Requirements:
+\begin{itemize}
+\item Uniquely identify pieces of software used by the research literature
+\item Connect authors and organizations to that software
+\item Connect various software versions together
+\end{itemize}
+
+\subsection{Domain group (e.g., ASCL, bioCADDIE), Libraries, and Archives (e.g., University library, laboratory archive, etc.) wants to build a catalog\slash registry of institutional or domain software}
+
+There are two different examples here: One is building a catalog/archive of software produced by those affiliated with the institution.
+The other is along the lines of Sayeed Choudhury's note that ``data are the new special collections.''
+An institution may choose to build a catalog/archive of many things within a single topic or subject in order to secure all the software on a certain topic or build a collection that may draw users to their establishment, much like special collections now do for university libraries and archives.
+
+\subsection{Repository showing scientific impact of holdings}
+
+A repository that archives and/or maintains a collection of software.
+The repository would like to address usage and impact of software in its holding.
+Usage would aid potential users whether the software is being actively maintained or developed or has been superseded.
+Both would help repository know how to direct resources, e.g., maintenance, training etc.
+
+Requirements:
+\begin{itemize}
+\item Code name ideally, a unique identifier
+\item Relationships to previous versions
+\item Connect to repository
+\item Connect to research
+\end{itemize}
+
+similar to Funder case
+
+mineable via repository name, code name, DOI, or citation.\niemnote{complete?}
+
+\subsection{Funder who wants to know how software they funded has been used}
+
+steps/requirements\niemnote{add more to this?}
+
+\subsection{Researcher gets credit for software development at the academic\slash governmental institution, in professional career, etc.}
+
+\niemnote{add description?}
+
+Requirements for researcher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in an official CV, in a departmental/institutional review report, etc.
+\item Role in the software creation, that is linked to version or component
+\item Role in contributing to the software as a ``package'' (not just lines of code) development of benchmarks, testing, documentation, tutorials etc.
+\end{itemize}
+
+\subsection{Researcher who wants to ``reproduce'' another person\slash group's analysis}
+
+\niemnote{add description?}
+
+Requirements for researcher:
+\begin{itemize}
+\item Name of software
+\item Location/repository for the exact release that was used 
+\item DOI or other persistent handle for that specific release
+\item Release has all components necessary for reproducing the work (note that this ideally also means sample inputs and outputs)
+\end{itemize}
+
+\subsection{Researcher who benchmarks someone else's software with or without modification on one or many hardware platforms for publication}
+
+\niemnote{add description?}
+
+Requirements for researcher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Software version number and release date
+\item Location/repository
+\item Citable DOI of software or paper recommended for citation
+\item Format for citing software in source code or citation metadata file
+\end{itemize}
+
+Possible steps:
+\begin{enumerate} \setcounter{enumi}{0}
+\item Software developers create CITATION file and associate with source code release/repository.
+\item Researcher finds and uses software in the development of new software.
+\item Researcher identifies citation metadata file (e.g., CITATION file) associated with downloaded/installed software source code or in online repository/published location.
+CITATION file includes necessary citation metadata.
+CITATION file may include BibTeX entry, suggested citation format.
+\item Researcher cites software in source code, documentation, or other metadata-containing file.
+\end{enumerate}
+
+\subsection{Reference management system used by researchers to author a manuscript}
+
+\niemnote{add description?}
+
+Requirements for reference manager:
+\begin{itemize}
+\item Names of software authors/contributors
+\item Software version number and release date
+\item Location/repository
+\item Citable DOI of software or paper recommended for citation
+\item Format for citing software in citation metadata file
+\item Citation metadata tags embedded in DOI landing page/software project page for easy ingest
+\end{itemize}
+
+Possible steps:
+\begin{enumerate}
+\item Reference management system such as EndNote, Mendeley, Zotero, etc. builds affordances for software references.
+\item Researcher finds software citation and adds it to their reference manager library, by (a) importing from the CITATION file (e.g., BibTeX, RIS), or (b) clicking on, e.g., an ``add to Zotero library'' widget in web browser.
+\item Researcher writes a paper and uses the reference manager to generate citations or bibliography.
+\end{enumerate}
+
+\subsection{Researcher who wants to find a piece of software to implement a task}
+
+This is the case where a research is looking for software to use but wants to understand whether it is being used in a scholarly fashion.
+For example, a researcher searches through a software repository and finds a package that might be useful.
+They look to find whether it has been used by others in the scientific literature.
+
+Requirements
+\begin{itemize}
+\item Either the software documentation page has a reference to existing literature that makes use of it. 
+\item There is a mechanism to look it up.
+\end{itemize}
+
+\subsection{Researcher wants to record the software that generated some data}
+
+This is the case where a researcher is using some software to perform an analysis, either of a physical sample or of data.
+The researcher needs to know which version was used, for example in case a bug was fixed.
+Note that knowing the software and its version is not sufficient to determine the ``conditions'' of the analysis, but they are essential.
+
+Requirement:
+The analysis, or the generated data, has information about the software used.
+
+\subsection{Researcher who wants to reproduce experience of use of a particular software implementation in context}
+
+\niemnote{This one was not adopted into table; should we keep it, remote here, or somehow indicate why it was not adopted?}
+
+Researcher is engaged in historical/cultural research -- e.g., study of video games as cultural artifacts.
+
+Requirements
+\begin{itemize}
+\item Name of software
+\item Software version number
+\item Documentation of the execution environment/context
+\item Location/repository for virtual machine (or equivalent) comprising both software and execution environment/context
+\item Persistent identifier associated with virtual machine instance (or equivalent) comprising both software and execution environment/context
+\end{itemize}
+
+Possible steps
+\begin{enumerate}
+\item Researcher obtains persistent ID from citation
+\item Research uses a persistent ID resolution service to resolve ID to a location of  an executable VM instance in a repository 
+\item Researcher obtains VM in the repository, executes it, and interacts with software 
+\end{enumerate}
+
+%References:
+%Use case examples and commentary
+%Library of congress report on software preservation:
+%http://www.digitalpreservation.gov/meetings/preservingsoftware2013.html
+%CASPAR Report on Software Significant Properties
+%http://sigsoft.dcc.rl.ac.uk/twiki/pub/Main/SigSoftTalks/SignificantPropertiesofSoftware-r1.doc
+%Examples of VM archiving and repositories
+%https://olivearchive.org/
+%https://archive.org/details/internetarcade
+
 \bibliographystyle{abbrv}

### 2016-07-17 sc-principles/rebuttal/rebuttal.tex
Starting to add some suggested responses

@@
 \textbf{Our response:}
+\asnote{This seems relatively straightforward?
+We had a workshop where we reviewed the principles with a group of FORCE11 participants (Dan - do we have a list of edits we made that day based on feedback?) and then presented (in ~5 mins) to the main FORCE11 session.
+I don't think the presentations/review at FORCE 2016 changed the principles significantly as they'd already been through pretty substantial review during the drafting process.}
 
@@
 \textbf{Our response:}
+\asnote{I think we just need to say that we thought it was important to draft the principles with the end-user(s) in mind, i.e.
+what different use cases/scenarios for software citation.
+The requirements table is an overview of these use cases and the metadata needs of each of these users/use cases.}
 
@@
 \textbf{Our response:}
+\asnote{Discussed above.}
 
@@
 \textbf{Our response:}
+\asnote{Perhaps we should be adding a short paragraph outlining how we believe that ultimately academic publishing should be moving towards sharing all of the products of researcher activity.
+This work is primarily about making software a citable entity which is just one part of a much larger change in the scholarly ecosystem.}
 
@@
 \textbf{Our response:}
+\asnote{The reviewer's suggested text sounds pretty good to me. Perhaps drop/change 'scientific' in that sentence?}
 
@@
 \textbf{Our response:}
+\asnote{We probably need to dig out some citations about improved/rich citation.
+I'm aware of some work that was going on in the PLOS Labs group about this but I'm not sure if that work was ever published.}
 
@@
 \textbf{Our response:}
+\asnote{Perhaps we can rephrase this to make the reviewer happier. IMO there's no obviously correct answer here.}
 

### 2016-07-18 sc-principles/rebuttal/rebuttal.tex
added some more comments to rebuttal

@@
 \emph{The reviewers did not agree on some aspects described in the paper, e.g.
-"if the software authors ask that a paper should be cited, that should typically be respected".
+``if the software authors ask that a paper should be cited, that should typically be respected''.
 I would like to see an appropriate discussion on these points, so as to clearly address these disagreements.}
@@
 
+\niemnote{Also, perhaps here we can discuss the long-term role of software citation for credit, reuse, reproducibility, etc.?}
+
 \item \textbf{Reviewer comment:}
 \emph{This paper doesn't have an experimental design in the strict sense, as it is more of a position paper than a research article.
-The PeerJ guidelines state that "the submission should clearly define the research question", which is currently not the case for this paper.
-However, I see a research question between the lines, which could be phrased as "how can citations be applied to software products to account for their increasing importance in scientific research?".
+The PeerJ guidelines state that ``the submission should clearly define the research question'', which is currently not the case for this paper.
+However, I see a research question between the lines, which could be phrased as ``how can citations be applied to software products to account for their increasing importance in scientific research?''.
 By making such a research question explicit, the PeerJ requirements can be met.}
@@
 \asnote{The reviewer's suggested text sounds pretty good to me. Perhaps drop/change 'scientific' in that sentence?}
+
+\niemnote{Agreed on both counts}
 
@@
 I think the main findings should be the principles stated in the beginning, and the use cases would then form the main method to arrive at these findings.
-Alternatively, one could label the principles as "assumptions" and apply the use case study to validate these assumptions.
+Alternatively, one could label the principles as ``assumptions'' and apply the use case study to validate these assumptions.
 In any case, I think the findings are valid but they should be labeled as preliminary, as it is not possible to arrive at a final conclusion for such kinds of questions.
@@
 \textbf{Our response:}
+
+\niemnote{If we need to formulate the paper as a research paper, then describing the principles as main findings and use cases\slash discussion as the methods to arrive at them seems an interesting way to do it.}
 
@@
 
+\niemnote{I think we have some citations about the importance of software to research we can add here (e.g., Hannay 2009), as well as the fact that research results depend on citation\slash proper version (Sufi 2014, Sandve 2013, Wilson 2014), and that errors in software or environment variations can effect results (Morin 2012, Soergel 2015).
+Although we probably want to avoid overlap with the challenge paper.}
+
 \item \textbf{Reviewer comment:}
@@
 
+\niemnote{I'm also not sure exactly what to do here; our original statement came about from the working group discussion.
+We could simply remove this sentence, since it might not need to be said explicitly.}
+
 \item \textbf{Reviewer comment:}
@@
 \textbf{Our response:}
+
+\niemnote{I think that we should make it clear that unique identifier includes the branch or version, for the first point.}
 

### 2016-07-18 sc-principles/rebuttal/rebuttal.tex
some new comments

@@
 I don't think the presentations/review at FORCE 2016 changed the principles significantly as they'd already been through pretty substantial review during the drafting process.}
+\katznote{we would have to go back through git versions.
+I know that there were changes to the principles themselves in the group phone calls, but not after that - ie, not in the workshop or general presentation.
+there were some smallish changes to the high level text describing each principle later.}
 
@@
 \niemnote{I think we have some citations about the importance of software to research we can add here (e.g., Hannay 2009), as well as the fact that research results depend on citation\slash proper version (Sufi 2014, Sandve 2013, Wilson 2014), and that errors in software or environment variations can effect results (Morin 2012, Soergel 2015).
-Although we probably want to avoid overlap with the challenge paper.}
+Although we probably want to avoid overlap with the challenge paper.} \katznote{basically, what is important is up the authors to define, and then reviewers to agree with, all in the context of community practices and `standards'}
 
@@
 \textbf{Our response:}
+
+\katznote{we're sorry if this seems weird to the reviewer, but rather than just pointing back to a previous part of the paper, it seems more clear to us to quote the basis by which we are making follow-on comments}
 
@@
 
+\katznote{this is a good point - we probably need to change our text a bit}
+
 \item \textbf{Reviewer comment:}

### 2016-07-19 sc-principles/rebuttal/rebuttal.tex
added more comments

@@
 \usepackage{url,breakurl}
+\usepackage{enumitem}
 
@@
 
-\usepackage{todonotes}
-
-\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
-
 %\date{}                                           % Activate to display a given date or no date
-
-\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
 
@@
 
-\emph{First of all, I would like to thank the reviewers, who have done a fantastic job and provided very insightful comments for improving the paper.
-I hope that the authors will take them into serious consideration in the revision.}
+{\itshape
+First of all, I would like to thank the reviewers, who have done a fantastic job and provided very insightful comments for improving the paper.
+I hope that the authors will take them into serious consideration in the revision.
 
-\emph{From my personal perspective, this is a very good contribution, which addresses one of the most important topics in the ``citation'' domain, i.e.
+From my personal perspective, this is a very good contribution, which addresses one of the most important topics in the ``citation'' domain, i.e.
 software citation.
-I totally agree with the reviewers about the readability of the paper, and I think this paper will deserve publication in PeerJ CS.}
+I totally agree with the reviewers about the readability of the paper, and I think this paper will deserve publication in PeerJ CS.
 
-\emph{However, there are some aspects that should be clarified in the revision. In particular:}
+However, there are some aspects that should be clarified in the revision. In particular:
+}
 
@@
 \section*{Reviewer \#1 (Tobias Kuhn):}
-\emph{This paper argues for the adoption of software citations, basically for the same reasons that data citations have been advocated: Because software packages have become so important in science, they should be appropriately credited and referenced, which can be achieved if we extend the concept of citations in scientific articles.}
 
-\emph{The paper tackles an important problem and provides a very valuable foundation for further discussions and future implementations.
-In particular, the six software citation principles presented in the beginning are very valuable.}
+{\itshape
+This paper argues for the adoption of software citations, basically for the same reasons that data citations have been advocated: Because software packages have become so important in science, they should be appropriately credited and referenced, which can be achieved if we extend the concept of citations in scientific articles.
 
-\emph{While I think this is a very useful paper, I see the following problems with the current version:}
+The paper tackles an important problem and provides a very valuable foundation for further discussions and future implementations.
+In particular, the six software citation principles presented in the beginning are very valuable.
+
+While I think this is a very useful paper, I see the following problems with the current version:
+}
 
@@
 
+\niemnote{I feel like I remember discussing this; perhaps we should just add it in the discussion? Certainly we'll still recommend DOI for reasons already mentioned.}
+
 \end{enumerate}
@@
 
-\item \textbf{Our response:}
+\textbf{Our response:}
+
+\niemnote{I think we all would agree that software is a research method and data are research findings, but I do think we need some explanation of the differences---if only to justify why software citation may be different than data citation}
 
@@
 \item \textbf{Reviewer comment:}
-\emph{Table 2, page 5: [Comment] This table contains a lot of information, and much food for debate.
-I was surprised by some omissions, which may mean that I wasn't fully understanding the concepts being presented:\\
-\\
-(a) I find it hard to imagine any of the use cases not benefiting from availability of a ``Description'', especially 1, 2, 9, and 16.\\
-(b) I would have expected ``Indexed citations'' to be useful for 5: getting credit.\\
-(c) For publishing a software paper, I would expect all of the basic requirements other than ``Indexed citations'' to be at least beneficial.\\
-(d) For 12: build a software catalog/registry, would there not be a role for indexed citations?\\
-(e) For 16: Store software try, I'd expect pretty much all of the Basic requirements other than ``Indexed citations'' to be relevant, and particularly ``Description'' and ``Keywords''.\\
-(f) I think there is one use-case missing, which I think is commonly faced by researchers: viz.
-``Evaluate suitability of software for a given task'' - I think this is distinct from discovery, and typically involves making a choice between different available software packages, or to use an existing package, modify an existing package or develop something new.}
+{\itshape
+Table 2, page 5: [Comment] This table contains a lot of information, and much food for debate.
+I was surprised by some omissions, which may mean that I wasn't fully understanding the concepts being presented:
+\begin{enumerate}[label={(\alph*)}]
+\item I find it hard to imagine any of the use cases not benefiting from availability of a ``Description'', especially 1, 2, 9, and 16.
+\item I would have expected ``Indexed citations'' to be useful for 5: getting credit.
+\item For publishing a software paper, I would expect all of the basic requirements other than ``Indexed citations'' to be at least beneficial.
+\item For 12: build a software catalog/registry, would there not be a role for indexed citations?
+\item For 16: Store software entry, I'd expect pretty much all of the Basic requirements other than ``Indexed citations'' to be relevant, and particularly ``Description'' and ``Keywords''.
+\item I think there is one use-case missing, which I think is commonly faced by researchers: viz.
+``Evaluate suitability of software for a given task'' - I think this is distinct from discovery, and typically involves making a choice between different available software packages, or to use an existing package, modify an existing package or develop something new.
+\end{enumerate}
+}
 
 \textbf{Our response:}
+
+\begin{enumerate}[label={(\alph*)}]
+\item \niemnote{I'm fine with adding description to these, unless we had a reason not to. I think the decisions on these were fairly arbitrary}
+\item \niemnote{I think I agree, or at least benefit}
+\item \niemnote{I think what we had originally makes sense, but perhaps there needs to be more explanation?}
+\item \niemnote{no, indexed citations not really relevant for catalog/registry, unless it wants to display citations of software---but that is use case 4}
+\item \niemnote{no, I don't think those things would be stored---because they aren't included in the citation itself}
+\item \niemnote{Is this different than 7, ``Find software to implement task''? I suppose it's not exactly the same, and I'm fine with adding a new one to appease the reviewer}
+\end{enumerate}
 
@@
 \textbf{Our response:}
+
+\niemnote{The table collects use cases that relate to or depend on software citation, but certainly each use case has additional dependencies that are not and cannot be contained in the table.
+So, perhaps we can rephrase somewhere to emphasize that the table requirements are only for the citation of software related to each use case, and not meant to be exhaustive about everything actually required for the use case beyond citation?}
 

### 2016-07-20 sc-principles/rebuttal/rebuttal.tex
added some rebuttal draft responses and comments

@@
 \item \textbf{Reviewer comment:}
-\emph{line 88: ``which will be presented ...'' > ``which was presented ...''}
+\emph{line 88: ``which will be presented ...'' to ``which was presented ...''}
 
@@
 \textbf{Our response:}
+Principle (6) states that ``Software identification should be as specific as necssary'', and we consider that the specific version includes the branch.
 
-\niemnote{I think that we should make it clear that unique identifier includes the branch or version, for the first point.}
+We agree that (3) is not relevant in the context of citation, and have removed it from the discussion.
+However, it is a valid identifier that could be used in other contexts.
 
@@
 \textbf{Our response:}
+Software as a service is a fact, and we need a method to cite it---even though some metadata may not exist or be available, we need to do the best we can.
+The same issues would hold for commerical software or that available only as an executable.
+If you can't find the name of software, should be able to find the name of the service\slash endpoint or function being used.
+
+Regarding accessibility, ideally we would like the source to be available, but if not we still need a way to cite.
+In this case, the Accessibility principle would apply to the service itself, rather than the source.
 
@@
 \textbf{Our response:}
-
-\katznote{we're sorry if this seems weird to the reviewer, but rather than just pointing back to a previous part of the paper, it seems more clear to us to quote the basis by which we are making follow-on comments}
+We feel it is more clear to a reader to quote the basis by which we are making follow-on comments, rather than pointing back to a previous part of the paper.
 
@@
 \niemnote{I think we all would agree that software is a research method and data are research findings, but I do think we need some explanation of the differences---if only to justify why software citation may be different than data citation}
+\niemnote{Dan: work on this}
 
@@
 \textbf{Our response:}
-\niemnote{The formatting will change in the final published paper, so this might not need to be addressed. Or, we can bring in some of the commented statements from above the table}
+We agree that this would be easier to read, but believe the formatting will change in the final published paper.
 
@@
 \item \niemnote{no, I don't think those things would be stored---because they aren't included in the citation itself}
-\item \niemnote{Is this different than 7, ``Find software to implement task''? I suppose it's not exactly the same, and I'm fine with adding a new one to appease the reviewer}
+\item We intended ``find'' to mean more than discovery, but since the appendix was not present in the first submission this was not clear.
+We have altered the use case to be ``Find/choose software to implement task'' to make it clear that this use case goes beyond discovery.
 \end{enumerate}
@@
 
-\katznote{this is a good point - we probably need to change our text a bit}
+\katznote{this is a good point - I probably need to change our text a bit}
 
@@
 So, perhaps we can rephrase somewhere to emphasize that the table requirements are only for the citation of software related to each use case, and not meant to be exhaustive about everything actually required for the use case beyond citation?}
+The requirements in the use cases table only indicate the metadata needed for citation related to each use case, and not necessarily everything needed to complete each use case.
+For example, use case 7 ``Benchmark software'' does not include information about actual benchmarking, or the need for similar hardware, etc.
 
@@
 \textbf{Our response:}
-Agreed; this change has been been.
+Agreed; this change has been made.
 

### 2016-07-20 sc-principles/software-citation-principles.tex
removed discussion of identifier for latest version

@@
 While we often need to cite a specific version of software, we may also need a way to cite the
-software in general, or the latest version, and to link multiple releases together, perhaps for
+software in general and to link multiple releases together, perhaps for
 the purpose of understanding citations to the software.  The principles in \S\ref{sec:principles} are

### 2016-07-20 sc-principles/software-citation-principles.tex
softened recommendation on paper citation

@@
 important to the work, then the software paper should also be cited.
-In addition, if the software authors ask that a paper should be cited, that should typically be respected and the paper cited \textit{in addition to} the software being cited.
-%There are situations where it is appropriate to cite the software, the software paper, or both.
+We believe that a request from the software authors to cite a paper should typically be respected, and the paper cited \textit{in addition to} the software.
 

### 2016-07-20 sc-principles/rebuttal/rebuttal.tex
added rebuttal comments, and update use case table

@@
 
+\niemnote{I will attempt this}
+
 \item \textbf{Comment:}
@@
 
+\niemnote{Move paragraph at end of Section 1 to Section 3, and expand/merge with paragraph after mission statement; find text from FORCE2016 presentation about use cases and explaining produre to get principles}
+\niemnote{I will do this}
+
 \item \textbf{Comment:}
@@
 \textbf{Our response:}
+We agree with Reviewer 2 on this issue, as this represents more of a position paper than an experimental paper.
+That said, we would be willing to add an explicit research question and discuss the ``experimental'' design if required, but feel this would be an artificial construct that doesn't add much to the paper.
 
@@
 \textbf{Our response:}
-\niemnote{I will add the full use cases text as an appendix}
+\niemnote{I will add the full use cases text as an appendix, and revise to match the numbering in the table}
+\niemnote{I will also add the FORCE16 comments and responses}
+
+\niemnote{Dan \& Arfon: could you assist with adding descriptions to some of the use cases that don't have one?}
 
@@
 \textbf{Our response:}
+We have either made appropriate changes or discussed these points.
 
@@
 \textbf{Our response:}
-\asnote{Discussed above.}
+
+\niemnote{This will be addressed somewhat by the changes discussed above; but Dan, could you help with this?}
 
@@
 
-\niemnote{Also, perhaps here we can discuss the long-term role of software citation for credit, reuse, reproducibility, etc.?}
+\niemnote{Dan, can you do this?}
 
@@
 \niemnote{Agreed on both counts}
+
+\niemnote{Let's say that}
+We agree that this paper is more of a position paper than research article, and thus that creating a research question would be an artificial construct.
+We are willing to make this change if required, but feel it wouldn't add much to the paper.
 
@@
 \niemnote{If we need to formulate the paper as a research paper, then describing the principles as main findings and use cases\slash discussion as the methods to arrive at them seems an interesting way to do it.}
+
+\niemnote{We agree that the main findings are the principles, and the use cases/discussion are the main to reach those.
+Don't agree that the principles are assumptions or preliminary, but reflect current knowledge and experience}
+
+\niemnote{I will work on this, but may need help}
 
@@
 
+\niemnote{Citations are going to go away---but format and how they fit in ecosystem may change.}
+
+\niemnote{Arfon: can you help here?}
+
 \item \textbf{Reviewer comment:}
@@
 
+\niemnote{Add discussion of both points; Dan's note is already there in the text. I will work on this.}
+
 \item \textbf{Reviewer comment:}
@@
 \textbf{Our response:}
-\asnote{Perhaps we can rephrase this to make the reviewer happier. IMO there's no obviously correct answer here.}
-
-\niemnote{I'm also not sure exactly what to do here; our original statement came about from the working group discussion.
-We could simply remove this sentence, since it might not need to be said explicitly.}
+We have modified the text to make it clear that this is a recommendation from the working group; the statement reflects an agreed position of the entire working group.
+The modification should make it clear that this is a recommendation only.
 
@@
 
-\niemnote{I feel like I remember discussing this; perhaps we should just add it in the discussion? Certainly we'll still recommend DOI for reasons already mentioned.}
+\niemnote{Arfon: can you add some discussion here?
+We can add a sentence addressing git hashes and other version control identifiers, and point out that while they may be sufficient in some ways (globally unique), they aren't necessarily resolvable on the Web, and different VCS use different identifiers, etc.
+In other words, they could be used (especially in the absence of other identifier like DOI), but not sufficient or recommended.}
 
@@
 \niemnote{I think we all would agree that software is a research method and data are research findings, but I do think we need some explanation of the differences---if only to justify why software citation may be different than data citation}
-\niemnote{Dan: work on this}
+
+\niemnote{Dan: work on this. Bring in text from CodeMeta paper and your blog post.}
 
@@
 \begin{enumerate}[label={(\alph*)}]
-\item \niemnote{I'm fine with adding description to these, unless we had a reason not to. I think the decisions on these were fairly arbitrary}
+\item Agreed for 1 and 2.
 \item \niemnote{I think I agree, or at least benefit}

### 2016-07-20 sc-principles/software-citation-principles.tex
added rebuttal comments, and update use case table

@@
 \midrule
-1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + &             &             & Researcher \\
-2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + &             &             & Researcher, software engineer \\
+1.\ Use software for a paper                     & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &   & + & + &   & Researcher \\
+2.\ Use software in\slash with new software      & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &   & + & + &   & Researcher, software engineer \\
 3.\ Contribute to software                       & \textbullet & \textbullet & \textbullet & + & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher, software engineer \\

### 2016-07-20 sc-principles/software-citation-principles.tex
Added FORCE2016 discussion as appendix

@@
 \usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage{csquotes}
 \usepackage{lmodern}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Software citation use cases}
 \label{app:usecases}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 CITATION file may include BibTeX entry, suggested citation format
-\item Researcher cites software appropriately, e.g. in methodology section; reference included in bibliography.
+\item Researcher cites software appropriately, e.g., in methodology section; reference included in bibliography.
 \end{enumerate}
@@
 
-%References:
-%Use case examples and commentary
-%Library of congress report on software preservation:
-%http://www.digitalpreservation.gov/meetings/preservingsoftware2013.html
-%CASPAR Report on Software Significant Properties
-%http://sigsoft.dcc.rl.ac.uk/twiki/pub/Main/SigSoftTalks/SignificantPropertiesofSoftware-r1.doc
-%Examples of VM archiving and repositories
-%https://olivearchive.org/
-%https://archive.org/details/internetarcade
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Feedback following FORCE16}
+\label{app:force16-feedback}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This appendix contains a record of comments made by the FORCE11 community on the draft Software Citation Principles, either directly via Hypothesis on the draft document\footnote{\url{https://www.force11.org/software-citation-principles}} posted following the FORCE2016 conference~\cite{force2016} or via GitHub issues\footnote{\url{https://github.com/force11/force11-scwg/issues}}, and the responses to these comments.
+
+\subsection{On unique identification:} I know this suggestion of a single unique identifier comes from the DOI perspective where it works pretty well, but I'm wondering if something different in the way of identification should be used for software.
+For creative works generally there is the FRBR model ( \url{https://en.wikipedia.org/wiki/Functional_Requirements_for_Bibliographic_Records}) which defines several levels for a creative entity - ``work'', ``expression'', ``manifestation'', and ``item''.
+I think something along these lines are particularly relevant for software - it is useful to be able to locate all uses of a particular piece of software no matter what version (the ``work'' level - software identified by a particular name and purpose over a period of time), but it is also important to specify the particular version used in any given work (``expression'' - the source code at the time of use) and in some cases also the platform (``manifestation'' - the compiled bytes including libraries, for example a docker image).
+``Item'' probably isn't relevant for software.
+That is, I think a software citation perhaps could use THREE distinct unique identifiers, one for the work itself, one for the specific version (source code), and possibly an additional one for the actual downloadable binary image that can be run.
+Rather than leave it implicit I think recognizing the different levels of citable record would be helpful here.
+\#{F11SC}
+
+\emph{Reply:} I interpret the requirement for ``global uniqueness'' as referring to the identifier itself.
+Two different people can have the same name (not globally unique) but cannot share a single ORCID (globally unique).
+Global uniqueness of the identifier does not preclude multiple identifiers pointing to the same person.
+I think the suggestion of differentiating between different software expressions\slash manifestations\slash items is a reasonable one, but I don't think it relaxes the requirement for identifiers to be globally unique.
+
+\textbf{Our response:}
+We agree that there are valid points here, but on balance we don't feel that the rewards from implementing this outweigh the practical challenges.
+
+\subsection{On accessibility:} Should this document address this in further detail?
+For example, ``permit and facilitate access'' could be explored further.
+Should this be done through open access licensing?
+repositories?
+Who's responsible for providing this access?
+
+I am also wondering if this is a separate issue since ``citing'' traditionally pointed to publications but did not necessarily address access.
+DOI, for example is stated, but doesn't guarantee ``access'', so does this simply restating point 3, or should it provide something new?
+
+\textbf{Our response:}
+We agree that accessibility should receive further attention, which the follow-on group focusing on implementation will provide.
+However, this is out of scope for the document outlining the principles.
+
+To the second point, accessibility provides information about access, but does not guarantee access itself (e.g., paywalled article).
+
+\subsection{On specificity:} I am wondering if this should be folded into number 3 ``Unique Identification''. Both seem to deal with the issue of identification and access.
+
+\textbf{Our response:} A unique software identifier \textbf{can} point to the specific version\slash variant of software, but it can also identify other things (collection of versions, repository, etc.), while this principle deals with the need to identify the specific version of software used (via citation).
+
+\subsection{On academic credit:} A lot of software that were developed by non-academic engineers also contribute to academic research indirectly.
+Their names and contributions should also be credited.
+So removing ``Academic'' makes more sense?
+
+\emph{Reply:} This is a good point, though I think academic and non-academic credit are different, so perhaps we can add to this regarding non-academic credit, rather than removing ``academic''.
+
+\emph{Reply:} I agree with Daniel on this. Keep Academic and add non-academic.
+
+\textbf{Our response:} We've made the bullet more general, just about credit, discussing academic credit and adding a sentence about non-academic credit as well.
+
+\subsection{On citations in text:} Although the focus here is on citations in the references, as a publisher, our experience is that most common practice of ``citation'' of data and software for authors is typically in the main body of the text.
+In order to encourage software to be treated and valued as a first-class research object, it is important that citations to it be positioned in the references as citations to articles and books are.
+However, it would be a missed opportunity if we did not leverage current practices of authors.
+This will also likely arise during implementation, as it has for the Data Citation Implementation Publisher Early Adopters Pilot.
+This could be addressed in future work on implementation.
+
+\textbf{Our response:} In the principles, we propose that software should be cited in the references list, to recognize the primary role of software in research.
+However, this practice is not mutually exclusive with \textbf{also} referencing\slash citing software in the main body of a paper---as long as the software is cited in the references.
+
+\subsection{On unique identification:} Clearer instructions will be needed for authors on which version to cite.
+For BioMed Central journals, we ask authors to cite two versions of the software, an archived version (e.g., on Zenodo) as well as the current version (e.g., on GitHub).
+This is to ensure accessibility.
+However, if repositories and archives were to include a persistent link to the current version of the software, publishers could then instruct authors to cite only software with a UID, which wouldn't point to a current version, but would point to the version(s) used and would be a more accurate version of scientific record.
+Related to this point is the idea of group object identifiers.
+A need for group identifiers has been identified in the area of data (e.g., in the case of meta-analyses), and one could also identify a use case for these in the case of software, collecting metadata around all versions of a given software package.
+See blog here (\url{https://blog.datacite.org/to-better-understand-research-communication-we-need-a-groid-group-object-identifier/}).
+
+\textbf{Our response:} We recommend citing the specific version of the software that was used.
+We expect that the unique identifier (e.g., DOI) will point to a landing page that directs to the repository\slash current version.
+However, this is more of a convenience issue that the software developers should address, rather than the author citing the software they used.
+
+\subsection{On future work:} For implementation we would recommend both consulting with adopters as well as developing metadata standards simultaneously rather than developing metadata standards and then pursuing early adopters implementation.
+The work early adopters are doing now for data citation will be able to be leveraged for software citation and the changes needed to do so could happen now.
+There is no need to wait on approval of new tagging for a specific metadata standard.
+Many publishers will have their own preferred metadata standards and so implementation could begin now with publishers, as long as we know what we want to capture.
+Future implementation groups might also consider levels of contribution.
+This is particularly relevant for software.
+Who is considered an author?
+For example, to what extent should authors of pull requests receive attribution?
+This might be considered in an FAQs group, or possibly an early adopters group.
+
+\textbf{Our response:} We agree that metadata standards should be developed with the input of adopters, and have updated this text accordingly.
+
+\subsection{Additional thoughts (not sure what section this applies to):} The principles do not address virtual machines.
+As these are becoming more common and relevant when addressing the reproducibility of research, it is important this ``form'' of software is acknowledged.
+The question remains in which cases should authors cite the current version, which the static archived version, and in which the virtual machine?
+In this way software is very much a unique evolving research object and might not fit perfectly into the same citation practices and structure as other research objects.
+In addition, software citation could possibly occur within the virtual machine.
+This could be added as a use case.
+
+\textbf{Our response:} We feel this has been addressed in Section 5.8, with the explicit addition of virtual machines in addition to executables and containers.
+This is also an issue that should be addressed further by the follow-on implementation working group.
+
+\subsection{On persistence of identifier vs.\ persistence of software:} The persistence principle outlined in (4) is a key element in making software citeable.
+Where software has become part of the record of science not only the identifier and metadata of the software should be persistent, it should also be the goal to keep a persistent copy of the source code, where applicable.
+This links with the accessibility principle (5).
+
+There are still many open questions about how to resolve package dependencies in the long term, therefore I would not make the persistent access to code a hard requirement but may add something more specific towards preserving the record of science.
+
+\textbf{Our response:} Our goal is for software citations to point to (persistent) archived source code, but we are not---nor could we---require this.
+
+\subsection{Granularity of the citation:} One of the key issues with any citation, whether document, individual, or software is the specificity of what is being cited.
+In the case of publications, there is almost zero specificity most of the time.
+
+It's very easy to cite an entire package even though one function was used.
+Part of this problem is being solved in the Python world through this project (\url{https://github.com/duecredit/duecredit}).
+
+Any citation should have the ability to specify more than just the obvious, but even the obvious would be a good starting point.
+
+The citation\slash url should therefore allow for greater specificity within a code base.
+In general though, a provenance record of the workflow would be significantly more useful than a citation from a research perspective.
+
+\textbf{Our response:} We agree that greater specificity is desirable in some cases, but we do not believe this rises to the level of what should be specified or discussed in the principles at this time.
+
+\subsection{``Software citations should permit \ldots access to the software itself'':} Under the ``Access'' header, the data declaration states that:
+
+\begin{displayquote}
+    Data citations should facilitate access to the data themselves
+\end{displayquote}
+
+Under the same header, the software declaration states:
+
+\begin{displayquote}
+    Software citations should permit and facilitate access to the software itself
+\end{displayquote}
+
+The addition of ``permit'' suggests that software citations should also grant the user with permission to access the software. Is this intentional?
+
+It doesn't seem like a good idea to make access a requirement for discovery, so ``permit'' might not be helpful in this sentence.
+
+\textbf{Our response:} To avoid confusion, we removed ``permit and'' from the accessibility principle.
+
+\subsection{Access to software: free vs commercial:} The section talks about software that is ``free'' as well as ``commercial'' software.
+I am not sure whether this is about free as in freedom (or just gratis or freely available), since it is compared with commercial software, which is unrelated in general, see \url{http://www.gnu.org/philosophy/words-to-avoid.html#Commercial}
+
+I suppose that ``free'' should be replaced by ``gratis'' and ``commercial'' be replaced by ``non-free'' in that section.
+
+\textbf{Our response:} We think this is sufficiently clear as written.
 

### 2016-07-20 sc-principles/software-citation-principles.tex
added mention of new appendix

@@
 practices in member disciplines; gather materials from workshops and other reports;
-review those materials, identifying overlaps and differences; and subsequently draft this resulting document,
-which was presented and discussed at the Force2016 Conference~\cite{force2016}
-in April 2016.  This discussion led to a second, final version, and we also
-plan to have a follow-on working group that will work with stakeholders to ensure that
-these principles impact the research process.
+review those materials, identifying overlaps and differences; and subsequently draft an initial version of this document.
+The draft Software Citation Principles document was discussed in a day-long workshop and presented at the FORCE2016 Conference~\cite{force2016} in April 2016.
+Members of the workshop and greater FORCE11 community gave feedback, which we recorded here in Appendix~\ref{app:force16-feedback}.
+This discussion led to a second, final version.
+We also plan to initiate a follow-on working group that will work with stakeholders to ensure that these principles impact the research process.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Feedback following FORCE16}
+\section{Feedback following FORCE2016}
 \label{app:force16-feedback}

### 2016-07-20 sc-principles/rebuttal/rebuttal.tex
response to editor comment about activities at FORCE2016

@@
 \textbf{Our response:}
-\asnote{This seems relatively straightforward?
-We had a workshop where we reviewed the principles with a group of FORCE11 participants (Dan - do we have a list of edits we made that day based on feedback?) and then presented (in ~5 mins) to the main FORCE11 session.
-I don't think the presentations/review at FORCE 2016 changed the principles significantly as they'd already been through pretty substantial review during the drafting process.}
-\katznote{we would have to go back through git versions.
-I know that there were changes to the principles themselves in the group phone calls, but not after that - ie, not in the workshop or general presentation.
-there were some smallish changes to the high level text describing each principle later.}
-
-\niemnote{I will attempt this}
+We have added some additional description of these activities in Section 2,
+and also a new appendix that records the feedback and discussion following
+the FORCE2016 workshop and presentation. The workshop and feedback following
+presentation to the broader FORCE11 community led to some edits to the use
+cases and discussion, but not the principles themselves.
 

### 2016-07-20 sc-principles/software-citation-principles.tex
rearranged use cases in appendix

@@
 Members of the workshop and greater FORCE11 community gave feedback, which we recorded here in Appendix~\ref{app:force16-feedback}.
-This discussion led to a second, final version.
+This discussion led to some changes in the use cases and discussion, although the principles themselves were not modified.
 We also plan to initiate a follow-on working group that will work with stakeholders to ensure that these principles impact the research process.
@@
 
-\subsection{Researcher who wants to publish about a piece of software}
-
-The research wants to publish about a version of software they have produced.
-A key part of this use case is to be able to connect the given narrative to a specific version of the software in questions and connect that in large story.
-
-Requirements:
-\begin{itemize}
-\item Name of software
-\item Names of software authors/contributors
-\item Location/repository
-\item Citable DOI of Software
-\item Links to older versions of software
-\end{itemize}
-
-\subsection{Publisher who wants to publish papers that cite software}
-
-\niemnote{add description?}
-
-Requirements for publisher:
-\begin{itemize}
-\item Name of software
-\item Names of software authors/contributors
-\item Location/repository
-\item Citable DOI of software
-\item Format for citing software in, e.g., JATS, as well as references in the text itself
-\end{itemize}
-
-\subsection{Indexer (e.g., Scopus, WoS, Scholar, MS Academic Search) who wants to build a catalog of software}
-
-Provide an index over the software that is used within the research domain. Track how that software is being used by different groups of researchers and to what ends.
-
-Requirements:
-\begin{itemize}
-\item Uniquely identify pieces of software used by the research literature
-\item Connect authors and organizations to that software
-\item Connect various software versions together
-\end{itemize}
-
-\subsection{Domain group (e.g., ASCL, bioCADDIE), Libraries, and Archives (e.g., University library, laboratory archive, etc.) wants to build a catalog\slash registry of institutional or domain software}
-
-There are two different examples here: One is building a catalog/archive of software produced by those affiliated with the institution.
-The other is along the lines of Sayeed Choudhury's note that ``data are the new special collections.''
-An institution may choose to build a catalog/archive of many things within a single topic or subject in order to secure all the software on a certain topic or build a collection that may draw users to their establishment, much like special collections now do for university libraries and archives.
-
-\subsection{Repository showing scientific impact of holdings}
-
-A repository that archives and/or maintains a collection of software.
-The repository would like to address usage and impact of software in its holding.
-Usage would aid potential users whether the software is being actively maintained or developed or has been superseded.
-Both would help repository know how to direct resources, e.g., maintenance, training etc.
-
-Requirements:
-\begin{itemize}
-\item Code name ideally, a unique identifier
-\item Relationships to previous versions
-\item Connect to repository
-\item Connect to research
-\end{itemize}
-
-similar to Funder case
-
-mineable via repository name, code name, DOI, or citation.\niemnote{complete?}
-
-\subsection{Funder who wants to know how software they funded has been used}
-
-steps/requirements\niemnote{add more to this?}
-
 \subsection{Researcher gets credit for software development at the academic\slash governmental institution, in professional career, etc.}
@@
 
+\subsection{Researcher who wants to find a piece of software to implement a task}
+
+This is the case where a research is looking for software to use but wants to understand whether it is being used in a scholarly fashion.
+For example, a researcher searches through a software repository and finds a package that might be useful.
+They look to find whether it has been used by others in the scientific literature.
+
+Requirements
+\begin{itemize}
+\item Either the software documentation page has a reference to existing literature that makes use of it.
+\item There is a mechanism to look it up.
+\end{itemize}
+
+\subsection{Publisher wants to publish a software paper}
+
+\niemnote{add description}
+
+\subsection{Publisher who wants to publish papers that cite software}
+
+\niemnote{add description?}
+
+Requirements for publisher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in, e.g., JATS, as well as references in the text itself
+\end{itemize}
+
+\subsection{Indexer (e.g., Scopus, WoS, Scholar, MS Academic Search) who wants to build a catalog of software}
+
+Provide an index over the software that is used within the research domain. Track how that software is being used by different groups of researchers and to what ends.
+
+Requirements:
+\begin{itemize}
+\item Uniquely identify pieces of software used by the research literature
+\item Connect authors and organizations to that software
+\item Connect various software versions together
+\end{itemize}
+
+\subsection{Domain group (e.g., ASCL, bioCADDIE), Libraries, and Archives (e.g., University library, laboratory archive, etc.) wants to build a catalog\slash registry of institutional or domain software}
+
+There are two different examples here: One is building a catalog/archive of software produced by those affiliated with the institution.
+The other is along the lines of Sayeed Choudhury's note that ``data are the new special collections.''
+An institution may choose to build a catalog/archive of many things within a single topic or subject in order to secure all the software on a certain topic or build a collection that may draw users to their establishment, much like special collections now do for university libraries and archives.
+
+\subsection{Repository showing scientific impact of holdings}
+
+A repository that archives and/or maintains a collection of software.
+The repository would like to address usage and impact of software in its holding.
+Usage would aid potential users whether the software is being actively maintained or developed or has been superseded.
+Both would help repository know how to direct resources, e.g., maintenance, training etc.
+
+Requirements:
+\begin{itemize}
+\item Code name ideally, a unique identifier
+\item Relationships to previous versions
+\item Connect to repository
+\item Connect to research
+\end{itemize}
+
+similar to Funder case
+
+mineable via repository name, code name, DOI, or citation.\niemnote{complete?}
+
+\subsection{Funder who wants to know how software they funded has been used}
+
+steps/requirements\niemnote{add more to this?}
+
+\subsection{Evaluator or funder wants to evaluate contributions of a researcher}
+
+\niemnote{add description}
+
 \subsection{Reference management system used by researchers to author a manuscript}
@@
 
-\subsection{Researcher who wants to find a piece of software to implement a task}
-
-This is the case where a research is looking for software to use but wants to understand whether it is being used in a scholarly fashion.
-For example, a researcher searches through a software repository and finds a package that might be useful.
-They look to find whether it has been used by others in the scientific literature.
-
-Requirements
-\begin{itemize}
-\item Either the software documentation page has a reference to existing literature that makes use of it.
-\item There is a mechanism to look it up.
+\subsection{Repository wants to publish mixed data\slash software packages}
+
+\niemnote{add description; came from \url{https://github.com/force11/force11-scwg/issues/128}}
+Domain and institutional data repositories have both data and software
+artifacts, and want to link these together in a provenance trace that can
+be cited. Sometimes the software is a separately identified artifact, but at
+other times software is included inside of data packages, and the researcher
+wants to cite the combined product.
+
+\subsection*{Use cases not adopted in the table:}
+
+\subsection{Researcher who wants to publish about a piece of software}
+
+\niemnote{this seems the same as use case 1}
+The research wants to publish about a version of software they have produced.
+A key part of this use case is to be able to connect the given narrative to a specific version of the software in questions and connect that in large story.
+
+Requirements:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of Software
+\item Links to older versions of software
 \end{itemize}
@@
 
+\niemnote{Is this distinct from 6? Or should it be a new one?}
 This is the case where a researcher is using some software to perform an analysis, either of a physical sample or of data.

### 2016-07-20 sc-principles/rebuttal/rebuttal.tex
adding some comments in the rebuttal

@@
 
-\niemnote{Dan \& Arfon: could you assist with adding descriptions to some of the use cases that don't have one?}
+\niemnote{Dan \& Arfon: could you assist with adding descriptions to some of the use cases that don't have one?} \katznote{In both the Google doc and the appendix I think?}
 
@@
 
-\niemnote{Dan, can you do this?}
+\niemnote{Dan, can you do this?} \katznote{I was hoping Arfon might do this one}
 
@@
 We are willing to make this change if required, but feel it wouldn't add much to the paper.
+\katznote{and maybe add}
+We have modeled our process on the work done by the FORCE2016 Data Citation group, which included publishing a PeerJ CS paper (\url{https://peerj.com/articles/cs-1/}) which has been written in a similar manner to our paper.
 

### 2016-07-20 sc-principles/software-citation-principles.tex
adding description to B.5 and google doc

@@
 
-\niemnote{add description?}
+This case describes the need for a research who has contributed to software (by design, software engineering, development, testing, patching, documentation, training, evangelizing, etc.)
+to have their software work recognized by their employer or colleagues for the purpose of career advancement and increased professional reputation.
 

### 2016-07-20 sc-principles/rebuttal/rebuttal.tex
fixing a comment from Kyle - please check to make sure this is correct
adding some sentences to the start of B and a note in the rebuttal to refer to this

@@
 
-\niemnote{Citations are going to go away---but format and how they fit in ecosystem may change.}
+\niemnote{The idea of citations is not going to go away---but format and how they fit in ecosystem may change.}
 
@@
 
+\katznote{also as a general comment, see the new intro to appendix B that states that some of this may be incomplete or wrong, but is included as a record to support the process that come up with the body of the document, and that Appendix B will be subject to further work in the future.}
+
 \item \textbf{Reviewer comment:}

### 2016-07-20 sc-principles/software-citation-principles.tex
adding description for B.10 and in google doc
adding some sentences to the start of B and a note in the rebuttal to refer to this

@@
 
-This appendix records an edited, extended description of the use cases discussed in~\S\ref{sec:use_cases}, originally found in~\cite{SC-Use-Cases}.
+This appendix records an edited, extended description of the use cases discussed in~\S\ref{sec:use_cases}, originally found in~\cite{SC-Use-Cases}.  This discussion
+is not fully complete, and in some cases, it may not be fully self-consistent, but it is
+part of this paper as a record of one of the inputs to the principles.  We expect that
+the follow-up Software Citation Implementation Group will further develop these uses
+case, including explaining in more detail how the software citation principles can be applied
+to each as part of working with the stakeholders to persuade them to actually implement
+the principles in their standard workflows.
 
@@
 
-\niemnote{add description?}
+This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
 

### 2016-07-21 sc-principles/software-citation-principles.tex
some wording fixes at beginning of appendix B

--- 2016-07-20 sc-principles/software-citation-principles.tex
@@
 
-This appendix records an edited, extended description of the use cases discussed in~\S\ref{sec:use_cases}, originally found in~\cite{SC-Use-Cases}.
+This appendix records an edited, extended description of the use cases
+discussed in~\S\ref{sec:use_cases}, originally found in~\cite{SC-Use-Cases}.
+This discussion is not fully complete, and in some cases, it may not be fully
+self-consistent, but it is part of this paper as a record of one of the inputs
+to the principles. We expect that the follow-on Software Citation
+Implementation Group will further develop these use cases, including explaining
+in more detail how the software citation principles can be applied
+to each as part of working with the stakeholders to persuade them to actually
+implement the principles in their standard workflows.
 

### 2016-07-21 sc-principles/software-citation-principles.tex
moving bits around and new section for better structure

@@
 
+\niemnote{Do we want a better introduction, to help with the structure of the paper?}
+
 The principles in this section are written fairly concisely, and discussed
@@
 \end{enumerate}
-
-These software citation principles were originally based on an adaptation of the
-FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were
-modified based on discussions of the FORCE11 Software Citation Working
-Group (see Appendix~\ref{app:wg_members} for members), information from the use
-cases in \S\ref{sec:use_cases}, and the related work in \S\ref{sec:related_work}.
-The adaptations have been made
-because software, while similar to data in terms of not traditionally having
-been cited in publications, is also different than data in that it can be used
-to express or explain concepts, it is updated more frequently, and it is
-executable.
-% \katznote{are there any other difference between software and data that should be stated here?}
-Also, while software can be considered a type of data, the converse is not
-generally true.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Process of creating principles}
+\label{sec:process}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 The FORCE11 Software Citation Working Group~\cite{f11scwg} was created in April
@@
 
+The software citation principles were originally based on an adaptation of the
+FORCE11 Data Citation Principles~\cite{data-citation-principles}, and
+modified based on discussions of the FORCE11 Software Citation Working
+Group (see Appendix~\ref{app:wg_members} for members), information from the use
+cases in \S\ref{sec:use_cases}, and the related work in \S\ref{sec:related_work}.
+The adaptations were made because software, while similar to data in terms of
+not traditionally having been cited in publications, is also different than
+data. Software can be used to express or explain concepts, it is typically
+updated more frequently than data\footnote{Datasets can of course be frequently
+updated, but each update likely represents the findings of a new or repeated
+study.}, and it is executable.
+\niemnote{Dan: can you add more here?}
+Also, while software can be considered a type of data, the converse is not
+generally true.
+
 The principles in this document should guide further development of software citation
@@
 
-We have documented and analyzed a set of use cases related to software citation
-in \cite{SC-Use-Cases}. Table~\ref{tab:use_cases} summarizes these use cases and
+We documented and analyzed a set of use cases related to software citation
+in \cite{SC-Use-Cases} (recorded in Appendix~\ref{app:usecases} for completeness).
+Table~\ref{tab:use_cases} summarizes these use cases and
 makes clear what the requirements are for software citation in each case.
-Each example represents a particular stakeholder performing an activity related to citing software, with the given metadata as information needed to do that.
-In that table,
-we use the following definitions:
+Each example represents a particular stakeholder performing an activity
+related to citing software, with the given metadata as information needed to
+do that. In that table, we use the following definitions:
 \begin{itemize}

### 2016-07-21 sc-principles/software-citation-principles.tex
edits to improve structure and flow

@@
 
-The group gathered members (see Appendix~\ref{app:wg_members}) in April and May 2015,
-and then began work in June, with a number of meetings
-and some off-line work by group members to gather materials documenting existing
-practices in member disciplines; gather materials from workshops and other reports;
-review those materials, identifying overlaps and differences; and subsequently draft an initial version of this document.
-The draft Software Citation Principles document was discussed in a day-long workshop and presented at the FORCE2016 Conference~\cite{force2016} in April 2016.
-Members of the workshop and greater FORCE11 community gave feedback, which we recorded here in Appendix~\ref{app:force16-feedback}.
-This discussion led to some changes in the use cases and discussion, although the principles themselves were not modified.
-We also plan to initiate a follow-on working group that will work with stakeholders to ensure that these principles impact the research process.
-
-The software citation principles were originally based on an adaptation of the
-FORCE11 Data Citation Principles~\cite{data-citation-principles}, and
-modified based on discussions of the FORCE11 Software Citation Working
-Group (see Appendix~\ref{app:wg_members} for members), information from the use
-cases in \S\ref{sec:use_cases}, and the related work in \S\ref{sec:related_work}.
+The group gathered members (see Appendix~\ref{app:wg_members}) in April and
+May 2015, and then began work in June. This materialize as a number of meetings
+and offline work by group members to document existing
+practices in member disciplines; gather materials from workshops and other
+reports; review those materials, identifying overlaps and differences;
+create a list of use cases related to software citation, recorded in
+Appendix~\ref{app:usecases}; and subsequently draft an initial version of this document.
+The draft Software Citation Principles document was discussed in a day-long
+workshop and presented at the FORCE2016 Conference~\cite{force2016} in April 2016.
+Members of the workshop and greater FORCE11 community gave feedback, which we
+recorded here in Appendix~\ref{app:force16-feedback}. This discussion led to
+some changes in the use cases and discussion, although the principles themselves
+were not modified. We also plan to initiate a follow-on working group that will
+work with stakeholders to ensure that these principles impact the research process.
+
+The process of creating the software citation principles began by adapting
+the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
+were then modified based on discussions of the FORCE11 Software Citation
+Working Group (see Appendix~\ref{app:wg_members} for members), information from
+the use cases in \S\ref{sec:use_cases}, and the related work in
+\S\ref{sec:related_work}.
 The adaptations were made because software, while similar to data in terms of

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
draft response to R1 #1

@@
 \niemnote{This will be addressed somewhat by the changes discussed above; but Dan, could you help with this?}
+We view the principles presented at the beginning as the main outcomes\slash results of this work, with the process of producing them (discussed in the new Section 3) based on a synthesis of information from the use cases and related work.
+Then, Section 6 (previously 5) discusses the ``results''.
+Our modifications to the structure of the paper, and changes we hope improve the logical flow, should help indicate these connections better.
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
finalized R1 #3 response

@@
 \textbf{Our response:}
-\asnote{The reviewer's suggested text sounds pretty good to me. Perhaps drop/change 'scientific' in that sentence?}
-
-\niemnote{Agreed on both counts}
-
-\niemnote{Let's say that}
 We agree that this paper is more of a position paper than research article, and thus that creating a research question would be an artificial construct.
 We are willing to make this change if required, but feel it wouldn't add much to the paper.
-\katznote{and maybe add}
-We have modeled our process on the work done by the FORCE2016 Data Citation group, which included publishing a PeerJ CS paper (\url{https://peerj.com/articles/cs-1/}) which has been written in a similar manner to our paper.
+
+We modeled our process on the work done by the FORCE2016 Data Citation group, which included publishing a \emph{PeerJ CS} paper (\url{https://peerj.com/articles/cs-1/}) which was written in a similar manner to our paper.
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
response for R1 #4

@@
 \textbf{Our response:}
-\niemnote{I will add the full use cases text as an appendix}
+We agree, and have added the full use cases text (with some edits for clarity and completion) as the new Appendix B for completeness.
+We also added the comments (with our responses) made publicly following the presentation of the draft principles document at the FORCE2016 confernece, as Appendix C.
 

### 2016-07-21 sc-principles/software-citation-principles.tex
added indexed citations as + to use case 5

@@
 4.\ Determine use\slash citations of software                   & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Researcher, software engineer \\
-5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & + &             & \textbullet & \textbullet &             &             &             &             & Researcher, software engineer \\
+5.\ Get credit for software development          & \textbullet & \textbullet & \textbullet & + &             & \textbullet & \textbullet & + &             &             &             & Researcher, software engineer \\
 6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher \\

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
completed response to R2 #8

@@
 \begin{enumerate}[label={(\alph*)}]
-\item Agreed for 1 and 2.
-\item \niemnote{I think I agree, or at least benefit}
-\item \niemnote{I think what we had originally makes sense, but perhaps there needs to be more explanation?}
-\item \niemnote{no, indexed citations not really relevant for catalog/registry, unless it wants to display citations of software---but that is use case 4}
-\item \niemnote{no, I don't think those things would be stored---because they aren't included in the citation itself}
+\item We agree for use cases 1 and 2, and have added ``Description'' as beneficial metadata.
+However, we feel that a description would not really be necessary for use case 9 from the perspective of a publisher, and that a citation manager would not really need or benefit much from the description.
+\item We agree that use case 5 could benefit from indexed citations, and added that as beneficial metadata.
+\item We feel that the metadata omitted would not be necessary from the perspective of the publisher, the stakeholder for this use case.
+\item We feel that indexed citations are not particularly relevant for a catalog or registry (based on input from adminstrators of such entities), unless they want to display citations of the software---but this would be use case 4.
+Thus, we prefer to leave this as is.
+\item We agree that indexed citations are not necessary for use case 16, and have removed this.
+We also agree that the use case could benefit from description and keywords, in the same way that citation managers likely contain an abstract and keywords for papers.
 \item We intended ``find'' to mean more than discovery, but since the appendix was not present in the first submission this was not clear.
@@
 
-\katznote{also as a general comment, see the new intro to appendix B that states that some of this may be incomplete or wrong, but is included as a record to support the process that come up with the body of the document, and that Appendix B will be subject to further work in the future.}
+Also, we have added the full use case discussion material as a new Appendix B; some of that content may be incomplete or wrong, but is included as a record to support the process that lead to the use cases material in the document body.
+The use cases will also be subject to further work by the follow-on implementation group.
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
suggested response for R2 #11

@@
 \textbf{Our response:}
+We have rephrased this statement to clarify our meaning.
 
-\niemnote{The table collects use cases that relate to or depend on software citation, but certainly each use case has additional dependencies that are not and cannot be contained in the table.
-So, perhaps we can rephrase somewhere to emphasize that the table requirements are only for the citation of software related to each use case, and not meant to be exhaustive about everything actually required for the use case beyond citation?}
-The requirements in the use cases table only indicate the metadata needed for citation related to each use case, and not necessarily everything needed to complete each use case.
-For example, use case 7 ``Benchmark software'' does not include information about actual benchmarking, or the need for similar hardware, etc.
+The requirements given in the use cases table only indicate the metadata needed for citation of software for each use case, and not necessarily all the requirements needed to actually complete each use case.
+For example, use case 7 ``Benchmark software'' does not include information about performing actual benchmarking, or the need for similar hardware, etc.
+\niemnote{how does this sound, in combination with the suggested rephrasing?}
 

### 2016-07-21 sc-principles/software-citation-principles.tex
removed indexed citations from use case 16
added keyword & description as + to use case 16
added suggested rephrasing
adding intro comment

@@
 \niemnote{Do we want a better introduction, to help with the structure of the paper?}
+\katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}
 
@@
 15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & + &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
-16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Citation manager \\
+16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & + & + & Citation manager \\
 17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + & + & + & Repository, library, archive \\
@@
 Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
+\niemnote{Perhaps rephrase to ``Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.''}
 %\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}

### 2016-07-21 sc-principles/software-citation-principles.tex
adding more about software/data differences

@@
 not traditionally having been cited in publications, is also different than
-data. Software can be used to express or explain concepts, it is typically
-updated more frequently than data\footnote{Datasets can of course be frequently
-updated, but each update likely represents the findings of a new or repeated
-study.}, and it is executable.
-\niemnote{Dan: can you add more here?}
-Also, while software can be considered a type of data, the converse is not
-generally true.
+data. While software can be considered a type of data, the converse is not
+generally true.  In addition:
+
+\begin{itemize}
+\item Software can be used to express or explain concepts, unlike data.
+\item Software is updated more frequently that papers or data\footnote{Datasets can of course be frequently updated, but each update likely represents the findings of a new or repeated study.}.
+\item Software is executable, unlike papers or data.
+\item Software suffers from a different type of bit rot than data --- software must be constantly maintained so that it continues to function as both the hardware and software environments that it is used on changes, as bugs are found and fixed, and as user requirements demand new features and capabilities.
+\item Software is frequently built to use other software, leading to complex dependencies, and these dependent software packages also are frequently changing.
+\item Software is generally smaller than data, so a number of the storage and preservation constraints on data don't apply to software.
+\item Software teams can be very large, and very multidisciplinary, and there can be many varied roles that are overlapping and sometimes ambiguous with respect to contributions, and which contributions rise to a level equivalent to paper authorship.
+\item The lifetime of software is generally not as long as that of data.
+\end{itemize}
 

### 2016-07-21 sc-principles/software-citation-principles.tex
accepting Kyle's suggestion in 6.1

@@
 In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification.
-Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
-\niemnote{Perhaps rephrase to ``Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.''}
+%Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
+%\niemnote{Perhaps rephrase to ``
+Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
+%''}
 %\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}

### 2016-07-21 sc-principles/software-citation-principles.tex
adding description for B.6 and in Google Doc

@@
 
-\niemnote{add description?}
+When a researcher wants to understand or verify a research results from another researcher, they
+would like to use the same software.  Note that accessing the exact same software is necessary but not sufficient for reproducibility.
 
@@
 \item DOI or other persistent handle for that specific release
-\item Release has all components necessary for reproducing the work (note that this ideally also means sample inputs and outputs)
+\item Release has all components necessary for reproducing the work (Note: this ideally also means sample inputs and outputs)
 \end{itemize}

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
adding response to reviewer 2 to data/software differences item

@@
 
-\niemnote{Dan: work on this. Bring in text from CodeMeta paper and your blog post.}
+\niemnote{Dan: work on this. Bring in text from CodeMeta paper and your blog post.} \katznote{done - and see suggested response below}
+
+We respectfully disagree with the reviewer about the value of this section.
+Since our work on software citation is being done following previous work on data citation (including the PeerJ CS paper on data citation (\url{https://peerj.com/articles/cs-1/}), and since our software citation principles have some difference from the data citation principles, we feel it is important to explain the high level reasons why differences might be expected.
+We do agree with the reviewer's specific point about release frequency, and have added a note about this.
+We have also expanded the set of differences to make the case more convincing.
 

### 2016-07-21 sc-principles/software-citation-principles.tex
fixing bullet in table to match text in B.7 Adding description in B.7 to match what was already in the Google Doc Suggesting that we merge B.7 with B.1, as suggested in Google Doc
changing text in 6.1 about citation vs provenance in response to reviewer 2 point 10, and changing the rebuttal to match

@@
 6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher \\
-7.\ Benchmark software                           & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher, software engineer \\
+7.\ Benchmark software                           & \textbullet & \textbullet &  \textbullet  &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher, software engineer \\
 8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & + & + & + & Researcher, software engineer \\
@@
 Note that some software which is or could be captured as part of data provenance may not be cited.
-Citation is a record of software that is important to a research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
-This implies that for a data research product, provenance data will include all cited software, but not necessarily vice versa.
-Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.
+Citation is partly a record of software that is important to a research outcome\footnote{Citation can be used for many purposes, including for software: which software has been used in the work, which software has influenced the work, which software is the work superseding, which software is the work disproving, etc.}, where provenance is a record of all steps (including software) used to generated particular data within the research process.
+This implies that for a data research product, provenance data will include some of the cited software.
+Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation for the software that was used in the work.
 The data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.
@@
 
-\niemnote{add description?}
+This case describes the need for a research who has contributed to software (by design, software engineering, development, testing, patching, documentation, training, evangelizing, etc.) to have their software work recognized by their employer or colleagues for the purpose of career advancement and increased professional reputation.
+
+\katznote{I think we should merge this with B.1, as suggested in the google doc - the bullets are all the same}
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
changing text in 6.1 about citation vs provenance in response to reviewer 2 point 10, and changing the rebuttal to match
small changes in response 11 to reviewer 2

@@
 
-\katznote{this is a good point - I probably need to change our text a bit}
+We agree with the reviewer.
+We had been thinking of citation only in the context of software that had been used, not the other uses of citation.
+We have changed the text to acknowledge these other uses of citation.
 
@@
 The requirements given in the use cases table only indicate the metadata needed for citation of software for each use case, and not necessarily all the requirements needed to actually complete each use case.
-For example, use case 7 ``Benchmark software'' does not include information about performing actual benchmarking, or the need for similar hardware, etc.
+For example, use case 7 ``Benchmark software'' does not include requirements for performing actual benchmarking, including hardware, performance monitoring software, performance modeling knowledge, etc.
 \niemnote{how does this sound, in combination with the suggested rephrasing?}
+\katznote{I've changed the text a little}
 

### 2016-07-21 sc-principles/software-citation-principles.tex
some wording edits in section 3 (software vs. data)

@@
 \S\ref{sec:related_work}.
-The adaptations were made because software, while similar to data in terms of
+We made the adaptations because software, while similar to data in terms of
 not traditionally having been cited in publications, is also different than
-data. While software can be considered a type of data, the converse is not
+data. Software can be considered a type of data, but the converse is not
 generally true. In addition:
@@
 \item Software can be used to express or explain concepts, unlike data.
-\item Software is updated more frequently that papers or data\footnote{Datasets can of course be frequently updated, but each update likely represents the findings of a new or repeated study.}.
+\item Software is updated more frequently that papers or data\footnote{Datasets
+can of course be frequently updated, but each update likely represents the
+findings of a new or repeated study.}.
 \item Software is executable, unlike papers or data.
-\item Software suffers from a different type of bit rot than data --- software must be constantly maintained so that it continues to function as both the hardware and software environments that it is used on changes, as bugs are found and fixed, and as user requirements demand new features and capabilities.
-\item Software is frequently built to use other software, leading to complex dependencies, and these dependent software packages also are frequently changing.
-\item Software is generally smaller than data, so a number of the storage and preservation constraints on data don't apply to software.
-\item Software teams can be very large, and very multidisciplinary, and there can be many varied roles that are overlapping and sometimes ambiguous with respect to contributions, and which contributions rise to a level equivalent to paper authorship.
+\item Software suffers from a different type of bit rot than data---software
+must be constantly maintained so that it continues to function as both the
+hardware and software environments that it is used on changes, as developers
+find and fix bugs, and as user requirements demand new features and capabilities.
+\item Software is frequently built to use other software, leading to complex
+dependencies, and these dependent software packages also frequently change.
+\item Software is generally smaller than data, so a number of the storage and
+preservation constraints on data don't apply to software.
+\item Software teams can be large and multidisciplinary, and with varied roles
+that overlap and sometimes make contributions ambiguous---some of which
+can rise to a level equivalent to paper authorship.
 \item The lifetime of software is generally not as long as that of data.
@@
 of software citation and see if it meets the principles.
-Please note that while we strive to offer practical guidelines that acknowledge the current incentive system of academic citation, a more modern system of assigning credit is sorely needed.
-It is not that academic software needs a separate system from academic papers, but that the need for credit for application software underscores the need to overhaul the system of credit for all research products.
+While we strive to offer practical guidelines that acknowledge the current
+incentive system of academic citation, a more modern system of assigning credit
+is sorely needed. It is not that academic software needs a separate credit
+system from that of academic papers, but that the need for credit for
+research software underscores the need to overhaul the system of credit for
+all research products.
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
finalized response to R2 #3

@@
 \textbf{Our response:}
+We do agree with the reviewer's specific point about release frequency, and have
+added a note about this. We have also expanded the set of differences to make
+the case more convincing.
 
-\niemnote{I think we all would agree that software is a research method and data are research findings, but I do think we need some explanation of the differences---if only to justify why software citation may be different than data citation}
-
-\niemnote{Dan: work on this. Bring in text from CodeMeta paper and your blog post.} \katznote{done - and see suggested response below}
-
-We respectfully disagree with the reviewer about the value of this section.
-Since our work on software citation is being done following previous work on data citation (including the PeerJ CS paper on data citation (\url{https://peerj.com/articles/cs-1/}), and since our software citation principles have some difference from the data citation principles, we feel it is important to explain the high level reasons why differences might be expected.
-We do agree with the reviewer's specific point about release frequency, and have added a note about this.
-We have also expanded the set of differences to make the case more convincing.
+However, we respectfully disagree with the reviewer about the value of this section.
+Since our work on software citation is being done following previous work on
+data citation (including the \emph{PeerJ CS} paper on data citation
+(\url{https://peerj.com/articles/cs-1/}), and since our software citation
+principles have some difference from the data citation principles, we feel it
+is important to explain the high level reasons why differences might be expected.
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
finalize response to R1 #1

@@
 \textbf{Our response:}
-
-\niemnote{This will be addressed somewhat by the changes discussed above; but Dan, could you help with this?}
 We view the principles presented at the beginning as the main outcomes\slash results of this work, with the process of producing them (discussed in the new Section 3) based on a synthesis of information from the use cases and related work.

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
finalize response to Editor #4

@@
 \textbf{Our response:}
-\niemnote{I will add the full use cases text as an appendix, and revise to match the numbering in the table}
-\niemnote{I will also add the FORCE16 comments and responses}
-
-\niemnote{Dan \& Arfon: could you assist with adding descriptions to some of the use cases that don't have one?} \katznote{In both the Google doc and the appendix I think?}
+We agree that adding the external use case material would help make the paper
+more complete, and have added it as a new Appendix B. In addition, we added the
+public\slash FORCE11 community feedback (and our responses) as a new Appendix C.
 

### 2016-07-21 sc-principles/rebuttal/rebuttal.tex
finalize response to Editor #2

@@
 \textbf{Our response:}
-\asnote{I think we just need to say that we thought it was important to draft the principles with the end-user(s) in mind, i.e.
-what different use cases/scenarios for software citation.
-The requirements table is an overview of these use cases and the metadata needs of each of these users/use cases.}
-
-\niemnote{Move paragraph at end of Section 1 to Section 3, and expand/merge with paragraph after mission statement; find text from FORCE2016 presentation about use cases and explaining produre to get principles}
-\niemnote{I will do this}
+We have made some changes that we hope improve the logical flow of the document,
+and also added a new section (3) that explicitly describes the process by which
+we developed the principles. This includes the role that the use cases played
+in this process.
 

### 2016-07-21 sc-principles/software-citation-principles.tex
focussed -> focused

@@
 Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation.
-Inspired by the activities of the FORCE11 working group focussed on data citation, this document summarizes the recommendations of the FORCE11 Software Citation Working Group and its activities between June 2015 and April 2016.
+Inspired by the activities of the FORCE11 working group focused on data citation, this document summarizes the recommendations of the FORCE11 Software Citation Working Group and its activities between June 2015 and April 2016.
 Based on a review of existing community practices, the goal of the working group was to produce a consolidated set of citation principles that may encourage broad adoption of a consistent policy for software citation across disciplines and venues.

### 2016-07-22 sc-principles/software-citation-principles.tex
adding a comment
adding response to comment

@@
 
-\niemnote{this seems the same as use case 1}
+\niemnote{this seems the same as use case 1} \katznote{I thought it was the same as 6}
+\niemnote{6 is about reproducing, while this seems to be writing a software paper}
 The research wants to publish about a version of software they have produced.

### 2016-07-23 sc-principles/software-citation-principles.tex
starting to try to improve intro to section 1
added some more text in response to R1 #9

@@
 \katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}
+
+\niemnote{How is this?}
+The main contribution of this document are the software citation principles,
+written fairly concisely in this section and discussed further later in the
+document (\S\ref{sec:discussion}).
 
@@
 
-Note that some software which is or could be captured as part of data provenance may not be cited.
-Citation is partly a record of software that is important to a research outcome\footnote{Citation can be used for many purposes, including for software: which software has been used in the work, which software has influenced the work, which software is the work superseding, which software is the work disproving, etc.}, where provenance is a record of all steps (including software) used to generated particular data within the research process.
-This implies that for a data research product, provenance data will include some of the cited software.
-Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation for the software that was used in the work.
-The data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.
+Some software which is, or could be, captured as part of data provenance may not be cited.
+Citation is partly a record of software important to a research outcome\footnote{Citation
+can be used for many purposes, including for software: which software has been used in
+the work, which software has influenced the work, which software is the work superseding,
+which software is the work disproving, etc.}, where provenance is a record of all steps
+(including software) used to generated particular data within the research process.
+This implies that for a data research product, provenance data will include some of
+the cited software.
+Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
+and thus may depend on the specific version
+used~\cite{Sandve:2013gh,Wilson:2014aa}---furthermore, errors in software or
+environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
+Similarly, the software metadata that is recorded as part of data provenance should be
+a superset of the metadata recorded as part of software citation for the software that
+was used in the work. The data recorded for reproducibility should also be a superset
+of the metadata recorded as part of software citation.
 These statements may also be true for software products.
-In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification.
-%Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
-%\niemnote{Perhaps rephrase to ``
+In general, we intend the software citation principles to cover the minimum of what is
+necessary for software citation for the purpose of software identification.
 Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
-%''}
 %\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}

### 2016-07-23 sc-principles/rebuttal/rebuttal.tex
finalized response to R2 #11
minor fix to response for R2 #10
edits and response to R1 #9

@@
 \textbf{Our response:}
-
-\niemnote{I think we have some citations about the importance of software to research we can add here (e.g., Hannay 2009), as well as the fact that research results depend on citation\slash proper version (Sufi 2014, Sandve 2013, Wilson 2014), and that errors in software or environment variations can effect results (Morin 2012, Soergel 2015).
-Although we probably want to avoid overlap with the challenge paper.} \katznote{basically, what is important is up the authors to define, and then reviewers to agree with, all in the context of community practices and `standards'}
-
-\niemnote{Add discussion of both points; Dan's note is already there in the text. I will work on this.}
+We have revised this discussion to better address these issues. In particular, we
+added some references that discuss the importance of specific software versions to
+research results (including data). For a particular study, it is up the authors to
+determine which software is important.
+\niemnote{how is this, including the new text?}
 
@@
 \textbf{Our response:}
-
 We agree with the reviewer.
@@
 For example, use case 7 ``Benchmark software'' does not include requirements for performing actual benchmarking, including hardware, performance monitoring software, performance modeling knowledge, etc.
-\niemnote{how does this sound, in combination with the suggested rephrasing?}
-\katznote{I've changed the text a little}
 

### 2016-07-23 sc-principles/software-citation-principles.tex
rearranged some text in 6.1
some edits at introduction

@@
 \katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}
-
 \niemnote{How is this?}
+
 The main contribution of this document are the software citation principles,
 written fairly concisely in this section and discussed further later in the
-document (\S\ref{sec:discussion}).
-
-The principles in this section are written fairly concisely, and discussed
-further later in this document (\S\ref{sec:discussion}). Here, for example,
-we do not define what software should be cited, but how it should be
-cited, and we talk about how such decisions might be made in the
-discussion section~(\S\ref{sec:discussion}).
+document (\S\ref{sec:discussion}). In addition, we also motivate the creation
+of these principles (\S\ref{sec:intro}), describe the process by which they
+were created (\S\ref{sec:process}), summarize use cases related to software
+citation (\S\ref{sec:use_cases}), and review related work
+(\S\ref{sec:related_work}). We also lay out the necessary future work
+(\S\ref{sec:futurework}).
+%The principles in this section are written fairly concisely, and discussed
+%further later in this document (\S\ref{sec:discussion}). Here, for example,
+%we do not define what software should be cited, but how it should be
+%cited, and we talk about how such decisions might be made in the
+%discussion section~(\S\ref{sec:discussion}).
 
@@
 (including software) used to generated particular data within the research process.
-This implies that for a data research product, provenance data will include some of
-the cited software.
 Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
@@
 environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
+This implies that for a data research product, provenance data will include some of
+the cited software.
 Similarly, the software metadata that is recorded as part of data provenance should be

### 2016-07-23 sc-principles/software-citation-principles.tex
minor wording edit

@@
 Some of the most actionable work has come from the UK Software Sustainability
-Institute (SSI) in the form of blog posts written by their community fellows:
-
-In a blog post from 2012, Jackson discusses some of the pitfalls of trying to
-cite software in publications~\cite{ssi-how-to-cite}. He includes useful
-guidance for when to consider citing software as well as some ways to help
+Institute (SSI) in the form of blog posts written by their community fellows.
+For examplel, in a blog post from 2012, Jackson discusses some of the pitfalls
+of trying to cite software in publications~\cite{ssi-how-to-cite}. He includes
+useful guidance for when to consider citing software as well as some ways to help
 ``convince'' journal editors to allow the inclusion of software citations.

### 2016-07-23 sc-principles/software-citation-principles.tex
comments on use case 7
edits to use case 13
added some detail to use case 14

@@
 \katznote{I think we should merge this with B.1, as suggested in the google doc - the bullets are all the same}
+\niemnote{the title and description are quite different---based on the description, this actually seems like 5}
 
@@
 Both would help repository know how to direct resources, e.g., maintenance, training etc.
+This is similar to the case of a funder wanting to know the impact of funded work.
 
@@
 \begin{itemize}
-\item Code name ideally, a unique identifier
+\item Code name, or a unique identifier
 \item Relationships to previous versions
@@
 
-similar to Funder case
-
-mineable via repository name, code name, DOI, or citation.\niemnote{complete?}
-
 \subsection{Funder who wants to know how software they funded has been used}
+
+This use case is similar to ``Repository showing scientific impact of holdings'',
+where a funder wants to find out the use and impact and software that they
+supported.
 

### 2016-07-23 sc-principles/software-citation-principles.tex
added description for use case 15

@@
 
-\niemnote{add description}
+In this use case, an evaluator (e.g., academic administrator) or funder wants
+to evaluate the contributions of a researcher who develops software. This case
+is related to those where researchers want to get credit for software
+development, or where organizations want to evaluate the impact of software
+itself.
 

### 2016-07-23 sc-principles/rebuttal/rebuttal.tex
change in response to Kyle's work on reviewer 1 item 9

@@
 research results (including data). For a particular study, it is up the authors to
-determine which software is important.
-\niemnote{how is this, including the new text?}
+determine which software is important, and peer-reviewers to decide if the authors have
+done this sufficiently well, according to community standards, similarly to how citation
+of papers in a paper is decided and judged.
+\niemnote{how is this, including the new text?} \katznote{I've added a bit to the response, and lightly edited the text.
+I also made a change in the text tied to reviewer 2 item 10 that I missed previously}
 

### 2016-07-23 sc-principles/software-citation-principles.tex
change in response to Kyle's work on reviewer 1 item 9

@@
 and thus may depend on the specific version
-used~\cite{Sandve:2013gh,Wilson:2014aa}---furthermore, errors in software or
+used~\cite{Sandve:2013gh,Wilson:2014aa}. Furthermore, errors in software or
 environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
@@
 Similarly, the software metadata that is recorded as part of data provenance should be
-a superset of the metadata recorded as part of software citation for the software that
-was used in the work. The data recorded for reproducibility should also be a superset
-of the metadata recorded as part of software citation.
+will overlap the metadata recorded as part of software citation for the software that
+was used in the work. The data recorded for reproducibility should also overlap
+the metadata recorded as part of software citation.
 These statements may also be true for software products.

### 2016-07-23 sc-principles/software-citation-principles.tex
light edit to sentence Dan edited, and finalized response to R1 #9

@@
 the cited software.
-Similarly, the software metadata that is recorded as part of data provenance should be
-will overlap the metadata recorded as part of software citation for the software that
+Similarly, the software metadata recorded as part of data provenance will
+overlap the metadata recorded as part of software citation for the software that
 was used in the work. The data recorded for reproducibility should also overlap
 the metadata recorded as part of software citation.
-These statements may also be true for software products.
+These statements may also be true for software products.\niemnote{Does the previous sentence still make sense?}
 In general, we intend the software citation principles to cover the minimum of what is

### 2016-07-24 sc-principles/software-citation-principles.tex
accepting change in intro, removing sentence that Kyle questioned in 6

--- 2016-07-23 sc-principles/software-citation-principles.tex
@@
 
-\niemnote{Do we want a better introduction, to help with the structure of the paper?}
-\katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}
-\niemnote{How is this?}
+%\niemnote{Do we want a better introduction, to help with the structure of the paper?}
+%\katznote{We need to keep this short, I think.  I'm not sure what else we can easily do.}
+%\niemnote{How is this?} \katznote{good}
 
@@
 the cited software.
-Similarly, the software metadata that is recorded as part of data provenance should be
-will overlap the metadata recorded as part of software citation for the software that
+Similarly, the software metadata recorded as part of data provenance will
+overlap the metadata recorded as part of software citation for the software that
 was used in the work. The data recorded for reproducibility should also overlap
 the metadata recorded as part of software citation.
-These statements may also be true for software products.
 In general, we intend the software citation principles to cover the minimum of what is

### 2016-07-24 sc-principles/software-citation-principles.tex
comment in B.7

@@
 
-This case describes the need for a research who has contributed to software (by design, software engineering, development, testing, patching, documentation, training, evangelizing, etc.)
+This case describes the need for a researcher who has contributed to software (by design, software engineering, development, testing, patching, documentation, training, evangelizing, etc.)
 to have their software work recognized by their employer or colleagues for the purpose of career advancement and increased professional reputation.
@@
 
-This case describes the need for a research who has contributed to software (by design, software engineering, development, testing, patching, documentation, training, evangelizing, etc.) to have their software work recognized by their employer or colleagues for the purpose of career advancement and increased professional reputation.
+This case describes the need for a researcher who has contributed to software (by design, software engineering, development, testing, patching, documentation, training, evangelizing, etc.) to have their software work recognized by their employer or colleagues for the purpose of career advancement and increased professional reputation.
 
 \katznote{I think we should merge this with B.1, as suggested in the google doc - the bullets are all the same}
-\niemnote{the title and description are quite different---based on the description, this actually seems like 5}
+\niemnote{the title and description are quite different---based on the description, this actually seems like 5} \katznote{somehow, the description is the same was copied from 5's - not sure why/how - but I don't think this matches the title at all.
+The title matches 1.
+It's a case of a researcher using someone else's software for a paper.
+Unlike 1, which is general in terms of the use, likely a new science results, here the use of a benchmarking study, but again, the outcome is a paper that needs to cite the software.
+I'm fairly confident we can delete this one entirely, as it is really subsumed within 1.
+If we do this, we should do it in the google doc too.}
 

### 2016-07-24 sc-principles/software-citation-principles.tex
copying desription for B.9 from the Google Doc
adding a little to B.14 and declaring it sufficiently done
adding description to B.16

@@
 
-\niemnote{add description}
+This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
+
+Requirements
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in JATS, for example, as well as references in the text itself
+\end{itemize}
 
@@
 where a funder wants to find out the use and impact and software that they
-supported.
-
-steps/requirements\niemnote{add more to this?}
+supported.  It is also similar to ``Researcher who wants to know who uses the researcher's software.''
 
@@
 
-\niemnote{add description?}
+Reference management systems may need to be updated to internally understand that their is a software reference type, and to be able to output references to software in common formats.
 

### 2016-07-25 sc-principles/software-citation-principles.tex
added acknowledgements per journal instructions

@@
 
+\section*{Acknowledgements}
+
+Work by Daniel S.~Katz was supported by the National Science Foundation (NSF) while working at the Foundation.
+Any opinion, finding, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the NSF.
+Work by Kyle E.~Niemeyer was supported in part by the NSF under grant ACI-1535065.
+
 \appendix

### 2016-07-25 sc-principles/software-citation-principles.tex
moved benchmark use case to those not adopted
also removed from the table

@@
 6.\ ``Reproduce'' analysis                       & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher \\
-7.\ Benchmark software                           & \textbullet & \textbullet &  \textbullet  &             & \textbullet & \textbullet & \textbullet &             & + & + &             & Researcher, software engineer \\
-8.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & + & + & + & Researcher, software engineer \\
-9.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             &             &             & Publisher \\
-10.\ Publish papers that cite software           & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Publisher \\
-11.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & + & + & + & Indexer \\
-12.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & + & + & Domain group, library, archive \\
-13.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Repository \\
-14.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder, policy maker \\
-15.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & + &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
-16.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & + & + & Citation manager \\
-17.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + & + & + & Repository, library, archive \\
+7.\ Find software to implement task              & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet & \textbullet & + & + & + & Researcher, software engineer \\
+8.\ Publish software paper                       & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             &             &             & Publisher \\
+9.\ Publish papers that cite software           & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             &             & Publisher \\
+10.\ Build catalog of software                   & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet & + & + & + & Indexer \\
+11.\ Build software catalog\slash registry       & \textbullet & \textbullet & \textbullet &             &             &             & \textbullet &             &             & + & + & Domain group, library, archive \\
+12.\ Show scientific impact of holdings          & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Repository \\
+13.\ Show how funded software has been used      & \textbullet & \textbullet &             &             &             &             &             & \textbullet &             &             &             & Funder, policy maker \\
+14.\ Evaluate contributions of researcher        & \textbullet &             & \textbullet & + &             & \textbullet &             & \textbullet &             &             &             & Evaluator, funder \\
+15.\ Store software entry                        & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             &             & + & + & Citation manager \\
+16.\ Publish mixed data\slash software packages  & \textbullet & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & + & + & + & Repository, library, archive \\
 \bottomrule
@@
 
-\subsection{Researcher who benchmarks someone else's software with or without modification on one or many hardware platforms for publication}
+\subsection{Researcher who wants to find a piece of software to implement a task}
+
+This is the case where a research is looking for software to use but wants to understand whether it is being used in a scholarly fashion.
+For example, a researcher searches through a software repository and finds a package that might be useful.
+They look to find whether it has been used by others in the scientific literature.
+
+Requirements
+\begin{itemize}
+\item Either the software documentation page has a reference to existing literature that makes use of it.
+\item There is a mechanism to look it up.
+\end{itemize}
+
+\subsection{Publisher wants to publish a software paper}
+
+This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
+
+Requirements
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in JATS, for example, as well as references in the text itself
+\end{itemize}
+
+\subsection{Publisher who wants to publish papers that cite software}
+
+This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
+
+Requirements for publisher:
+\begin{itemize}
+\item Name of software
+\item Names of software authors/contributors
+\item Location/repository
+\item Citable DOI of software
+\item Format for citing software in, e.g., JATS, as well as references in the text itself
+\end{itemize}
+
+\subsection{Indexer (e.g., Scopus, WoS, Scholar, MS Academic Search) who wants to build a catalog of software}
+
+Provide an index over the software that is used within the research domain. Track how that software is being used by different groups of researchers and to what ends.
+
+Requirements:
+\begin{itemize}
+\item Uniquely identify pieces of software used by the research literature
+\item Connect authors and organizations to that software
+\item Connect various software versions together
+\end{itemize}
+
+\subsection{Domain group (e.g., ASCL, bioCADDIE), Libraries, and Archives (e.g., University library, laboratory archive, etc.) wants to build a catalog\slash registry of institutional or domain software}
+
+There are two different examples here: One is building a catalog/archive of software produced by those affiliated with the institution.
+The other is along the lines of Sayeed Choudhury's note that ``data are the new special collections.''
+An institution may choose to build a catalog/archive of many things within a single topic or subject in order to secure all the software on a certain topic or build a collection that may draw users to their establishment, much like special collections now do for university libraries and archives.
+
+\subsection{Repository showing scientific impact of holdings}
+
+A repository that archives and/or maintains a collection of software.
+The repository would like to address usage and impact of software in its holding.
+Usage would aid potential users whether the software is being actively maintained or developed or has been superseded.
+Both would help repository know how to direct resources, e.g., maintenance, training etc.
+This is similar to the case of a funder wanting to know the impact of funded work.
+
+Requirements:
+\begin{itemize}
+\item Code name, or a unique identifier
+\item Relationships to previous versions
+\item Connect to repository
+\item Connect to research
+\end{itemize}
+
+\subsection{Funder who wants to know how software they funded has been used}
+
+This use case is similar to ``Repository showing scientific impact of holdings'',
+where a funder wants to find out the use and impact and software that they
+supported.  It is also similar to ``Researcher who wants to know who uses the researcher's software.''
+
+\subsection{Evaluator or funder wants to evaluate contributions of a researcher}
+
+In this use case, an evaluator (e.g., academic administrator) or funder wants
+to evaluate the contributions of a researcher who develops software. This case
+is related to those where researchers want to get credit for software
+development, or where organizations want to evaluate the impact of software
+itself.
+
+\subsection{Reference management system used by researchers to author a manuscript}
+
+Reference management systems may need to be updated to internally understand that their is a software reference type, and to be able to output references to software in common formats.
+
+Requirements for reference manager:
+\begin{itemize}
+\item Names of software authors/contributors
+\item Software version number and release date
+\item Location/repository
+\item Citable DOI of software or paper recommended for citation
+\item Format for citing software in citation metadata file
+\item Citation metadata tags embedded in DOI landing page/software project page for easy ingest
+\end{itemize}
+
+Possible steps:
+\begin{enumerate}
+\item Reference management system such as EndNote, Mendeley, Zotero, etc. builds affordances for software references.
+\item Researcher finds software citation and adds it to their reference manager library, by (a) importing from the CITATION file (e.g., BibTeX, RIS), or (b) clicking on, e.g., an ``add to Zotero library'' widget in web browser.
+\item Researcher writes a paper and uses the reference manager to generate citations or bibliography.
+\end{enumerate}
+
+\subsection{Repository wants to publish mixed data\slash software packages}
+
+\niemnote{add description; came from \url{https://github.com/force11/force11-scwg/issues/128}}
+Domain and institutional data repositories have both data and software
+artifacts, and want to link these together in a provenance trace that can
+be cited. Sometimes the software is a separately identified artifact, but at
+other times software is included inside of data packages, and the researcher
+wants to cite the combined product.
+
+\subsection*{Use cases not adopted in the table:}
+
+\subsection*{Researcher who benchmarks someone else's software with or without modification on one or many hardware platforms for publication}
 
@@
 
-\subsection{Researcher who wants to find a piece of software to implement a task}
-
-This is the case where a research is looking for software to use but wants to understand whether it is being used in a scholarly fashion.
-For example, a researcher searches through a software repository and finds a package that might be useful.
-They look to find whether it has been used by others in the scientific literature.
-
-Requirements
-\begin{itemize}
-\item Either the software documentation page has a reference to existing literature that makes use of it.
-\item There is a mechanism to look it up.
-\end{itemize}
-
-\subsection{Publisher wants to publish a software paper}
-
-This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
-
-Requirements
-\begin{itemize}
-\item Name of software
-\item Names of software authors/contributors
-\item Location/repository
-\item Citable DOI of software
-\item Format for citing software in JATS, for example, as well as references in the text itself
-\end{itemize}
-
-\subsection{Publisher who wants to publish papers that cite software}
-
-This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
-
-Requirements for publisher:
-\begin{itemize}
-\item Name of software
-\item Names of software authors/contributors
-\item Location/repository
-\item Citable DOI of software
-\item Format for citing software in, e.g., JATS, as well as references in the text itself
-\end{itemize}
-
-\subsection{Indexer (e.g., Scopus, WoS, Scholar, MS Academic Search) who wants to build a catalog of software}
-
-Provide an index over the software that is used within the research domain. Track how that software is being used by different groups of researchers and to what ends.
-
-Requirements:
-\begin{itemize}
-\item Uniquely identify pieces of software used by the research literature
-\item Connect authors and organizations to that software
-\item Connect various software versions together
-\end{itemize}
-
-\subsection{Domain group (e.g., ASCL, bioCADDIE), Libraries, and Archives (e.g., University library, laboratory archive, etc.) wants to build a catalog\slash registry of institutional or domain software}
-
-There are two different examples here: One is building a catalog/archive of software produced by those affiliated with the institution.
-The other is along the lines of Sayeed Choudhury's note that ``data are the new special collections.''
-An institution may choose to build a catalog/archive of many things within a single topic or subject in order to secure all the software on a certain topic or build a collection that may draw users to their establishment, much like special collections now do for university libraries and archives.
-
-\subsection{Repository showing scientific impact of holdings}
-
-A repository that archives and/or maintains a collection of software.
-The repository would like to address usage and impact of software in its holding.
-Usage would aid potential users whether the software is being actively maintained or developed or has been superseded.
-Both would help repository know how to direct resources, e.g., maintenance, training etc.
-This is similar to the case of a funder wanting to know the impact of funded work.
-
-Requirements:
-\begin{itemize}
-\item Code name, or a unique identifier
-\item Relationships to previous versions
-\item Connect to repository
-\item Connect to research
-\end{itemize}
-
-\subsection{Funder who wants to know how software they funded has been used}
-
-This use case is similar to ``Repository showing scientific impact of holdings'',
-where a funder wants to find out the use and impact and software that they
-supported.  It is also similar to ``Researcher who wants to know who uses the researcher's software.''
-
-\subsection{Evaluator or funder wants to evaluate contributions of a researcher}
-
-In this use case, an evaluator (e.g., academic administrator) or funder wants
-to evaluate the contributions of a researcher who develops software. This case
-is related to those where researchers want to get credit for software
-development, or where organizations want to evaluate the impact of software
-itself.
-
-\subsection{Reference management system used by researchers to author a manuscript}
-
-Reference management systems may need to be updated to internally understand that their is a software reference type, and to be able to output references to software in common formats.
-
-Requirements for reference manager:
-\begin{itemize}
-\item Names of software authors/contributors
-\item Software version number and release date
-\item Location/repository
-\item Citable DOI of software or paper recommended for citation
-\item Format for citing software in citation metadata file
-\item Citation metadata tags embedded in DOI landing page/software project page for easy ingest
-\end{itemize}
-
-Possible steps:
-\begin{enumerate}
-\item Reference management system such as EndNote, Mendeley, Zotero, etc. builds affordances for software references.
-\item Researcher finds software citation and adds it to their reference manager library, by (a) importing from the CITATION file (e.g., BibTeX, RIS), or (b) clicking on, e.g., an ``add to Zotero library'' widget in web browser.
-\item Researcher writes a paper and uses the reference manager to generate citations or bibliography.
-\end{enumerate}
-
-\subsection{Repository wants to publish mixed data\slash software packages}
-
-\niemnote{add description; came from \url{https://github.com/force11/force11-scwg/issues/128}}
-Domain and institutional data repositories have both data and software
-artifacts, and want to link these together in a provenance trace that can
-be cited. Sometimes the software is a separately identified artifact, but at
-other times software is included inside of data packages, and the researcher
-wants to cite the combined product.
-
-\subsection*{Use cases not adopted in the table:}
-
-\subsection{Researcher who wants to publish about a piece of software}
+\subsection*{Researcher who wants to publish about a piece of software}
 
@@
 
-\subsection{Researcher wants to record the software that generated some data}
+\subsection*{Researcher wants to record the software that generated some data}
 
@@
 
-\subsection{Researcher who wants to reproduce experience of use of a particular software implementation in context}
+\subsection*{Researcher who wants to reproduce experience of use of a particular software implementation in context}
 

### 2016-07-25 sc-principles/software-citation-principles.tex
added text explaining decision on benchmarking use case

@@
 
-\katznote{I think we should merge this with B.1, as suggested in the google doc - the bullets are all the same}
-\niemnote{the title and description are quite different---based on the description, this actually seems like 5} \katznote{somehow, the description is the same was copied from 5's - not sure why/how - but I don't think this matches the title at all.
-The title matches 1.
-It's a case of a researcher using someone else's software for a paper.
-Unlike 1, which is general in terms of the use, likely a new science results, here the use of a benchmarking study, but again, the outcome is a paper that needs to cite the software.
-I'm fairly confident we can delete this one entirely, as it is really subsumed within 1.
-If we do this, we should do it in the google doc too.}
-
 Requirements for researcher:
@@
 
+After review of this use case, we decided that based on the title this falls under
+use case 1, where a researcher uses someone else's software for a paper.
+Unlike use case 1, which is general in terms of the use of software, here the
+use leads to a benchmarking study---but the outcome in both cases is a paper
+that needs to cite the software.
+
 \subsection*{Researcher who wants to publish about a piece of software}

### 2016-07-25 sc-principles/software-citation-principles.tex
updated description for use case 8
finishing up use case about reproducing experience

@@
 
-This case asks what information regarding software is needed for a publisher who wants to publish papers that cite that software.
+This case asks what information regarding software is needed for a publisher who wants to publish a paper describing that software.
 
@@
 
-\niemnote{This one was not adopted into table; should we keep it, remote here, or somehow indicate why it was not adopted?}
-
 Researcher is engaged in historical/cultural research -- e.g., study of video games as cultural artifacts.
 
-Requirements
+Requirements:
 \begin{itemize}
@@
 
-Possible steps
+Possible steps:
 \begin{enumerate}
@@
 \end{enumerate}
+
+Due to overlaps with use case 6, reproducing analysis, we decided not to include
+this use case in the table.
 

### 2016-07-25 sc-principles/software-citation-principles.tex
removed note from use case 16

@@
 
-\niemnote{add description; came from \url{https://github.com/force11/force11-scwg/issues/128}}
 Domain and institutional data repositories have both data and software

### 2016-07-25 sc-principles/rebuttal/rebuttal.tex
light edit to sentence Dan edited, and finalized response to R1 #9
note on R1 #2 response

@@
 \niemnote{Dan, can you do this?} \katznote{I was hoping Arfon might do this one}
+\niemnote{I think we have something along these lines now at the end of section 3, although perhaps could say more.}
 
@@
 of papers in a paper is decided and judged.
-\niemnote{how is this, including the new text?} \katznote{I've added a bit to the response, and lightly edited the text.
-I also made a change in the text tied to reviewer 2 item 10 that I missed previously}
 

### 2016-07-25 sc-principles/software-citation-principles.tex
slight ack update
adding comments to some use cases, filling out description and steps for B.3

@@
 
-Work by Daniel S.~Katz was supported by the National Science Foundation (NSF) while working at the Foundation.
+Work by Daniel S.~Katz was supported in part by the National Science Foundation (NSF) while working at the Foundation.
 Any opinion, finding, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the NSF.
@@
 
-maybe a variant of above\niemnote{add more to this?}
+A researcher wants to contribute to someone else's software in the manner in which their contributions will be accepted and recognized.
+
+Possible steps:
+\begin{enumerate}
+\item Researcher finds information about the software, and how contributors will be recognized
+\item Researcher possibly submit a Contributor License Agreement (CLA) or Copyright Assignment Agreement (CAA) to allow the contributed content to be distributed with the software being contributed to
+\item Researcher contributes to the software
+\item Software maintainers accept contribution, recognize researcher's contribution, and update the software metadata as appropriate
+\end{enumerate}
 
@@
 \niemnote{this seems the same as use case 1} \katznote{I thought it was the same as 6}
-\niemnote{6 is about reproducing, while this seems to be writing a software paper}
+\niemnote{6 is about reproducing, while this seems to be writing a software paper} \katznote{ok, I guess this is like 1, but in 1, the research is using someone else's software, and here, the researcher is using their own software.
+Maybe this doesn't matter so much.
+So should we just add a comment that this is the same as 1 other than the fact that the software developer and paper author are mostly the same here?}
 The research wants to publish about a version of software they have produced.
@@
 
-\niemnote{Is this distinct from 6? Or should it be a new one?}
+\niemnote{Is this distinct from 6? Or should it be a new one?} \katznote{I think this is the same as 1, other than in 1 the research output is a paper, while here, the output is a dataset.}
 This is the case where a researcher is using some software to perform an analysis, either of a physical sample or of data.

### 2016-07-25 sc-principles/software-citation-principles.tex
finishing up use case about publishing software

@@
 
-\niemnote{this seems the same as use case 1} \katznote{I thought it was the same as 6}
-\niemnote{6 is about reproducing, while this seems to be writing a software paper} \katznote{ok, I guess this is like 1, but in 1, the research is using someone else's software, and here, the researcher is using their own software.
-Maybe this doesn't matter so much.
-So should we just add a comment that this is the same as 1 other than the fact that the software developer and paper author are mostly the same here?}
 The research wants to publish about a version of software they have produced.
@@
 \end{itemize}
+
+This is similar to use case 1, other than the fact that the software developer(s) and
+paper author(s) will likely be the same person\slash people here.
 

### 2016-07-25 sc-principles/software-citation-principles.tex
finishing use case about generating data

@@
 
-\niemnote{Is this distinct from 6? Or should it be a new one?} \katznote{I think this is the same as 1, other than in 1 the research output is a paper, while here, the output is a dataset.}
 This is the case where a researcher is using some software to perform an analysis, either of a physical sample or of data.
@@
 The analysis, or the generated data, has information about the software used.
+
+This is also similar to use case 1, except in that case the research output is
+a paper, while here the output is a dataset.
 

### 2016-07-25 sc-principles/software-citation-principles.tex
editing wording for explanation of historical use case

@@
 
-Researcher is engaged in historical/cultural research -- e.g., study of video games as cultural artifacts.
+Researcher is engaged in historical/cultural research---e.g., study of video games as cultural artifacts.
 
@@
 
-Due to overlaps with use case 6, reproducing analysis, we decided not to include
-this use case in the table.
+This overlaps with use case 6 (reproducing analysis), and so we decided not to
+include this as a distinct use case.
 

### 2016-07-25 sc-principles/rebuttal/rebuttal.tex
finalizing response to R1 #5

@@
 \textbf{Our response:}
+We agree that the main findings of the paper are the software citation principles, and as described in our response to point 1 above we have updated the introduction and made some structural changes to make this clear.
 
-\niemnote{If we need to formulate the paper as a research paper, then describing the principles as main findings and use cases\slash discussion as the methods to arrive at them seems an interesting way to do it.}
-
-\niemnote{We agree that the main findings are the principles, and the use cases/discussion are the main to reach those.
-Don't agree that the principles are assumptions or preliminary, but reflect current knowledge and experience}
-
-\niemnote{I will work on this, but may need help}
+However, we respectfully disagree that the principles are preliminary, but that they refelct current knowledge and experience.
 

### 2016-07-26 sc-principles/software-citation-principles.tex
fixed quotation marks
further  editing wording for explanation of historical use case

--- 2016-07-25 sc-principles/software-citation-principles.tex
@@
 \item Researcher finds and uses software in the development of new software.
-\item Researcher identifies citation metadata file (e.g., ''CITATION'' file) associated with downloaded/installed software source code or in online repository/published location.
+\item Researcher identifies citation metadata file (e.g., ``CITATION'' file) associated with downloaded/installed software source code or in online repository/published location.
 CITATION file includes necessary citation metadata.
@@
 
-Researcher is engaged in historical/cultural research -- e.g., study of video games as cultural artifacts.
+Researcher is engaged in historical/cultural research --- e.g., study of video games as cultural artifacts.
 
@@
 
-Due to overlaps with use case 6, reproducing analysis, we decided not to include
-this use case in the table.
+This overlaps use case 6 (reproducing analysis), and so we decided not to
+include this as a distinct use case.
 

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex
added note on long-term vision question
update to change in Reviewer 1 response, fixing a spelling error, changing a --- to a new sentence

@@
 We agree, and have added the full use cases text (with some edits for clarity and completion) as the new Appendix B for completeness.
-We also added the comments (with our responses) made publicly following the presentation of the draft principles document at the FORCE2016 confernece, as Appendix C.
+We also added the comments (with our responses) made publicly following the presentation of the draft principles document at the FORCE2016 conference, as Appendix C.
 
@@
 \textbf{Our response:}
-We agree that the main findings of the paper are the software citation principles, and as described in our response to point 1 above we have updated the introduction and made some structural changes to make this clear.
+We agree that the main findings of the paper are the software citation principles, and as described in our response to point 1 above, we have updated the introduction and made some structural changes to make this clear.
 
-However, we respectfully disagree that the principles are preliminary, but that they refelct current knowledge and experience.
+However, we respectfully don't agree that the principles are preliminary, but would describe them as reflecting the group's current knowledge and experience.
+We agree that ``only the future can tell whether software citations really work out,'' but this is true of most papers: they capture the current state of scientific knowledge in an area.
 
@@
 \textbf{Our response:}
-Thank you for pointing this error out---we have corrected this statement to indicate that the past discussion led to a second version (what was submitted).
+Thank you for pointing this error out. We have corrected this statement to indicate that the past discussion led to a second version (what was submitted).
 
@@
 \niemnote{Arfon: can you help here?}
+\niemnote{I think we have something on this at the end of section 3}
 

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex
Adding some language about software citations being only a part of a more complete solution for citing research outputs.

@@
 \textbf{Our response:}
-\asnote{Perhaps we should be adding a short paragraph outlining how we believe that ultimately academic publishing should be moving towards sharing all of the products of researcher activity.
-This work is primarily about making software a citable entity which is just one part of a much larger change in the scholarly ecosystem.}
-
-\niemnote{Dan, can you do this?} \katznote{I was hoping Arfon might do this one}
-\niemnote{I think we have something along these lines now at the end of section 3, although perhaps could say more.}
+Ultimately we believe that software citation is just a part of a more complete solution for the sharing and citation of all of the scholarly research products.
+We have updated the language in the first paragraph of ``Motivations''
 

### 2016-07-26 sc-principles/software-citation-principles.tex
fixing spaces around em-dash
(final) wording fix on historical use case
Adding some language about software citations being only a part of a more complete solution for citing research outputs.

@@
 challenge is not just the textual citation of software in a paper, but the more
-general identification of software used within the research process.
+general identification of software used within the research process. This work is
+focused on making software a citable entity throught the scholarly ecosystem. While
+software products are a relatively small fraction of the sum total of
+research output, this work together with other efforts such as the Force11
+Data Citation Principles~\cite{data-citation-principles} collectively represent
+an effort to better describe all the outputs of research.
 
@@
 
-Researcher is engaged in historical/cultural research --- e.g., study of video games as cultural artifacts.
+Researcher is engaged in historical/cultural research, e.g., a study of video games as cultural artifacts.
 

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex
Adding citation to original Transitive Credit paper

@@
 \textbf{Our response:}
-\asnote{We probably need to dig out some citations about improved/rich citation.
-I'm aware of some work that was going on in the PLOS Labs group about this but I'm not sure if that work was ever published.}
+In addition to the new language in the ``motivations'' section describing a what a more complete `fix' might look like we have included an additional sentence and citation to existing work on this topic towards the end of section 3.
 
-\niemnote{The idea of citations is not going to go away---but format and how they fit in ecosystem may change.}
-
-\niemnote{Arfon: can you help here?}
-\niemnote{I think we have something on this at the end of section 3}
+\asnote{Added a self-citation (for Dan) here... :-\ Not sure if that will be well-received?}
 

### 2016-07-26 sc-principles/software-citation-principles.tex
Adding citation to original Transitive Credit paper

@@
 Data Citation Principles~\cite{data-citation-principles} collectively represent
-an effort to better describe all the outputs of research.
+an effort to better describe all of the outputs of research.
 
@@
 research software underscores the need to overhaul the system of credit for
-all research products.
+all research products. One possible solution for a more complete description of the
+citations and associated credit is discussed in Katz~\cite{transitive_credit_original}.
 

### 2016-07-26 sc-principles/software-citation-principles.tex
all of the outputs -> all outputs

@@
 Data Citation Principles~\cite{data-citation-principles} collectively represent
-an effort to better describe all of the outputs of research.
+an effort to better describe all outputs of research.
 

### 2016-07-26 sc-principles/software-citation-principles.tex
added 2nd transitive credit ref to text

@@
 all research products. One possible solution for a more complete description of the
-citations and associated credit is discussed in Katz~\cite{transitive_credit_original}.
+citations and associated credit is the transitive credit proposed by
+Katz and Smith~\cite{transitive_credit_original,transitive_credit_json-ld}.
 

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex
finalize response to R1 #8

@@
 \textbf{Our response:}
-In addition to the new language in the ``motivations'' section describing a what a more complete `fix' might look like we have included an additional sentence and citation to existing work on this topic towards the end of section 3.
-
-\asnote{Added a self-citation (for Dan) here... :-\ Not sure if that will be well-received?}
+We have added new language to the Motivation section (\S 2) describing what a
+more complete solution might look like, and also included an additional sentence
+and citation to existing work on this topic towards the end of \S 3.
 

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex
edit response to R1 #2

@@
 \textbf{Our response:}
-Ultimately we believe that software citation is just a part of a more complete solution for the sharing and citation of all of the scholarly research products.
-We have updated the language in the first paragraph of ``Motivations''
+Ultimately we believe that software citation is just a part of a more complete
+solution for the sharing and citation of all products of scholarly research.
+We have updated the language in the first paragraph of \S 2 to reflect this.
 

### 2016-07-26 sc-principles/software-citation-principles.tex
some wording fix at beginning of motivation

@@
 general identification of software used within the research process. This work is
-focused on making software a citable entity throught the scholarly ecosystem. While
-software products are a relatively small fraction of the sum total of
-research output, this work together with other efforts such as the Force11
-Data Citation Principles~\cite{data-citation-principles} collectively represent
-an effort to better describe all outputs of research.
+focused on making software a citable entity in the scholarly ecosystem. While
+software products represent a small fraction of the sum total of
+research output, this work together with other efforts such as the FORCE11
+Data Citation Principles~\cite{data-citation-principles,10.7717/peerj-cs.1}
+collectively represent an effort to better describe (and cite) all outputs of
+research.
 

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex
Adding rebuttal for Git SHA recommendation. Still need to add new text to paper

--- 2016-07-26 sc-principles/rebuttal/rebuttal.tex
@@
 \textbf{Our response:}
+While we believe there is value in including the explicit version (e.g.
+Git SHA1 hash, Subversion revision number) of the software in a software citation we believe there are a number of reasons that this is not sufficient alone for the purposes of software citation:
 
-\niemnote{Arfon: can you add some discussion here?
-We can add a sentence addressing git hashes and other version control identifiers, and point out that while they may be sufficient in some ways (globally unique), they aren't necessarily resolvable on the Web, and different VCS use different identifiers, etc.
-In other words, they could be used (especially in the absence of other identifier like DOI), but not sufficient or recommended.}
+- Version numbers are not guaranteed to be permanent: projects can be migrated to new version control systems (e.g.
+SVN to Git).
+In addition, it is possible to overwrite versions (e.g.
+Git force-pushing).
+- Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL
+
+``Unique identifier'' refers to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL.
+
+We recommend that software citations capture the versions of the code expressed as (ideally) a SemVer number (http://semver.org/) as these are points in time at which the authors of the code made the explicit decision to release the code.
+
+A SemVer release number together with a DOI (and associated archive in e.g.
+Zenodo) is we feel the is the most robust & future-proof solution.
+That said, in the absence of `recommended' identifiers such as a DOI) a Git SHA together with a link to a Git repository would be preferable to not having this information.
+
+We would also like to point out that there is a follow-on working group that will be producing recommended implementations for software citations in a particular setting.
+Ultimately the decision to (optionally) include e.g.
+Git SHAs are left for the implementation group to consider.
 

### 2016-07-29 sc-principles/rebuttal/rebuttal.tex
Adding discussion about versions to main text

@@
 - Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL
-
-``Unique identifier'' refers to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL.
+- When we talk about a ``Unique identifier'' in the document, we're refering to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL and we don't believe a repository address with a version number is sufficient.
 
@@
 Git SHAs are left for the implementation group to consider.
+
+We have updated the paper text to reflect some of these ideas.
 

### 2016-07-29 sc-principles/software-citation-principles.tex
is focused -> focuses
Adding discussion about versions to main text

@@
 challenge is not just the textual citation of software in a paper, but the more
-general identification of software used within the research process. This work is
-focused on making software a citable entity in the scholarly ecosystem. While
+general identification of software used within the research process. This work
+focuses on making software a citable entity in the scholarly ecosystem. While
 software products represent a small fraction of the sum total of
@@
 
+While we believe there is value in including the explicit version (e.g.
+Git SHA1 hash, Subversion revision number) of the software in any software citation we believe there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:
+
+\begin{enumerate}
+\item Version numbers/commit references are not guaranteed to be permanent.
+Projects can be migrated to new version control systems (e.g.
+SVN to Git).
+In addition, it is possible to overwrite/clobber particular version (e.g.
+force-pushing in the case of Git).
+\item A repository address and version number doesn't guarantee that the software is available at a particular (resolvable) URL.
+\item A particular version number/commit reference may not represent a `preferred' point at which to cite the software from the perspective of the package authors.
+\end{enumerate}
+
 Note that the ``unique'' in a UID means that it points to a unique, specific software. However, multiple UIDs might point to the same software.

### 2016-07-29 sc-principles/rebuttal/rebuttal.tex
cleaning up rebuttal text, fixing spelling, etc.

--- 2016-07-26 sc-principles/rebuttal/rebuttal.tex
@@
 Then, Section 6 (previously 5) discusses the ``results''.
-Our modifications to the structure of the paper, and changes we hope improve the logical flow, should help indicate these connections better.
+Our modifications to the structure of the paper, and changes that we hope improve the logical flow, should help make these connections more clear.
 
@@
 solution for the sharing and citation of all products of scholarly research.
-We have updated the language in the first paragraph of \S 2 to reflect this.
+We have updated the language in the first paragraph of \S2 to reflect this.
 
@@
 \textbf{Our response:}
-We have added new language to the Motivation section (\S 2) describing what a
+We have added new language to the Motivation section (\S2) describing what a
 more complete solution might look like, and also included an additional sentence
-and citation to existing work on this topic towards the end of \S 3.
+and citation to existing work on this topic towards the end of \S3.
 
@@
 \textbf{Our response:}
-Principle (6) states that ``Software identification should be as specific as necssary'', and we consider that the specific version includes the branch.
+Principle (6) states that ``Software identification should be as specific as necessary'', and we consider that the specific version includes the branch.
 
@@
 Software as a service is a fact, and we need a method to cite it---even though some metadata may not exist or be available, we need to do the best we can.
-The same issues would hold for commerical software or that available only as an executable.
+The same issues would hold for commercial software or that available only as an executable.
 If you can't find the name of software, should be able to find the name of the service\slash endpoint or function being used.
@@
 \textbf{Our response:}
-We feel it is more clear to a reader to quote the basis by which we are making follow-on comments, rather than pointing back to a previous part of the paper.
+We feel it is more clear to a reader to quote the basis by which we are making follow-on comments, rather than just pointing back to a previous part of the paper.
 
@@
 While we believe there is value in including the explicit version (e.g.
-Git SHA1 hash, Subversion revision number) of the software in a software citation we believe there are a number of reasons that this is not sufficient alone for the purposes of software citation:
+Git SHA1 hash, Subversion revision number) of the software in a software citation, we also believe there are a number of reasons that this, by itself, is not sufficient for the purposes of software citation:
 
@@
 - Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL
+- When we talk about a ``Unique identifier'' in the document, we're referring to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL; we don't believe a repository address with a version number is sufficient.
 
-``Unique identifier'' refers to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL.
+We recommend that software citations capture the versions of the code expressed as (ideally) a semantic version (SemVer) number (http://semver.org/).
+These are points in time at which the authors of the code made the explicit decision to release the code.
 
-We recommend that software citations capture the versions of the code expressed as (ideally) a SemVer number (http://semver.org/) as these are points in time at which the authors of the code made the explicit decision to release the code.
-
-A SemVer release number together with a DOI (and associated archive in e.g.
-Zenodo) is we feel the is the most robust & future-proof solution.
+We feel a SemVer release number together with a DOI (and associated archive in, e.g., Zenodo) is the most robust and future-proof solution.
 That said, in the absence of `recommended' identifiers such as a DOI) a Git SHA together with a link to a Git repository would be preferable to not having this information.
 
-We would also like to point out that there is a follow-on working group that will be producing recommended implementations for software citations in a particular setting.
-Ultimately the decision to (optionally) include e.g.
-Git SHAs are left for the implementation group to consider.
+We would also like to point out that there will be a follow-on working group that will produce recommended implementations for software citations in a particular setting.
+Ultimately, the decision to (optionally) include something like Git SHAs is left for that implementation group to consider.
+
+We have updated the paper text to reflect some of these ideas.
 
@@
 \item We feel that the metadata omitted would not be necessary from the perspective of the publisher, the stakeholder for this use case.
-\item We feel that indexed citations are not particularly relevant for a catalog or registry (based on input from adminstrators of such entities), unless they want to display citations of the software---but this would be use case 4.
+\item We feel that indexed citations are not particularly relevant for a catalog or registry (based on input from administrators of such entities), unless they want to display citations of the software---but this would be use case 4.
 Thus, we prefer to leave this as is.

### 2016-07-29 sc-principles/software-citation-principles.tex
minor cleanup of new text small changes in my affiliation

@@
 \thanks{{}$^{1}$ GitHub, Inc., San Francisco, CA, USA}
-\thanks{{}$^{2}$ National Center for Supercomputing Applications \& Electrical and Computer Engineering Department \& School of Information Sciences, University of Illinois at Urbana--Champaign, Urbana, IL, USA}
+\thanks{{}$^{2}$ National Center for Supercomputing Applications (NCSA) \& Electrical and Computer Engineering (ECE) Department \& School of Information Sciences (iSchool), University of Illinois at Urbana--Champaign, Urbana, IL, USA}
 \thanks{{}$^{3}$ School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
@@
 
-While we believe there is value in including the explicit version (e.g.
-Git SHA1 hash, Subversion revision number) of the software in any software citation we believe there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:
+While we believe there is value in including the explicit version (e.g., Git SHA1 hash, Subversion revision number) of the software in any software citation, we believe there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:
 
@@
 \item Version numbers/commit references are not guaranteed to be permanent.
-Projects can be migrated to new version control systems (e.g.
-SVN to Git).
-In addition, it is possible to overwrite/clobber particular version (e.g.
-force-pushing in the case of Git).
+Projects can be migrated to new version control systems (e.g., SVN to Git).
+In addition, it is possible to overwrite/clobber particular version (e.g., force-pushing in the case of Git).
 \item A repository address and version number doesn't guarantee that the software is available at a particular (resolvable) URL.

### 2016-07-30 sc-principles/rebuttal/rebuttal.tex
revise final response to R1 #14 (unique identifier)

@@
 \textbf{Our response:}
-While we believe there is value in including the explicit version (e.g.
-Git SHA1 hash, Subversion revision number) of the software in a software citation, we also believe there are a number of reasons that this, by itself, is not sufficient for the purposes of software citation:
+While we agree there is value in including the explicit version (e.g., Git SHA1 hash, Subversion revision number) of the software in a software citation, we also believe there are a number of reasons that this, by itself, is not sufficient for the purposes of software citation:
 
-- Version numbers are not guaranteed to be permanent: projects can be migrated to new version control systems (e.g.
-SVN to Git).
-In addition, it is possible to overwrite versions (e.g.
-Git force-pushing).
-- Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL
-- When we talk about a ``Unique identifier'' in the document, we're referring to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL; we don't believe a repository address with a version number is sufficient.
+\begin{itemize}
+\item Version numbers are not guaranteed to be permanent: projects can be migrated to new version control systems (e.g., SVN to Git).
+In addition, it is possible to overwrite versions (e.g., Git force-pushing).
+\item Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL.
+\item When we talk about a ``Unique identifier'' in the document, we are referring to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL; we do not believe a repository address with a version number is sufficient.
+\end{itemize}
 
-We recommend that software citations capture the versions of the code expressed as (ideally) a semantic version (SemVer) number (http://semver.org/).
+We recommend that software citations capture the versions of the code expressed as (ideally) a semantic version (SemVer) number (\url{http://semver.org/}).
 These are points in time at which the authors of the code made the explicit decision to release the code.
-
 We feel a SemVer release number together with a DOI (and associated archive in, e.g., Zenodo) is the most robust and future-proof solution.
-That said, in the absence of `recommended' identifiers such as a DOI) a Git SHA together with a link to a Git repository would be preferable to not having this information.
+That said, in the absence of ``recommended'' identifiers (such as a DOI) a Git SHA together with a link to a Git repository would be preferable to not having this information.
 

### 2016-07-30 sc-principles/software-citation-principles.tex
spelling fix
final minor edits

@@
 Institute (SSI) in the form of blog posts written by their community fellows.
-For examplel, in a blog post from 2012, Jackson discusses some of the pitfalls
+For example, in a blog post from 2012, Jackson discusses some of the pitfalls
 of trying to cite software in publications~\cite{ssi-how-to-cite}. He includes
@@
 
-While we believe there is value in including the explicit version (e.g., Git SHA1 hash, Subversion revision number) of the software in any software citation, we believe there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:
+While we believe there is value in including the explicit version (e.g., Git SHA1 hash, Subversion revision number) of the software in any software citation, there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:
 
@@
 Projects can be migrated to new version control systems (e.g., SVN to Git).
-In addition, it is possible to overwrite/clobber particular version (e.g., force-pushing in the case of Git).
-\item A repository address and version number doesn't guarantee that the software is available at a particular (resolvable) URL.
-\item A particular version number/commit reference may not represent a `preferred' point at which to cite the software from the perspective of the package authors.
+In addition, it is possible to overwrite/clobber a particular version (e.g., force-pushing in the case of Git).
+\item A repository address and version number does not guarantee that the software is available at a particular (resolvable) URL.
+\item A particular version number/commit reference may not represent a ``preferred'' point at which to cite the software from the perspective of the package authors.
 \end{enumerate}
 
-Note that the ``unique'' in a UID means that it points to a unique, specific software. However, multiple UIDs might point to the same software.
+Note that the ``unique'' in a UID means that it points to a unique, specific software version. However, multiple UIDs might point to the same software.
 This is not recommended, but is possible.

### 2016-07-30 sc-principles/software-citation-principles.tex
materialize -> materialized

@@
 The group gathered members (see Appendix~\ref{app:wg_members}) in April and
-May 2015, and then began work in June. This materialize as a number of meetings
+May 2015, and then began work in June. This materialized as a number of meetings
 and offline work by group members to document existing

### 2016-08-02 sc-principles/software-citation-principles.tex
per journal instruction, removed funding info from acks

@@
 
-Work by Daniel S.~Katz was supported in part by the National Science Foundation (NSF) while working at the Foundation.
-Any opinion, finding, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the NSF.
-Work by Kyle E.~Niemeyer was supported in part by the NSF under grant ACI-1535065.
+Any opinion, finding, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the NSF.
 

### 2016-08-02 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
adding source for submitted version of WSSSPE4 lightning talk as latex still need to update

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Lightning Talk: Software Citation: Process, Principles, and Implementation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{Daniel S. Katz, }
%\IEEEauthorblockA{University of Illinois Urbana-Champaign\\
%Urbana, Illinois, USA\\
%Email: d.katz@ieee.org}
%\and
%\IEEEauthorblockN{Kyle E. Niemeyer}
%\IEEEauthorblockA{Oregon State University\\
%Corvallis, Oregon, USA\\
%Email: }
%\and
%\IEEEauthorblockN{Arfon M. Smith}
%\IEEEauthorblockA{GitHub, Inc.\\
%San Francisco, California, USA\\
%Email: }
%\and
%\IEEEauthorblockN{{FORCE11 Software Citation Working Group}}
%\IEEEauthorblockA{FORCE11}
%}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
\author{\IEEEauthorblockN{Daniel S. Katz\IEEEauthorrefmark{1},
Kyle E. Niemeyer\IEEEauthorrefmark{2},
Arfon M. Smith\IEEEauthorrefmark{3}, 
FORCE11 Software Citation Working Group\IEEEauthorrefmark{4}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}University of Illinois Urbana-Champaign\\
Urbana, Illinois, USA\\
Email: d.katz@ieee.org}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Oregon State University\\
Corvallis, Oregon, USA\\
Email: Kyle.Niemeyer@oregonstate.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}GitHub, Inc.\\
San Francisco, California, USA\\
Email: arfon@github.com}
\IEEEauthorblockA{\IEEEauthorrefmark{4}FORCE11\\
\url{https://www.force11.org/group/software-citation-working-group}}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation. Inspired by the activities of the FORCE11 working group focused on data citation, FORCE11 started a Software Citation Working Group (SCWG). The group initially sought members, and currently has about 55--60 members. The working group reviewed existing community practices, developed a set of use cases, and drafted a software citation principles document.

This presentation will discuss the principles (in brief: importance, credit and attribution, unique identification, persistence, accessibility, and specificity), how they will impact the practice of research, and they can be implemented by researchers, publishers, librarians and others who build and maintain repositories, scholars of science, university administrators, and research funders. It should also spark discussion in Track 2 of WSSSPE4 about both the next steps related to software citation and the community goals related to software credit, reproducibility, and sustainability. 
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Lightning Talk}
% no \IEEEPARstart
%This demo file is intended to serve as a ``starter file''
%for IEEE conference papers produced under \LaTeX\ using
%IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
Software is a critical part of modern research and yet there is little support across the scholarly ecosystem for its acknowledgement and citation. Inspired by the activities of the FORCE11 working group focused on data citation, FORCE11 started a Software Citation Working Group (SCWG). The group initially sought members, and currently has about 55–60 people (researchers, developers, publishers, repositories, librarians) as members, including members of WSSSPE3 software credit breakout group who joined en masse in October 2015.

The working group reviewed existing community practices, including those of groups such as the Software Sustainability Institute, WSSSPE, Project CRediT, Ontosoft, and CodeMeta, and in domains such as astronomy and astrophysics, life sciences, geosciences. The group then developed a set of use cases (collaborative via a Google Doc \cite{bib1}). Finally, the group drafted a software citation principles document, starting with the FORCE11 data citation principles, and then updated the equivalent software principles based on the software use cases and related work, a set of working group discussions, community feedback and review of the draft, and feedback and discussion in a one-day workshop at FORCE2016 in April 2016.

Seventeen discrete use cases related to software citation were established to help understand the necessary requirements for citation. These involve 13 stakeholder types, including researchers, research software engineers, publishers, indexers, domain groups, libraries, archives, repositories, funders, policy makers, evaluators, and citation managers. The use cases helped identify the basic metadata needed in a citation of software: unique identifier, software name, author(s), version number, release date, and location. Interestingly, only unique identifier was needed for all use cases.
Beyond the principles of software citation themselves, the document contains extensive discussion about the principles and related topics. These topics include which unique identifiers should be used (DOIs are recommended), what software should be cited, the role of software papers, how to cite derived software, basic elements of a citation format in reference lists, the elimination of citation limits, the types of software that should be cited (all), and what an identifier should resolve to. In addition, the document contains a discussion of past and other work related to software citation, both in specific domains and the general research software community.

The FORCE11 Software Citation Working Group then made a set of recommendations for Software Citation Principles at the full FORCE2016 meeting April 2016. This has the goal of encouraging broad adoption of a consistent policy for software citation across disciplines and venues. The group also presented a discussion of the motivations for developing the principles, reviews of existing community practice, and a discussion of the requirements these principles would place upon different stakeholders.

Working examples and possible technical solutions for how these principles can be implemented will be discussed in a separate paper that is being developed.

After the FORCE2016 events, the working group modified the principles document to reflect issues that were raised, and then published the draft document on the FORCE11 website \cite{bib2} and called for public comments, which were made via Hypothes.is (\url{http://hypothes.is}) annotations, GitHub issues, and emails. All comments were compiled into a document on GitHub \cite{bib3}, and explanation of the working group chairs’ responses (e.g., document changes, further discussion, etc.) were also recorded there. All previous work by the group was also done in the open, and recorded on GitHub, so that all decisions/discussions can be traced.

The final draft paper was then submitted to  PeerJ Computer Science , published as a PeerJ preprint \cite{bib4}, and posted on the FORCE11 website \cite{bib2}. Initial review comments were provided July 8, 2016, and the working group chairs are currently discussing them. Our goal is to have a final version of the paper accepted by WSSSPE4.

The remaining plans of the Software Citation Working Group involve promotion and distribution of the Software Citation Principles, specifically an endorsement effort to get both individuals and organizations to sign on to the principles and related, creation of some publicity material such as an infographic and 1--3 slides. After this, the Software Citation Working Group will have completed its work and will end.

Next, we expect FORCE11 to spin up of a new working group focused on implementing the software citation principles. This group will work with institutions, publishers, funders, researchers, etc., to implement the principles established by the first group, and will write an implementation examples paper.

This presentation will discuss the principles (in brief: importance, credit and attribution, unique identification, persistence, accessibility, and specificity), how they will impact the practice of research, and they can be implemented by researchers, publishers, librarians and others who build and maintain repositories, scholars of science, university administrators, and research funders. It should also spark discussion in Track 2 of WSSSPE4 about both the next steps related to software citation and the community goals related to software credit, reproducibility, and sustainability.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

%\section{Conclusion}
%The conclusion goes here.
%

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{bib1}
FORCE11 Software Citation Working Group, ``Software citation use cases,'' \url{https://docs.google.com/document/d/1dS0SqGoBIFwLB5G3HiLLEOSAAgMdo8QPEpjYUaWCvIU}, 2016. Accessed: 2016-07-10.
\bibitem{bib2} 
A.~M.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group ``Software Citation Principles,'' FORCE2016 Website, \url{https://www.force11.org/software?citation?principles}, 2016. Accessed: 2016-07-10.
\bibitem{bib3} 
FORCE11 Software Citation Working Group, GitHub repository, \url{https://github.com/force11/force11-scwg}.  Accessed: 2016-07-10.
\bibitem{bib4} 
A.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group, “Software Citation Principles,” PeerJ Preprints 4:e2169v2, 2106. \url{https://doi.org/10.7287/peerj.preprints.2169v2}.

\end{thebibliography}

% that's all folks
\end{document}

### 2016-08-02 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
updating text+URL to show resubmitted version of principles paper fixing columns on last page

@@
 
+\usepackage{flushend} % equalizes columns on last page
+
 % correct bad hyphenation here
@@
 
-The final draft paper was then submitted to  PeerJ Computer Science , published as a PeerJ preprint \cite{bib4}, and posted on the FORCE11 website \cite{bib2}.
-Initial review comments were provided July 8, 2016, and the working group chairs are currently discussing them.
+The final draft paper was then submitted to  PeerJ Computer Science, published as a PeerJ preprint, and posted on the FORCE11 website \cite{bib2}.
+Initial review comments were returned July 8, 2016, and the working group chairs recently (August 1, 2016) submitted a revised version, and updated the PeerJ preprint \cite{bib4}.
 Our goal is to have a final version of the paper accepted by WSSSPE4.
@@
 A.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group, “Software Citation Principles,” PeerJ Preprints 4:e2169v2, 2106.
-\url{https://doi.org/10.7287/peerj.preprints.2169v2}.
+\url{https://doi.org/10.7287/peerj.preprints.2169v3}.
 

### 2016-08-02 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
fixing bib items

@@
 \bibitem{bib2} 
-A.~M.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group ``Software Citation Principles,'' FORCE2016 Website, \url{https://www.force11.org/software?citation?principles}, 2016.
+A.~M.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group ``Software Citation Principles,'' FORCE2016 Website, \url{https://www.force11.org/software-citation-principles}, 2016.
 Accessed: 2016-07-10.
@@
 \bibitem{bib4} 
-A.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group, “Software Citation Principles,” PeerJ Preprints 4:e2169v2, 2106.
+A.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group, “Software Citation Principles,” \textit{PeerJ Preprints} 4:e2169v2, 2106.
 \url{https://doi.org/10.7287/peerj.preprints.2169v3}.

### 2016-08-04 sc-principles/software-citation-principles.tex

@@
 
-Any opinion, finding, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the NSF.
+While D.~S.~Katz prepared this material while employed at the NSF, any opinion, finding, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the NSF.
 

### 2016-08-04 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
small wording change
small edit for -- adding ack fixing last bib entry spacing

--- 2016-08-02 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
@@
 % Basically, \url{my_url_here}.
+\usepackage{hyperref}
 
@@
 % to submit to, of course. )
+
+\usepackage[keeplastbox]{flushend} % equalizes columns on last page
 
@@
 Inspired by the activities of the FORCE11 working group focused on data citation, FORCE11 started a Software Citation Working Group (SCWG).
-The group initially sought members, and currently has about 55–60 people (researchers, developers, publishers, repositories, librarians) as members, including members of WSSSPE3 software credit breakout group who joined en masse in October 2015.
+The group initially sought members, and currently has about 55--60 people (researchers, developers, publishers, repositories, librarians) as members, including members of WSSSPE3 software credit breakout group who joined en masse in October 2015.
 
@@
 
-After the FORCE2016 events, the working group modified the principles document to reflect issues that were raised, and then published the draft document on the FORCE11 website \cite{bib2} and called for public comments, which were made via Hypothes.is (\url{http://hypothes.is}) annotations, GitHub issues, and emails.
-All comments were compiled into a document on GitHub \cite{bib3}, and explanation of the working group chairs’ responses (e.g., document changes, further discussion, etc.) were also recorded there.
+After the FORCE2016 events, the working group modified the principles document to reflect issues that were raised, and then published the draft document on the FORCE11 website~\cite{bib2} and called for public comments, which were made via Hypothes.is (\url{http://hypothes.is}) annotations, GitHub issues, and emails.
+All comments were compiled into a document on GitHub~\cite{bib3}, and explanation of the working group chairs' responses (e.g., document changes, further discussion, etc.) were also recorded there.
 All previous work by the group was also done in the open, and recorded on GitHub, so that all decisions/discussions can be traced.
 
-The final draft paper was then submitted to  PeerJ Computer Science , published as a PeerJ preprint \cite{bib4}, and posted on the FORCE11 website \cite{bib2}.
-Initial review comments were provided July 8, 2016, and the working group chairs are currently discussing them.
-Our goal is to have a final version of the paper accepted by WSSSPE4.
+The final draft paper was then submitted to \textit{PeerJ Computer Science}, published in \textit{PeerJ Preprints}, and posted on the FORCE11 website~\cite{bib2}.
+Initial review comments were returned 8 July 2016, and the working group chairs recently (1 August 2016) submitted a revised version, and updated the \textit{PeerJ Preprints} document~\cite{bib4}.
+Our goal is to have a final version of the paper accepted and published by WSSSPE4.
 
@@
 
-The authors would like to thank...
+The authors would like to thank all members of the FORCE11 Software Citation Group (listed in~\cite{bib3}), the members of the FORCE11 executive committee for suggesting this activity and supporting our work on it, and the WSSSPE community (\url{http://wssspe.researchcomputing.org.uk}) for also suggesting this activity and providing a group of active participants over 3 meetings and the set of FORCE11 working group activities.
 
@@
 \bibitem{bib2}
-A.~M.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group ``Software Citation Principles,'' FORCE2016 Website, \url{https://www.force11.org/software?citation?principles}, 2016.
+A.~M.~Smith, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group ``Software Citation Principles,'' FORCE2016 Website, \url{https://www.force11.org/software-citation-principles}, 2016.
 Accessed: 2016-07-10.
@@
 \bibitem{bib4}
-A.~Smith AM, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group, “Software Citation Principles,” PeerJ Preprints 4:e2169v2, 2106.
-\url{https://doi.org/10.7287/peerj.preprints.2169v2}.
+A.~M.~Smith, D.~S.~Katz, K.~E.~Niemeyer, and FORCE11 Software Citation Working Group, ``Software Citation Principles,'' \textit{PeerJ Preprints} 4:e2169v3, 2016.
+\url{https://doi.org/10.7287/peerj.preprints.2169v3}.
 

### 2016-08-04 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
adding funding stuff in ack

@@
 
+Work by Daniel S.~Katz was supported in part by the National Science Foundation (NSF) while working at the Foundation.
+Any opinion, finding, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the NSF.
+Work by Kyle E.~Niemeyer was supported in part by the NSF under grant ACI-1535065.
+
 % trigger a \newpage just before the given reference

### 2016-08-12 meetings/WSSSPE4-lightning-talk/WSSSPE4-lightning-talk.tex
final adding copying and formatting authors and column lengths on last page

@@
 
-\usepackage[keeplastbox]{flushend} % equalizes columns on last page
+%\usepackage[keeplastbox]{flushend} % equalizes columns on last page
 
@@
 Arfon M. Smith\IEEEauthorrefmark{3},
-FORCE11 Software Citation Working Group\IEEEauthorrefmark{4}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}University of Illinois Urbana-Champaign\\
-Urbana, Illinois, USA\\
+FORCE11 Software Citation Working Group\IEEEauthorrefmark{4}\\}
+\vspace{0.2cm}
+\IEEEauthorblockA{\IEEEauthorrefmark{1}University of Illinois Urbana-Champaign,
+Urbana, Illinois, USA,
 Email: d.katz@ieee.org}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Oregon State University\\
-Corvallis, Oregon, USA\\
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Oregon State University,
+Corvallis, Oregon, USA,
 Email: Kyle.Niemeyer@oregonstate.edu}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}GitHub, Inc.\\
-San Francisco, California, USA\\
+\IEEEauthorblockA{\IEEEauthorrefmark{3}GitHub, Inc.,
+San Francisco, California, USA,
 Email: arfon@github.com}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}FORCE11\\
+\IEEEauthorblockA{\IEEEauthorrefmark{4}FORCE11,
 \url{https://www.force11.org/group/software-citation-working-group}}}
@@
 %\IEEEspecialpapernotice{(Invited Paper)}
+
+\IEEEoverridecommandlockouts
+\IEEEpubid{\makebox[\columnwidth]{This work is licensed under a \href{https://creativecommons.org/licenses/by/4.0/}{CC-BY-4.0 license}. \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
 
@@
 
+{
 This presentation will discuss the principles (in brief: importance, credit and attribution, unique identification, persistence, accessibility, and specificity), how they will impact the practice of research, and they can be implemented by researchers, publishers, librarians and others who build and maintain repositories, scholars of science, university administrators, and research funders.
-It should also spark discussion in Track 2 of WSSSPE4 about both the next steps related to software citation and the community goals related to software credit, reproducibility, and sustainability.
+It should also spark discussion in Track 2 of WSSSPE4 about both the next steps related to software
+\unskip\parfillskip 0pt \par }
+
+\vfill \break
+
+\noindent citation and the community goals related to software credit, reproducibility, and sustainability.
 

### 2016-08-16 sc-principles/rebuttal-2/rebuttal.tex
added second rebuttal with reviewer comments

\pdfoutput=1
\pdfminorversion=4
\documentclass{article}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\geometry{letterpaper}
\usepackage{graphicx, caption}
\usepackage{latexsym,url,amscd, amssymb, amsmath}
\usepackage{epstopdf}

\usepackage{multicol}
\usepackage{booktabs}
\usepackage{fancyvrb}

\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}

\usepackage[english]{babel}
\usepackage{textcomp}

\usepackage{siunitx}
\sisetup{group-separator={,},
         detect-all,
         binary-units,
         list-units = single,
         range-units = single,
         tophrase = --,
         per-mode = symbol-or-fraction,
         separate-uncertainty = true
%        scientific-notation = fixed
}
\DeclareSIUnit\atm{atm}
\DeclareSIUnit\CA{CA}
\DeclareSIUnit\rpm{rpm}

%to highlight text
\usepackage{soul}
\usepackage[usenames, dvipsnames]{xcolor}
\newcommand{\hly}[1]{{\sethlcolor{yellow}\hl{#1}}}
\newcommand{\hlb}[1]{{\sethlcolor{SkyBlue}\hl{#1}}}
\newcommand{\hlg}[1]{{\sethlcolor{green}\hl{#1}}}

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{}{;}
\usepackage[linkcolor=blue, citecolor=blue, colorlinks=true, breaklinks=true]{hyperref}

\usepackage{url,breakurl}
\usepackage{enumitem}

\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}} % Dan
\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}} % Kyle
\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}} %Arfon

%\date{}                                           % Activate to display a given date or no date

\setlength{\parindent}{0in}
\setlength{\parskip}{0.4cm}

%%% BEGIN DOCUMENT
\begin{document}

We thank the editor and reviewers for their helpful comments, and have addressed their requests and questions here.
Specific changes or additions can be found in the marked revision.
% in response to the Editor, Reviewer \#1, and Reviewer \#2 are respectively highlighted in the marked manuscript in \hly{yellow}, \hlb{blue}, and \hlg{green}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Editor:}

{\itshape
First of all, I would like to thank the authors for having addressed all my
comments and reviewers' suggestions in their revision (and clearly explained
their modifications in their response letter), and thanks again to the
reviewers for their wonderful job.

While the paper is now in a good shape for being published, there are just
minor issues that should be properly addressed before having it accepted
for the final publication. Such issues are highlighted by the reviewer, and
concern the new parts the authors have added, in particular:

\begin{enumerate}
\item the comparison between software and data;
\item the assignment of an identifier for explicit versions of a software by
means of repository capabilities (e.g. hashes).
\end{enumerate}

This is not a huge amount of work to add, as far as I can see, but a clear
discussion (even supported by appropriate references) would be beneficial
for the article.

Thanks again for your wonderful work.
}

\textbf{Our response:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Reviewer \#1 (Tobias Kuhn):}

{\itshape
The authors followed most of my suggestions, and where they did not, they
mostly presented convincing arguments. I like in particular that there is
now an appendix with the use cases.

However, there are two aspects of the latest version that I find not very
convincing:
}

\begin{enumerate}

\item \textbf{Reviewer comment:}
\emph{The listed differences between software and data (lines 100--114) are not
convincing in my opinion:}

\begin{itemize}

\item \emph{``'Software can be used to express or explain concepts, unlike
data'': I don't understand this.}

\item \emph{``'Software is updated more frequently that papers or data''
(typo: ``''that'' instead of ``than''): This was questioned also by the
second reviewer, and I don't think the issue is resolved. The footnote is
unsupported and seems to partly contradict the main point. (I can imagine that
maybe only *popular* software tends to be updated frequently, whereas
*popular* dataset are often not.)}

\item \emph{``'Software suffers from a different type of bit rot than data'':
I am not convinced that this is really a different *type* of bit rot.
In particular datasets in proprietary formats seem to suffer from exactly the
same problems.}

\item \emph{``Software is frequently built to use other software'': True,
though similar things can occur in datasets too, in particular with respect to
Linked Data, where datasets typically use a number of third-party ontologies and
vocabularies.}

\item \emph{``Software is generally smaller than data'': Another unsupported
claim, and I am not fully convinced that this is true. There must be a large
number of very small spreadsheets of data. So, on average data might be smaller
than software. The difference is certainly true on the upper end of the scale
though: the largest datasets are larger than the largest pieces of software.}

\item{``Software teams can be large and multidisciplinary, \ldots'': I think
this applies to data too.}

\item{``The lifetime of software is generally not as long as that of data.'':
Again unsupported, and again I can imagine this to be false. Many datasets, in
particular small ones, might be stillbirths (i.e.\ they are never used). The
same might be true for software, but possibly to a lesser degree. And software
might commonly evolve, whereas data might more often be replaced by a
different and newer dataset, in which case evolving software would live
longer.}

\end{itemize}

\item \textbf{Reviewer comment:}
\emph{I am also not convinced by the arguments against hashes as identifiers:}

\begin{itemize}

\item \emph{``commit references are not guaranteed to be permanent'':
identifiers consisting of strong hash values *are* guaranteed to be permanent
in the sense that they can forever only represent what they were meant to
represent, i.e. the bits and bytes on which the hash was calculated (which, in
the case of Git, is the entire version history).}

\item \emph{``A repository address and version number does not guarantee that
the software is available at a particular (resolvable) URL''. It depends. A URL
like \url{https://github.com/torvalds/linux/commit/a0cba2179ea4c1820fce2ee046b6ed90ecc56196}
*does* come with a hash and can be resolved to get to the actual software.
There is no guarantee that github.com will be available forever, but this is
also true for publisher sites and therefore DOIs.}

\item \emph{``A particular version number/commit reference may not represent a
``preferred'' point at which to cite the software from the perspective of the
package authors.'': Hashes certainly don't prevent you from defining preferred
versions for future reference. But they come with the *possibility* to identify
all versions, which I would in fact consider an important feature. If I want to
make my research perfectly reproducible, I should be able to cite the exact
version that I used, whether that happens to be a preferred version or not.}

\item \emph{``When we talk about a ``Unique identifier'' in the document, we
are referring to unique, persistent, and machine-actionable identifiers such as
a DOI, ARK, or PURL; we do not believe a repository address with a version
number is sufficient.'' [from the rebuttal letter]: Identifiers that include
hashes are in important ways *more* unique, *more* persistent, and *more*
machine-actionable than DOIs or PURLs. Strong hashing algorithms guarantee
uniqueness (no other input can be constructed for the same hash), are in
practice probably more persistent (for example, there exist independent
complete copies of all GitHub content, so even if github.com closes for good,
the cited software could still be found), and can be automatically verified
against their supposed content. (The above holds for *strong* hash functions,
and Git's SHA-1 is no longer considered strong - even though not a single
collision has been found until now - but this doesn't invalidate the general
points.)}

\end{itemize}

\end{enumerate}

%\bibliography{}
%\bibliographystyle{}

\end{document}

### 2016-08-16 sc-principles/rebuttal-2/rebuttal.tex
added entries for us to respond
draft response on address/permanence

@@
 
+\textbf{Our response:}
+
 \item \emph{``'Software is updated more frequently that papers or data''
@@
 
+\textbf{Our response:}
+
 \item \emph{``'Software suffers from a different type of bit rot than data'':
@@
 
+\textbf{Our response:}
+
 \item \emph{``Software is frequently built to use other software'': True,
@@
 vocabularies.}
+
+\textbf{Our response:}
 
@@
 
+\textbf{Our response:}
+
 \item{``Software teams can be large and multidisciplinary, \ldots'': I think
 this applies to data too.}
+
+\textbf{Our response:}
 
@@
 longer.}
+
+\textbf{Our response:}
 
@@
 
+\textbf{Our response:}
+It is possible to overwrite\slash clobber a particular commit (e.g, using a
+git force-push), and so access to that hash may not be persistent.\niemnote{is this true?}
+
 \item \emph{``A repository address and version number does not guarantee that
@@
 also true for publisher sites and therefore DOIs.}
+
+\textbf{Our response:}
+We agree that a repository URL combined with hash can be used to get to
+the specific version of software referenced. However, this approach to citation
+applies only to git, which not all research software uses. As a result, multiple
+citation solutions would be needed. We prefer to recommend a practice that is
+independent of the software development tools used---archiving a software
+version and registering a DOI does not require the use of Git, or in fact any
+version control system at all (though we of course recommend it).
+
+Regarding the peristence of GitHub vs.\ the DOI system, DOIs are an ISO standard
+governed by a nonprofit foundation with a specific goal of persistence. Beyond
+that, there are robust disaster recovery solutions in place for archivers and
+publishers---the persistence of scholarly objects (and the DOI system) is
+more certain than that of GitHub.
+
+Furthermore, one of the goals behind creating the software citation principles
+is increasing the prominence of software as an equal research product to
+publications and data, which both use DOIs as a standard identifier and locator.
+Thus, we feel it is important (in combination with our other stated reasons) to
+recommend a similar standard for software; while other solutions (such as URL +
+hash) certainly exist, suggesting a different---if equivalently suitable---standard
+would serve to propagate the idea that software is a lesser research output.
+\niemnote{this might be a general reply on URL+hash vs. DOI}
 
@@
 version that I used, whether that happens to be a preferred version or not.}
+
+\textbf{Our response:}
 
@@
 
+\textbf{Our response:}
+
 \end{itemize}

### 2016-08-16 sc-principles/rebuttal-2/rebuttal.tex
added a general response about hash vs DOI

@@
 
+\textbf{Our response:}
+As a general comment, we prefer to recommend DOIs associated with specific,
+released versions of
+software for a number of reasons (not specific to VCS used, more guaranteed
+persistence, same persistent identifier used for publications and data).
+However, there are certain sitatuions where this may not be possible:
+(1) the software developers did not register a DOI or release specific
+versions, (2) the version of software used does not match what is available
+to cite (perhaps they used the dev/prerelease version), etc. In those cases,
+falling back on URL+hash would be an approriate way to cite the software used.
+
+We have updated the manuscript to better match this position.
+\niemnote{need to actually do this}
+
 \begin{itemize}

### 2016-08-16 sc-principles/rebuttal-2/rebuttal.tex
added some additional draft responses on hash vs DOI

@@
 \textbf{Our response:}
+As stated in our general response above and reflected in our changes to the
+manuscript, we agree hashes (combined with a
+repository URL) could be used when citing a version of software that does not
+have an associated DOI.
 
@@
 \textbf{Our response:}
+\niemnote{Something like:} We chose DOIs as the primary unique identifier\slash
+locator to satisfy the principles of unique identification, persistence, and
+accessibility. While strong hashing algorithms (which Git does not use) are
+unique, they cannot currently be used with GitHub repo URLs (e.g.). In addition,
+we prefer a persistent solution that doesn't depend on outside solutions if
+(e.g.) GitHub goes away---the DOI system ensures persistence regardless of the
+original ``owner''\slash naming authority.
 

### 2016-08-17 sc-principles/rebuttal-2/rebuttal.tex
Moving some text around and expanding on hashes as identifiers section

@@
 As a general comment, we prefer to recommend DOIs associated with specific,
-released versions of
-software for a number of reasons (not specific to VCS used, more guaranteed
-persistence, same persistent identifier used for publications and data).
+released versions of software for a number of reasons: (not specific to VCS used,
+more guaranteed persistence, same persistent identifier used for publications and data).
 However, there are certain sitatuions where this may not be possible:
@@
 falling back on URL+hash would be an approriate way to cite the software used.
+
+Furthermore, one of the goals behind creating the software citation principles
+is increasing the prominence of software as an equal research product to
+publications and data, which both use DOIs as a standard identifier and locator.
+Thus, we feel it is important (in combination with our other stated reasons) to
+recommend a similar standard for software; while other solutions (such as URL +
+hash) certainly exist, suggesting a different---if equivalently suitable---standard
+would serve to propagate the idea that software is a lesser research output.
+
+We would also like to draw the reviewer's attention to the fact that this paper
+describes the result of the inital Software Citation Principles working group. A
+followup effort around the implementation of these principles (in a variety of
+settings) will likely provide more detailed guidance and discussion around what
+exactly a software citation should look like and the data it should capture such
+as a Git SHA/REF.
 

### 2016-08-17 sc-principles/rebuttal-2/rebuttal.tex
Removing comment
Adding some additional text to the paper to reflect our comments about URLs+commit refs

@@
 \niemnote{need to actually do this}
-
+\asnote{I've taken a first pass at this.}
 \begin{itemize}
@@
 the specific version of software referenced. However, this approach to citation
-applies only to git, which not all research software uses. As a result, multiple
+applies only to Git, which not all research software uses. As a result, multiple
 citation solutions would be needed. We prefer to recommend a practice that is
@@
 would serve to propagate the idea that software is a lesser research output.
-\niemnote{this might be a general reply on URL+hash vs. DOI}
 
@@
 have an associated DOI.
+
+\asnote{Should we say that this is something we expect the implementation WG to
+work on?}
 

### 2016-08-17 sc-principles/software-citation-principles.tex
Adding some additional text to the paper to reflect our comments about URLs+commit refs

@@
 
-While we believe there is value in including the explicit version (e.g., Git SHA1 hash, Subversion revision number) of the software in any software citation, there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:
+While we believe there is value in including the explicit version (e.g., Git SHA1 hash, Subversion revision number) of the software in any software citation, there are a number of reasons that a commit reference together with a repository URL is not recommended for the purposes of software citation:
 
@@
 In addition, it is possible to overwrite/clobber a particular version (e.g., force-pushing in the case of Git).
-\item A repository address and version number does not guarantee that the software is available at a particular (resolvable) URL.
+\item A repository address and version number does not guarantee that the software is available at a particular (resolvable) URL, especially as it is possible for authors to remove their content from e.g.
+GitHub.
 \item A particular version number/commit reference may not represent a ``preferred'' point at which to cite the software from the perspective of the package authors.
 \end{enumerate}
+
+We recognize however that there are certain situations where it may not be possible for follow the recommended best-practice.
+For example, if (1) the software authors did not register a DOI and/or release a specific version or (2) the version of the
+software used does not match what is available to cite. In those cases, falling back on a combination of the repository URL and
+version number/commit hash would be an appropriate way to cite the software used.
 

### 2016-08-21 sc-principles/rebuttal-2/rebuttal.tex
a few minor edits, removed notes on general url + hash comment

@@
 As a general comment, we prefer to recommend DOIs associated with specific,
-released versions of software for a number of reasons: (not specific to VCS used,
-more guaranteed persistence, same persistent identifier used for publications and data).
+released versions of software for a number of reasons: not specific to VCS used,
+more guaranteed persistence, same persistent identifier used for publications and data.
 However, there are certain sitatuions where this may not be possible:
@@
 We have updated the manuscript to better match this position.
-\niemnote{need to actually do this}
-\asnote{I've taken a first pass at this.}
+
 \begin{itemize}

### 2016-08-21 sc-principles/software-citation-principles.tex
a few minor edits, removed notes on general url + hash comment

@@
 In addition, it is possible to overwrite/clobber a particular version (e.g., force-pushing in the case of Git).
-\item A repository address and version number does not guarantee that the software is available at a particular (resolvable) URL, especially as it is possible for authors to remove their content from e.g.
-GitHub.
+\item A repository address and version number does not guarantee that the software is available at a particular (resolvable) URL, especially as it is possible for authors to remove their content from, e.g., GitHub.
 \item A particular version number/commit reference may not represent a ``preferred'' point at which to cite the software from the perspective of the package authors.
@@
 
-We recognize however that there are certain situations where it may not be possible for follow the recommended best-practice.
-For example, if (1) the software authors did not register a DOI and/or release a specific version or (2) the version of the
-software used does not match what is available to cite. In those cases, falling back on a combination of the repository URL and
+We recognize that there are certain situations where it may not be possible for follow the recommended best-practice.
+For example, if (1) the software authors did not register a DOI and/or release a specific version, or (2) the version of the
+software used does not match what is available to cite.
+In those cases, falling back on a combination of the repository URL and
 version number/commit hash would be an appropriate way to cite the software used.

### 2016-08-21 sc-principles/rebuttal-2/rebuttal.tex
cleaned up rebuttal response on hashes

@@
 It is possible to overwrite\slash clobber a particular commit (e.g, using a
-git force-push), and so access to that hash may not be persistent.\niemnote{is this true?}
+git force-push), or remove a repository entirely, and so access to that hash
+may not be persistent.
 
@@
 have an associated DOI.
-
-\asnote{Should we say that this is something we expect the implementation WG to
-work on?}
+We do expect the follow-on implementation working group to work further on these solutions.
 
@@
 \textbf{Our response:}
-\niemnote{Something like:} We chose DOIs as the primary unique identifier\slash
+The working group focused on DOIs as the primary unique identifier\slash
 locator to satisfy the principles of unique identification, persistence, and

### 2016-08-21 sc-principles/software-citation-principles.tex
changing software vs data discussion to point to github doc

--- 2016-08-17 sc-principles/software-citation-principles.tex
@@
 \S\ref{sec:related_work}.
-We made the adaptations because software, while similar to data in terms of
-not traditionally having been cited in publications, is also different than
-data. Software can be considered a type of data, but the converse is not
-generally true. In addition:
-
-\begin{itemize}
-\item Software can be used to express or explain concepts, unlike data.
-\item Software is updated more frequently that papers or data\footnote{Datasets
-can of course be frequently updated, but each update likely represents the
-findings of a new or repeated study.}.
-\item Software is executable, unlike papers or data.
-\item Software suffers from a different type of bit rot than data---software
-must be constantly maintained so that it continues to function as both the
-hardware and software environments that it is used on changes, as developers
-find and fix bugs, and as user requirements demand new features and capabilities.
-\item Software is frequently built to use other software, leading to complex
-dependencies, and these dependent software packages also frequently change.
-\item Software is generally smaller than data, so a number of the storage and
-preservation constraints on data don't apply to software.
-\item Software teams can be large and multidisciplinary, and with varied roles
-that overlap and sometimes make contributions ambiguous---some of which
-can rise to a level equivalent to paper authorship.
-\item The lifetime of software is generally not as long as that of data.
-\end{itemize}
+
+We made the adaptations because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data.
+In the context of research (e.g., in science), the term ``data'' usually refers to electronic records of observations made in the course of a research study (``raw data'') or to information derived from such observations by some form of processing (``processed data''), as well as the output of simulation or modeling software (``simulated data'').
+Some confusion about the distinction between software and data comes in part from the much wider sense that the term ``data'' has in computing and information science, where it refers to anything that can be processed by a computer.
+In that sense, software is just a special kind of data.
+Because of this, citing software is not the same as citing data.  A more general discussion about these distinctions in currently underway~\cite{software-vs-data}.
 

### 2016-08-21 sc-principles/rebuttal-2/rebuttal.tex
adding point about software vs data

--- 2016-08-17 sc-principles/rebuttal-2/rebuttal.tex
@@
 
-\textbf{Our response:}
-
 \item \emph{``'Software is updated more frequently that papers or data''
@@
 
-\textbf{Our response:}
-
 \item \emph{``'Software suffers from a different type of bit rot than data'':
@@
 
-\textbf{Our response:}
-
 \item \emph{``Software is frequently built to use other software'': True,
@@
 vocabularies.}
-
-\textbf{Our response:}
 
@@
 
-\textbf{Our response:}
-
 \item{``Software teams can be large and multidisciplinary, \ldots'': I think
 this applies to data too.}
-
-\textbf{Our response:}
 
@@
 
-\textbf{Our response:}
+\end{itemize}
 
-\end{itemize}
+\textbf{Our response:} We, and the overall working group, believe there are significant difference
+between software and data, and that these lead to differences in the corresponding citation principles
+as well as their application.
+However, we agree that there has not been sufficient discussion to document these
+differences in a concise, well-structured, and referenced manner.
+For this reason, we have removed the list of differences in the paper and point
+to an ongoing discussion where a large set of community members are working
+to complete this structured documentation with evidence of each point.  An example
+of one difference is that appropriate licenses for software are different than appropriate
+licenses for data (i.e., for open software, an OSI-approved license, while for open data, a CC license).
 

### 2016-08-21 sc-principles/software-citation-principles.tex
minor wording edit, and in -> is, for software vs data

@@
 
-We made the adaptations because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data.
-In the context of research (e.g., in science), the term ``data'' usually refers to electronic records of observations made in the course of a research study (``raw data'') or to information derived from such observations by some form of processing (``processed data''), as well as the output of simulation or modeling software (``simulated data'').
-Some confusion about the distinction between software and data comes in part from the much wider sense that the term ``data'' has in computing and information science, where it refers to anything that can be processed by a computer.
-In that sense, software is just a special kind of data.
-Because of this, citing software is not the same as citing data.  A more general discussion about these distinctions in currently underway~\cite{software-vs-data}.
+We made the adaptations because software, while similar to data in terms of not
+traditionally having been cited in publications, is also different than data.
+In the context of research (e.g., in science), the term ``data'' usually refers
+to electronic records of observations made in the course of a research study
+(``raw data'') or to information derived from such observations by some form
+of processing (``processed data''), as well as the output of simulation or
+modeling software (``simulated data'').
+Some confusion about the distinction between software and data comes in part
+from the much wider scope of the term ``data'' in computing and
+information science, where it refers to anything that can be processed by a
+computer. In that sense, software is just a special kind of data.
+Because of this, citing software is not the same as citing data. A more general
+discussion about these distinctions is currently
+underway~\cite{software-vs-data}.
 

### 2016-08-21 sc-principles/rebuttal-2/rebuttal.tex
fixed formatting of reviewer comment 1

@@
 \begin{itemize}
+{\itshape
 
-\item \emph{``'Software can be used to express or explain concepts, unlike
-data'': I don't understand this.}
+\item ``Software can be used to express or explain concepts, unlike
+data'': I don't understand this.
 
-\item \emph{``'Software is updated more frequently that papers or data''
-(typo: ``''that'' instead of ``than''): This was questioned also by the
+\item ``Software is updated more frequently that papers or data''
+(typo: ``that'' instead of ``than''): This was questioned also by the
 second reviewer, and I don't think the issue is resolved. The footnote is
@@
 maybe only *popular* software tends to be updated frequently, whereas
-*popular* dataset are often not.)}
+*popular* dataset are often not.)
 
-\item \emph{``'Software suffers from a different type of bit rot than data'':
+\item ``Software suffers from a different type of bit rot than data'':
 I am not convinced that this is really a different *type* of bit rot.
 In particular datasets in proprietary formats seem to suffer from exactly the
-same problems.}
+same problems.
 
-\item \emph{``Software is frequently built to use other software'': True,
+\item ``Software is frequently built to use other software'': True,
 though similar things can occur in datasets too, in particular with respect to
 Linked Data, where datasets typically use a number of third-party ontologies and
-vocabularies.}
+vocabularies.
 
-\item \emph{``Software is generally smaller than data'': Another unsupported
+\item ``Software is generally smaller than data'': Another unsupported
 claim, and I am not fully convinced that this is true. There must be a large
@@
 than software. The difference is certainly true on the upper end of the scale
-though: the largest datasets are larger than the largest pieces of software.}
+though: the largest datasets are larger than the largest pieces of software.
 
-\item{``Software teams can be large and multidisciplinary, \ldots'': I think
-this applies to data too.}
+\item ``Software teams can be large and multidisciplinary, \ldots'': I think
+this applies to data too.
 
-\item{``The lifetime of software is generally not as long as that of data.'':
+\item ``The lifetime of software is generally not as long as that of data.'':
 Again unsupported, and again I can imagine this to be false. Many datasets, in
@@
 different and newer dataset, in which case evolving software would live
-longer.}
+longer.
+}
 

### 2016-08-27 sc-principles/software-citation-principles.tex
turning line numbers off
removing extra outline at the end of section 3

@@
 \usepackage{lineno}
-\linenumbers
+%\linenumbers
 
@@
 citation (\S\ref{sec:use_cases}), and review related work
-(\S\ref{sec:related_work}). We also lay out the necessary future work
-(\S\ref{sec:futurework}).
+(\S\ref{sec:related_work}). We also lay out  
+the work needed to lead to these
+software citation principles being applied (\S\ref{sec:futurework}).
 %The principles in this section are written fairly concisely, and discussed
@@
 Katz and Smith~\cite{transitive_credit_original,transitive_credit_json-ld}.
-
-In the next section~(\S\ref{sec:use_cases}), we provide some detailed context in
-which software citation is important, by means of use cases. In
-\S\ref{sec:related_work}, we summarize and analyze a large amount of previous
-work and thinking in this area. In \S\ref{sec:discussion}, we discuss issues
-related to the principles stated in \S\ref{sec:principles}, and finally, in
-\S\ref{sec:futurework} we discuss the work needed to lead to these
-software citation principles being applied.
 

### 2016-09-10 sc-principles/software-citation-principles.tex

@@
 \item \textbf{Credit and Attribution}: \label{principle:credit}
-Software citations should facilitate giving scholarly credit and normative
-and legal attribution to all contributors to the software, recognizing
+Software citations should facilitate giving scholarly credit and normative,
+legal attribution to all contributors to the software, recognizing
 that a single style or mechanism of attribution may not be applicable to
@@
 (e.g., audio and  video lectures).  Research knowledge is embedded in these
-components. And papers and books themselves are also becoming increasingly
+components. Papers and books themselves are also becoming increasingly
 digital, allowing them to become executable and reproducible. As we move towards
@@
 digital products, the characteristics and properties that developed for books
-and papers need to be applied to all digital products and possibly adjusted.
+and papers need to be applied to, and possibly adjusted for,  all digital products.
 Here, we are concerned specifically with the citation of software products. The
@@
 simple names in the full-text, URLs in footnotes, and different kinds of
-mentions in references lists: project names or websites, user manuals,
+mentions in references lists: project names or websites, user manuals, or
 publications that describe or introduce the software~\cite{howison2015jasist}.
@@
 \item Reproducibility: Citation of specific software used is necessary for
-reproducibility, but is not sufficient. Additional information such as
+reproducibility, although not sufficient. Additional information such as
 configurations and platform issues are also needed.
@@
 some changes in the use cases and discussion, although the principles themselves
-were not modified. We also plan to initiate a follow-on working group that will
+were not modified. We also plan to initiate a follow-on implementation working group that will
 work with stakeholders to ensure that these principles impact the research process.
@@
 mechanisms and systems, and the reader should be able to look at any particular example
-of software citation and see if it meets the principles.
+of software citation to see if it meets the principles.
 While we strive to offer practical guidelines that acknowledge the current
@@
 For example, if the version number and release date are not available, the download date can be used.
-And the contact name\slash email is an alternative to the location\slash repository.
+Similarly the contact name\slash email is an alternative to the location\slash repository.
 
@@
 specification (e.g., it is not machine readable) this does offer a solution for
-authors wishing to give explicit instructions to potential citing authors and as
+authors wishing to give explicit instructions to potential citing authors, and as
 noted in the motivation section (\S\ref{sec:intro}), there is evidence that
@@
 
-Perhaps in recognition of the broad range of research domains struggling with
+Perhaps in light of the broad range of research domains struggling with
 the challenge of better recognizing the role of software, funders and agencies
@@
 \emph{SoftwareX}~\cite{softwareX}, \emph{Open Research Computation}~\cite{ORC}, and
-the \emph{Journal of Open Research Software} allow for submissions entirely
+the \emph{Journal of Open Research Software}~\cite{jors} allow for submissions entirely
 focused on research software.
@@
 CRediT~\cite{casrai-credit}, the OpenRIF Contribution Role Ontology~\cite{Gutzman2016}, Ontosoft~\cite{ontosoft}, RRR\slash JISC guidelines~\cite{JISC2015},
-or the terms and classes defined at \href{https://schema.org}{Schema.org} related to the
+or the terms and classes defined at \href{https://schema.org}{schema.org} related to the
 \href{https://schema.org/SoftwareApplication}{\texttt{SoftwareApplication}} class.   In addition,
@@
 However, this is really analogous to how any research builds on other research, where each research product just cites those products that it directly builds on, not those that it indirectly builds on.
-Understanding these chains of knowledge and credit have been part of the history of science field for some time, though more recent work is suggesting more nuanced evaluation of the credit chains~\cite{casrai-credit, transitive_credit_json-ld}.
+Understanding these chains of knowledge and credit have been part of the history of science field for some time, though more recent work suggests more nuanced evaluation of the credit chains~\cite{casrai-credit, transitive_credit_json-ld}.
 
@@
 The Unique Identification principle~(\ref{principle:uid}) calls for ``a method for identification that is machine actionable, globally unique, interoperable, and recognized by a community.''
-What this means for data is discussed in detail in the ``Unique Identification'' section of a report by the FORCE11 Data Citation Implementation Group (DCIG)~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
+What this means for data is discussed in detail in the ``Unique Identification'' section of a report by the FORCE11 Data Citation Implementation Group~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
 This report also lists examples of identifiers that match these criteria including DOIs, PURLs, Handles, ARKS, and NBNs.
@@
 
-We recognize that there are certain situations where it may not be possible for follow the recommended best-practice.
+We recognize that there are certain situations where it may not be possible to follow the recommended best-practice.
 For example, if (1) the software authors did not register a DOI and/or release a specific version, or (2) the version of the
@@
 
-There are at least three different potential relationships between identifiers and versions of software.
+There are at least three different potential relationships between identifiers and versions of software:
 \begin{enumerate}
@@
 If the software is free, the metadata will likely provide an identifier that can be resolved to a URL pointing to the specific version of the software being cited.
-For commercial software, the metadata should still provide information on how to access the specific software, but this may be a company's product number or a link to a web site that allows the software be purchased.
+For commercial software, the metadata should still provide information on how to access the specific software, but this may be a company's product number or a link to a website that allows the software be purchased.
 As stated in the Persistence principle~(\ref{principle:persistence}), we recognize that the software version may no longer be available, but it still should be cited along with information about how it was accessed.
